### 2021-10-01 paper/main.tex
initial draft of paper

\documentclass[runningheads]{llncs}

% \usepackage[numbers]{natbib}
% \renewcommand\bibsection{\section*{\refname}\small\renewcommand\bibnumfmt[1]{##1.}}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts,amssymb}
\usepackage{hyperref}
% \usepackage{xspace}
% \usepackage{amsthm}
% \usepackage{enumitem} % topsep

\usepackage{tikz}
% TikZ libraries
\usetikzlibrary{myautomata}
\usetikzlibrary{fit, backgrounds}

\newenvironment{nstabbing}
  {\setlength{\topsep}{0pt}%
   \setlength{\partopsep}{0pt}%
   \tabbing}
  {\endtabbing}

%{{{ (La)TeX definitions

\newcommand{\Nset}{{\mathbb{N}_0}}

\newcommand{\todo}[1]{\textcolor{red}{#1}}
\def\Inf{\ensuremath{\mathsf{Inf}}}
\def\Fin{\ensuremath{\mathsf{Fin}}}
\def\ttrue{\ensuremath{\mathsf{t}}}
\def\ffalse{\ensuremath{\mathsf{f}}}
\def\CQ{\mathcal{C}}
\def\BigO#1{\ensuremath{\mathcal{O}\left(#1\right)}}
\newcommand{\B}{\texorpdfstring{\ensuremath{\mathds{B}}}{B}}
\def\isempty{\textsc{is\_empty}\xspace}
\def\sccsof{\textsc{SCCs\_of}\xspace}
\def\issccempty{\textsc{is\_SCC\_empty}\xspace}
\def\marksof{\textsc{marks\_of}\xspace}
\def\remove{\textsc{remove}\xspace}

\def\Mocc{M_{\mathrm{occur}}}
\def\Mevery{M_{\mathrm{every}}}
\newcommand{\G}{\mathbf{G}}
\newcommand{\CA}{\ensuremath{\mathcal{A}}}
\newcommand{\CB}{\ensuremath{\mathcal{B}}}
\def\init{\iota}
\def\lang{L} % FIX this
\newcommand{\cM}{\mathcal{M}}
\newcommand{\bscc}{\mathcal{B}}
\newcommand{\sched}{\mathfrak{s}}
\newcommand{\Act}{\mathit{Act}}
\newcommand{\NP}{\textsf{NP}}
\newcommand{\POLY}{\textsf{P}}
\newcommand{\class}{\ensuremath{\mathcal{C}}}
\newcommand{\finally}{\mathsf{F}}
\newcommand{\globally}{\mathsf{G}}
% \newcommand{\finally}{\mathbf{F}}
% \newcommand{\globally}{\mathbf{G}}

% Remove space from footnote after interpunction. Usage:
% \punct{.}\footnote{footnotetext}
\newlength{\spc} % declare a variable to save spacing value
\newcommand{\punct}[1]{%
  \settowidth{\spc}{#1}% set value of \spc variable to the width of #1 argument
  \addtolength{\spc}{-1.8\spc}% subtract from \spc about two (1.8) of its values making its magnitude negative
  #1% print the optional argument
  \hspace*{\spc}% print an additional negative spacing stored in \spc after #1
}

%}}}

\begin{document}

%{{{ Title and authors

\title{Reducing Acceptance Condition\\ of Emerson-Lei Automata}
\authorrunning{T.~Schwarzová and J. Strej\v{c}ek}
 \author{Tereza Schwarzová \and Jan Strej\v{c}ek}
 \institute{%
   Masaryk University, Brno, Czech Republic\\ \email{\{xschwar3,strejcek\}@mail.muni.cz}
 }

%}}}

\maketitle

%{{{ Abstract

\begin{abstract}
  We present an QBF-based algorithm that reduces the number of
  acceptance sets of transition-based Emerson-Lei automata.
\end{abstract}

%}}}

\section{Introduction}
\begin{itemize}
\item Emerson-Lei automata jsou v principu staré, ale v poslední době
  znovuobjevene~\cite{babiak.15.cav} a je o tom ted spousta
  clanku~\cite{baier.19.atva} a nastroju
\item algoritmy jsou obvykle citlive na akceptacni podminku (citace)
\item cilem prace je zjednodusit akceptacni podminku (zejmnena
  redukovat pocet akceptacnich znacek) bez zmeny struktury automatu
\end{itemize}

\section{Preliminaries}
Definice TELA, popsat QBF

\section{Reduction based on acceptance sets relations}
Puvodni telatko

\section{QBF-based reduction}
\begin{itemize}
\item obecny popis konstrukce formule
\item (level 3)
\item level 4
\end{itemize}

\section{Implementation and experimental evaluation}

\section{Conclusions}

%\bibliographystyle{abbrvnat}
\bibliographystyle{plain}
\bibliography{mc}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% End:

### 2021-10-13 paper/main.tex
tacc nejde vytexovat??

@@
 \def\remove{\textsc{remove}\xspace}
+\def\false{\mathit{false}}
+\def\true{\mathit{true}}
 
@@
 \section{Preliminaries}
-Definice TELA, popsat QBF
 
-\section{Reduction based on acceptance sets relations}
-Puvodni telatko
+\subsection{TELA}
+A \emph{transition-based Emerson-Lei automaton} over alphabet $\Sigma$ is a tuple $\mathcal{A} = \left(  \mathcal{Q}, M, \Sigma,q_{0}, \delta, \varphi \right)$, where
+\begin{itemize}
+
+    \item $\mathcal{Q}$ is a finite set of \emph{states},
+    \item $M$ is a finite set of \emph{acceptance marks},
+    \item $\Sigma$ is a finite \emph{alphabet},
+    \item $q_{0}$ is an \emph{initial state},
+    \item $\delta \subseteq \mathcal{Q} \times \Sigma \times 2^{M} \times \mathcal{Q}$ is a transition relation,
+    \item $\varphi$ is the acceptance condition constructed by the following abstract syntax equation, where $m \in M$:
+
+    \begin{center}
+        $\varphi ::= \true \mid \false \mid \Inf(m) \mid \Fin(m) \mid \left( \varphi \land \varphi \right) \mid \left( \varphi \lor \varphi \right)$.\\
+    \end{center} 
+\end{itemize}
+
+A run $\pi$ of $\mathcal{A}$ on word $u = u_{0}u_{1}u_{2}\dots \in \Sigma^{\omega}$ is an infinite sequence of transitions $\pi = (q_{0}, u_{0}, M_{0}, q_{1})(q_{2}, u_{1}, M_{1}, q_{2}) \dots \in \delta^{\omega}$.
+
+Let $M(\pi)$ denote set of all acceptance marks, that occur in run $\pi$ in infinitely many transitions.
+The run satisfies $\Inf(m)$ $\mathit{iff} \ m \in M(\pi)$ and it satisfies $\Fin(m) \ \mathit{iff} \ m \ \not \in M(\pi)$.
+Run $\pi$ is \emph{accepting} if it satisfies the formula $\varphi$. A set of words, which are accepted by $\mathcal{A}$, represent a language $\mathcal{L}(\mathcal{A})$.
+
+Sometimes the set of transitions with the same acceptance mark $m\in M$ is referred to as \emph{acceptance set} $Z(m)$.
+
+\begin{itemize}
+    \item define DNF
+    \item define SCC
+\end{itemize}
+
+\subsection{Quantified Boolean formula}
+\emph{Quantified Boolean formulae} (QBFs) is an extension of propositional logic -- some Boolean variables in the formula can be quantified with \emph{universal} or \emph{existential} quantifiers.
+Formula $\Phi$ is a QBF formula in \emph{prenex normal form} $\iff$ $\Phi = Q_{1}x_{1}...Q_{n}x_{n} .
+\varphi$, where $Q_{i} \in \{\forall, \exists\}$, $x_{i}$ is a quantified Boolean variable and $\varphi$ is quantifier-free Boolean formula.
+
+\smallskip 
+Let $\Phi$ be a QBF.
+Semantically we define the \emph{universal} quantifier as $\forall x .
+\Phi = \Phi [x \rightarrow \true] \land \Phi [x \rightarrow \false]$ and the \emph{existential} quantifier as $\exists x .
+\Phi = \Phi [x \rightarrow \true] \lor \Phi [x \rightarrow \false]$.
+
+Let $X$ be a finite set of variables. Assignment $\mathcal{A}_{X}$ $:X \rightarrow \{\false, \true\}$ is a total mapping of variables defined on set $X$ to Boolean values.
+Let $\Phi$ be a QBF. If there is an assignment $\mathcal{A}_{X}$ that evaluates formula $\Phi$ as $\true$, then $\Phi$ is $\mathit{satisfiable}$ (SAT).
+A set $S \subseteq X$ is a \emph{model} of $\Phi$ if the assignment $\mathcal{A}_{X}$ maps $\true$ to every element of $S$.
+If there is no such assignment that evaluates the formula as $\true$, we say the formula is $\mathit{unsatisfiable}$ (UNSAT).
+
+\section{Reduction based on acceptance sets relations / SCC-based reduction}
+In this section, the simplification strategy is to find relations between acceptance sets and reduce the redundant ones. SCC-based simplification is the coarsest simplification we propose.
+
+Let $\mathcal{A}$ be an original automaton and let $\psi$ be an acceptance condition of $\mathcal{A}$.
+Every run of $\mathcal{A}$ has an infinite suffix that takes place within one SCC $S_{i}$.
+Thus the evaluation of $\psi$ depends purely on the SCC $S_{i}$ and we can optimize the acceptance condition for each SCC separately.
+:w
+This optimization consists of removing redundant terms from acceptance condition and relabeling of acceptance marks on the transitions.
+The state and transition structure of the SCC does not change.
+This way we obtain set of simplified acceptance conditions that we merge into new acceptance formula $\psi'$.
+The automaton with relabeled transitons and new acceptance condition $\psi'$ we denote $\mathcal{A}'$.
+Finally, we ensure that automaton $\mathcal{A'}$ is equivalent to the original automaton $\mathcal{A}$.
+
+\subsection{$S_{i}$ simplification}
+At first, we remove the unused acceptance marks and transform the SCC's acceptance condition $\psi_{i}$ to \emph{disjunctive normal form} (DNF), so we have a unified shape of the condition.
+If TELA acceptance condition is in DNF, two distinct terms in one disjunct can only appear in three possible forms:
+
+\begin{itemize}
+    \item $\Inf\tacc{2}{i} \land \Inf\tacc{4}{j}$
+    \item $\Fin\tacc{8}{i} \land \Fin\tacc{4}{j}$
+    \item $\Fin\tacc{8}{i} \land \Inf\tacc{4}{j}$
+\end{itemize}
+
+Since formula $\psi_{i}$ is in DNF, we can represent it as a set of disjuncts $\overline{\psi_{i}} = \{D_1, D_2, \ldots D_k\}$.
+
+Let $D_{k} \in \overline{\psi_{i}}$ be a disjunct of formula $\overline{\psi_{i}}$ and $C_{j} \in D_{k}$ a conjunct of disjunct $D_{k}$.
+Furthermore let  $\tacc{0}{i}, \tacc{4}{j}, \tacc{8}{k}, \tacc{2}{l} \in M$ be distinct acceptance marks that occur in $\mathcal{A}$.
+
+In the next section, some properties of an automaton allow us to substitute a Boolean value $\true$ or $\false$ for a particular term of the formula $\psi_i$.
+The consequences of this substitution are divided into a number of cases. We represent it on the set-format of formula we just defined.
+
+\begin{itemize}
+    \item If $C_{j}$ is substituted by $\true$, the conjunct is omitted from $D_{k}$. Thus $D_{k} = D_{k} \setminus \{C_{j}\}$.
+    \item If $C_{j}$ is substituted by $\false$, the conjunct causes that the evaluates of the whole $D_{k}$ is $\false$. Thus $\overline{\psi_{i}} = \overline{\psi_{i}} \setminus \{D_{k}\}$.
+
+\end{itemize}
+
+That being said, we can define the reduction techniques.
 

### 2021-10-13 paper/main.tex
paper| telatko section

@@
     \item define DNF
-    \item define SCC
+
 \end{itemize}
+
+\subsection{Strongly connected component}
+Let us consider a TELA $\mathcal{A} = \left(  \mathcal{Q}, M, \Sigma,q_{0}, \delta, \varphi \right)$.
+\emph{Strongly connected component} (SCC) of $\mathcal{A}$ is each maximal set of states $S \subseteq Q$ such that any two states $q_{1}, q_{2} \in S$ are connected by a sequence of transitions.
+Let us enumerate the SCCs of $\mathcal{A}$.
+Set of all transitions of the SCC $S_{i} \text{ is } \delta_{S_{i}} = \delta \cap (S_{i} \times \Sigma \times 2^M \times S_{i})$.
+A run of SCC $S_{i}$ is a run of automaton $\mathcal{A}$ with the infinite suffix contained within $S_{i}$.
 
@@
 \subsection{$S_{i}$ simplification}
+\label{section:simplification}
 At first, we remove the unused acceptance marks and transform the SCC's acceptance condition $\psi_{i}$ to \emph{disjunctive normal form} (DNF), so we have a unified shape of the condition.
@@
     \item If $C_{j}$ is substituted by $\true$, the conjunct is omitted from $D_{k}$. Thus $D_{k} = D_{k} \setminus \{C_{j}\}$.
-    \item If $C_{j}$ is substituted by $\false$, the conjunct causes that the evaluates of the whole $D_{k}$ is $\false$. Thus $\overline{\psi_{i}} = \overline{\psi_{i}} \setminus \{D_{k}\}$.
+    \item If $C_{j}$ is substituted by $\false$, the conjunct causes that the whole $D_{k}$ evaluates to $\false$. Thus $\overline{\psi_{i}} = \overline{\psi_{i}} \setminus \{D_{k}\}$.
 
@@
 That being said, we can define the reduction techniques.
+
+\subsubsection{Acceptance formula modifications\\} 
+\label{subsection:acc_modif}
+Since we optimize for each SCC separately, the acceptance formula corresponding to SCC $S_{i}$ can contain terms with acceptance marks that are not present on the transitions of $S_{i}$.
+Let $\tacc{8}{k} \in M$ be an acceptance mark that is not present on any edge of $S_{i}$.
+Then any term that contains this acceptance mark in $\overline{\psi_{i}}$ can be immediately substituted with a boolean value and thus removed from $\overline{\psi_{i}}$.
+\medskip
+
+Therefore acceptance mark $\tacc{8}{k}$ is not visited in any run of $S_{i}$ and thus every term in form $\Fin\tacc{8}{k}$ is substituted with $\true$ and every term $\Inf\tacc{8}{k}$ is substituted with $\false$.
+
+On the contrary, let $\tacc{4}{j}$ be an acceptance mark that is present on every transition of $S_{i}$.
+That means that at least one transition with $\tacc{4}{j}$ is visited infinitely often by a run of $\mathcal{S_{i}}$.
+Thus every term $\Fin\tacc{4}{j}$ is substituted by $\false$ and every term in form $\Inf\tacc{4}{j}$ is substituted by $\true$.
+
+\subsubsection{$\Inf$ conjuncts\\}
+\label{subsection:simpl_inf}
+Reduction of a conjunct of $\Inf\tacc{8}{k}$ form  is based on the inclusion of the sets of edges labeled with $\tacc{8}{k}$ and set of edges labeled with $\tacc{4}{j}$.
+If all transitions that contain an acceptance mark $\tacc{8}{k}$ also contain $\tacc{4}{j}$, we can remove $\Inf\tacc{4}{j}$ from all disjuncts, where it occurs together with $\Inf\tacc{8}{k}$.
+More formally, if the following conditions hold:
+\begin{itemize}
+    \item $Z(\tacc{8}{k}) \subseteq Z(\tacc{4}{j})$
+    \item $\Inf\tacc{8}{k}, \Inf\tacc{4}{j} \in D_{k}$
+\end{itemize}
+then we can we can substitute $\Inf\tacc{4}{j}$ with Boolean value $\true$. 
+This modification does not change the language because it does not affect accepting runs of $S_{i}$.
+The transitions can create an accepting run that has the potential to satisfy $D_{k}$ only if they are labeled with both  $\tacc{8}{k}$  and $\tacc{4}{j}$.
+Since $Z(\tacc{8}{k}) \subseteq Z(\tacc{4}{j})$, one can notice that that if a run satisfies $\Inf\tacc{8}{k}$ then  $\Inf\tacc{4}{j}$ is satisfied as well.
+Therefore if we remove $\Inf \tacc{4}{j}$ from $\overline{\psi_{i}}$, this modification does not affect the language.
+
+\subsubsection{$\Fin$ conjuncts\\}
+Similarly as in previous section, this reduction is based on inclusion.
+If following conditions hold:
+\begin{itemize}
+    \item $Z(\tacc{8}{k}) \subseteq Z(\tacc{4}{j})$
+    \item $\Fin\tacc{8}{k}, \Fin\tacc{4}{j} \in D_{k}$
+\end{itemize}
+then we can we can substitute $\Fin\tacc{8}{k}$ with boolean value $\true$.
+This modification does not change the language because from if $\tacc{4}{j}$ is visited finitely often, then also $\tacc{8}{k}$ is visited finitely often.
+\medskip
+
+We can merge two conjuncts into one if the conjuncts always occur in the same disjuncts in $\Fin$ form.
+We can do so regardless of the position of the acceptance marks that are contained in the conjuncts.
+If the following condition is met:
+\begin{itemize}
+    \item $\forall D_{k}: \Fin\tacc{8}{k} \in D_{k} \Longleftrightarrow \Fin\tacc{4}{j} \in D_{k}$
+\end{itemize}
+We introduce a fresh acceptance mark $\tacc{2}{l}$, such that $Z(\tacc{2}{l}) = Z(\tacc{8}{k}) \cup Z(\tacc{4}{j})$.
+ Finally, we can update $D_{k}$:
+\begin{align*}
+    D_{k} = (D_{k} \setminus \{ \Fin\tacc{8}{k}, \Fin\tacc{4}{j} \}) \cup \{\Fin\tacc{2}{l}\}
+\end{align*}
+
+This update basically means that $\Fin\tacc{8}{k}, \Fin\tacc{4}{j}$ are substituted by $\true$ and new conjunct $\Fin\tacc{2}{l}$ is added to $D_{k}$.
+The acceptance mark $\tacc{2}{j}$ is placed on every edge of the SCC $S_{i}$ that is labeled with $\tacc{8}{k}$ or $\tacc{4}{j}$.
+This modification does not change the language because it does not change the acceptance runs of $S_{i}$. It only relabels the acceptance marks on edges that can be seen finitely often.
+
+\subsubsection{$\Fin \land \Inf$ conjuncts\\}
+There are two cases when we can reduce one of the pair of $\Inf\tacc{0}{i}$ and $\Fin\tacc{2}{l}$ conjuncts.
+
+At first, if every transition labeled with  $\tacc{0}{i}$ is also labeled with $\tacc{2}{l}$, the conjunction of $\Inf\tacc{0}{i} \land \Fin\tacc{2}{l}$ is never $\true$ because every run that visits $\tacc{2}{l}$ also visits $\tacc{0}{i}$.
+Therefore we can reduce the whole disjunct $D_{k}$.
+If the following conditions are met:
+\begin{itemize}
+    \item $Z(\tacc{0}{i}) \subseteq Z(\tacc{2}{l})$
+    \item $\Inf\tacc{0}{i}, \Fin\tacc{2}{l} \in D_{k}$
+\end{itemize}
+We substitute the whole disjunct $D_{k}$ for $\false$. In other words $\overline{\psi_{i}} = \overline{\psi_{i}} \setminus \{D_{K}\}$.
+
+Finally we can simplify $\overline{\psi_{i}}$, if the edges labeled with $\tacc{0}{i}$ and edges labeled with $\tacc{2}{l}$ are complementary.
+More formally if:
+\begin{itemize}
+    \item  $\Inf\tacc{0}{i}$ and $\Fin\tacc{2}{l} \in D_{k}$
+    \item $Z(\tacc{0}{i}) \cap Z(\tacc{2}{l}) = \emptyset \land Z(\tacc{0}{i}) \cup Z(\tacc{2}{l}) = \delta_{S_{i}} $
+\end{itemize}
+Then we can substitute $\true$ for $\Inf\tacc{0}{i}$.
+This reduction does not change the language because if during a run of $S_{i}$ all transitions labeled with $\tacc{2}{l}$ are visited only finitely often, then necessarily at least one transition labeled with $\tacc{0}{i}$ is visited infinitely often.
+Therefore removing $\Inf\tacc{0}{i}$ from $\overline{\psi_{i}}$ does not affect the language, since the validity of $\Fin\tacc{2}{l}$ implies the validity of $\Inf\tacc{0}{i}$.
+
+\subsection{Merge of $\psi_{i}$}
+After applying reduction techniques from the previous section, each $S_{i}$ has its own acceptance condition $\psi_{i}$.
+The goal of this section is to merge these acceptance conditions into one formula $\psi'$ with emphasis on obtaining a formula with the minimal number of acceptance sets.
+
+At first, the algorithm finds $\psi_{i}$ with the greatest number of disjuncts and states it as a base of the new acceptance condition $\psi'$.
+Now the algorithm continues by the successive merging of $\psi'$ with the unmerged formula $\psi_{i}$ that has the greatest number of disjuncts.
+This repeats until all formulae $\psi_{i}$ are merged.
+This process updates the form of $\psi'$ until it reaches its final form.
+The process of merging two formulae consists of two phases.
+In the first phase a suitable pairing of disjuncts is found and in the second phase, these disjuncts are merged.
+
+\begin{enumerate}
+	\item Let $\psi'$ denote the future acceptance condition of $\mathcal{A}'$ and $\psi_{i}$ is the acceptance condition of $S_{i}$ chosen to be merged.
+Then the algorithm uses \emph{linear sum assignment} to determine which disjunct of $\psi_{i}$ is merged with which disjunct of $\psi'$.
+The process of using \emph{linear sum assignment} and the pairing procedure is described in detail in work by T.
+Šťastná \cite{ts_simpl}.
+    \item Let $D_{K} \in \overline{\psi_{i}}$ be a disjunct paired with $D_{L} \in \overline{\psi'}$.
+All conjuncts of $D_{L}$ are initially labeled as unused.
+Now the algorithm maps every conjunct $C_{k} \in D_{K}$ to a suitable conjunct $C_{l} \in D_{L}$ and labels $C_{l}$ as used.
+\footnote{The indexes differ to emphasize the fact that they are not equal.} A conjunct $C_{l}$ is suitable for $C_{k}$ is it is on the same type ($\Fin$ or $\Inf$), it is labeled as unused.
+If no suitable $C_{l}$ is found, the algorithm adds $C_{k}$ to disjunct $D_{L}$ and marks it as used.
+\end{enumerate}
+
+In an acceptance mark $m_{l} \in M$ occurs in $\psi'$ in more than one conjunct, we need to check whether all of these conjuncts are mapped to conjuncts form $\psi_{i}$ with the same acceptance mark $m_{k} \in M$.
+If not, we resolve this conflict by replacing all additional occurrences of $m_{l}$ with fresh acceptance mark $m_{n}$ and we place $m_{n}$ on the exact transition of $\mathcal{A}'$ where $m_{l}$ is placed.
+Finally, in the end, the algorithm removes every acceptance mark that is not present in $\psi'$ from the edge of $\mathcal{A}$.
+
+\medskip
+
+\begin{example}
+Consider $\overline{\psi'} =  \{ \{ \Fin\tacc{0}{0}, \Inf\tacc{1}{1} \} , \{\Fin\tacc{0}{0} \} \}$ \\ and $\overline{\psi_{i}} = \{  \{\Fin\tacc{4}{4} \}, \{ \Fin\tacc{2}{2}, \Inf\tacc{3}{3} \} \} $.
+We obtain the pairing $(1,2), (2,1)$, meaning that the first disjunct of $\overline{\psi_{i}}$ is merged wit the second disjunct of $\overline{\psi'}$ and second disjunct of $\overline{\psi_{i}}$ is merged with the first disjunct of $\overline{\psi'}$.
+Then $\Fin\tacc{0}{0}$ is mapped to $\Fin\tacc{2}{2}$, $\Inf\tacc{1}{1} $ is mapped to $\Inf\tacc{3}{3}$ and $\Fin\tacc{0}{0}$ is again mapped to $\Fin\tacc{4}{4}$ which causes a conflict.
+Therefore we need to replace the second occurrence of $\tacc{0}{0}$ with fresh mark $\tacc{5}{5}$.
+The set-representation of formula $\psi'$ after merging is $\psi' =  \{ \{ \Fin\tacc{0}{0}, \Inf\tacc{1}{1} \} , \{\Fin\tacc{5}{5} \} \}$.
+
+\end{example}
+
+Unfortunately, sometimes when the optimal mapping was not found, the procedure produces an automaton with a more complex acceptance condition than the acceptance condition of the input automaton.
+Therefore we propose and implement an optimization, which prevents such a behavior and in general, produces smaller acceptance condition than the original one.
+
+\begin{itemize}
+    \item The selection of base of $\psi'$ is based on two keys.
+The primary key is cardinality (number of clauses) of $\overline{\psi_{i}}$, the secondary key is cardinality of clauses (number of terms in clause).
+    \item We order the disjunct in every $\overline{\psi_{i}}$ in descending order, where the key is the cardinality of a particular disjunct.
+    \item The DNF conversion of the input acceptance condition can lead to an exponential blowup of the formula, where some acceptance marks occur more than once.
+Since the reduction of acceptance sets is basically denoting the redundant ones as $\true$ or $\false$, we can as well easily obtain simplified acceptance formula in CNF (we simply denote the same acceptance sets as $\true$ or $\false$ as we did in DNF).
+According to the shape of the input acceptance condition, we choose the suitable (shorter) normal form and perform the merging algorithm on it.
+\end{itemize}
+By choosing the CNF over DNF in the cases where CNF is more natural (shorter), we prevent the formula to contain some acceptance marks more than once (at least, we do not create them by conversion of $\psi$ to DNF).
+This way, we prevent adding new acceptance marks to the formula when resolving conflicts of acceptance marks which leads to better results.
+If the original formula contains an acceptance mark that is present in more than one disjunct, then by ordering the disjuncts in every $\overline{\psi_{i}}$, the \emph{linear sum assignment} returns more convenient pairing.
+Meaning that if \emph{linear sum assignment} finds two equal assignments, it respects the order and the disjuncts with the  acceptance mark that occurs more than once are paired with disjuncts that also contain acceptance mark that occurs more than once.
+
+\begin{example}
+	Consider an automaton $\mathcal{A}$ in Figure \ref{fig:L1_before}.
+In this example, we demonstrate the simplification procedure described in Section \ref{section:simplification} and obtain simplified acceptance conditions $\psi_{i}$ for each SCC $S_{i}$.
+The simplified acceptance conditions $\overline{\psi_{i}}$ are displayed in the set-format we introduced earlier because we make use of it afterward during the merge.
+Then we merge these formulae into the acceptance condition of the simplified automaton $\mathcal{A}'$.
+We enumerate the SCCs $S_{1}, S_{2}, S_{3}$.
+(The indices correspond with the numbers inside the states in Figure \ref{fig:L1_before}.) And we assign an accepting condition $\psi_{1}$ to $S_{1}$, $\psi_{2}$ to $S_{2}$ and $\psi_{3}$ to $S_{3}$.
+    \begin{enumerate}
+	    \item We simplify $\psi_{1}$ according to the placement of acceptance marks on the edges in $S_{1}$.
+At first we notice, that the formula $\overline{\psi_{1}}$ contains acceptance marks that are not present on the edges of $S_{1}$.
+Therefore modify $\overline{\psi_{1}}$ as described in Subsection \ref{subsection:acc_modif}.
+We substitute $\true$ for $\Fin\tacc{3}{3}$ and $\false$ for $\Inf\tacc{1}{1}$ and $\Inf\tacc{0}{0}$.
+(This substitution deletes the whole disjunct $D_{1}$.) Then we notice that the acceptance marks are in a position which allows us to perform the simplification described in first part of subsection 3.1.4.
+We substitute $\true $ for $\Inf\tacc{4}{4}$ and obtain the final form of $\overline{\psi_{1}} = \{\{ \Fin\tacc{2}{2}\} \}$.
+	    \item Similarly as in previous case, we notice that acceptance marks $\tacc{2}{2}$, $\tacc{3}{3}$ and $\tacc{4}{4}$ are not present on the edges of $S_{2}$.
+Therefore we substitute $\true$ for both $\Fin\tacc{2}{2} \text{ and } \Fin\tacc{3}{3}$ and $\false$ for $\Inf\tacc{4}{4}$ (and thus remove the whole disjunct $D_{2}$).
+Then we use the procedure described in Subsection \ref{subsection:simpl_inf} and remove $\Inf\tacc{1}{1}$ from the disjunct $D_{1}$ and obtain $\overline{\psi_{2}} = \{ \{\Inf\tacc{0}{0}\} \}$.
+        \item Finally, we yet again remove marks that are not present in $S_{3}$ and perform the simplification described in subsection 3.1.3.
+We substitute $\true $ for $\Fin\tacc{2}{2} \text{ and } \Fin\tacc{3}{3}$ and add new term $\Fin\tacc{5}{5}$ into $D_{2}$ and place $\tacc{5}{5}$ on the edges that are labeled with $\tacc{2}{2}$ or $\tacc{3}{3}$.
+The simplified acceptance condition is $\overline{\psi_{3}} = \{\{ \Fin\tacc{5}{5}, \Inf\tacc{4}{4}\}\}$.
+    \end{enumerate}
+
+	Now we proceed to the merge of the formulae of the SCCs $\psi_{i}$.
+As a base of the new acceptance condition $\psi'$ we choose the formula $\psi_{3}$ because it has the highest number of disjuncts compared to the other formulae $\psi_i$, so $\overline{\psi'}=\{ \{ \Fin\tacc{5}{5}, \Inf\tacc{4}{4} \}\}$.
+Now we merge  $\psi_{1}$ and $\psi_{2}$ with $\psi'$.
+The order is in this particular case irrelevant because $\overline{\psi_{1}}$ has the same number of disjuncts as $\overline{\psi_{2}}$ (and both disjuncts have the same number of conjuncts).
+So we merge $\overline{\psi_{1}}$ at first because of its lower index.
+Since both $\overline{\psi'}$ and $\overline{\psi_{1}}$ have only one disjunct, the only possible pairing that \emph{linear sum assignment
+    } can give us is $(0,0)$, meaning that $\{ \Fin\tacc{2}{2}\}$ is merged with $\{ \Fin\tacc{5}{5}, \Inf\tacc{4}{4}\}$.
+Then we need to map an unused conjunct  $\Fin\tacc{5}{5}$ to a conjunct of the same type which is $\Fin\tacc{2}{2}$.
+Similarly in case of $\overline{\psi_{2}}$ we map $\Inf\tacc{4}{4}$ on $\Inf\tacc{0}{0}$.
+This way we produce $\mathcal{A}'$ displayed in Figure  \ref{fig:L1_middle} with the acceptance condition $ \Fin\tacc{5}{5} \land \Inf\tacc{4}{4}$.
+Observe that the automaton $\mathcal{A}'$ is not equivalent to the original $\mathcal{A}$.
+For example in Figure \ref{fig:L1_middle} is the highlighted loop not accepting but in the original automaton in Figure \ref{fig:L1_before} the loop accepting is.
+To reach the equivalence between the two automata, we need to \emph{restore equivalence} which is described in the next section.
+\end{example}
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+			\node[flabel] at (2,5) {$( \Inf\tacc{0}{0} \land \Inf\tacc{1}{1}) \lor (\Fin\tacc{2}{2} \land \Fin\tacc{3}{3} \land \Inf\tacc{4}{4})$};
+
+			\node[state, initial] (x) at (0,0) {0};
+			\node[state] (y) at (2,2) {1};
+			\node[state] (z) at (2,-1.5) {2};
+			\node[state] (w) at (5,-1.5) {3};
+
+			\path[->]
+				(x) edge[]  (y)
+				(x) edge[]  (z)
+				(z) edge[] (w)
+
+            	(z) edge[loop, in = 75, out = 105, looseness=6] (z)
+            	(z) edge[loop, in = 72, out = 108, looseness=9] (z)
+            	(z) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc1] {1}(z)
+            	(z) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc0, pos=0.6] {0} node[accset,collacc1, pos=0.4] {1}(z)
+
+            	(w) edge[loop, in = 75, out = 105, looseness=6] (w)
+            	(w) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc3] {3}(w)
+            	(w) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc2] {2}(w)
+            	(w) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4}(w)
+
+            	(y) edge[loop, in = 75, out = 105, looseness=6] node[accset,collacc4] {4}(y)
+            	(y) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc4] {4}(y)
+            	(y) edge[loop, in = 70, out = 110, looseness=12] node[accset,collacc2] {2} (y)
+            	(y) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc2] {2}(y)
+
+			;
+		\end{tikzpicture}
+    \caption{TELA before simplification Level 1}
+    \label{fig:L1_before}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+			\node[flabel] at (2,5) {$ \Fin\tacc{5}{5} \land \Inf\tacc{4}{4}$};
+
+			\node[state, initial] (x) at (0,0) {0};
+			\node[state] (y) at (2,2) {1};
+			\node[state] (z) at (2,-1.5) {2};
+			\node[state] (w) at (5,-1.5) {3};
+
+			\path[->]
+				(x) edge[]  (y)
+				(x) edge[]  (z)
+				(z) edge[] (w)
+
+            	(z) edge[loop, in = 75, out = 105, looseness=6] (z)
+            	(z) edge[loop, in = 72, out = 108, looseness=9] (z)
+            	(z) edge[loop, in = 70, out = 110, looseness=12]  (z)
+            	(z) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4} (z)
+
+            	(w) edge[loop, in = 75, out = 105, looseness=6] (w)
+            	(w) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc5] {5}(w)
+            	(w) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc5] {5}(w)
+            	(w) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4}(w)
+
+            	(y) edge[loop, in = 75, out = 105, looseness=6, color=red] (y)
+            	(y) edge[loop, in = 72, out = 108, looseness=9] (y)
+            	(y) edge[loop, in = 70, out = 110, looseness=12] node[accset,collacc5] {5} (y)
+            	(y) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc5] {5}(y)
+
+			;
+		\end{tikzpicture}
+    \caption{TELA after simplification and merging in Level 1}
+    \label{fig:L1_middle}
+\end{figure}
+
+\subsection{Restoring equivalence}
+Replacing the input acceptance condition with $\psi'$ in $\mathcal{A}$ might change the language, so it is no longer equivalent to the language recognized by the original automaton $\mathcal{A}$. 
+ By now, each SCC $S_{i}$ is adjusted to the acceptance condition $\psi_{i}$ which might be different than $\psi'$. 
+ Therefore, for each SCC we need to restore equivalence with the new acceptance condition $\psi'$.
+ The idea is that the only terms of $\psi'$ that can affect accepting runs on SCC $S_{i}$ are the ones that have been mapped on some terms in $\psi_{i}$.
+From now on, we refer to them as \emph{used terms}.
+We need to make sure that the other terms (the ones that have not been mapped on any term of $\psi_{i}$) can't affect any accepting run of $\mathcal{A}'$.
+From now on, we call these terms \emph{unused terms}.
+We influence the evaluation of the  \emph{unused terms} in the following way:
+\begin{itemize}
+     \item If we want to make term in the $\Fin$ form always $\false$, we place it's acceptance mark on every edge of the SCC.
+     \item If we want to make term in the $\Inf$ form always $\true$, we place it's acceptance mark on every edge of the SCC.
+\end{itemize}
+ Keep in mind that by now, any acceptance mark that is not in $\psi'$ is not present on any edge of $\mathcal{A}'$.
+Therefore any $\Fin$ term is already always $\true$ if its acceptance mark is not present on any edge of the SCC.
+Dually, $\Inf$ term is already always $\false$ if it is not present on any edge of the SCC.
+\medskip
+ Now we distinguish the cases when $\psi'$ is in DNF or $\psi'$ is in CNF. 
+\begin{itemize}
+     \item If $\psi'$ is in DNF, we need to make $\false $ every disjunct that does not contain any \emph{used term}.
+Further, every \emph{unused term} that occurs in the same disjunct as any \emph{used term} needs to be made $\true$.
+     \item Dually, if $\psi'$ is in CNF, every conjunct that does not contain any \emph{used term} needs to be made $\true$ and every \emph{unused term} that occurs in a conjunct with any \emph{used term} needs to be made $\false$.
+
+\end{itemize}
+
+\begin{example}
+	Consider the example from the Figure \ref{fig:L1_middle}, where the DNF acceptance condition $\psi' = (\Fin\tacc{5}{5} \lor \Inf\tacc{4}{4})$ and the acceptance condition of SCC $S_{1}$ $\psi_{1} = (\Fin\tacc{2}{2})$.
+The term $\Fin\tacc{5}{5}$ is mapped to $\Fin\tacc{2}{2}$.
+Then we need to enforce that the \emph{unused term} $\Inf\tacc{4}{4}$ has no effect on the evaluation of $\psi'$ by the run with the infinite suffix contained within the SCC $S_{1}$.
+Otherwise the result will not be equivalent (see the highlited loop in Figure \ref{fig:L1_middle}).
+Therefore we make term $\Inf\tacc{4}{4}$ $\true$ by adding $\tacc{4}{4}$ on every edge of the SCC $S_{i}$.
+The result can be seen in Figure \ref{fig:L1_done}.
+\end{example}
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+			\node[flabel] at (2,5) {$ \Fin\tacc{5}{5} \land \Inf\tacc{4}{4}$};
+
+			\node[state, initial] (x) at (0,0) {0};
+			\node[state] (y) at (2,2) {1};
+			\node[state] (z) at (2,-1.5) {2};
+			\node[state] (w) at (5,-1.5) {3};
+
+			\path[->]
+				(x) edge[]  (y)
+				(x) edge[]  (z)
+				(z) edge[] (w)
+
+            	(z) edge[loop, in = 75, out = 105, looseness=6] (z)
+            	(z) edge[loop, in = 72, out = 108, looseness=9] (z)
+            	(z) edge[loop, in = 70, out = 110, looseness=12]  (z)
+            	(z) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4} (z)
+
+            	(w) edge[loop, in = 75, out = 105, looseness=6] (w)
+            	(w) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc5] {5}(w)
+            	(w) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc5] {5}(w)
+            	(w) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4}(w)
+
+            	(y) edge[loop, in = 75, out = 105, looseness=6] node[accset,collacc4] {4} (y)
+            	(y) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc4] {4}(y)
+            	(y) edge[loop, in = 70, out = 110, looseness=12] node[accset,collacc4, pos=0.4] {4} node[accset,collacc5, pos=0.6] {5}(y)
+            	(y) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4, pos=0.4] {4} node[accset,collacc5, pos=0.6] {5}(y)
+
+			;
+		\end{tikzpicture}
+    \caption{Equivalent TELA after simplification Level 1}
+    \label{fig:L1_done}
+\end{figure}
 

### 2021-10-13 paper/main.tex
abstract added!

@@
 \begin{abstract}
-  We present an QBF-based algorithm that reduces the number of
-  acceptance sets of transition-based Emerson-Lei automata.
+  The formalism called \emph{transition-based Emerson-Lei automata
+    (TELA)} generalizes many traditional kinds of automata over
+  infinite words including Büchi, co-Büchi, Rabin, Streett, and parity
+  automata. Transitions in TELA are marked with acceptance marks and
+  accepting formulas are boolean combinations of terms saying that a
+  particular mark has to be visited infinitely or finitely often in an
+  accepting run. Algorithms processing these automata can be very
+  sensitive to the number of different acceptance marks. We introduce
+  two techniques reducing the number of acceptance marks in a TELA
+  without altering its structure and the represented language. One
+  technique represents a standard approach based on relations among
+  acceptance marks in individual strongly connected components of the
+  automaton. The other technique constructs quantified boolean formula
+  (QBF) queries that ask a QBF solver for an acceptance condition with
+  fewer acceptance marks and the placement of these marks in the
+  automaton structure. Both techniques are implemented and experiments
+  show that the number of acceptance marks in TELA produced by
+  state-of-the-art tools ltl3tela, Rabinizer~4, and SPOT can be
+  sometimes reduced. In the case of Rabinizer~4, we reduced the
+  cummulative number of acceptance marks to less than one half.
 \end{abstract}

### 2021-10-13 paper/main.tex
folding marks added
zacatek intra, opravy v bibu

@@
 
-%{{{ Title and authors
+%{{{ title and authors
 
-\title{Reducing Acceptance Condition\\ of Emerson-Lei Automata}
+\title{Reducing Acceptance Marks \\ in Emerson-Lei Automata}
 \authorrunning{T.~Schwarzová and J. Strej\v{c}ek}
@@
 
-%{{{ Abstract
+%{{{ abstract
 
@@
 
+%{{{ introduction
+
 \section{Introduction}
+
+Automata over infinite words (also known as $\omega$-automata) are now
+used mainly in specification, verification, analysis, monitoring, and
+synthesis of various systems with infinite behaviour. The first and
+the most common kind of $\omega$-automata is Büchi
+automaton~\cite{buchi.60.clmps} introduced in 1960. Other kinds of
+$\omega$-automata have been defined \ldots
+
+\todo{Honza je tu!}
+
+Emerson-Lei automata~\cite{emerson.87.scp} were
+introduced Automata over infinite words are studied for decades and
+used in many areas of computer science including verification,
+controller synthesis, m
+
+Emerson-Lei automata~\cite{emerson.87.scp} were introduced in 
 \begin{itemize}
@@
 \end{itemize}
+
+%}}}
+%{{{ preliminaries
 
@@
 If there is no such assignment that evaluates the formula as $\true$, we say the formula is $\mathit{unsatisfiable}$ (UNSAT).
+
+%}}}
+%{{{ standard reductions
 
@@
 The process of using \emph{linear sum assignment} and the pairing procedure is described in detail in work by T.
-Šťastná \cite{ts_simpl}.
+Šťastná \cite{stastna.19.bc}.
     \item Let $D_{K} \in \overline{\psi_{i}}$ be a disjunct paired with $D_{L} \in \overline{\psi'}$.
@@
 
+%}}}
+%{{{ QBF-based reduction
+
 \section{QBF-based reduction}
@@
 
+%}}}
+%{{{ implementation and experimental evaluation
+
 \section{Implementation and experimental evaluation}
 
+%}}}
+%{{{ conclusion
+
 \section{Conclusions}
+
+%}}}
 

### 2021-10-14 paper/main.tex
Intro

--- 2021-10-13 paper/main.tex
@@
   acceptance marks in individual strongly connected components of the
-  automaton. The other technique constructs quantified boolean formula
-  (QBF) queries that ask a QBF solver for an acceptance condition with
-  fewer acceptance marks and the placement of these marks in the
-  automaton structure. Both techniques are implemented and experiments
-  show that the number of acceptance marks in TELA produced by
-  state-of-the-art tools ltl3tela, Rabinizer~4, and SPOT can be
-  sometimes reduced. In the case of Rabinizer~4, we reduced the
+  automaton. The other technique constructs \emph{quantified boolean
+    formula (QBF)} queries that ask a QBF solver for an acceptance
+  condition with fewer acceptance marks and the placement of these
+  marks in the automaton structure. Both techniques are implemented
+  and experiments show that the number of acceptance marks in TELA
+  produced by state-of-the-art tools ltl3tela, Rabinizer~4, and SPOT
+  can be sometimes reduced. In the case of Rabinizer~4, we reduced the
   cummulative number of acceptance marks to less than one half.
@@
 
+%{{{ introduction
+
 \section{Introduction}
+
+% Automata over infinite words (also known as $\omega$-automata) are now
+% used in specification, verification, analysis, monitoring, and
+% synthesis of various systems with infinite behaviour. The first and
+% the most common kind of $\omega$-automata is Büchi
+% automaton~\cite{buchi.60.clmps} introduced in 1960. As deterministic
+% Büchi automata are less expressive than nondeterministic ones,
+% deterministic $\omega$-automata are typically considered with other
+% acceptance condition like Rabin, Streett, Muller, or parity. In 1987,
+% Emerson and Lei~\cite{emerson.87.scp} introduced $\omega$-automata
+% where the acceptance condition can be an arbitrary combination of
+% acceptance primitives saying that a certain set of states should be
+% visited finitely often or infinitely often. In 2015, the same kind of
+% acceptance condition was implemented in the \emph{Hanoi omega-automata
+%   format (HOAF)}~\cite{babiak.15.cav}, with the difference that finiteness
+
+In 1987, Emerson and Lei~\cite{emerson.87.scp} introduced automata
+over infinite words where acceptance conditions are arbitrary
+combinations of acceptance primitives saying that a certain set of
+states should be visited finitely often or infinitely often. In 2015,
+the same kind of acceptance condition was described in the \emph{Hanoi
+  omega-automata format (HOAF)}~\cite{babiak.15.cav}. The only
+difference is that the acceptance primitives talk about finitely or
+infinitely often visited acceptance marks rather than sets of states.
+Acceptance marks are placed on transitions and each mark identifies
+the set of transitions containing this mark. Hence, these automata are
+called \emph{transition-based Emerson-Lei automata (TELA)} and they
+generalize many traditional kinds of automata over infinite words
+including Büchi, co-Büchi, Rabin, Streett, Muller or parity
+automata.
+
+TELA have attracted a lot of attention during the last five years.
+Their popularity comes probably from the fact that these automata can
+often use less states than equivalent automata with traditional
+acceptance conditions. Further, algorithms handling TELA can
+automatically handle all automata with traditional acceptance
+conditions. TELA automata can be obtained for example by translating
+LTL formulas by Delag~\cite{muller.17.gandalf},
+ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav}, or
+SPOT~\cite{duret.16.atva2}. There are also some algorithms processing
+these automata, for example the emptiness check~\cite{baier.19.atva}
+or translation of TELA to parity automata~\cite{renkin.20.atva}.
+
+\todo{Nechcem do experimentu pridat Delag?}
+
+Algorithms processing TELA are often sensitive to the number of
+acceptance marks more than to other parts of the automaton. For
+example, the transformation of TELA to parity automata based on
+\emph{color appearance record}~\cite{renkin.20.atva} transforms a TELA
+with $s$ states and $m$ acceptance marks into a parity automaton with
+up to $s\cdot m!$ states. Further, the emptiness
+check~\cite{baier.19.atva} is exponential in the number of acceptance
+marks that appear in acceptance primitives saying that a mark has to
+be visited finitely often, while it is only polynomial in other
+measures of the input automaton.
+
+These observations motivate our study of techniques reducing the
+number of acceptance marks without altering the structure of the
+automaton. We present basically two reduction techniques. The first
+one is based on relations among individual transition sets defined by
+various acceptance marks in each \emph{strongly connected component
+  (SCC)}. This technique can be seen as an extension of the
+\emph{SCC-base simplifications of acceptance
+  conditions}~\cite{babiak.13.spin} designed for
+\emph{transition-based generalized Büchi automata (TGBA)} and
+implemented in SPOT. Probably the main difference between the original
+technique for TGBA and the one for TELA is in the final step when
+simplified acceptance conditions for each SCC are merged together into
+one reduced acceptance condition.
+
+The second technique is completely different. For a given TELA and
+parameters $k,n$, it produces a \emph{quantified boolean formula
+  (QBF)} which is satisfiable if and only if there exists an
+equivalent automaton with the identical structure, $n$ acceptance
+marks, and an acceptance formula in \todo{conjunctive?}  normal form
+with $k$ clauses. The placement of the marks on transitions and the
+acceptance formula can be obtained from a model of the
+formula. Besides this formula, we describe also the construction of a
+simpler formula such its satisfiability implies existence of an
+equivalent automaton with the identical structure and $n$ acceptacne
+marks, but not vice versa.
+
+We have implemented both reduction techniques in a tool called
+\texttt{telatko}. We show that the tool can reduce acceptance marks in
+automata produced by ltl3tela~\cite{major.19.atva},
+Rabinizer~4~\cite{kretinsky.18.cav}, and SPOT~\cite{duret.16.atva2}.
+While the reduction is relatively modest for TELA produced by ltl3tela
+and SPOT, it is substantial for automata produced by Rabinizer~4.
+
+\subsubsection{Related results}
+We are not aware about any existing research focused on simplification
+of acceptance formulas of TELA or reduction of the number of its
+acceptance marks without increasing the number of states.
+
+Casares, Colcombet, and Fijalkow very recently introduced a structure
+called \emph{alternating cycle decomposition
+  (ACD)}~\cite{casares.21.icalp} which compactly represents the
+information about all accepting and non-accepting automata cycles.  We
+expect that ACD could potentialy be used to reduce the number of
+acceptance marks or to simplify the acceptance condition. However,
+such a reduction is not obvious. Moreover, the construction of ACD
+seems to be computationally expensive.
+
+\subsubsection{Structure of the paper} The next section introduces
+basic terms used in the paper. Section~\ref{sec:standard} presents the
+reduction technique based on relations among individual acceptance
+marks on each SCC. The QBF-based technique is explained in
+Section~\ref{sec:qbf}. Implementation is briefly described in
+Section~\ref{sec:experiments}, which also shows the experimental
+results. Finally, Section~\ref{sec:conclusion} suggests other
+applications of our QBF-reduction technique and closes the paper.
+
+%}}}
+%{{{ preliminaries
+
+\section{Preliminaries}\label{sec:prelim}
+
+In this section we recall the basic terms around TELA and QBF.
+
+\begin{definition}[TELA]
+  A \emph{transition-based Emerson-Lei automaton (TELA)} is a tuple
+  $\mathcal{A} =(Q,M,\Sigma,q_{0},\delta,\varphi)$, where
   \begin{itemize}
-\item Emerson-Lei automata jsou v principu staré, ale v poslední době
-  znovuobjevene~\cite{babiak.15.cav} a je o tom ted spousta
-  clanku~\cite{baier.19.atva} a nastroju
-\item algoritmy jsou obvykle citlive na akceptacni podminku (citace)
-\item cilem prace je zjednodusit akceptacni podminku (zejmnena
-  redukovat pocet akceptacnich znacek) bez zmeny struktury automatu
-\end{itemize}
-
-\section{Preliminaries}
-
-\subsection{TELA}
-A \emph{transition-based Emerson-Lei automaton} over alphabet $\Sigma$ is a tuple $\mathcal{A} = \left(  \mathcal{Q}, M, \Sigma,q_{0}, \delta, \varphi \right)$, where
-\begin{itemize}
-
-    \item $\mathcal{Q}$ is a finite set of \emph{states},
+  \item $Q$ is a finite set of \emph{states},
   \item $M$ is a finite set of \emph{acceptance marks},
   \item $\Sigma$ is a finite \emph{alphabet},
-    \item $q_{0}$ is an \emph{initial state},
-    \item $\delta \subseteq \mathcal{Q} \times \Sigma \times 2^{M} \times \mathcal{Q}$ is a transition relation,
-    \item $\varphi$ is the acceptance condition constructed by the following abstract syntax equation, where $m \in M$:
+  \item $q_{0}\in Q$ is an \emph{initial state},
+  \item $\delta \subseteq Q \times \Sigma \times 2^{M} \times Q$ is a
+    \emph{transition relation}, and
+  \item $\varphi$ is the \emph{acceptance condition} constructed by
+    the following abstract syntax equation, where $m$ ranges over $M$:
 
@@
 \end{itemize}
+\end{definition}
+
+\todo{Honza je tu a jde spat (je 5 rano pryc).}
 
@@
 
-\section{Reduction based on acceptance sets relations / SCC-based reduction}
+%}}}
+%{{{ standard reductions
+
+\section{Reduction based on acceptance marks relations}\label{sec:standard}
+
 In this section, the simplification strategy is to find relations between acceptance sets and reduce the redundant ones. SCC-based simplification is the coarsest simplification we propose.
@@
 
-\section{QBF-based reduction}
+%}}}
+%{{{ QBF-based reduction
+
+\section{QBF-based reduction}\label{sec:qbf}
+
 \begin{itemize}
@@
 
-\section{Implementation and experimental evaluation}
+%}}}
+%{{{ implementation and experimental evaluation
 
-\section{Conclusions}
+\section{Implementation and experimental evaluation}\label{sec:experiments}
+
+%}}}
+%{{{ conclusion
+
+\section{Conclusions}\label{sec:conclusion}
+
+%}}}
 

### 2021-10-14 paper/main.tex
some fixes in intro

@@
 parameters $k,n$, it produces a \emph{quantified boolean formula
-  (QBF)} which is satisfiable if and only if there exists an
-equivalent automaton with the identical structure, $n$ acceptance
-marks, and an acceptance formula in \todo{conjunctive?}  normal form
-with $k$ clauses. The placement of the marks on transitions and the
+  (QBF)} which is satisfiable if and only if there exists an automaton
+with the same structure, $n$ acceptance marks, an acceptance formula
+in \todo{conjunctive?}  normal form with $k$ clauses, and the same set
+of accepting runs. The placement of the marks on transitions and the
 acceptance formula can be obtained from a model of the
@@
 simpler formula such its satisfiability implies existence of an
-equivalent automaton with the identical structure and $n$ acceptacne
-marks, but not vice versa.
+automaton with the same structure, $n$ acceptance marks, and the same
+set of accepting runs, but not vice versa.
 

### 2021-10-14 paper/main.tex
small fixes and note in conclusion

--- 2021-10-14 paper/main.tex
@@
 The second technique is completely different. For a given TELA and
-parameters $k,n$, it produces a \emph{quantified boolean formula
-  (QBF)} which is satisfiable if and only if there exists an
-equivalent automaton with the identical structure, $n$ acceptance
-marks, and an acceptance formula in \todo{conjunctive?}  normal form
-with $k$ clauses. The placement of the marks on transitions and the
+parameters $c,m$, it produces a \emph{quantified boolean formula
+  (QBF)} which is satisfiable if and only if there exists an automaton
+with the same structure, $m$ acceptance marks, an acceptance formula
+in disjunctive normal form with $c$ clauses, and the same set
+of accepting runs. The placement of the marks on transitions and the
 acceptance formula can be obtained from a model of the
@@
 simpler formula such its satisfiability implies existence of an
-equivalent automaton with the identical structure and $n$ acceptacne
-marks, but not vice versa.
+automaton with the same structure, $m$ acceptance marks, and the same
+set of accepting runs, but not vice versa.
 
@@
 
+chceme sem napsat, ze QBF-based reduction se da pouzit napr. k
+nalezeni akceptacni formule s $c$ klauzulemi a $m$ znackami v CNF,
+pripadne k nalezeni ekvivalentniho paritniho automat s danou maximalni
+paritou $p$ a stejnou strukturou.
+
 %}}}

### 2021-10-14 paper/main.tex
QBF z BP

@@
 \usetikzlibrary{fit, backgrounds}
+\usetikzlibrary{extshapes}
 
@@
 
+The aim of this section is to create a QBF-formula for given $n$ that guarantees the existence of equivalent acceptance condition of a given automaton with $n$ acceptance marks. 
+\smallskip
+
+Let $\mathcal{A}$ with the acceptance condition $\psi$ be an original automaton on the input.
+Let $\mathcal{A}'$ denote automaton with simplified acceptance condition $\psi'$ equivalent to the original automaton $\mathcal{A}$.
+We encode $\psi$ and the placement of acceptance marks into QBF  $\varphi$.
+
+Let $\varphi$ be a QBF formula in \emph{prenex normal form}. Formula $\varphi$
+is composed of quantified prefix $\overrightarrow{Q}$ and three subformulae $\varphi_{O}, \varphi_{N} $ and $\varphi_{C}$.
+The exact shape of the formula is $\varphi = \overrightarrow{Q} .
+\  \varphi_{C} \implies ( \varphi_{O} \Longleftrightarrow \varphi_{N})$.
+Subformula $\varphi_{O}$ expresses properties of $\psi$.
+Subformula $\varphi_{N}$ expresses properties of $\psi'$, namely the number of acceptance marks and a shape of $\psi'$.
+Before we describe the role of $\varphi_{C}$, we need to introduce \emph{relevant set}.
+
+\smallskip
+
+Evaluation of a run of $\mathcal{A}$ is dependent on set of edges that are contained infinitely within the run.
+The set of these edges is from now on referred to as \emph{relevant set}, since only these edges are relevant for the evaluation of $\psi$.
+The \emph{relevant set} is a subset of $\delta_{S_{i}}$ for some SCC $S_{i}$ and any two edges from the \emph{relevant set} are connected by a sequence of edges that also belong to the \emph{relevant set}.
+The purpose of subformula $\varphi_{C}$ is to describe some properties of \emph{relevant sets}.
+For each level of simplification, $\varphi_{C}$ describes these properties more specifically.
+Naturally, the more we specify these properties, the more complex the formula $\varphi_{C}$ is.
+
+\smallskip
+
+Afterwards the formula $\varphi$ is created, we use a QBF-solver to find a model of $\varphi$.
+If no such model exists ($\varphi$ is UNSAT), we know that automaton $\mathcal{A}'$ with $n$ acceptance marks and equivalent to $\mathcal{A}$ does not exist.
+Otherwise, we use the model of $\varphi$ for decoding the automaton $\mathcal{A}'$.
+The structure of $\mathcal{A}'$ remains the same as the structure of $\mathcal{A}$.
+The only thing we change is the acceptance condition $\psi'$ and the placement of acceptance marks on the edges.
+
+\subsection{Motivation}
+Based on the fundamentals given in the previous section, the simplification Level 1 is based on a relationship between acceptance sets of an automaton.
+To achieve finer simplification levels, the relation between the inner structure of the automaton and its acceptance condition needs to be taken into consideration.
+
+As the acceptance condition and the inner structure differ for every specific automaton, finding algorithmic ways of reducing acceptance marks would be complicated.
+Instead, we encode the given automaton into a QBF-formula and let the solver find the redundant marks for us.
+
+\subsection{Encoding}
+ In this section we propose encoding of the relevant parts of a given automaton into QBF-formula.
+\subsubsection{Intuition}
+The goal of the procedure is to find acceptance condition for a particular automaton with less acceptance marks.
+Whilst doing that, we need to ensure that the new automaton stays equivalent to the original automaton because the algorithm not only changes the acceptance condition it also changes the placement of acceptance marks on the edges.
+To do so, we need to ensure that every run on $\mathcal{A}'$ is evaluated the same way as run on $\mathcal{A}$ (these two automata have the same structure).
+
+Since we work with transition-based automata, the evaluation of an acceptance condition is dependent on the edges.
+Intuitively we want formula $\varphi$ to say that for every \emph{relevant set}, the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
+Unfortunately, to identify the exact edges that belong to the \emph{relevant set} of a run is not trivial, since we need to take into consideration every infinite suffix of every run on $\mathcal{A}$.
+Instead we let the formula $\varphi$ say that for every \emph{selection of edges} the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
+For now the \emph{selection of edges} can contain any edges of the automaton.
+Later we restrict the \emph{selection of edges} as we force it to have some properties of \emph{relevant set}.
+
+To express this in the QBF formula, we need to introduce variables that represent edges of the automaton.
+
+Let us enumerate set of edges $\delta$.
+Every edge of the automaton is represented by variable $e_{i}$, where the index corresponds with the enumeration, thus $e_{i}$ represents the edge $t_{i} \in \delta$.
+If such a variable is evaluated as $\true$, it means that the edge is contained in the \emph{selection of edges}.
+Analogously if a variable $e_{i}$ is evaluated as $\false$, it means that the edge is not contained in the \emph{selection of edges}.
+
+\smallskip
+
+If we want to consider every possible \emph{selection of edges}, we need to consider every possible evaluation of every variable that represents an edge.
+
+\todo{tohle bych odstranila}
+\begin{example}
+   Let $t_{1}, t_{2} \in \delta$ be the edges of the automaton from the Figure \ref{fig:elval_of_edges}.
+Since the automaton has two edges, we get four possible evaluation of variables representing the edges.
+    \begin{enumerate}
+        \item $e_{1}$ is $\true \land e_{2}$ is $\true \Longleftrightarrow t_{1}, t_{2}$ are in the \emph{selection of edges}.
+        \item $e_{1}$ is $\true \land e_{2}$ is $\false \Longleftrightarrow t_{1}$ is in the \emph{selection of edges} and $t_{2}$ is not in the \emph{selection of edges}.
+        \item $e_{1}$ is $\false \land e_{2}$ is $\true \Longleftrightarrow t_{1}$ is not in the \emph{selection of edges} and $t_{2}$ is in the \emph{selection of edges}.
+        \item $e_{1}$ is $\false \land e_{2}$ is $\false \Longleftrightarrow t_{1}, t_{2}$ are not in the \emph{selection of edges}.
+    \end{enumerate}
+\end{example}
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+			\node[state, initial] (z) at (0,0) {};
+
+			\path[->]
+
+            	(z) edge[loop, in = 70, out = 110, looseness=6] node [above, scale=0.7] {1} (z)
+            	(z) edge[loop, in = 68, out = 112, looseness=10] node [above, scale=0.7] {2} (z)
+			;
+		\end{tikzpicture}
+    \caption{TELA with enumerated edges}
+    \label{fig:elval_of_edges}
+\end{figure}
+
+Hereby the expressive power of quantified Boolean formula comes to use.
+If we universally quantify every variable $e_{i}$ that represents an edge $t_{i} \in \delta$, then we consider the situation when $t_{i}$ is contained in the \emph{selection of edges} and the situation when $t_{i}$ is not contained in the \emph{selection of edges}.
+
+That being said, we can proceed to the encoding of some information about automaton into the QBF-formula $\varphi$.
+As mentioned before, the formula $\varphi$ expresses that for every \emph{selection of edges}, the acceptance condition $\psi$ is evaluated as $\true$ $\mathit{iff}$ the acceptance condition $\psi'$ is evaluated as true.
+
+The subformula $\varphi_{O}$ encodes the shape of the acceptance condition $\psi$ and information which edges are labeled with which acceptance marks.
+The evaluation of $\varphi_{O}$ depends purely on the evaluation of variables that represent edges of the automaton.
+It is encoded in such way that the subformula $\varphi_{O}$ is satisfied iff the edges of \emph{selection of edges} satisfy $\psi$.
+The details of $\varphi_{O}$ encoding are described in Subsection \ref{subsection:original}.
+
+\smallskip
+
+The subformula $\varphi_{N}$ contains information about the number of acceptance marks we want in the new acceptance formula $\psi'$.
+Further, it contains free variables.
+The model of these free variables gives us the information about the shape of $\psi'$ and the placement of acceptance marks on the edges of $\mathcal{A}'$.
+The details of $\varphi_{N}$ encoding are described in Subsection \ref{subsection:new}.
+
+The quantification of all variables that represent the edges of the automaton ensures that we consider every possible \emph{selection of edges}, as described in Section \ref{subsection:quant_pref}. 
+\begin{center}
+    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . ( \varphi_{O} \Longleftrightarrow \varphi_{N} ) $
+\end{center}
+However, one can clearly notice that some edges will never be visited infinitely often in the same run.
+(For example, when some of the edges occur in different strongly connected components than the others.) In other words, these edges will never belong to any \emph{relevant set} together.
+Therefore it is possible to eliminate every evaluation of QBF, where variables representing such edges are evaluated as $\true$.
+It is natural to express this constrain as an implication.
+Let $\varphi_{C}$ be a formula describing that the \emph{selection of edges} has a potential be a \emph{relevant set}.
+The QBF representation is:
+\begin{center}
+    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . \varphi_{C} \implies (\varphi_{O} \Longleftrightarrow \varphi_{N})$
+\end{center}
+The matter of eliminating various combinations of edges is described in more detail in Subection \ref{subsection:connection}.
+
+\subsubsection{Variables}
+In this section, we describe the syntax and intended semantics of variables, parameters, and auxiliary letters used in this section.
+Hereby is the overview of the parameters used when building the formula:
+\begin{itemize}
+    \item K -- The number of acceptance marks in $\psi'$.
+    \item C -- the number of clauses in formula $\psi'$.
+\end{itemize}
+
+These parameters are derived from the original automaton $\mathcal{A}$. As the algorithm iterates (see Subection \ref{subsection:procedure} )  the parameter $K$ decreases.
+
+\smallskip
+
+The auxiliary letters have the following function:
+\begin{itemize}
+    \item $k$ -- the number of particular acceptance mark
+    \item $c$ -- the number of particular clause
+    \item $i$ -- the number of particular edge 
+\end{itemize}
+
+Hereby is the overview of the variables used when building the QBF formula $\varphi$ :
+\begin{itemize}
+    \item $$f_{i,k} = \begin{cases}
+    1& \text{edge }t_{i} \text{ is labeled with the acceptance mark }k\\
+    0& \text{else}
+\end{cases}$$
+\item $$p_{c,k} = \begin{cases}
+    1& \Inf(k) \text{ is in the clause }c\\
+    0& \text{else}
+\end{cases}$$
+\item $$n_{c,k}= \begin{cases}
+    1& \Fin(k) \text{ is in the clause }c\\
+    0& \text{else}
+\end{cases}$$
+\item $$e_{i} \;= \begin{cases}
+    1& \text{The edge } t_{i} \text{ is in the \emph{selection of edges} }\\
+    0& \text{else}
+\end{cases}$$
+\end{itemize}
+
+\subsection{QBF decoding}
+After we build $\varphi$, we use QBF-solver to find satisfying assignment of variables.
+If such a valuation exists, we can build a new automaton $\mathcal{A}'$ with the same structure and fewer acceptance marks.
+This new automaton is equivalent to the original one $\mathcal{A}$ -- this is guaranteed by the QBF-formula $\varphi$.
+There are two possible outcomes from the QBF-solver:
+\begin{enumerate}
+    \item $\varphi$ is UNSAT -- a model of $\varphi$ does not exist.
+That means that the acceptance condition of $\mathcal{A}$ cannot be simplified for given C and K.
+Thus we return automaton $\mathcal{A}$ and the information that $\mathcal{A}$ is the automaton with the simplest acceptance condition possible.
+    \item $\varphi$ is SAT -- QBF-solver provides (returns) a model of unquantified variables.
+Namely to $p_{c, k},n_{c,k} \text{ and } f_{i,k}, \text{ for each }1 \leq c \leq C, 1 \leq k \leq K, 1 \leq i \leq |\delta|$.
+    Firstly, we decode the variables $f_{i,k}$ and use them to identify new placement of the acceptance marks.
+Specifically, we create a new automaton $\mathcal{A}'$ with the same structure as $\mathcal{A}$.
+Then  we add acceptance mark $\tacc{5}{k}$ on edge $t_{i}$ for every $f_{i,k}$ variable evaluated as $\true$.
+
+    The new acceptance condition of automaton $\mathcal{A}'$ is then build from variables $p_{c,k}$ and $n_{c,k}$.
+The second part of $\varphi_{N}$ was encoded to produce new acceptance formula in DNF.
+(See the alternation between Boolean operators  in order $\lor, \land$ in Section \ref{subsection:new}.) Thus the shape of the new acceptance condition is set as DNF skeleton, and we simply add terms according to the evaluation of variables.
+If $p_{c,k}$ is $\true$, we add term $\Inf\tacc{5}{k}$ to the disjunct $c$.
+Similarly, if $n_{c,k}$ is $\true$, then we add $\Fin\tacc{5}{k}$ to the disjunct $c$.
+This way, we process every variable that QBF-solver gives us and produce the new automaton $\mathcal{A}'$
+    equivalent to the original one.
+
+\end{enumerate}
+
+\subsection{Quantified prefix}
+\label{subsection:quant_pref}
+The aim of the quantified prefix in formula $\varphi$ is to consider every possible selection of edges.
+As mentioned in the previous sections, edge $t_{i} \in \delta$ belongs to the \emph{selection of edges} if the variable $e_{i}$ that represents the edge $t_{i}$ is evaluated as $\true$.
+Therefore if we want to consider every possible \emph{selection of edges}, we need to consider every evaluation of the variables representing edges.
+We use a universal quantifier to express this.
+The universal quantifier states that the formula has to hold for both evaluations of the quantified variable.
+
+The QBF prefix that expresses the universal quantification for every variable representing an edge has the following shape:
+\begin{align*}
+    \overrightarrow{Q} = \forall\{e_{1}, \ldots ,e_{|\delta|} \}
+\end{align*}
+
+The advantage of QBF is that it is able to express a statement in shorter formula then if we tried to express the statement in standard Boolean formula with the use of conjunction.
+
+\begin{example}
+Consider the QBF $\theta = \forall e_{1} \forall e_{2} .
+(e_1 \lor e_2) $.
+The QBF $\theta $ transformed to standard Boolean formula is $\theta' = (e_{1} \lor e_{2}) \land (  e_{1} \lor \lnot e_{2}) \land  (  \lnot e_{1} \lor e_{2}) \land  (  \lnot e_{1} \lor \lnot e_{2}) .$
+\end{example}
+
+\subsection{Original acceptance formula}
+\label{subsection:original}
+
+As mentioned before, the satisfiability of $\varphi_{O}$ is purely resultant of assessment of $e_{i}, 1 \leq i \leq |\delta|$ variables representing edges in automaton.
+
+Let us enumerate the edges of the original automaton $\mathcal{A}$ and let $Z(m), m \in M$ be the set of edges labeled with mark $m$. 
+The encoding of original acceptance condition terms into $\varphi_{O}$ is defined as follows:
+%%davat tam rovnitko nebo nejak sipku??
+\begin{align*}
+    &\varphi(\Inf(m))=\bigvee\limits_{t_{i} \in Z(m)} e_{i}\\
+    &\varphi(\Fin(m)) = \bigwedge\limits_{t_{i} \in Z(m)} \lnot e_{i}
+\end{align*}
+\begin{example}
+
+Consider the automaton in Figure \ref{fig:oficial_acc}.
+The set $Z_{0} = \{t_3, t_2\},\, Z_{1} = \{t_1\}$, where $t_{1}, t_{2}, t_{3} \in \delta$.
+Thus the encoded $\varphi_{O} = e_1 \land (\lnot e_2 \land \lnot e_3)$.
+Therefore if variable $e_{1}=\true$, $e_{2}=\false$ and $e_{3}=\false$, acceptance formula is $\true$.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+\node[cstate, initial] (g) {};
+\node[flabel, above] at (0, 1.5) {$\Fin\taccsq{0}{0}\land\Inf\tacc{1}{1}$};
+%
+\path[->]
+    (g) edge[loop, in=330, out=30, looseness=6]node[right, pos=0.4, scale=0.8] {$1$} node[accset,collacc1, pos=0.6] {1}(g)
+	(g) edge[loop, in=80, out=140, looseness=6] node[above, pos= 0.4, scale=0.8] {$2$} node[accsquare, collacc0, pos=0.6]{0}(g)
+	(g) edge[loop, in=220, out=280, looseness=6] node[below, pos=0.6, scale=0.8] {$3$} node[accsquare, collacc0, pos=0.4] {0} (g)
+	;
+\end{tikzpicture}
+    \caption{TELA with edges labeled with their numbers}
+    \label{fig:oficial_acc}
+\end{figure}
+
+\end{example}
+
+\subsection{New acceptance formula}
+\label{subsection:new}
+The new acceptance formula $\varphi_{N}$ determines the shape of the future acceptance condition of automaton, and the assignment of variables used in $\varphi_{N}$ determines new positions of acceptance marks on the transitions.
+
+The formula $\varphi_{N}$ is divided into two parts.
+The first part of the formula (4.1) expresses a simple constrain that the acceptance set that occurs in one clause cannot be in $\Inf$ and simultaneously in $\Fin$ form.
+
+The second part (4.2) of $\varphi_{N}$ is divided into two subformulas.
+The first subformula expresses that if the term $\Inf(k)$ is present in clause $c$, then at least one edge of the \emph{selection of edges} needs to be labeled with the acceptance mark $k$.
+
+Similarly, the second subformula implies that if the term $\Fin(k)$ is present in clause $c$, then none of the \emph{selected edges} can be labeled with the acceptance mark $k$.
+
+Using the variables presented in previous sections, we encode $\varphi_{N}$:
+\begin{align}
+    \varphi_{N} = & \bigwedge\limits_{1 \leq c \leq C}\;\bigwedge\limits_{1 \leq k \leq K} (\lnot p_{c,k} \lor \lnot n_{c,k}) \; \land \\
+    &\bigvee\limits_{1 \leq c \leq C} \; \bigwedge\limits_{1 \leq k \leq K} \Bigg[\big(  p_{c, k} \implies \bigvee\limits_{t_{i} \in \delta}(f_{i,k} \land e_{i})  \big) \land \\& \;\;\;\quad\quad\quad\quad\quad\quad\big(  n_{c, k} \implies \bigwedge\limits_{t_{i} \in \delta}(\lnot f_{i,k} \lor \lnot e_{i})  \big)\Bigg]
+\end{align}
+
+Observe the alternation of $\lor \land$ from the second part of the formula (4.2). This alternation determines that the formula $\psi'$ is in DNF.
+
+\subsection{Connection formula}
+\label{subsection:connection}
+The connection formula distinguishes the remaining three levels of simplification. Previously we denoted it as $\varphi_{C}$.
+
+Formula $\varphi_{C}$ defines requirements on the edges that are together in \emph{selection of edges}.
+These requirements are based on some properties of these edges.
+The more restrictive these requirements are, the more properties of \emph{relevant sets} are possessed by the \emph{selection of edges}.
+In other words, formula $\varphi_{C}$ holds if the \emph{selection of edges} has the required property.
+If not, the formula $\varphi_{C}$ is $\false$, and then the whole formula $\varphi$ is trivially satisfied.
+Formula $\varphi_{C}$ consists of two parts.
+
+The first part expresses the requirement that the \emph{selection of edges} is \emph{continuous}.
+Loosely said, this means that if a transition from \emph{selection of edges} enters some state $q$ then another transition that belongs to the \emph{selection of edges} exits the state $q$.
+This holds in the opposite direction as well.
+Thus if a transition from \emph{selection of edges} exits some state $q$ then another transition that belongs to the \emph{selection of edges} enters the state $q$.
+This is naturally expressed by equivalence.
+Note that if the selected transition is a self-loop, the destination and the source of the edge is the same state.
+
+The Boolean formula for the described statement:
+\begin{center}
+    $\varphi_{C_{1}}=\bigwedge\limits_{q \in Q} \Bigg( ( \bigvee\limits_{t_{i} \in \delta \cap \{ q \} \times  \Sigma \times  2^M \times Q} e_i ) \Longleftrightarrow( \bigvee\limits_{t_{j} \in \delta  \cap \ Q \times  \Sigma \times 2^M \times \{q\}} e_{j} ) \Bigg)$
+\end{center}
+
+\begin{example}
+   Consider the state $q$ in Figure \ref{fig:inNout} this part of formula is $(e_1 \lor e_2) \Longleftrightarrow (e_3 \lor e_4 \lor e_5)$.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+        \node[state] (q) at (0, 0) {$q$};
+
+	    \path[->]
+		(q) edge node[above, pos=0.4, scale=0.8] {$4$} ++(1.5, 0)
+		(q) edge node[above, pos=0.4, scale=0.8] {$5$} ++(1.3, 1)
+		(q) edge node[above, pos=0.4, scale=0.8] {$3$} ++(1.3, -1)
+	;
+	\path[<-]
+	(q) edge node[above, pos=0.6, scale=0.8] {$2$} ++(-1.3, -0.5)
+	(q) edge node[above, pos=0.6, scale=0.8] {$1$} ++(-1.3, 0.5)
+	;
+    \end{tikzpicture}
+    \caption{State $q$ and its entering and exiting edges}
+    \label{fig:inNout}
+\end{figure}
+\end{example}
+
+Even though we ensured that the selected transitions create continuous paths, the paths might not be connected.
+
+\begin{example}
+Consider the automaton in Figure \ref{fig:continuousNotConnected}.
+The transitions $t_1$ and $t_2$ are loops (the simplest cycles), but are not connected (each is located in different SCC).
+Thus $\{ t_{1}, t_{2}\}$ is not a \emph{relevant set}.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+			\node[cstate, initial] (i) at (0, 0) {};
+			\node[cstate] (j) at (2.5, 1) {};
+			\node[cstate] (k) at (2.5, -1) {};
+			\path[->]
+
+			(i) edge  (k)
+			(i) edge (j)
+			(j) edge[loop right, draw=red] node[right] {$1$} (j)
+			(k) edge[loop right, draw=red] node[right] {$2$} (k)
+
+		;
+    \end{tikzpicture}
+    \caption{Not connected cycles}
+    \label{fig:continuousNotConnected}
+\end{figure}
+\end{example}
+
+The second part of $\varphi_{C}$ specifies the properties of \emph{relevant set} that the edges that belong together to \emph{selected edges} need to have.
+We divided these properties into two levels Level 2, Level 3, Level 4, all described in ensuing sections.
+
+\subsubsection{SCC accuracy}
+The SCC accuracy corresponds with the Level of simplification 2. At this level, the formula specifies that the edges from one \emph{selection of edges} belong to the same SCC.
+ Thus straightforwardly translated into QBF formula:
+\begin{center}
+    $\varphi_{C_{2}} = \bigvee\limits_{\delta_{S_{i}} \subseteq \delta}\Bigg( ( \bigvee\limits_{t_{i} \in \delta_{S_{i}}} e_i ) \land  ( \bigwedge\limits_{t_{j} \in \delta \setminus \delta_{S_{i}}} \lnot e_j  ) \Bigg)$
+\end{center}
+
+Finally, the connection formula is the conjunction of the two parts:
+\begin{center}
+    $\varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}$.
+\end{center}
+
+\begin{example}
+This example works as a motivation for simplification Level 3.
+For simplification Level 2, edges number $1$ and number $3$ in Figure \ref{fig:connected_motivation} can be in the same \emph{selection of edges} because they are placed in the same SCC.
+Obviously, since they are not connected, they cannot form \emph{relevant set}.
+Therefore we can specify a new level of simplification -- Level 3, which forces all edges from the \emph{selection of edges} to be connected.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+    \node[state, initial] (g) {$0$};
+    \node[state, right=2cm of g](q){$1$};
+    %
+    \path[->, auto]
+    (g) edge[bend right] node[below, scale=0.8] {$2$}(q)
+    (q) edge[bend right]node[above, scale=0.8] {$4$} (g)
+    (q) edge[loop, in=65, out=115, looseness=10, draw=red] node[ scale=0.8] {$3$}(q)
+    (g) edge[loop, in=65, out=115, looseness=10, draw=red] node[scale=0.8] {$1$}(g)
+	;
+
+\end{tikzpicture}
+
+    \caption{Motivation for simplification L3}
+    \label{fig:connected_motivation}
+\end{figure}
+
+\end{example}
+
+\subsubsection{CYCLES accuracy}
+The simplification Level 3 ensures that \emph{selection of edges} is connected, but we can go one step further and specify that \emph{selection of edges} is in fact a \emph{relevant set}.
+
+The intuition by simplification Level 4 is that the states of an automaton are again divided into two groups -- the states that belong to \emph{selected states} and the states do not belong to \emph{selected states}.
+We consider three possible scenarios that can occur while the states are divided into these two groups:
+\begin{enumerate}
+    \item The source of edge $t_{i}$ is in different group than the destination of edge $t_{i}$.
+    \item Source and destination of edge $t_{i}$ are in the \emph{selected states}.
+    \item Source and destination of edge $t_{i}$ are not in \emph{selected states}.
+\end{enumerate}
+Cases two and three do not violate the requirement that all connected states need to be in one group.
+Case one is a little more complicated.
+If an edge has a source in \emph{selected states} and destination not in the \emph{selected states}, we need to secure that the sequence of the edges returns to \emph{selected states}.
+In other words, there exists another edge that has a source in the not \emph{selected states} and destination in the \emph{selected states}.
+(Dually the other way around.)
+
+In order to express this in the QBF-formula we introduce a new set of quantified variables $\{ w_{1}, \ldots , w_{|Q|}$. The variables have the following semantics:
+\begin{align*}
+    w_{q} = \begin{cases}
+    1& \text{state } q \in Q \text{ is in the \emph{selected group}}\\
+    0& \text{else}
+    \end{cases}
+\end{align*}
+Since we want to ensure that for every \emph{selection of edges} (for every evaluation of variables that represent the edges) there exist such a groups we update the quantified prefix  $\overrightarrow{Q}$ in following way:
+\begin{align*}
+    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
+\end{align*}
+
+\smallskip
+
+Since we want the statement to hold for every \emph{selection of edges}, we place the existentially quantified variables $\{w_{1}, \ldots, w_{|Q|} \}$ behind the universally quantified variables $\{e_{1}, \ldots, e_{|\delta|} \} $.
+At first we update the quantified prefix:
+\begin{align*}
+    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
+\end{align*}
+
+And then we can create the $\varphi_{C}$ formula:
+\begin{align*}
+    \varphi_{C} =  \Bigg[ &\bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (w_{q} \land w_{r}) \Bigg) \lor \\
+    &  \bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma  \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (\lnot w_{q} \land \lnot w_{r}) \Bigg) \lor \\ \Bigg( &\bigvee \limits_{\substack{t_{i} \in \{q\} \times \Sigma \times 2^M  \times \{ r \}, \\ q \in Q, r \in Q}} ( e_{i} \land w_{q} \land \lnot w_{r} )  \land \bigvee \limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} (e_{i} \land \lnot  w_{q} \land  w_{r}) \Bigg) \Bigg]
+\end{align*}
+
+\begin{example}
+Figure \ref{fig:cycles_sensitivity} displays an example of an automaton reduced by Level 4.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+
+    \node[state, initial] (g) {$0$};
+
+    \node[flabel] at (1.5, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{1}{1} \land \Inf\tacc{2}{2}$};
+    \node[state, right=2cm of g](q){$1$};
+
+    \path[->, auto]
+    (g) edge[bend right] node [accset,collacc0, pos=0.5]{0}node [below, pos=0.2, scale=0.8] {$2$}(q)
+    (q) edge[bend right] node[above, pos=0.8, scale=0.8] {$4$}(g)
+    (q) edge[loop, in=75, out=105, looseness=8]node[left, pos=0.1, scale=0.8]{5} (q)
+    (q) edge[loop, in=70, out=110, looseness=12] node[accset,collacc1] {1}node[left, pos=0.2, scale=0.8]{6} (q)
+    (g) edge[loop, in=75, out=105, looseness=8] node[accset,collacc2] {2} node[left, pos=0.1, scale=0.8]{1}(g)
+    (g) edge[loop, in=70, out=110, looseness=12] node[accset,collacc2, pos=0.4] {2} node[accset,collacc1, pos=0.7] {1} node[left, pos=0.2, scale=0.8]{3}(g)
+	;
+
+	\node[state, initial, right=4cm of g] (o){$0$};
+
+    \node[flabel] at (6.7, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{3}{3}$};
+    \node[state, right=2cm of o](p){$1$};
+
+    \path[->, auto]
+    (o) edge[bend right] node [accset,collacc0, pos=0.5]{0}(p)
+    (p) edge[bend right](o)
+    (p) edge[loop, in=75, out=105, looseness=8] (p)
+    (p) edge[loop, in=70, out=110, looseness=12] (p)
+    (o) edge[loop, in=75, out=105, looseness=8] (o)
+    (o) edge[loop, in=70, out=110, looseness=12] node[accset,collacc3] {3}(o)
+
+	;
+
+\end{tikzpicture}
+
+    \caption{TELA reduced with the CYCLES sensitivity}
+    \label{fig:cycles_sensitivity}
+\end{figure}
+\end{example}
+
+Similarly as in the previous section the connection formula is the conjunction of the two parts:
+\begin{align*}
+    \varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}.
+\end{align*}
+
+\subsection{Simplification procedure}
+\label{subsection:procedure}
+
+Now we describe our simplification algorithm taking as input an arbitrary TELA $\mathcal{A}$ with an acceptance condition $\psi$, required level of simplification $L$ and timeout period $T$.
+\begin{enumerate}
+    \item Convert acceptance condition $ \psi$ to DNF.
+    \item Apply the algorithm for simplification Level 1 described in Section \ref{sec:standard} on $\mathcal{A}$.
+    \item If $L = 1$ return automaton $\mathcal{A}$.
+     \item The initial value of $C$ is the number of clauses in $\psi$. Similarly, the initial value of $K$ is the number of acceptance sets of $\psi$ decreased by 1.
+    \item Encode $\mathcal{A}$ into QBF formula $\varphi$ according to the required level of simplification $L$.
+    \item Run a QBF-solver on $\varphi$ for $T$ seconds.
+    \begin{enumerate}
+        \item If the QBF-solver evaluates $\varphi$ as UNSAT, return automaton $\mathcal{A}'$.
+        \item If the timeout $T$ expires before the QBF-solver evaluates $\varphi$, return automaton $\mathcal{A}'$.
+        \item If the QBF-solver evaluates $\varphi$ as SAT. Decode the assignment into next TELA $\mathcal{A}'$.
+        \begin{enumerate}
+            \item If $K=1$, try to reduce $\varphi$ to $\true$ or $\false$, the simplest form of acceptance condition and return $\mathcal{A}$.
+            \item Decrease $K$ by 1 and go to (b).
+
+    \end{enumerate}
+
+    \end{enumerate}
+\end{enumerate}
+
+\todo{kde je optimalni misto, kam napsat:}
+For simplicity, we do not display letters on the edges of the automata in Figures. We can afford to omit them because they are not important for our simplification techniques.
+
 \begin{itemize}

### 2021-10-14 paper/main.tex
Level 3 tweak

@@
 \label{subsection:connection}
-The connection formula distinguishes the remaining three levels of simplification. Previously we denoted it as $\varphi_{C}$.
+The connection formula distinguishes the remaining two levels of simplification. Previously we denoted it as $\varphi_{C}$.
 
@@
 If not, the formula $\varphi_{C}$ is $\false$, and then the whole formula $\varphi$ is trivially satisfied.
-Formula $\varphi_{C}$ consists of two parts.
+Formula $\varphi_{C}$ consists of two parts $\varphi_{C_{1}}$ and $\varphi_{C_{2}}$.
 
@@
 The second part of $\varphi_{C}$ specifies the properties of \emph{relevant set} that the edges that belong together to \emph{selected edges} need to have.
-We divided these properties into two levels Level 2, Level 3, Level 4, all described in ensuing sections.
+We divided these properties into two levels Level 2 and Level 3, all described in ensuing sections.
 
@@
 Obviously, since they are not connected, they cannot form \emph{relevant set}.
-Therefore we can specify a new level of simplification -- Level 3, which forces all edges from the \emph{selection of edges} to be connected.
+Therefore we can specify a new level of simplification -- Level 3, which forces all edges from the \emph{selection of edges} to be cycles.
 \begin{figure}[h]
@@
 \subsubsection{CYCLES accuracy}
-The simplification Level 3 ensures that \emph{selection of edges} is connected, but we can go one step further and specify that \emph{selection of edges} is in fact a \emph{relevant set}.
-
-The intuition by simplification Level 4 is that the states of an automaton are again divided into two groups -- the states that belong to \emph{selected states} and the states do not belong to \emph{selected states}.
+The simplification Level 2 ensures that \emph{selection of edges} is placed in one SCC, but we can go further and specify that \emph{selection of edges} is in fact a \emph{relevant set}.
+
+The intuition by simplification Level 3 is that the states of an automaton are  divided into two groups -- the states that belong to the group of \emph{selected states} and the states do not belong to the group of \emph{selected states}.
+Then we require that the states that belong to the same group are connected.
 We consider three possible scenarios that can occur while the states are divided into these two groups:
@@
 \smallskip
-
+\todo{neoptimalni}
 Since we want the statement to hold for every \emph{selection of edges}, we place the existentially quantified variables $\{w_{1}, \ldots, w_{|Q|} \}$ behind the universally quantified variables $\{e_{1}, \ldots, e_{|\delta|} \} $.

### 2021-10-14 paper/main.tex
preliminaries polished

@@
 \newcommand{\cM}{\mathcal{M}}
+\newcommand{\mA}{\mathcal{A}}
 \newcommand{\bscc}{\mathcal{B}}
@@
   infinite words including Büchi, co-Büchi, Rabin, Streett, and parity
-  automata. Transitions in TELA are marked with acceptance marks and
-  accepting formulas are boolean combinations of terms saying that a
-  particular mark has to be visited infinitely or finitely often in an
-  accepting run. Algorithms processing these automata can be very
-  sensitive to the number of different acceptance marks. We introduce
-  two techniques reducing the number of acceptance marks in a TELA
-  without altering its structure and the represented language. One
-  technique represents a standard approach based on relations among
-  acceptance marks in individual strongly connected components of the
-  automaton. The other technique constructs \emph{quantified boolean
-    formula (QBF)} queries that ask a QBF solver for an acceptance
-  condition with fewer acceptance marks and the placement of these
-  marks in the automaton structure. Both techniques are implemented
-  and experiments show that the number of acceptance marks in TELA
-  produced by state-of-the-art tools ltl3tela, Rabinizer~4, and SPOT
-  can be sometimes reduced. In the case of Rabinizer~4, we reduced the
-  cummulative number of acceptance marks to less than one half.
+  automata. Transitions in a TELA are marked with acceptance marks and
+  its accepting formula is a positive Boolean combinations of terms
+  saying that a particular mark has to be visited infinitely or
+  finitely often in an accepting run. Algorithms processing these
+  automata can be very sensitive to the number of different acceptance
+  marks. We introduce two techniques reducing the number of acceptance
+  marks in a TELA without altering its structure and the represented
+  language. One technique represents a standard approach based on
+  relations among acceptance marks in individual strongly connected
+  components of the automaton. The other technique constructs
+  \emph{quantified Boolean formula (QBF)} queries that ask a QBF
+  solver for an acceptance condition with fewer acceptance marks and
+  the placement of these marks in the automaton structure. Both
+  techniques are implemented and experiments show that the number of
+  acceptance marks in TELA produced by state-of-the-art tools
+  ltl3tela, Rabinizer~4, and SPOT can be sometimes reduced. In the
+  case of Rabinizer~4, we reduced the cummulative number of acceptance
+  marks to less than one half.
 \end{abstract}
@@
 Their popularity comes probably from the fact that these automata can
-often use less states than equivalent automata with traditional
-acceptance conditions. Further, algorithms handling TELA can
-automatically handle all automata with traditional acceptance
-conditions. TELA automata can be obtained for example by translating
-LTL formulas by Delag~\cite{muller.17.gandalf},
-ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav}, or
+often use less states than equivalent automata with simpler acceptance
+conditions. Further, algorithms handling TELA can automatically handle
+all automata with traditional acceptance conditions. TELA automata can
+be obtained for example by translating LTL formulas by
+Delag~\cite{muller.17.gandalf}, ltl3tela~\cite{major.19.atva},
+Rabinizer~4~\cite{kretinsky.18.cav}, or
 SPOT~\cite{duret.16.atva2}. There are also some algorithms processing
@@
 The second technique is completely different. For a given TELA and
-parameters $c,m$, it produces a \emph{quantified boolean formula
+parameters $c,m$, it produces a \emph{quantified Boolean formula
   (QBF)} which is satisfiable if and only if there exists an automaton
@@
   A \emph{transition-based Emerson-Lei automaton (TELA)} is a tuple
-  $\mathcal{A} =(Q,M,\Sigma,q_{0},\delta,\varphi)$, where
+  $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$, where
   \begin{itemize}
@@
   \item $\Sigma$ is a finite \emph{alphabet},
-  \item $q_{0}\in Q$ is an \emph{initial state},
+  \item $q_I\in Q$ is an \emph{initial state},
   \item $\delta \subseteq Q \times \Sigma \times 2^{M} \times Q$ is a
     \emph{transition relation}, and
-  \item $\varphi$ is the \emph{acceptance condition} constructed by
-    the following abstract syntax equation, where $m$ ranges over $M$:
-
-    \begin{center}
-        $\varphi ::= \true \mid \false \mid \Inf(m) \mid \Fin(m) \mid \left( \varphi \land \varphi \right) \mid \left( \varphi \lor \varphi \right)$.\\
-    \end{center} 
+  \item $\varphi$ is the \emph{acceptance condition} constructed
+    according to the following abstract syntax equation, where $m$
+    ranges over $M$:
+    $$\varphi ::= \true \mid \false \mid \Inf(m) \mid \Fin(m) \mid \left( \varphi \land \varphi \right) \mid \left( \varphi \lor \varphi \right).$$
 \end{itemize}
@@
 
-\todo{Honza je tu a jde spat (je 5 rano pryc).}
-
-A run $\pi$ of $\mathcal{A}$ on word $u = u_{0}u_{1}u_{2}\dots \in \Sigma^{\omega}$ is an infinite sequence of transitions $\pi = (q_{0}, u_{0}, M_{0}, q_{1})(q_{2}, u_{1}, M_{1}, q_{2}) \dots \in \delta^{\omega}$.
-
-Let $M(\pi)$ denote set of all acceptance marks, that occur in run $\pi$ in infinitely many transitions.
-The run satisfies $\Inf(m)$ $\mathit{iff} \ m \in M(\pi)$ and it satisfies $\Fin(m) \ \mathit{iff} \ m \ \not \in M(\pi)$.
-Run $\pi$ is \emph{accepting} if it satisfies the formula $\varphi$. A set of words, which are accepted by $\mathcal{A}$, represent a language $\mathcal{L}(\mathcal{A})$.
-
-Sometimes the set of transitions with the same acceptance mark $m\in M$ is referred to as \emph{acceptance set} $Z(m)$.
-
-\begin{itemize}
-    \item define DNF
-
-\end{itemize}
-
-\subsection{Strongly connected component}
-Let us consider a TELA $\mathcal{A} = \left(  \mathcal{Q}, M, \Sigma,q_{0}, \delta, \varphi \right)$.
-\emph{Strongly connected component} (SCC) of $\mathcal{A}$ is each maximal set of states $S \subseteq Q$ such that any two states $q_{1}, q_{2} \in S$ are connected by a sequence of transitions.
-Let us enumerate the SCCs of $\mathcal{A}$.
-Set of all transitions of the SCC $S_{i} \text{ is } \delta_{S_{i}} = \delta \cap (S_{i} \times \Sigma \times 2^M \times S_{i})$.
-A run of SCC $S_{i}$ is a run of automaton $\mathcal{A}$ with the infinite suffix contained within $S_{i}$.
-
-\subsection{Quantified Boolean formula}
-\emph{Quantified Boolean formulae} (QBFs) is an extension of propositional logic -- some Boolean variables in the formula can be quantified with \emph{universal} or \emph{existential} quantifiers.
-Formula $\Phi$ is a QBF formula in \emph{prenex normal form} $\iff$ $\Phi = Q_{1}x_{1}...Q_{n}x_{n} .
-\varphi$, where $Q_{i} \in \{\forall, \exists\}$, $x_{i}$ is a quantified Boolean variable and $\varphi$ is quantifier-free Boolean formula.
-
-\smallskip 
-Let $\Phi$ be a QBF.
-Semantically we define the \emph{universal} quantifier as $\forall x .
-\Phi = \Phi [x \rightarrow \true] \land \Phi [x \rightarrow \false]$ and the \emph{existential} quantifier as $\exists x .
-\Phi = \Phi [x \rightarrow \true] \lor \Phi [x \rightarrow \false]$.
-
-Let $X$ be a finite set of variables. Assignment $\mathcal{A}_{X}$ $:X \rightarrow \{\false, \true\}$ is a total mapping of variables defined on set $X$ to Boolean values.
-Let $\Phi$ be a QBF. If there is an assignment $\mathcal{A}_{X}$ that evaluates formula $\Phi$ as $\true$, then $\Phi$ is $\mathit{satisfiable}$ (SAT).
-A set $S \subseteq X$ is a \emph{model} of $\Phi$ if the assignment $\mathcal{A}_{X}$ maps $\true$ to every element of $S$.
-If there is no such assignment that evaluates the formula as $\true$, we say the formula is $\mathit{unsatisfiable}$ (UNSAT).
+%A transition $(p,a,M',q)\in\delta$ is a transition leading from state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$
+A run $\pi$ of $\mA$ over an infinite word
+$u = u_{0}u_{1}u_{2}\dots \in \Sigma^{\omega}$ is an infinite sequence
+of adjacent transitions
+$\pi=(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots\in\delta^{\omega}$ where
+$q_0=q_I$. Let $M(\pi)$ denote set of all acceptance marks that occur
+in run $\pi$ in infinitely many transitions. Run $\pi$ is
+\emph{accepting} iff it satisfies the formula $\varphi$, where $\pi$
+satisfies $\Inf(m)$ iff $m\in M(\pi)$ and $\pi$ satisfies $\Fin(m)$
+iff $m\not\in M(\pi)$. The \emph{language} of automaton $\mA$ is the
+set $L(\mA)=\{u\in \Sigma^{\omega}\mid\textrm{there is an accepting run of }\mA\textrm{ over }u\}$.
+% such that $\mA$ has
+% an accepting run over $u$.
+
+An acceptance formula $\varphi$ is in \emph{disjunctive normal formal
+  (DNF)} if it is a disjunction of clauses, where each clause is a
+conjunction of some terms of the form $\Fin(m)$ or $\Inf(m)$.  Formula
+$\false$ corresponds to the disjunction of zero clauses and formula
+$\true$ is the clause with zero terms. Each acceptance formula can be
+transformed into an equivalent formula in DNF.
+
+For each mark $m\in M$, let $\delta_m$ denotes the set of all
+transitions marked with $m$, i.e.,
+$\delta_m=\{(p,a,M',q)\in\delta\mid m\in M'\}$. Further, a \emph{path}
+from a state $p$ to a state $q$ is a finite sequence of adjacent
+transitions of the form
+$(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
+such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
+is called a nontrivial \emph{strongly connected component (SCC)} if
+for each $p,q\in S$ there is a path from $p$ to $q$. An SCC S is
+\emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
+the rest of this paper, SCC always refers to a maximal SCC. Given an
+SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
+denotes the set of all transitions between states in $S$.
+
+We often use acceptance marks $\tacc{1}{1},\tacc{2}{2},\ldots\in M$
+and we write $\Inf\tacc{4}{j}$ and $\Fin\tacc{8}{k}$ instead of
+$\Inf(\tacc{4}{j})$ and $\Fin(\tacc{8}{k})$, respectively. In our
+graphical representation of automata,
+\begin{tikzpicture}[smallautomaton]
+  \node[state] (p) at (0,0) {$p$};
+  \node[state] (q) at (1.8,0) {$q$};
+  \path [->] (p) edge node[above,pos=.5,overlay] {$a$} node[accset,collacc8, pos=0.65] {k} node[accset,collacc4, pos=0.3] {j} (q);
+\end{tikzpicture}
+denotes the transition $(p,a,\{\tacc{4}{j},\tacc{8}{k}\},q)\in\delta$.
+
+% Finaly, \emph{loop} is a path from a state $p$ to the same state.
+
+\bigskip
+
+\emph{Quantified Boolean formulas (QBFs)} are Boolean formulas
+extended with universal or existential quantification over
+propositional variables. We assume that subformulas of the form
+$\forall x.\psi$ and $\exists x.\psi$ do not contain another
+quantification of variable $x$ inside $\psi$. The semantics of
+$\forall x.\psi$ and $\exists x.\psi$ is given by equivalences
+\[
+  \begin{array}{rcl}
+    \forall x.\psi &~\equiv~& \psi[x\rightarrow\true]~\wedge~\psi[x\rightarrow\false]\\
+    \exists x.\psi &~\equiv~& \psi[x\rightarrow\true]~\vee~   \psi[x\rightarrow\false]
+  \end{array}
+\]
+where $\psi[x\rightarrow \rho]$ denotes the formula $\psi$ with all
+occurrences of $x$ simultaneously replaced by $\rho$. The equivalences
+imply that each QBF can be transformed into an equivalent Boolean
+formula. However, the size of this Boolean formula can be exponential
+in the size of the original QBF. Let $V$ be the set of all
+propositional variables. A mapping $\mu:V\rightarrow\{\true,\false\}$
+is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
+assignment of the equivalent Boolean formula. A QBF formula is
+\emph{satisfiable} iff it has a model.
+
+A QBF $\varphi$ is in the \emph{prenex normal form} if it has the form
+$\varphi = Q_{1}x_{1}...Q_{n}x_{n} . \psi$, where
+$Q_1,Q_2,\ldots,Q_n \in \{\forall, \exists\}$ and $\psi$ is a Boolean
+formula.
 

### 2021-10-14 paper/main.tex
begining os section 3 modified

@@
 \newcommand{\mA}{\mathcal{A}}
+\newcommand{\mB}{\mathcal{B}}
 \newcommand{\bscc}{\mathcal{B}}
@@
 \subsubsection{Related results}
-We are not aware about any existing research focused on simplification
+We are not aware about any existing research aimed at simplification
 of acceptance formulas of TELA or reduction of the number of its
@@
     according to the following abstract syntax equation, where $m$
-    ranges over $M$:
-    $$\varphi ::= \true \mid \false \mid \Inf(m) \mid \Fin(m) \mid \left( \varphi \land \varphi \right) \mid \left( \varphi \lor \varphi \right).$$
+    ranges over $M$.
+    $$\varphi ::= \true \mid \false \mid \Inf(m) \mid \Fin(m) \mid \left( \varphi \land \varphi \right) \mid \left( \varphi \lor \varphi \right)$$
 \end{itemize}
@@
 iff $m\not\in M(\pi)$. The \emph{language} of automaton $\mA$ is the
-set $L(\mA)=\{u\in \Sigma^{\omega}\mid\textrm{there is an accepting run of }\mA\textrm{ over }u\}$.
+set
+$L(\mA)=\{u\in \Sigma^{\omega}\mid\textrm{there is an accepting run of
+}\mA\textrm{ over }u\}$. Two automata $\mA,\mB$ are \emph{equivalent}
+if $L(\mA)=L(\mB)$.
 % such that $\mA$ has
@@
 
-For each mark $m\in M$, let $\delta_m$ denotes the set of all
-transitions marked with $m$, i.e.,
-$\delta_m=\{(p,a,M',q)\in\delta\mid m\in M'\}$. Further, a \emph{path}
-from a state $p$ to a state $q$ is a finite sequence of adjacent
-transitions of the form
+A \emph{path} from a state $p$ to a state $q$ is a finite sequence of
+adjacent transitions of the form
 $(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
@@
 SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
-denotes the set of all transitions between states in $S$.
+denotes the set of all transitions between states in $S$. Further, for
+each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
+marked with $m$, i.e., $\delta_m=\{(p,a,M',q)\in\delta\mid m\in
+M'\}$. The sets $\delta_m$ are called \emph{acceptance sets}.
+Finally, given an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we
+denote the set $\delta_{S,m}=\delta_S\cap\delta_m$.
 
@@
 graphical representation of automata,
-\begin{tikzpicture}[smallautomaton]
+\begin{tikzpicture}[smallautomaton,baseline=-2pt]
   \node[state] (p) at (0,0) {$p$};
@@
 $\varphi = Q_{1}x_{1}...Q_{n}x_{n} . \psi$, where
-$Q_1,Q_2,\ldots,Q_n \in \{\forall, \exists\}$ and $\psi$ is a Boolean
+$Q_1,\ldots,Q_n \in \{\forall, \exists\}$ and $\psi$ is a Boolean
 formula.
@@
 
-In this section, the simplification strategy is to find relations between acceptance sets and reduce the redundant ones. SCC-based simplification is the coarsest simplification we propose.
-
-Let $\mathcal{A}$ be an original automaton and let $\psi$ be an acceptance condition of $\mathcal{A}$.
-Every run of $\mathcal{A}$ has an infinite suffix that takes place within one SCC $S_{i}$.
-Thus the evaluation of $\psi$ depends purely on the SCC $S_{i}$ and we can optimize the acceptance condition for each SCC separately.
-:w
-This optimization consists of removing redundant terms from acceptance condition and relabeling of acceptance marks on the transitions.
-The state and transition structure of the SCC does not change.
-This way we obtain set of simplified acceptance conditions that we merge into new acceptance formula $\psi'$.
-The automaton with relabeled transitons and new acceptance condition $\psi'$ we denote $\mathcal{A}'$.
-Finally, we ensure that automaton $\mathcal{A'}$ is equivalent to the original automaton $\mathcal{A}$.
-
-\subsection{$S_{i}$ simplification}
-\label{section:simplification}
-At first, we remove the unused acceptance marks and transform the SCC's acceptance condition $\psi_{i}$ to \emph{disjunctive normal form} (DNF), so we have a unified shape of the condition.
-If TELA acceptance condition is in DNF, two distinct terms in one disjunct can only appear in three possible forms:
-
-\begin{itemize}
-    \item $\Inf\tacc{2}{i} \land \Inf\tacc{4}{j}$
-    \item $\Fin\tacc{8}{i} \land \Fin\tacc{4}{j}$
-    \item $\Fin\tacc{8}{i} \land \Inf\tacc{4}{j}$
-\end{itemize}
+This section presents a reduction technique based on relations between
+acceptance sets restricted to individual SCCs. It employs the fact
+that acceptance of a run always depends only on a single SCC, namely
+the SCC that contains the whole run except some finite prefix.
+
+Let $\mA$ be a TELA with the acceptance condition $\varphi$. We assume
+that $\varphi$ is in DNF. This technique has two steps.
+\begin{enumerate}
+\item For each SCC $S$, we make a copy $\varphi_S$ of $\varphi$ and
+  simplify it according to relations between acceptance sets
+  restricted to $\delta_S$. 
+\item We collect the simplified acceptance conditions
+  $\varphi_{S_1},\ldots,\varphi_{S_k}$ of all SCCs and merge them into
+  one acceptance condition $\varphi'$. Additionally, we sometimes add
+  acceptance marks to all transitions of some SCCs to make the
+  resulting automaton with acceptance formula $\varphi'$ equivalent to
+  the original automaton.
+\end{enumerate}
+
+% Let $\mathcal{A}$ be an original automaton and let $\psi$ be an
+% acceptance condition of $\mathcal{A}$. Every run of $\mathcal{A}$ has
+% an infinite suffix that takes place within one SCC $S_{i}$. Thus the
+% evaluation of $\psi$ depends purely on the SCC $S_{i}$ and we can
+% optimize the acceptance condition for each SCC separately. This
+% optimization consists of removing redundant terms from acceptance
+% condition and relabeling of acceptance marks on the transitions. The
+% state and transition structure of the SCC does not change. This way we
+% obtain set of simplified acceptance conditions that we merge into new
+% acceptance formula $\psi'$. The automaton with relabeled transitons
+% and new acceptance condition $\psi'$ we denote $\mathcal{A}'$.
+% Finally, we ensure that automaton $\mathcal{A'}$ is equivalent to the
+% original automaton $\mathcal{A}$.
+
+\subsection{Simplification of $\varphi_S$ for an SCC $S$}\label{ssec:step1}
+Recall that $\varphi_S$ is in DNF. Hence, we can represent it as a set
+$\varphi_S=\{D_1,\ldots,D_k\}$ of disjunctive clauses.
 

### 2022-02-21 paper/main.tex
moving towards SAT submission

@@
 
-\newenvironment{nstabbing}
-  {\setlength{\topsep}{0pt}%
-   \setlength{\partopsep}{0pt}%
-   \tabbing}
-  {\endtabbing}
-
 %{{{ (La)TeX definitions
@@
 \newcommand{\Nset}{{\mathbb{N}_0}}
-
 \newcommand{\todo}[1]{\textcolor{red}{#1}}
@@
 \def\Fin{\ensuremath{\mathsf{Fin}}}
-\def\ttrue{\ensuremath{\mathsf{t}}}
-\def\ffalse{\ensuremath{\mathsf{f}}}
-\def\CQ{\mathcal{C}}
-\def\BigO#1{\ensuremath{\mathcal{O}\left(#1\right)}}
-\newcommand{\B}{\texorpdfstring{\ensuremath{\mathds{B}}}{B}}
-\def\isempty{\textsc{is\_empty}\xspace}
-\def\sccsof{\textsc{SCCs\_of}\xspace}
-\def\issccempty{\textsc{is\_SCC\_empty}\xspace}
-\def\marksof{\textsc{marks\_of}\xspace}
-\def\remove{\textsc{remove}\xspace}
 \def\false{\mathit{false}}
 \def\true{\mathit{true}}
-
-\def\Mocc{M_{\mathrm{occur}}}
-\def\Mevery{M_{\mathrm{every}}}
-\newcommand{\G}{\mathbf{G}}
-\newcommand{\CA}{\ensuremath{\mathcal{A}}}
-\newcommand{\CB}{\ensuremath{\mathcal{B}}}
-\def\init{\iota}
-\def\lang{L} % FIX this
-\newcommand{\cM}{\mathcal{M}}
 \newcommand{\mA}{\mathcal{A}}
 \newcommand{\mB}{\mathcal{B}}
-\newcommand{\bscc}{\mathcal{B}}
-\newcommand{\sched}{\mathfrak{s}}
-\newcommand{\Act}{\mathit{Act}}
-\newcommand{\NP}{\textsf{NP}}
-\newcommand{\POLY}{\textsf{P}}
-\newcommand{\class}{\ensuremath{\mathcal{C}}}
-\newcommand{\finally}{\mathsf{F}}
-\newcommand{\globally}{\mathsf{G}}
-% \newcommand{\finally}{\mathbf{F}}
-% \newcommand{\globally}{\mathbf{G}}
-
-% Remove space from footnote after interpunction. Usage:
-% \punct{.}\footnote{footnotetext}
-\newlength{\spc} % declare a variable to save spacing value
-\newcommand{\punct}[1]{%
-  \settowidth{\spc}{#1}% set value of \spc variable to the width of #1 argument
-  \addtolength{\spc}{-1.8\spc}% subtract from \spc about two (1.8) of its values making its magnitude negative
-  #1% print the optional argument
-  \hspace*{\spc}% print an additional negative spacing stored in \spc after #1
-}
-
 %}}}
@@
 
-\title{Reducing Acceptance Marks \\ in Emerson-Lei Automata}
-\authorrunning{T.~Schwarzová and J. Strej\v{c}ek}
- \author{Tereza Schwarzová \and Jan Strej\v{c}ek}
+\title{Reducing Acceptance Marks\\ in Emerson-Lei Automata by QBF-solving}
+\authorrunning{J.~Major, T.~Schwarzová and J. Strej\v{c}ek}
+ \author{Juraj Major \and Tereza Schwarzová \and Jan Strej\v{c}ek}
  \institute{%
-   Masaryk University, Brno, Czech Republic\\ \email{\{xschwar3,strejcek\}@mail.muni.cz}
+   Masaryk University, Brno, Czech Republic\\ \email{\{major,xschwar3,strejcek\}@fi.muni.cz}
  }
@@
   automata. Transitions in a TELA are marked with acceptance marks and
-  its accepting formula is a positive Boolean combinations of terms
+  its accepting formula is a positive Boolean combination of terms
   saying that a particular mark has to be visited infinitely or
@@
   automata can be very sensitive to the number of different acceptance
-  marks. We introduce two techniques reducing the number of acceptance
-  marks in a TELA without altering its structure and the represented
-  language. One technique represents a standard approach based on
-  relations among acceptance marks in individual strongly connected
-  components of the automaton. The other technique constructs
-  \emph{quantified Boolean formula (QBF)} queries that ask a QBF
-  solver for an acceptance condition with fewer acceptance marks and
-  the placement of these marks in the automaton structure. Both
-  techniques are implemented and experiments show that the number of
-  acceptance marks in TELA produced by state-of-the-art tools
-  ltl3tela, Rabinizer~4, and SPOT can be sometimes reduced. In the
-  case of Rabinizer~4, we reduced the cummulative number of acceptance
-  marks to less than one half.
+  marks. We introduce a new technique reducing the number of
+  acceptance marks in a TELA with use of QBF solving. The technique
+  produces QBF formulas
+
+  \textbf{Honza je tu}
+
+  without altering its structure and the
+  represented language. The technique constructs a \emph{quantified
+    Boolean formula (QBF)}
+
+  whose satisfiebility impli a QBF formula parametrized
+  which is satisfiable only if
+
+  solver for an
+  acceptance condition with fewer acceptance marks and the placement
+  of these marks in the automaton structure. Both techniques are
+  implemented and experiments show that the number of acceptance marks
+  in TELA produced by state-of-the-art tools Delag, ltl3tela, Rabinizer~4,
+  and SPOT can be sometimes reduced. In the case of Rabinizer~4, we
+  reduced the cummulative number of acceptance marks to less than one
+  half.
 \end{abstract}

### 2022-02-21 paper/main_lipics.tex
moving towards SAT submission

--- 2021-10-14 paper/main.tex
@@
-\documentclass[runningheads]{llncs}
-
-% \usepackage[numbers]{natbib}
-% \renewcommand\bibsection{\section*{\refname}\small\renewcommand\bibnumfmt[1]{##1.}}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts,amssymb}
-\usepackage{hyperref}
-% \usepackage{xspace}
-% \usepackage{amsthm}
-% \usepackage{enumitem} % topsep
+\documentclass[a4paper,UKenglish,cleveref, autoref, anonymous, thm-restate]{lipics-v2021}
+%This is a template for producing LIPIcs articles. 
+%See lipics-v2021-authors-guidelines.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+%for section-numbered lemmas etc., use "numberwithinsect"
+%for enabling cleveref support, use "cleveref"
+%for enabling autoref support, use "autoref"
+%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
+%for enabling thm-restate support, use "thm-restate"
+%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
+%for producing a PDF according the PDF/A standard, add "pdfa"
+
+%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
+%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g.
+when preparing a pre-final version to be uploaded to arXiv or another public repository% \usepackage[numbers]{natbib}
+
+\bibliographystyle{plainurl}% the mandatory bibstyle
 
@@
 
-\newenvironment{nstabbing}
-  {\setlength{\topsep}{0pt}%
-   \setlength{\partopsep}{0pt}%
-   \tabbing}
-  {\endtabbing}
+% \newenvironment{nstabbing}
+%   {\setlength{\topsep}{0pt}%
+%    \setlength{\partopsep}{0pt}%
+%    \tabbing}
+%   {\endtabbing}
 
@@
 \def\Fin{\ensuremath{\mathsf{Fin}}}
-\def\ttrue{\ensuremath{\mathsf{t}}}
-\def\ffalse{\ensuremath{\mathsf{f}}}
-\def\CQ{\mathcal{C}}
-\def\BigO#1{\ensuremath{\mathcal{O}\left(#1\right)}}
-\newcommand{\B}{\texorpdfstring{\ensuremath{\mathds{B}}}{B}}
-\def\isempty{\textsc{is\_empty}\xspace}
-\def\sccsof{\textsc{SCCs\_of}\xspace}
-\def\issccempty{\textsc{is\_SCC\_empty}\xspace}
-\def\marksof{\textsc{marks\_of}\xspace}
-\def\remove{\textsc{remove}\xspace}
 \def\false{\mathit{false}}
 \def\true{\mathit{true}}
-
-\def\Mocc{M_{\mathrm{occur}}}
-\def\Mevery{M_{\mathrm{every}}}
-\newcommand{\G}{\mathbf{G}}
-\newcommand{\CA}{\ensuremath{\mathcal{A}}}
-\newcommand{\CB}{\ensuremath{\mathcal{B}}}
-\def\init{\iota}
-\def\lang{L} % FIX this
-\newcommand{\cM}{\mathcal{M}}
 \newcommand{\mA}{\mathcal{A}}
 \newcommand{\mB}{\mathcal{B}}
-\newcommand{\bscc}{\mathcal{B}}
-\newcommand{\sched}{\mathfrak{s}}
-\newcommand{\Act}{\mathit{Act}}
-\newcommand{\NP}{\textsf{NP}}
-\newcommand{\POLY}{\textsf{P}}
-\newcommand{\class}{\ensuremath{\mathcal{C}}}
-\newcommand{\finally}{\mathsf{F}}
-\newcommand{\globally}{\mathsf{G}}
-% \newcommand{\finally}{\mathbf{F}}
-% \newcommand{\globally}{\mathbf{G}}
-
-% Remove space from footnote after interpunction. Usage:
-% \punct{.}\footnote{footnotetext}
-\newlength{\spc} % declare a variable to save spacing value
-\newcommand{\punct}[1]{%
-  \settowidth{\spc}{#1}% set value of \spc variable to the width of #1 argument
-  \addtolength{\spc}{-1.8\spc}% subtract from \spc about two (1.8) of its values making its magnitude negative
-  #1% print the optional argument
-  \hspace*{\spc}% print an additional negative spacing stored in \spc after #1
-}
 
@@
 
+%{{{ title and authors
+
+\title{Reducing Acceptance Marks \\ in Emerson-Lei Automata}
+\author{Tereza Schwarzov\'{a}}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
+\author{Jan Strej\v{c}ek}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{strejcek@fi.muni.cz}{https://orcid.org/0000-0001-5873-403X}{}
+\author{Juraj Major}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{}{}
+\authorrunning{T. Schwarzov\'{a}, J. Strej\v{c}ek, and J. Major}
+\Copyright{Tereza Schwarzov\'{a}, Jan Strej\v{c}ek, and Juraj Major}
+\ccsdesc[500]{Theory of computation~Automata over infinite objects}
+\ccsdesc[500]{Theory of computation~Logic}
+\keywords{Emerson-Lei automata, TELA, automata reduction, QBF, \todo{telatko?}}
+\funding{This work has been supported by the Czech Science Foundation grant GA19-24397S.}
+
+%}}}
+
 \begin{document}
-
-%{{{ title and authors
-
-\title{Reducing Acceptance Marks \\ in Emerson-Lei Automata}
-\authorrunning{T.~Schwarzová and J. Strej\v{c}ek}
- \author{Tereza Schwarzová \and Jan Strej\v{c}ek}
- \institute{%
-   Masaryk University, Brno, Czech Republic\\ \email{\{xschwar3,strejcek\}@mail.muni.cz}
- }
-
-%}}}
 
@@
 
-%\bibliographystyle{abbrvnat}
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
 \bibliography{mc}

### 2022-02-21 paper/main.tex
back to lipics, new abstract

@@
-\documentclass[runningheads]{llncs}
-
-% \usepackage[numbers]{natbib}
-% \renewcommand\bibsection{\section*{\refname}\small\renewcommand\bibnumfmt[1]{##1.}}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts,amssymb}
-\usepackage{hyperref}
-% \usepackage{xspace}
-% \usepackage{amsthm}
-% \usepackage{enumitem} % topsep
+\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
+%\documentclass[a4paper,UKenglish,cleveref, autoref, anonymous, thm-restate]{lipics-v2021}
+%
+%This is a template for producing LIPIcs articles. 
+%See lipics-v2021-authors-guidelines.pdf for further information.
+%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
+%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
+%for section-numbered lemmas etc., use "numberwithinsect"
+%for enabling cleveref support, use "cleveref"
+%for enabling autoref support, use "autoref"
+%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
+%for enabling thm-restate support, use "thm-restate"
+%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
+%for producing a PDF according the PDF/A standard, add "pdfa"
+
+%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
+%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g.
+when preparing a pre-final version to be uploaded to arXiv or another public repository% \usepackage[numbers]{natbib}
+
+\bibliographystyle{plainurl}% the mandatory bibstyle
 
@@
 \newcommand{\mB}{\mathcal{B}}
+
 %}}}
 
+%{{{ title and authors
+
+\title{Reducing Acceptance Marks\\ in Emerson-Lei Automata by QBF-solving}
+\author{Tereza Schwarzov\'{a}}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
+\author{Jan Strej\v{c}ek}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{strejcek@fi.muni.cz}{https://orcid.org/0000-0001-5873-403X}{}
+\author{Juraj Major}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{}{}
+\authorrunning{T. Schwarzov\'{a}, J. Strej\v{c}ek, and J. Major}
+\Copyright{Tereza Schwarzov\'{a}, Jan Strej\v{c}ek, and Juraj Major}
+\ccsdesc[500]{Theory of computation~Logic}
+\ccsdesc[500]{Theory of computation~Automata over infinite objects}
+\keywords{Emerson-Lei automata, TELA, automata reduction, QBF, \todo{telatko?}}
+\funding{This work has been supported by the Czech Science Foundation grant GA19-24397S.}
+
+%}}}
+
 \begin{document}
 
-%{{{ title and authors
-
-\title{Reducing Acceptance Marks\\ in Emerson-Lei Automata by QBF-solving}
-\authorrunning{J.~Major, T.~Schwarzová and J. Strej\v{c}ek}
- \author{Juraj Major \and Tereza Schwarzová \and Jan Strej\v{c}ek}
- \institute{%
-   Masaryk University, Brno, Czech Republic\\ \email{\{major,xschwar3,strejcek\}@fi.muni.cz}
- }
-
-%}}}
-
 \maketitle
@@
 \begin{abstract}
-  The formalism called \emph{transition-based Emerson-Lei automata
-    (TELA)} generalizes many traditional kinds of automata over
+  \emph{Transition-based Emerson-Lei automata (TELA)} become a popular
+  formalism that generalizes many traditional kinds of automata over
   infinite words including Büchi, co-Büchi, Rabin, Streett, and parity
-  automata. Transitions in a TELA are marked with acceptance marks and
-  its accepting formula is a positive Boolean combination of terms
+  automata. Transitions in a TELA are labelled with acceptance marks
+  and its accepting formula is a positive Boolean combination of terms
   saying that a particular mark has to be visited infinitely or
   finitely often in an accepting run. Algorithms processing these
-  automata can be very sensitive to the number of different acceptance
-  marks. We introduce a new technique reducing the number of
-  acceptance marks in a TELA with use of QBF solving. The technique
-  produces QBF formulas
-
-  \textbf{Honza je tu}
-
-  without altering its structure and the
-  represented language. The technique constructs a \emph{quantified
-    Boolean formula (QBF)}
-
-  whose satisfiebility impli a QBF formula parametrized
-  which is satisfiable only if
-
-  solver for an
-  acceptance condition with fewer acceptance marks and the placement
-  of these marks in the automaton structure. Both techniques are
-  implemented and experiments show that the number of acceptance marks
-  in TELA produced by state-of-the-art tools Delag, ltl3tela, Rabinizer~4,
-  and SPOT can be sometimes reduced. In the case of Rabinizer~4, we
-  reduced the cummulative number of acceptance marks to less than one
-  half.
+  automata are often very sensitive to the number of different
+  acceptance marks. We introduce a new technique reducing the number
+  of different acceptance marks in a TELA with use of \emph{quantified
+    Boolean formulas (QBF)}: if a constructed QBF formula is
+  satisfiable then its model encodes an equivalent accepting formula
+  with less acceptance marks. In fact, we present three formula
+  constructions trading the speed of the technique for higher
+  reduction efficiency. The technique is implemented and evaluated on
+  TELA produced by state-of-the-art tools Delag, ltl3tela,
+  Rabinizer~4, and SPOT. The technique reduced some acceptance marks
+  in automata produced by all these tools. In the case of Rabinizer~4,
+  the technique reduced the cummulative number of acceptance marks to
+  less than one half. 
 \end{abstract}
@@
 %}}}
+
+\todo{Nezapomenout na clanky o redukci automatu pomoci SATu - on virtue of patience?}
 

### 2022-02-22 qbf/parser.py
Z3, level1 merge refactoring

@@ -1,4 +1,5 @@
 from telatko2.classes import *
+from z3 import is_true
 
 
 def create_acc(variables):
@@ -158,24 +159,48 @@ def max_set_num(acc):
                 num = con.num
     return num
 
+def parse_limboole(model):
+    # list of variables that are true and are denoted as p/n - acceptance
+    # condition variables
+    condition_vars = list(filter(lambda var: (
+        var[0] == "p" or var[0] == "n") and var[-1] == '1', model))
+    # list of variables that are true and denoted as f - acceptance sets
+    acc_set_vars = list(
+        filter(lambda var: var[0] == "f" and var[-1] == '1', model))
+    #print("cond vars: ", condition_vars)
+    return condition_vars, acc_set_vars
 
-def process_variables(aut, variables):
+def parse_z3(model):
+    #TO DO TOMMOROW
+
+    condition_vars = []
+    acc_set_vars = []
+    for t in model.decls():
+        if is_true(model[t]):
+            if str(t)[0] == 'p' or str(t)[0] == 'n':
+                condition_vars.append(str(t))
+            elif str(t)[0] == 'f':
+                acc_set_vars.append(str(t))
+    #print("cond vars: ", condition_vars)
+    #print("acc set vars: ", acc_set_vars )
+    return condition_vars, acc_set_vars
+
+
+def process_variables(aut, model, qbf_solver):
     """
     Filters SAT-variables that are true and splits them. Then calls functions to process these variables.
     Args:
         aut: input automata
-        variables: output from SAT-solver
+        model: output from SAT-solver
+        qbf_solver : type of solver (solver output)
 
     Returns:
 
     """
-    # list of variables that are true and are denoted as p/n - acceptance
-    # condition variables
-    condition_vars = list(filter(lambda var: (
-        var[0] == "p" or var[0] == "n") and var[-1] == '1', variables))
-    # list of variables that are true and denoted as f - acceptance sets
-    acc_set_vars = list(
-        filter(lambda var: var[0] == "f" and var[-1] == '1', variables))
+    if qbf_solver == 'limboole':
+        condition_vars, acc_set_vars = parse_limboole(model)
+    else:
+        condition_vars, acc_set_vars = parse_z3(model)
 
     # returns acceptance condition [[MarkType]]
     acc = create_acc(prepare_acc_vars(condition_vars))

### 2022-02-22 qbf/q_main.py

@@ -3,6 +3,7 @@ import random
 from qbf.parser import *
 from qbf.formula import *
 from qbf.optimization import *
+from qbf.z3_formula import *
 
 
 def edge_dictionary(aut):
@@ -79,112 +80,8 @@ def scc_info(aut):
     return scc_state_inf, scc_edg
 
 
-def SAT_output(name, quantified, formula):
-    """
-    Prints formula into file(for QBF solver) named satfile.
-    Args:
-        quantified: Universal quantificators of formula
-        formula: SATformula - Boolean formula
-
-    Returns:
-
-    """
-    f = open(name, "w")
-    f.write(str(quantified) + str(formula))
-    f.close()
 
 
-def w_quant(aut):
-    """
-    Creates existential variables w_
-    Args:
-        aut:
-
-    Returns:
-
-    """
-    exist_formula = ""
-    num_states = aut.num_states()
-    for i in range(num_states):
-        exist_formula += "?w_" + str(i)
-    return exist_formula
-
-
-def laso2(aut, inner_edges, scc_edg):
-    """
-    QBF CYCLES ve statistikach
-    Formula that makes sure that the edges are cycles.
-    Unluckily the complexity is too high to compute jeden pitomej automat s 12 stavama.
-    Args:
-        aut:
-        inner_edges:
-        scc_edg:
-
-    Returns:
-
-    """
-
-    main_dis = SATformula("|")
-
-    con3 = SATformula("&")
-    dis1 = SATformula("|")
-    dis2 = SATformula("|")
-    con1 = SATformula("&")
-    con2 = SATformula("&")
-
-    for e in inner_edges:
-        src = str(e.src)
-        dst = str(e.dst)
-        """
-        if e.src == e.dst:
-            continue
-        """
-
-        # part one
-        impl = SATformula("->")
-        impl.add_subf(SATformula("e_" + str(aut.edge_number(e))))
-        c = SATformula("&")
-        c.add_subf(SATformula("w_" + src))
-        c.add_subf(SATformula("w_" + dst))
-        impl.add_subf(c)
-        con1.add_subf(impl)
-
-        # part two
-        impl2 = SATformula("->")
-        impl2.add_subf(SATformula("e_" + str(aut.edge_number(e))))
-        c1 = SATformula("&")
-        c1.add_subf(SATformula("!w_" + src))
-        c1.add_subf(SATformula("!w_" + dst))
[…]

### 2022-02-22 qbf/z3_formula.py

@@ -1,0 +1,317 @@
+from telatko2.classes import *
+from z3 import *
+from qbf.q_main import FormulaCreator
+
+def foo():
+    p1 = Bool('p1')
+    p2 = Bool('p1')
+
+    #q = Bool('q')
+    F = ForAll([p1, p2], Or(And(Not(p1),Not(p2)), And(p1,p2)))
+    #solver = Solver()
+    #solver.add(F)
+    result = solver.check()
+    print(F, result, solver.model())
+
+class Z3_f_ctor(FormulaCreator):
+    def __init__(self,  edge_dict, scc_edg, scc_state_info
+                  ,inner_edges_nums,C, K, inner_edges
+                  , mode, qbf_solver):
+        super().__init__( edge_dict, scc_edg, scc_state_info
+                      ,inner_edges_nums,C, K, inner_edges
+                      , mode, qbf_solver)
+
+        self.univ_vars = []
+        self.exist_vars = []
+
+
+
+    def get_qbf_formula(self, aut):
+        # adds unif. quantified variables to
+        self.quant_all()
+        if self.mode > 2:
+
+            self.w_quant(aut)
+
+        if self.mode == 4:
+
+            laso = self.laso_cycles(aut)
+            laso = simplify(laso)
+            in_out = self.in_n_out()
+            laso = And(laso, in_out)
+            #assert(False)
+
+        else:
+            laso = self.laso_f(aut)
+
+        old = self.old_formula(ACC_DNF(aut.get_acceptance().to_dnf()))
+        new = self.new_formula()
+        old = simplify(old)
+        #print(new)
+        new = simplify(new)
+
+        laso = simplify(laso)
+
+        eq = self.equivalence_f(old, new)
+        impl = Implies(laso, eq)
+        inf_not_fin = self.inf_is_not_fin_clause()
+        inf_or_fin = self.inf_or_fin_f()
+        inf_not_fin = simplify(inf_not_fin)
+        inf_or_fin = simplify(inf_or_fin)
+        #print(inf_not_fin, inf_or_fin)
+        #assert(False)
+
+        formula = And(impl, inf_not_fin, inf_or_fin)
+        if self.exist_vars:
+            formula = ForAll(self.exist_vars, formula)
+
+        formula = ForAll(self.univ_vars, formula)
+
+        return formula
+
+    def laso_cycles(self, aut):
+        main_dis = []
+        con1, con2, con3 = [], [], []
+        dis1, dis2 = [], []
[…]

### 2022-02-22 telatko2/playground.py

@@ -60,7 +60,7 @@ def get_acc(aut):
     Returns PACC (dnf) if dnf formula is shorter else PACC_CNF(cnf)
     """
     dnf_acc = ACC_DNF(aut.get_acceptance().to_dnf())
-    return dnf_acc
+    #return dnf_acc
     cnf_acc = ACC_CNF(aut.get_acceptance().to_cnf())
 
     #print((cnf_acc),"\n", (dnf_acc))
@@ -199,6 +199,26 @@ def try_tt_ff(aut):
     return last_eq_aut, False
 
 
+def map_accs(merged_f, short_accs):
+
+    for acc in short_accs:
+        # list_of_unmapped_dependencies.append({})
+        if acc.sat is not None or not acc.formula:
+
+            continue
+
+        get_mapping(merged_f, acc)
+
+        test_mapping(acc)
+
+
+def merge_accs(aut, merged_f, short_accs, sccs):
+    for acc in short_accs:
+        if acc.sat is not None or not acc.formula:
+            continue
+        duplicate_marks(aut, merged_f, acc, sccs)
+
+
 def process_automaton(aut):
     orig = spot.automaton(aut.to_str())
     spot.cleanup_acceptance_here(aut)
@@ -236,20 +256,37 @@ def process_automaton(aut):
     # disjunct with highest cardinality
     # final acceptance formula for whole automaton to be
     merged_f = copy.deepcopy(nempty_short_accs[0])
+    merged_f.set_merged_f()
 
     # shift numbers of acceptance sets in future final formula to prevent
     # conflict
     shift_first_acc(merged_f, m)
+    merged_f.get_dependencies()
+    #print(merged_f, merged_f.dependencies)
 
+    # rename to merge_logs
+
+    map_accs(merged_f, short_accs)
+
+    unused_clauses_depend = resolve_dependencies(
+        aut, merged_f, short_accs, sccs)
+
+    merge_accs(aut, merged_f, short_accs, sccs)
+
+    #print("unused clauses dependent literals: ", unused_clauses_depend)
+
+    # return aut
+    """
     # list of numbers of acc sets with more than one occurance
     dependence = get_dependencies(merged_f)
-    #print("dependencies:", dependence)
+    print("dependencies:", dependence)
 
 
     #[{ACCMark.num : ACCMark.num}]
     list_of_logs = []
     list_of_logs = []
     # numbers of dependent acc sets that are not mapped
[…]

### 2022-02-22 telatko2/restore_equivalence.py

@@ -207,7 +207,7 @@ def restore_never_accepting(aut, scc, merged_f):
             make_expr_false(aut, scc, expr)
 
 
-def new_make_equiv(aut, accs, sccs, merge_logs, merged_f, depend_logs):
+def new_make_equiv(aut, accs, sccs, merged_f, depend_logs):
     """
     Makes automata equivalent with the new acceptance formula
     :param aut:
@@ -222,7 +222,6 @@ def new_make_equiv(aut, accs, sccs, merge_logs, merged_f, depend_logs):
     for i in range(len(accs)):
         scc = sccs[i]
         acc = accs[i]
-        m_log = merge_logs[i]
         dep_log = depend_logs[i]
         if acc.sat is True:
             restore_always_accepting(aut, scc, merged_f)
@@ -231,9 +230,16 @@ def new_make_equiv(aut, accs, sccs, merge_logs, merged_f, depend_logs):
         else:
             used = []
             unused = []
+            for clause in acc:
+                for lit in clause:
+                    used.append(merged_f.formula[lit.map_clause_index][lit.map_literal_index].num)
+            unused = [num for num in all_set_nums if num not in used]
+
+            """
             for num in all_set_nums:
                 if num in m_log.values():
                     used.append(num)
                 else:
                     unused.append(num)
+            """
             restore_maybe_accepting(aut, scc, merged_f, unused, dep_log)

### 2022-02-22 paper/main.tex
progress in intro

@@
 \ccsdesc[500]{Theory of computation~Automata over infinite objects}
-\keywords{Emerson-Lei automata, TELA, automata reduction, QBF, \todo{telatko?}}
+\keywords{Emerson-Lei automata, TELA, automata reduction, QBF, telatko}
 \funding{This work has been supported by the Czech Science Foundation grant GA19-24397S.}
@@
   saying that a particular mark has to be visited infinitely or
-  finitely often in an accepting run. Algorithms processing these
-  automata are often very sensitive to the number of different
+  finitely often. % in an accepting run.
+  Algorithms processing these automata are often very sensitive to the
+  number of %different
   acceptance marks. We introduce a new technique reducing the number
-  of different acceptance marks in a TELA with use of \emph{quantified
-    Boolean formulas (QBF)}: if a constructed QBF formula is
-  satisfiable then its model encodes an equivalent accepting formula
-  with less acceptance marks. In fact, we present three formula
-  constructions trading the speed of the technique for higher
-  reduction efficiency. The technique is implemented and evaluated on
-  TELA produced by state-of-the-art tools Delag, ltl3tela,
-  Rabinizer~4, and SPOT. The technique reduced some acceptance marks
-  in automata produced by all these tools. In the case of Rabinizer~4,
-  the technique reduced the cummulative number of acceptance marks to
-  less than one half. 
+  of %different
+  acceptance marks in TELA with use of \emph{quantified Boolean
+    formulas (QBF)}: if a constructed QBF formula is satisfiable then
+  its model encodes an equivalent accepting formula with less
+  acceptance marks. In fact, we present three formula constructions
+  trading the speed of the technique for higher reduction
+  efficiency. We evaluated our reduction technique on TELA produced by
+  state-of-the-art tools Delag, ltl3tela, Rabinizer~4, and SPOT. The
+  technique reduced some acceptance marks in automata produced by all
+  these tools. In the case of Rabinizer~4, the cumulative number of
+  acceptance marks was reduced to less than one half.
 \end{abstract}
@@
 %}}}
-
-\todo{Nezapomenout na clanky o redukci automatu pomoci SATu - on virtue of patience?}
 
@@
 generalize many traditional kinds of automata over infinite words
-including Büchi, co-Büchi, Rabin, Streett, Muller or parity
-automata.
+including Büchi, co-Büchi, Rabin, Streett, and parity automata.
 
@@
 these automata, for example the emptiness check~\cite{baier.19.atva}
-or translation of TELA to parity automata~\cite{renkin.20.atva}.
-
-\todo{Nechcem do experimentu pridat Delag?}
+or translation of TELA to parity automata~\cite{renkin.20.atva}. \todo{Doplnit!}
 
@@
 measures of the input automaton.
+
+\todo{Honza je tu!}
 
@@
 \texttt{telatko}. We show that the tool can reduce acceptance marks in
-automata produced by ltl3tela~\cite{major.19.atva},
-Rabinizer~4~\cite{kretinsky.18.cav}, and SPOT~\cite{duret.16.atva2}.
-While the reduction is relatively modest for TELA produced by ltl3tela
-and SPOT, it is substantial for automata produced by Rabinizer~4.
+automata produced by Delag~\cite{muller.17.gandalf},
+ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav},
+and SPOT~\cite{duret.16.atva2}.  While the reduction is relatively
+modest for TELA produced by ltl3tela and SPOT, it is substantial for
+automata produced by Rabinizer~4.
 
@@
 acceptance marks without increasing the number of states.
+
+There are some SAT-based approaches aimed to reduce the number of
+states of automata over infinite words. Namely, there are reductions
+designed for nondeterministic Büchi automata~\cite{ehlers.10.spin},
+deterministic Büchi automata~\cite{ehlers.10.sat}, and deterministic
+generalized Büchi automata~\cite{baarir.14.forte}. Further, there is
+also a SAT-based approach that transforms a deterministic TELA
+automaton to an equivalent automaton with a given acceptance condition
+and a given number of states~\cite{baarir.15.lpar} (if such an
+automaton exists). Note that all these techniques are usually very
+slow and their authors typically suggest to use them only for specific
+purposes like looking for cases where some automata construction can
+be improved.
 

### 2022-02-23 paper/main.tex
intro and part of prelim. polished

@@
 
-\title{Reducing Acceptance Marks\\ in Emerson-Lei Automata by QBF-solving}
+\title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving}
 \author{Tereza Schwarzov\'{a}}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
@@
   acceptance marks in TELA with use of \emph{quantified Boolean
-    formulas (QBF)}: if a constructed QBF formula is satisfiable then
-  its model encodes an equivalent accepting formula with less
-  acceptance marks. In fact, we present three formula constructions
-  trading the speed of the technique for higher reduction
+    formulas (QBF)}.
+  % : if a constructed QBF formula is satisfiable then
+  % its model encodes an equivalent accepting formula with less
+  % acceptance marks.
+  In fact, we present three formula constructions
+  trading speed for higher reduction
   efficiency. We evaluated our reduction technique on TELA produced by
-  state-of-the-art tools Delag, ltl3tela, Rabinizer~4, and SPOT. The
+  state-of-the-art tools Delag, ltl3tela, Rabinizer~4, and Spot. The
   technique reduced some acceptance marks in automata produced by all
@@
 these automata, for example the emptiness check~\cite{baier.19.atva}
-or translation of TELA to parity automata~\cite{renkin.20.atva}. \todo{Doplnit!}
+or translation of TELA to parity
+automata~\cite{renkin.20.atva,casares.22.tacas}.
 
@@
 
-\todo{Honza je tu!}
-
-These observations motivate our study of techniques reducing the
+These observations motivate our study of a technique reducing the
 number of acceptance marks without altering the structure of the
-automaton. We present basically two reduction techniques. The first
-one is based on relations among individual transition sets defined by
-various acceptance marks in each \emph{strongly connected component
-  (SCC)}. This technique can be seen as an extension of the
-\emph{SCC-base simplifications of acceptance
-  conditions}~\cite{babiak.13.spin} designed for
-\emph{transition-based generalized Büchi automata (TGBA)} and
-implemented in SPOT. Probably the main difference between the original
-technique for TGBA and the one for TELA is in the final step when
-simplified acceptance conditions for each SCC are merged together into
-one reduced acceptance condition.
-
-The second technique is completely different. For a given TELA and
-parameters $c,m$, it produces a \emph{quantified Boolean formula
-  (QBF)} which is satisfiable if and only if there exists an automaton
-with the same structure, $m$ acceptance marks, an acceptance formula
-in disjunctive normal form with $c$ clauses, and the same set
-of accepting runs. The placement of the marks on transitions and the
+automaton. This paper presents a technique heavily based on
+\emph{quantified Boolean formulas (QBF)}. For a given TELA and
+parameters $c,m$, it produces a QBF which is satisfiable if and only
+if there exists an automaton with the same structure, $m$ acceptance
+marks, an acceptance formula in disjunctive normal form with $c$
+clauses, and the same set of accepting runs as the original
+automaton. The placement of the marks on transitions and the
 acceptance formula can be obtained from a model of the
-formula. Besides this formula, we describe also the construction of a
-simpler formula such its satisfiability implies existence of an
-automaton with the same structure, $m$ acceptance marks, and the same
-set of accepting runs, but not vice versa.
-
-We have implemented both reduction techniques in a tool called
+formula. Besides this formula, we describe also the construction of
+two simpler formulas such that their satisfiability implies the
+existence of an automaton with the same structure, $m$ acceptance
+marks, and the same set of accepting runs, but not vice versa.
+
+We have implemented our reduction technique in a tool called
 \texttt{telatko}. We show that the tool can reduce acceptance marks in
@@
 ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav},
-and SPOT~\cite{duret.16.atva2}.  While the reduction is relatively
-modest for TELA produced by ltl3tela and SPOT, it is substantial for
-automata produced by Rabinizer~4.
-
-\subsubsection{Related results}
-We are not aware about any existing research aimed at simplification
-of acceptance formulas of TELA or reduction of the number of its
-acceptance marks without increasing the number of states.
-
-There are some SAT-based approaches aimed to reduce the number of
-states of automata over infinite words. Namely, there are reductions
-designed for nondeterministic Büchi automata~\cite{ehlers.10.spin},
-deterministic Büchi automata~\cite{ehlers.10.sat}, and deterministic
-generalized Büchi automata~\cite{baarir.14.forte}. Further, there is
-also a SAT-based approach that transforms a deterministic TELA
-automaton to an equivalent automaton with a given acceptance condition
-and a given number of states~\cite{baarir.15.lpar} (if such an
-automaton exists). Note that all these techniques are usually very
-slow and their authors typically suggest to use them only for specific
-purposes like looking for cases where some automata construction can
-be improved.
+and Spot~\cite{duret.16.atva2}.  While the reduction is relatively
+modest for TELA produced by ltl3tela and Spot, it is substantial for
+automata produced by Rabinizer~4. \todo{A jak vyjde Delag? Fixnout i
+  Abstract.}
+
+\paragraph*{Related results}
+We are not aware about any existing research primarily aimed at
+simplification of acceptance formulas of TELA or reduction of the
+number of its acceptance marks without increasing the number of
+states. There exists only a SAT-based approach that transforms a
+deterministic TELA automaton to an equivalent automaton with a given
+acceptance condition and a given number of
+states~\cite{baarir.15.lpar} (if such an automaton exists). Further,
+there are some SAT-based approaches aimed to reduce the number of
+states of automata over infinite words. More preceisely, there are
+reductions designed for nondeterministic Büchi
+automata~\cite{ehlers.10.spin}, deterministic Büchi
+automata~\cite{ehlers.10.sat}, and deterministic generalized Büchi
+automata~\cite{baarir.14.forte}. Note that all these techniques are
+usually very slow and their authors typically suggest to use them only
+for specific purposes like looking for cases where some automata
+construction can be improved.
 
@@
 acceptance marks or to simplify the acceptance condition. However,
-such a reduction is not obvious. Moreover, the construction of ACD
-seems to be computationally expensive.
-
-\subsubsection{Structure of the paper} The next section introduces
-basic terms used in the paper. Section~\ref{sec:standard} presents the
-reduction technique based on relations among individual acceptance
-marks on each SCC. The QBF-based technique is explained in
-Section~\ref{sec:qbf}. Implementation is briefly described in
-Section~\ref{sec:experiments}, which also shows the experimental
-results. Finally, Section~\ref{sec:conclusion} suggests other
-applications of our QBF-reduction technique and closes the paper.
+such a reduction is not obvious.
+%Moreover, the construction of ACD seems to be computationally expensive.
+
+\paragraph*{Structure of the paper} The next section introduces basic
+terms used in the paper. Section~\ref{sec:qbf} explains the
+construction of three mentioned QBF formulas. The reduction algorithm
+based on these formulas is described together with its implementation
+in Section~\ref{sec:algimpl}. Section~\ref{sec:experiments} shows the
+experimental results. Finally, Section~\ref{sec:conclusion} suggests
+other applications of our QBF-reduction technique and closes the
+paper.
 
@@
 $q_0=q_I$. Let $M(\pi)$ denote set of all acceptance marks that occur
-in run $\pi$ in infinitely many transitions. Run $\pi$ is
+in infinitely many transitions of run $\pi$. Run $\pi$ is
 \emph{accepting} iff it satisfies the formula $\varphi$, where $\pi$
@@
 % an accepting run over $u$.
+
+\todo{Jan is here}
 
@@
 %}}}
+%{{{ QBF-based reduction
+
+\section{QBF-based reduction}\label{sec:qbf}
+
+The aim of this section is to create a QBF-formula for given $n$ that guarantees the existence of equivalent acceptance condition of a given automaton with $n$ acceptance marks. 
+\smallskip
+
+Let $\mathcal{A}$ with the acceptance condition $\psi$ be an original automaton on the input.
+Let $\mathcal{A}'$ denote automaton with simplified acceptance condition $\psi'$ equivalent to the original automaton $\mathcal{A}$.
+We encode $\psi$ and the placement of acceptance marks into QBF  $\varphi$.
+
+Let $\varphi$ be a QBF formula in \emph{prenex normal form}. Formula $\varphi$
+is composed of quantified prefix $\overrightarrow{Q}$ and three subformulae $\varphi_{O}, \varphi_{N} $ and $\varphi_{C}$.
+The exact shape of the formula is $\varphi = \overrightarrow{Q} .
+\  \varphi_{C} \implies ( \varphi_{O} \Longleftrightarrow \varphi_{N})$.
+Subformula $\varphi_{O}$ expresses properties of $\psi$.
+Subformula $\varphi_{N}$ expresses properties of $\psi'$, namely the number of acceptance marks and a shape of $\psi'$.
+Before we describe the role of $\varphi_{C}$, we need to introduce \emph{relevant set}.
+
+\smallskip
+
+Evaluation of a run of $\mathcal{A}$ is dependent on set of edges that are contained infinitely within the run.
+The set of these edges is from now on referred to as \emph{relevant set}, since only these edges are relevant for the evaluation of $\psi$.
+The \emph{relevant set} is a subset of $\delta_{S_{i}}$ for some SCC $S_{i}$ and any two edges from the \emph{relevant set} are connected by a sequence of edges that also belong to the \emph{relevant set}.
+The purpose of subformula $\varphi_{C}$ is to describe some properties of \emph{relevant sets}.
+For each level of simplification, $\varphi_{C}$ describes these properties more specifically.
+Naturally, the more we specify these properties, the more complex the formula $\varphi_{C}$ is.
+
+\smallskip
+
+Afterwards the formula $\varphi$ is created, we use a QBF-solver to find a model of $\varphi$.
+If no such model exists ($\varphi$ is UNSAT), we know that automaton $\mathcal{A}'$ with $n$ acceptance marks and equivalent to $\mathcal{A}$ does not exist.
+Otherwise, we use the model of $\varphi$ for decoding the automaton $\mathcal{A}'$.
+The structure of $\mathcal{A}'$ remains the same as the structure of $\mathcal{A}$.
+The only thing we change is the acceptance condition $\psi'$ and the placement of acceptance marks on the edges.
+
+\subsection{Motivation}
+Based on the fundamentals given in the previous section, the simplification Level 1 is based on a relationship between acceptance sets of an automaton.
+To achieve finer simplification levels, the relation between the inner structure of the automaton and its acceptance condition needs to be taken into consideration.
+
+As the acceptance condition and the inner structure differ for every specific automaton, finding algorithmic ways of reducing acceptance marks would be complicated.
+Instead, we encode the given automaton into a QBF-formula and let the solver find the redundant marks for us.
+
+\subsection{Encoding}
+ In this section we propose encoding of the relevant parts of a given automaton into QBF-formula.
+\subsubsection{Intuition}
+The goal of the procedure is to find acceptance condition for a particular automaton with less acceptance marks.
+Whilst doing that, we need to ensure that the new automaton stays equivalent to the original automaton because the algorithm not only changes the acceptance condition it also changes the placement of acceptance marks on the edges.
+To do so, we need to ensure that every run on $\mathcal{A}'$ is evaluated the same way as run on $\mathcal{A}$ (these two automata have the same structure).
+
+Since we work with transition-based automata, the evaluation of an acceptance condition is dependent on the edges.
+Intuitively we want formula $\varphi$ to say that for every \emph{relevant set}, the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
+Unfortunately, to identify the exact edges that belong to the \emph{relevant set} of a run is not trivial, since we need to take into consideration every infinite suffix of every run on $\mathcal{A}$.
+Instead we let the formula $\varphi$ say that for every \emph{selection of edges} the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
+For now the \emph{selection of edges} can contain any edges of the automaton.
+Later we restrict the \emph{selection of edges} as we force it to have some properties of \emph{relevant set}.
+
+To express this in the QBF formula, we need to introduce variables that represent edges of the automaton.
+
+Let us enumerate set of edges $\delta$.
+Every edge of the automaton is represented by variable $e_{i}$, where the index corresponds with the enumeration, thus $e_{i}$ represents the edge $t_{i} \in \delta$.
+If such a variable is evaluated as $\true$, it means that the edge is contained in the \emph{selection of edges}.
+Analogously if a variable $e_{i}$ is evaluated as $\false$, it means that the edge is not contained in the \emph{selection of edges}.
+
+\smallskip
+
+If we want to consider every possible \emph{selection of edges}, we need to consider every possible evaluation of every variable that represents an edge.
+
+\todo{tohle bych odstranila}
+\begin{example}
+   Let $t_{1}, t_{2} \in \delta$ be the edges of the automaton from the Figure \ref{fig:elval_of_edges}.
+Since the automaton has two edges, we get four possible evaluation of variables representing the edges.
+    \begin{enumerate}
+        \item $e_{1}$ is $\true \land e_{2}$ is $\true \Longleftrightarrow t_{1}, t_{2}$ are in the \emph{selection of edges}.
+        \item $e_{1}$ is $\true \land e_{2}$ is $\false \Longleftrightarrow t_{1}$ is in the \emph{selection of edges} and $t_{2}$ is not in the \emph{selection of edges}.
+        \item $e_{1}$ is $\false \land e_{2}$ is $\true \Longleftrightarrow t_{1}$ is not in the \emph{selection of edges} and $t_{2}$ is in the \emph{selection of edges}.
+        \item $e_{1}$ is $\false \land e_{2}$ is $\false \Longleftrightarrow t_{1}, t_{2}$ are not in the \emph{selection of edges}.
+    \end{enumerate}
+\end{example}
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+			\node[state, initial] (z) at (0,0) {};
+
+			\path[->]
+
+            	(z) edge[loop, in = 70, out = 110, looseness=6] node [above, scale=0.7] {1} (z)
+            	(z) edge[loop, in = 68, out = 112, looseness=10] node [above, scale=0.7] {2} (z)
+			;
+		\end{tikzpicture}
+    \caption{TELA with enumerated edges}
+    \label{fig:elval_of_edges}
+\end{figure}
+
+Hereby the expressive power of quantified Boolean formula comes to use.
+If we universally quantify every variable $e_{i}$ that represents an edge $t_{i} \in \delta$, then we consider the situation when $t_{i}$ is contained in the \emph{selection of edges} and the situation when $t_{i}$ is not contained in the \emph{selection of edges}.
+
+That being said, we can proceed to the encoding of some information about automaton into the QBF-formula $\varphi$.
+As mentioned before, the formula $\varphi$ expresses that for every \emph{selection of edges}, the acceptance condition $\psi$ is evaluated as $\true$ $\mathit{iff}$ the acceptance condition $\psi'$ is evaluated as true.
+
+The subformula $\varphi_{O}$ encodes the shape of the acceptance condition $\psi$ and information which edges are labeled with which acceptance marks.
+The evaluation of $\varphi_{O}$ depends purely on the evaluation of variables that represent edges of the automaton.
+It is encoded in such way that the subformula $\varphi_{O}$ is satisfied iff the edges of \emph{selection of edges} satisfy $\psi$.
+The details of $\varphi_{O}$ encoding are described in Subsection \ref{subsection:original}.
+
+\smallskip
+
+The subformula $\varphi_{N}$ contains information about the number of acceptance marks we want in the new acceptance formula $\psi'$.
+Further, it contains free variables.
+The model of these free variables gives us the information about the shape of $\psi'$ and the placement of acceptance marks on the edges of $\mathcal{A}'$.
+The details of $\varphi_{N}$ encoding are described in Subsection \ref{subsection:new}.
+
+The quantification of all variables that represent the edges of the automaton ensures that we consider every possible \emph{selection of edges}, as described in Section \ref{subsection:quant_pref}. 
+\begin{center}
+    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . ( \varphi_{O} \Longleftrightarrow \varphi_{N} ) $
+\end{center}
+However, one can clearly notice that some edges will never be visited infinitely often in the same run.
+(For example, when some of the edges occur in different strongly connected components than the others.) In other words, these edges will never belong to any \emph{relevant set} together.
+Therefore it is possible to eliminate every evaluation of QBF, where variables representing such edges are evaluated as $\true$.
+It is natural to express this constrain as an implication.
+Let $\varphi_{C}$ be a formula describing that the \emph{selection of edges} has a potential be a \emph{relevant set}.
+The QBF representation is:
+\begin{center}
+    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . \varphi_{C} \implies (\varphi_{O} \Longleftrightarrow \varphi_{N})$
+\end{center}
+The matter of eliminating various combinations of edges is described in more detail in Subection \ref{subsection:connection}.
+
+\subsubsection{Variables}
+In this section, we describe the syntax and intended semantics of variables, parameters, and auxiliary letters used in this section.
+Hereby is the overview of the parameters used when building the formula:
+\begin{itemize}
+    \item K -- The number of acceptance marks in $\psi'$.
+    \item C -- the number of clauses in formula $\psi'$.
+\end{itemize}
+
+These parameters are derived from the original automaton $\mathcal{A}$. As the algorithm iterates (see Subection \ref{subsection:procedure} )  the parameter $K$ decreases.
+
+\smallskip
+
+The auxiliary letters have the following function:
+\begin{itemize}
+    \item $k$ -- the number of particular acceptance mark
+    \item $c$ -- the number of particular clause
+    \item $i$ -- the number of particular edge 
+\end{itemize}
+
+Hereby is the overview of the variables used when building the QBF formula $\varphi$ :
+\begin{itemize}
+    \item $$f_{i,k} = \begin{cases}
+    1& \text{edge }t_{i} \text{ is labeled with the acceptance mark }k\\
+    0& \text{else}
+\end{cases}$$
+\item $$p_{c,k} = \begin{cases}
+    1& \Inf(k) \text{ is in the clause }c\\
+    0& \text{else}
+\end{cases}$$
+\item $$n_{c,k}= \begin{cases}
+    1& \Fin(k) \text{ is in the clause }c\\
+    0& \text{else}
+\end{cases}$$
+\item $$e_{i} \;= \begin{cases}
+    1& \text{The edge } t_{i} \text{ is in the \emph{selection of edges} }\\
+    0& \text{else}
+\end{cases}$$
+\end{itemize}
+
+\subsection{QBF decoding}
+After we build $\varphi$, we use QBF-solver to find satisfying assignment of variables.
+If such a valuation exists, we can build a new automaton $\mathcal{A}'$ with the same structure and fewer acceptance marks.
+This new automaton is equivalent to the original one $\mathcal{A}$ -- this is guaranteed by the QBF-formula $\varphi$.
+There are two possible outcomes from the QBF-solver:
+\begin{enumerate}
+    \item $\varphi$ is UNSAT -- a model of $\varphi$ does not exist.
+That means that the acceptance condition of $\mathcal{A}$ cannot be simplified for given C and K.
+Thus we return automaton $\mathcal{A}$ and the information that $\mathcal{A}$ is the automaton with the simplest acceptance condition possible.
+    \item $\varphi$ is SAT -- QBF-solver provides (returns) a model of unquantified variables.
+Namely to $p_{c, k},n_{c,k} \text{ and } f_{i,k}, \text{ for each }1 \leq c \leq C, 1 \leq k \leq K, 1 \leq i \leq |\delta|$.
+    Firstly, we decode the variables $f_{i,k}$ and use them to identify new placement of the acceptance marks.
+Specifically, we create a new automaton $\mathcal{A}'$ with the same structure as $\mathcal{A}$.
+Then  we add acceptance mark $\tacc{5}{k}$ on edge $t_{i}$ for every $f_{i,k}$ variable evaluated as $\true$.
+
+    The new acceptance condition of automaton $\mathcal{A}'$ is then build from variables $p_{c,k}$ and $n_{c,k}$.
+The second part of $\varphi_{N}$ was encoded to produce new acceptance formula in DNF.
+(See the alternation between Boolean operators  in order $\lor, \land$ in Section \ref{subsection:new}.) Thus the shape of the new acceptance condition is set as DNF skeleton, and we simply add terms according to the evaluation of variables.
+If $p_{c,k}$ is $\true$, we add term $\Inf\tacc{5}{k}$ to the disjunct $c$.
+Similarly, if $n_{c,k}$ is $\true$, then we add $\Fin\tacc{5}{k}$ to the disjunct $c$.
+This way, we process every variable that QBF-solver gives us and produce the new automaton $\mathcal{A}'$
+    equivalent to the original one.
+
+\end{enumerate}
+
+\subsection{Quantified prefix}
+\label{subsection:quant_pref}
+The aim of the quantified prefix in formula $\varphi$ is to consider every possible selection of edges.
+As mentioned in the previous sections, edge $t_{i} \in \delta$ belongs to the \emph{selection of edges} if the variable $e_{i}$ that represents the edge $t_{i}$ is evaluated as $\true$.
+Therefore if we want to consider every possible \emph{selection of edges}, we need to consider every evaluation of the variables representing edges.
+We use a universal quantifier to express this.
+The universal quantifier states that the formula has to hold for both evaluations of the quantified variable.
+
+The QBF prefix that expresses the universal quantification for every variable representing an edge has the following shape:
+\begin{align*}
+    \overrightarrow{Q} = \forall\{e_{1}, \ldots ,e_{|\delta|} \}
+\end{align*}
+
+The advantage of QBF is that it is able to express a statement in shorter formula then if we tried to express the statement in standard Boolean formula with the use of conjunction.
+
+\begin{example}
+Consider the QBF $\theta = \forall e_{1} \forall e_{2} .
+(e_1 \lor e_2) $.
+The QBF $\theta $ transformed to standard Boolean formula is $\theta' = (e_{1} \lor e_{2}) \land (  e_{1} \lor \lnot e_{2}) \land  (  \lnot e_{1} \lor e_{2}) \land  (  \lnot e_{1} \lor \lnot e_{2}) .$
+\end{example}
+
+\subsection{Original acceptance formula}
+\label{subsection:original}
+
+As mentioned before, the satisfiability of $\varphi_{O}$ is purely resultant of assessment of $e_{i}, 1 \leq i \leq |\delta|$ variables representing edges in automaton.
+
+Let us enumerate the edges of the original automaton $\mathcal{A}$ and let $Z(m), m \in M$ be the set of edges labeled with mark $m$. 
+The encoding of original acceptance condition terms into $\varphi_{O}$ is defined as follows:
+%%davat tam rovnitko nebo nejak sipku??
+\begin{align*}
+    &\varphi(\Inf(m))=\bigvee\limits_{t_{i} \in Z(m)} e_{i}\\
+    &\varphi(\Fin(m)) = \bigwedge\limits_{t_{i} \in Z(m)} \lnot e_{i}
+\end{align*}
+\begin{example}
+
+Consider the automaton in Figure \ref{fig:oficial_acc}.
+The set $Z_{0} = \{t_3, t_2\},\, Z_{1} = \{t_1\}$, where $t_{1}, t_{2}, t_{3} \in \delta$.
+Thus the encoded $\varphi_{O} = e_1 \land (\lnot e_2 \land \lnot e_3)$.
+Therefore if variable $e_{1}=\true$, $e_{2}=\false$ and $e_{3}=\false$, acceptance formula is $\true$.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+\node[cstate, initial] (g) {};
+\node[flabel, above] at (0, 1.5) {$\Fin\taccsq{0}{0}\land\Inf\tacc{1}{1}$};
+%
+\path[->]
+    (g) edge[loop, in=330, out=30, looseness=6]node[right, pos=0.4, scale=0.8] {$1$} node[accset,collacc1, pos=0.6] {1}(g)
+	(g) edge[loop, in=80, out=140, looseness=6] node[above, pos= 0.4, scale=0.8] {$2$} node[accsquare, collacc0, pos=0.6]{0}(g)
+	(g) edge[loop, in=220, out=280, looseness=6] node[below, pos=0.6, scale=0.8] {$3$} node[accsquare, collacc0, pos=0.4] {0} (g)
+	;
+\end{tikzpicture}
+    \caption{TELA with edges labeled with their numbers}
+    \label{fig:oficial_acc}
+\end{figure}
+
+\end{example}
+
+\subsection{New acceptance formula}
+\label{subsection:new}
+The new acceptance formula $\varphi_{N}$ determines the shape of the future acceptance condition of automaton, and the assignment of variables used in $\varphi_{N}$ determines new positions of acceptance marks on the transitions.
+
+The formula $\varphi_{N}$ is divided into two parts.
+The first part of the formula (4.1) expresses a simple constrain that the acceptance set that occurs in one clause cannot be in $\Inf$ and simultaneously in $\Fin$ form.
+
+The second part (4.2) of $\varphi_{N}$ is divided into two subformulas.
+The first subformula expresses that if the term $\Inf(k)$ is present in clause $c$, then at least one edge of the \emph{selection of edges} needs to be labeled with the acceptance mark $k$.
+
+Similarly, the second subformula implies that if the term $\Fin(k)$ is present in clause $c$, then none of the \emph{selected edges} can be labeled with the acceptance mark $k$.
+
+Using the variables presented in previous sections, we encode $\varphi_{N}$:
+\begin{align}
+    \varphi_{N} = & \bigwedge\limits_{1 \leq c \leq C}\;\bigwedge\limits_{1 \leq k \leq K} (\lnot p_{c,k} \lor \lnot n_{c,k}) \; \land \\
+    &\bigvee\limits_{1 \leq c \leq C} \; \bigwedge\limits_{1 \leq k \leq K} \Bigg[\big(  p_{c, k} \implies \bigvee\limits_{t_{i} \in \delta}(f_{i,k} \land e_{i})  \big) \land \\& \;\;\;\quad\quad\quad\quad\quad\quad\big(  n_{c, k} \implies \bigwedge\limits_{t_{i} \in \delta}(\lnot f_{i,k} \lor \lnot e_{i})  \big)\Bigg]
+\end{align}
+
+Observe the alternation of $\lor \land$ from the second part of the formula (4.2). This alternation determines that the formula $\psi'$ is in DNF.
+
+\subsection{Connection formula}
+\label{subsection:connection}
+The connection formula distinguishes the remaining two levels of simplification. Previously we denoted it as $\varphi_{C}$.
+
+Formula $\varphi_{C}$ defines requirements on the edges that are together in \emph{selection of edges}.
+These requirements are based on some properties of these edges.
+The more restrictive these requirements are, the more properties of \emph{relevant sets} are possessed by the \emph{selection of edges}.
+In other words, formula $\varphi_{C}$ holds if the \emph{selection of edges} has the required property.
+If not, the formula $\varphi_{C}$ is $\false$, and then the whole formula $\varphi$ is trivially satisfied.
+Formula $\varphi_{C}$ consists of two parts $\varphi_{C_{1}}$ and $\varphi_{C_{2}}$.
+
+The first part expresses the requirement that the \emph{selection of edges} is \emph{continuous}.
+Loosely said, this means that if a transition from \emph{selection of edges} enters some state $q$ then another transition that belongs to the \emph{selection of edges} exits the state $q$.
+This holds in the opposite direction as well.
+Thus if a transition from \emph{selection of edges} exits some state $q$ then another transition that belongs to the \emph{selection of edges} enters the state $q$.
+This is naturally expressed by equivalence.
+Note that if the selected transition is a self-loop, the destination and the source of the edge is the same state.
+
+The Boolean formula for the described statement:
+\begin{center}
+    $\varphi_{C_{1}}=\bigwedge\limits_{q \in Q} \Bigg( ( \bigvee\limits_{t_{i} \in \delta \cap \{ q \} \times  \Sigma \times  2^M \times Q} e_i ) \Longleftrightarrow( \bigvee\limits_{t_{j} \in \delta  \cap \ Q \times  \Sigma \times 2^M \times \{q\}} e_{j} ) \Bigg)$
+\end{center}
+
+\begin{example}
+   Consider the state $q$ in Figure \ref{fig:inNout} this part of formula is $(e_1 \lor e_2) \Longleftrightarrow (e_3 \lor e_4 \lor e_5)$.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+        \node[state] (q) at (0, 0) {$q$};
+
+	    \path[->]
+		(q) edge node[above, pos=0.4, scale=0.8] {$4$} ++(1.5, 0)
+		(q) edge node[above, pos=0.4, scale=0.8] {$5$} ++(1.3, 1)
+		(q) edge node[above, pos=0.4, scale=0.8] {$3$} ++(1.3, -1)
+	;
+	\path[<-]
+	(q) edge node[above, pos=0.6, scale=0.8] {$2$} ++(-1.3, -0.5)
+	(q) edge node[above, pos=0.6, scale=0.8] {$1$} ++(-1.3, 0.5)
+	;
+    \end{tikzpicture}
+    \caption{State $q$ and its entering and exiting edges}
+    \label{fig:inNout}
+\end{figure}
+\end{example}
+
+Even though we ensured that the selected transitions create continuous paths, the paths might not be connected.
+
+\begin{example}
+Consider the automaton in Figure \ref{fig:continuousNotConnected}.
+The transitions $t_1$ and $t_2$ are loops (the simplest cycles), but are not connected (each is located in different SCC).
+Thus $\{ t_{1}, t_{2}\}$ is not a \emph{relevant set}.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+			\node[cstate, initial] (i) at (0, 0) {};
+			\node[cstate] (j) at (2.5, 1) {};
+			\node[cstate] (k) at (2.5, -1) {};
+			\path[->]
+
+			(i) edge  (k)
+			(i) edge (j)
+			(j) edge[loop right, draw=red] node[right] {$1$} (j)
+			(k) edge[loop right, draw=red] node[right] {$2$} (k)
+
+		;
+    \end{tikzpicture}
+    \caption{Not connected cycles}
+    \label{fig:continuousNotConnected}
+\end{figure}
+\end{example}
+
+The second part of $\varphi_{C}$ specifies the properties of \emph{relevant set} that the edges that belong together to \emph{selected edges} need to have.
+We divided these properties into two levels Level 2 and Level 3, all described in ensuing sections.
+
+\subsubsection{SCC accuracy}
+The SCC accuracy corresponds with the Level of simplification 2. At this level, the formula specifies that the edges from one \emph{selection of edges} belong to the same SCC.
+ Thus straightforwardly translated into QBF formula:
+\begin{center}
+    $\varphi_{C_{2}} = \bigvee\limits_{\delta_{S_{i}} \subseteq \delta}\Bigg( ( \bigvee\limits_{t_{i} \in \delta_{S_{i}}} e_i ) \land  ( \bigwedge\limits_{t_{j} \in \delta \setminus \delta_{S_{i}}} \lnot e_j  ) \Bigg)$
+\end{center}
+
+Finally, the connection formula is the conjunction of the two parts:
+\begin{center}
+    $\varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}$.
+\end{center}
+
+\begin{example}
+This example works as a motivation for simplification Level 3.
+For simplification Level 2, edges number $1$ and number $3$ in Figure \ref{fig:connected_motivation} can be in the same \emph{selection of edges} because they are placed in the same SCC.
+Obviously, since they are not connected, they cannot form \emph{relevant set}.
+Therefore we can specify a new level of simplification -- Level 3, which forces all edges from the \emph{selection of edges} to be cycles.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+    \node[state, initial] (g) {$0$};
+    \node[state, right=2cm of g](q){$1$};
+    %
+    \path[->, auto]
+    (g) edge[bend right] node[below, scale=0.8] {$2$}(q)
+    (q) edge[bend right]node[above, scale=0.8] {$4$} (g)
+    (q) edge[loop, in=65, out=115, looseness=10, draw=red] node[ scale=0.8] {$3$}(q)
+    (g) edge[loop, in=65, out=115, looseness=10, draw=red] node[scale=0.8] {$1$}(g)
+	;
+
+\end{tikzpicture}
+
+    \caption{Motivation for simplification L3}
+    \label{fig:connected_motivation}
+\end{figure}
+
+\end{example}
+
+\subsubsection{CYCLES accuracy}
+The simplification Level 2 ensures that \emph{selection of edges} is placed in one SCC, but we can go further and specify that \emph{selection of edges} is in fact a \emph{relevant set}.
+
+The intuition by simplification Level 3 is that the states of an automaton are  divided into two groups -- the states that belong to the group of \emph{selected states} and the states do not belong to the group of \emph{selected states}.
+Then we require that the states that belong to the same group are connected.
+We consider three possible scenarios that can occur while the states are divided into these two groups:
+\begin{enumerate}
+    \item The source of edge $t_{i}$ is in different group than the destination of edge $t_{i}$.
+    \item Source and destination of edge $t_{i}$ are in the \emph{selected states}.
+    \item Source and destination of edge $t_{i}$ are not in \emph{selected states}.
+\end{enumerate}
+Cases two and three do not violate the requirement that all connected states need to be in one group.
+Case one is a little more complicated.
+If an edge has a source in \emph{selected states} and destination not in the \emph{selected states}, we need to secure that the sequence of the edges returns to \emph{selected states}.
+In other words, there exists another edge that has a source in the not \emph{selected states} and destination in the \emph{selected states}.
+(Dually the other way around.)
+
+In order to express this in the QBF-formula we introduce a new set of quantified variables $\{ w_{1}, \ldots , w_{|Q|}$. The variables have the following semantics:
+\begin{align*}
+    w_{q} = \begin{cases}
+    1& \text{state } q \in Q \text{ is in the \emph{selected group}}\\
+    0& \text{else}
+    \end{cases}
+\end{align*}
+Since we want to ensure that for every \emph{selection of edges} (for every evaluation of variables that represent the edges) there exist such a groups we update the quantified prefix  $\overrightarrow{Q}$ in following way:
+\begin{align*}
+    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
+\end{align*}
+
+\smallskip
+\todo{neoptimalni}
+Since we want the statement to hold for every \emph{selection of edges}, we place the existentially quantified variables $\{w_{1}, \ldots, w_{|Q|} \}$ behind the universally quantified variables $\{e_{1}, \ldots, e_{|\delta|} \} $.
+At first we update the quantified prefix:
+\begin{align*}
+    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
+\end{align*}
+
+And then we can create the $\varphi_{C}$ formula:
+\begin{align*}
+    \varphi_{C} =  \Bigg[ &\bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (w_{q} \land w_{r}) \Bigg) \lor \\
+    &  \bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma  \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (\lnot w_{q} \land \lnot w_{r}) \Bigg) \lor \\ \Bigg( &\bigvee \limits_{\substack{t_{i} \in \{q\} \times \Sigma \times 2^M  \times \{ r \}, \\ q \in Q, r \in Q}} ( e_{i} \land w_{q} \land \lnot w_{r} )  \land \bigvee \limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} (e_{i} \land \lnot  w_{q} \land  w_{r}) \Bigg) \Bigg]
+\end{align*}
+
+\begin{example}
+Figure \ref{fig:cycles_sensitivity} displays an example of an automaton reduced by Level 4.
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[automaton]
+
+    \node[state, initial] (g) {$0$};
+
+    \node[flabel] at (1.5, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{1}{1} \land \Inf\tacc{2}{2}$};
+    \node[state, right=2cm of g](q){$1$};
+
+    \path[->, auto]
+    (g) edge[bend right] node [accset,collacc0, pos=0.5]{0}node [below, pos=0.2, scale=0.8] {$2$}(q)
+    (q) edge[bend right] node[above, pos=0.8, scale=0.8] {$4$}(g)
+    (q) edge[loop, in=75, out=105, looseness=8]node[left, pos=0.1, scale=0.8]{5} (q)
+    (q) edge[loop, in=70, out=110, looseness=12] node[accset,collacc1] {1}node[left, pos=0.2, scale=0.8]{6} (q)
+    (g) edge[loop, in=75, out=105, looseness=8] node[accset,collacc2] {2} node[left, pos=0.1, scale=0.8]{1}(g)
+    (g) edge[loop, in=70, out=110, looseness=12] node[accset,collacc2, pos=0.4] {2} node[accset,collacc1, pos=0.7] {1} node[left, pos=0.2, scale=0.8]{3}(g)
+	;
+
+	\node[state, initial, right=4cm of g] (o){$0$};
+
+    \node[flabel] at (6.7, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{3}{3}$};
+    \node[state, right=2cm of o](p){$1$};
+
+    \path[->, auto]
+    (o) edge[bend right] node [accset,collacc0, pos=0.5]{0}(p)
+    (p) edge[bend right](o)
+    (p) edge[loop, in=75, out=105, looseness=8] (p)
+    (p) edge[loop, in=70, out=110, looseness=12] (p)
+    (o) edge[loop, in=75, out=105, looseness=8] (o)
+    (o) edge[loop, in=70, out=110, looseness=12] node[accset,collacc3] {3}(o)
+
+	;
+
+\end{tikzpicture}
+
+    \caption{TELA reduced with the CYCLES sensitivity}
+    \label{fig:cycles_sensitivity}
+\end{figure}
+\end{example}
+
+Similarly as in the previous section the connection formula is the conjunction of the two parts:
+\begin{align*}
+    \varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}.
+\end{align*}
+
+\subsection{Simplification procedure}
+\label{subsection:procedure}
+
+Now we describe our simplification algorithm taking as input an arbitrary TELA $\mathcal{A}$ with an acceptance condition $\psi$, required level of simplification $L$ and timeout period $T$.
+\begin{enumerate}
+    \item Convert acceptance condition $ \psi$ to DNF.
+    \item Apply the algorithm for simplification Level 1 described in Section \ref{sec:standard} on $\mathcal{A}$.
+    \item If $L = 1$ return automaton $\mathcal{A}$.
+     \item The initial value of $C$ is the number of clauses in $\psi$. Similarly, the initial value of $K$ is the number of acceptance sets of $\psi$ decreased by 1.
+    \item Encode $\mathcal{A}$ into QBF formula $\varphi$ according to the required level of simplification $L$.
+    \item Run a QBF-solver on $\varphi$ for $T$ seconds.
+    \begin{enumerate}
+        \item If the QBF-solver evaluates $\varphi$ as UNSAT, return automaton $\mathcal{A}'$.
+        \item If the timeout $T$ expires before the QBF-solver evaluates $\varphi$, return automaton $\mathcal{A}'$.
+        \item If the QBF-solver evaluates $\varphi$ as SAT. Decode the assignment into next TELA $\mathcal{A}'$.
+        \begin{enumerate}
+            \item If $K=1$, try to reduce $\varphi$ to $\true$ or $\false$, the simplest form of acceptance condition and return $\mathcal{A}$.
+            \item Decrease $K$ by 1 and go to (b).
+
+    \end{enumerate}
+
+    \end{enumerate}
+\end{enumerate}
+
+\todo{kde je optimalni misto, kam napsat:}
+For simplicity, we do not display letters on the edges of the automata in Figures. We can afford to omit them because they are not important for our simplification techniques.
+
+\begin{itemize}
+\item obecny popis konstrukce formule
+\item (level 3)
+\item level 4
+\end{itemize}
+
+%}}}
+%{{{ algorithm and implementation
+
+\section{Reduction algorithm and its implementation}\label{sec:algimpl}
+Jak se nastavuji inicialni parametry a jak postupne klesaji.
+Jak prechazim od nizsich levelu k vyssim.
+Vyuzivame incremental QBF solving?
+
+%}}}
+%{{{ experimental evaluation
+
+\section{Experimental evaluation}\label{sec:experiments}
+
+%}}}
+%{{{ conclusion
+
+\section{Conclusions}\label{sec:conclusion}
+
+chceme sem napsat, ze QBF-based reduction se da pouzit napr. k
+nalezeni akceptacni formule s $c$ klauzulemi a $m$ znackami v CNF,
+pripadne k nalezeni ekvivalentniho paritniho automat s danou maximalni
+paritou $p$ a stejnou strukturou.
+
+%}}}
+
+%\bibliographystyle{plain}
+\bibliography{mc}
+
+\end{document}
+
 %{{{ standard reductions
@@
 %}}}
-%{{{ QBF-based reduction
-
-\section{QBF-based reduction}\label{sec:qbf}
-
-The aim of this section is to create a QBF-formula for given $n$ that guarantees the existence of equivalent acceptance condition of a given automaton with $n$ acceptance marks. 
-\smallskip
-
-Let $\mathcal{A}$ with the acceptance condition $\psi$ be an original automaton on the input.
-Let $\mathcal{A}'$ denote automaton with simplified acceptance condition $\psi'$ equivalent to the original automaton $\mathcal{A}$.
-We encode $\psi$ and the placement of acceptance marks into QBF  $\varphi$.
-
-Let $\varphi$ be a QBF formula in \emph{prenex normal form}. Formula $\varphi$
-is composed of quantified prefix $\overrightarrow{Q}$ and three subformulae $\varphi_{O}, \varphi_{N} $ and $\varphi_{C}$.
-The exact shape of the formula is $\varphi = \overrightarrow{Q} .
-\  \varphi_{C} \implies ( \varphi_{O} \Longleftrightarrow \varphi_{N})$.
-Subformula $\varphi_{O}$ expresses properties of $\psi$.
-Subformula $\varphi_{N}$ expresses properties of $\psi'$, namely the number of acceptance marks and a shape of $\psi'$.
-Before we describe the role of $\varphi_{C}$, we need to introduce \emph{relevant set}.
-
-\smallskip
-
-Evaluation of a run of $\mathcal{A}$ is dependent on set of edges that are contained infinitely within the run.
-The set of these edges is from now on referred to as \emph{relevant set}, since only these edges are relevant for the evaluation of $\psi$.
-The \emph{relevant set} is a subset of $\delta_{S_{i}}$ for some SCC $S_{i}$ and any two edges from the \emph{relevant set} are connected by a sequence of edges that also belong to the \emph{relevant set}.
-The purpose of subformula $\varphi_{C}$ is to describe some properties of \emph{relevant sets}.
-For each level of simplification, $\varphi_{C}$ describes these properties more specifically.
-Naturally, the more we specify these properties, the more complex the formula $\varphi_{C}$ is.
-
-\smallskip
-
-Afterwards the formula $\varphi$ is created, we use a QBF-solver to find a model of $\varphi$.
-If no such model exists ($\varphi$ is UNSAT), we know that automaton $\mathcal{A}'$ with $n$ acceptance marks and equivalent to $\mathcal{A}$ does not exist.
-Otherwise, we use the model of $\varphi$ for decoding the automaton $\mathcal{A}'$.
-The structure of $\mathcal{A}'$ remains the same as the structure of $\mathcal{A}$.
-The only thing we change is the acceptance condition $\psi'$ and the placement of acceptance marks on the edges.
-
-\subsection{Motivation}
-Based on the fundamentals given in the previous section, the simplification Level 1 is based on a relationship between acceptance sets of an automaton.
-To achieve finer simplification levels, the relation between the inner structure of the automaton and its acceptance condition needs to be taken into consideration.
-
-As the acceptance condition and the inner structure differ for every specific automaton, finding algorithmic ways of reducing acceptance marks would be complicated.
-Instead, we encode the given automaton into a QBF-formula and let the solver find the redundant marks for us.
-
-\subsection{Encoding}
- In this section we propose encoding of the relevant parts of a given automaton into QBF-formula.
-\subsubsection{Intuition}
-The goal of the procedure is to find acceptance condition for a particular automaton with less acceptance marks.
-Whilst doing that, we need to ensure that the new automaton stays equivalent to the original automaton because the algorithm not only changes the acceptance condition it also changes the placement of acceptance marks on the edges.
-To do so, we need to ensure that every run on $\mathcal{A}'$ is evaluated the same way as run on $\mathcal{A}$ (these two automata have the same structure).
-
-Since we work with transition-based automata, the evaluation of an acceptance condition is dependent on the edges.
-Intuitively we want formula $\varphi$ to say that for every \emph{relevant set}, the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
-Unfortunately, to identify the exact edges that belong to the \emph{relevant set} of a run is not trivial, since we need to take into consideration every infinite suffix of every run on $\mathcal{A}$.
-Instead we let the formula $\varphi$ say that for every \emph{selection of edges} the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
-For now the \emph{selection of edges} can contain any edges of the automaton.
-Later we restrict the \emph{selection of edges} as we force it to have some properties of \emph{relevant set}.
-
-To express this in the QBF formula, we need to introduce variables that represent edges of the automaton.
-
-Let us enumerate set of edges $\delta$.
-Every edge of the automaton is represented by variable $e_{i}$, where the index corresponds with the enumeration, thus $e_{i}$ represents the edge $t_{i} \in \delta$.
-If such a variable is evaluated as $\true$, it means that the edge is contained in the \emph{selection of edges}.
-Analogously if a variable $e_{i}$ is evaluated as $\false$, it means that the edge is not contained in the \emph{selection of edges}.
-
-\smallskip
-
-If we want to consider every possible \emph{selection of edges}, we need to consider every possible evaluation of every variable that represents an edge.
-
-\todo{tohle bych odstranila}
-\begin{example}
-   Let $t_{1}, t_{2} \in \delta$ be the edges of the automaton from the Figure \ref{fig:elval_of_edges}.
-Since the automaton has two edges, we get four possible evaluation of variables representing the edges.
-    \begin{enumerate}
-        \item $e_{1}$ is $\true \land e_{2}$ is $\true \Longleftrightarrow t_{1}, t_{2}$ are in the \emph{selection of edges}.
-        \item $e_{1}$ is $\true \land e_{2}$ is $\false \Longleftrightarrow t_{1}$ is in the \emph{selection of edges} and $t_{2}$ is not in the \emph{selection of edges}.
-        \item $e_{1}$ is $\false \land e_{2}$ is $\true \Longleftrightarrow t_{1}$ is not in the \emph{selection of edges} and $t_{2}$ is in the \emph{selection of edges}.
-        \item $e_{1}$ is $\false \land e_{2}$ is $\false \Longleftrightarrow t_{1}, t_{2}$ are not in the \emph{selection of edges}.
-    \end{enumerate}
-\end{example}
-
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-			\node[state, initial] (z) at (0,0) {};
-
-			\path[->]
-
-            	(z) edge[loop, in = 70, out = 110, looseness=6] node [above, scale=0.7] {1} (z)
-            	(z) edge[loop, in = 68, out = 112, looseness=10] node [above, scale=0.7] {2} (z)
-			;
-		\end{tikzpicture}
-    \caption{TELA with enumerated edges}
-    \label{fig:elval_of_edges}
-\end{figure}
-
-Hereby the expressive power of quantified Boolean formula comes to use.
-If we universally quantify every variable $e_{i}$ that represents an edge $t_{i} \in \delta$, then we consider the situation when $t_{i}$ is contained in the \emph{selection of edges} and the situation when $t_{i}$ is not contained in the \emph{selection of edges}.
-
-That being said, we can proceed to the encoding of some information about automaton into the QBF-formula $\varphi$.
-As mentioned before, the formula $\varphi$ expresses that for every \emph{selection of edges}, the acceptance condition $\psi$ is evaluated as $\true$ $\mathit{iff}$ the acceptance condition $\psi'$ is evaluated as true.
-
-The subformula $\varphi_{O}$ encodes the shape of the acceptance condition $\psi$ and information which edges are labeled with which acceptance marks.
-The evaluation of $\varphi_{O}$ depends purely on the evaluation of variables that represent edges of the automaton.
-It is encoded in such way that the subformula $\varphi_{O}$ is satisfied iff the edges of \emph{selection of edges} satisfy $\psi$.
-The details of $\varphi_{O}$ encoding are described in Subsection \ref{subsection:original}.
-
-\smallskip
-
-The subformula $\varphi_{N}$ contains information about the number of acceptance marks we want in the new acceptance formula $\psi'$.
-Further, it contains free variables.
-The model of these free variables gives us the information about the shape of $\psi'$ and the placement of acceptance marks on the edges of $\mathcal{A}'$.
-The details of $\varphi_{N}$ encoding are described in Subsection \ref{subsection:new}.
-
-The quantification of all variables that represent the edges of the automaton ensures that we consider every possible \emph{selection of edges}, as described in Section \ref{subsection:quant_pref}. 
-\begin{center}
-    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . ( \varphi_{O} \Longleftrightarrow \varphi_{N} ) $
-\end{center}
-However, one can clearly notice that some edges will never be visited infinitely often in the same run.
-(For example, when some of the edges occur in different strongly connected components than the others.) In other words, these edges will never belong to any \emph{relevant set} together.
-Therefore it is possible to eliminate every evaluation of QBF, where variables representing such edges are evaluated as $\true$.
-It is natural to express this constrain as an implication.
-Let $\varphi_{C}$ be a formula describing that the \emph{selection of edges} has a potential be a \emph{relevant set}.
-The QBF representation is:
-\begin{center}
-    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . \varphi_{C} \implies (\varphi_{O} \Longleftrightarrow \varphi_{N})$
-\end{center}
-The matter of eliminating various combinations of edges is described in more detail in Subection \ref{subsection:connection}.
-
-\subsubsection{Variables}
-In this section, we describe the syntax and intended semantics of variables, parameters, and auxiliary letters used in this section.
-Hereby is the overview of the parameters used when building the formula:
-\begin{itemize}
-    \item K -- The number of acceptance marks in $\psi'$.
-    \item C -- the number of clauses in formula $\psi'$.
-\end{itemize}
-
-These parameters are derived from the original automaton $\mathcal{A}$. As the algorithm iterates (see Subection \ref{subsection:procedure} )  the parameter $K$ decreases.
-
-\smallskip
-
-The auxiliary letters have the following function:
-\begin{itemize}
-    \item $k$ -- the number of particular acceptance mark
-    \item $c$ -- the number of particular clause
-    \item $i$ -- the number of particular edge 
-\end{itemize}
-
-Hereby is the overview of the variables used when building the QBF formula $\varphi$ :
-\begin{itemize}
-    \item $$f_{i,k} = \begin{cases}
-    1& \text{edge }t_{i} \text{ is labeled with the acceptance mark }k\\
-    0& \text{else}
-\end{cases}$$
-\item $$p_{c,k} = \begin{cases}
-    1& \Inf(k) \text{ is in the clause }c\\
-    0& \text{else}
-\end{cases}$$
-\item $$n_{c,k}= \begin{cases}
-    1& \Fin(k) \text{ is in the clause }c\\
-    0& \text{else}
-\end{cases}$$
-\item $$e_{i} \;= \begin{cases}
-    1& \text{The edge } t_{i} \text{ is in the \emph{selection of edges} }\\
-    0& \text{else}
-\end{cases}$$
-\end{itemize}
-
-\subsection{QBF decoding}
-After we build $\varphi$, we use QBF-solver to find satisfying assignment of variables.
-If such a valuation exists, we can build a new automaton $\mathcal{A}'$ with the same structure and fewer acceptance marks.
-This new automaton is equivalent to the original one $\mathcal{A}$ -- this is guaranteed by the QBF-formula $\varphi$.
-There are two possible outcomes from the QBF-solver:
-\begin{enumerate}
-    \item $\varphi$ is UNSAT -- a model of $\varphi$ does not exist.
-That means that the acceptance condition of $\mathcal{A}$ cannot be simplified for given C and K.
-Thus we return automaton $\mathcal{A}$ and the information that $\mathcal{A}$ is the automaton with the simplest acceptance condition possible.
-    \item $\varphi$ is SAT -- QBF-solver provides (returns) a model of unquantified variables.
-Namely to $p_{c, k},n_{c,k} \text{ and } f_{i,k}, \text{ for each }1 \leq c \leq C, 1 \leq k \leq K, 1 \leq i \leq |\delta|$.
-    Firstly, we decode the variables $f_{i,k}$ and use them to identify new placement of the acceptance marks.
-Specifically, we create a new automaton $\mathcal{A}'$ with the same structure as $\mathcal{A}$.
-Then  we add acceptance mark $\tacc{5}{k}$ on edge $t_{i}$ for every $f_{i,k}$ variable evaluated as $\true$.
-
-    The new acceptance condition of automaton $\mathcal{A}'$ is then build from variables $p_{c,k}$ and $n_{c,k}$.
-The second part of $\varphi_{N}$ was encoded to produce new acceptance formula in DNF.
-(See the alternation between Boolean operators  in order $\lor, \land$ in Section \ref{subsection:new}.) Thus the shape of the new acceptance condition is set as DNF skeleton, and we simply add terms according to the evaluation of variables.
-If $p_{c,k}$ is $\true$, we add term $\Inf\tacc{5}{k}$ to the disjunct $c$.
-Similarly, if $n_{c,k}$ is $\true$, then we add $\Fin\tacc{5}{k}$ to the disjunct $c$.
-This way, we process every variable that QBF-solver gives us and produce the new automaton $\mathcal{A}'$
-    equivalent to the original one.
-
-\end{enumerate}
-
-\subsection{Quantified prefix}
-\label{subsection:quant_pref}
-The aim of the quantified prefix in formula $\varphi$ is to consider every possible selection of edges.
-As mentioned in the previous sections, edge $t_{i} \in \delta$ belongs to the \emph{selection of edges} if the variable $e_{i}$ that represents the edge $t_{i}$ is evaluated as $\true$.
-Therefore if we want to consider every possible \emph{selection of edges}, we need to consider every evaluation of the variables representing edges.
-We use a universal quantifier to express this.
-The universal quantifier states that the formula has to hold for both evaluations of the quantified variable.
-
-The QBF prefix that expresses the universal quantification for every variable representing an edge has the following shape:
-\begin{align*}
-    \overrightarrow{Q} = \forall\{e_{1}, \ldots ,e_{|\delta|} \}
-\end{align*}
-
-The advantage of QBF is that it is able to express a statement in shorter formula then if we tried to express the statement in standard Boolean formula with the use of conjunction.
-
-\begin{example}
-Consider the QBF $\theta = \forall e_{1} \forall e_{2} .
-(e_1 \lor e_2) $.
-The QBF $\theta $ transformed to standard Boolean formula is $\theta' = (e_{1} \lor e_{2}) \land (  e_{1} \lor \lnot e_{2}) \land  (  \lnot e_{1} \lor e_{2}) \land  (  \lnot e_{1} \lor \lnot e_{2}) .$
-\end{example}
-
-\subsection{Original acceptance formula}
-\label{subsection:original}
-
-As mentioned before, the satisfiability of $\varphi_{O}$ is purely resultant of assessment of $e_{i}, 1 \leq i \leq |\delta|$ variables representing edges in automaton.
-
-Let us enumerate the edges of the original automaton $\mathcal{A}$ and let $Z(m), m \in M$ be the set of edges labeled with mark $m$. 
-The encoding of original acceptance condition terms into $\varphi_{O}$ is defined as follows:
-%%davat tam rovnitko nebo nejak sipku??
-\begin{align*}
-    &\varphi(\Inf(m))=\bigvee\limits_{t_{i} \in Z(m)} e_{i}\\
-    &\varphi(\Fin(m)) = \bigwedge\limits_{t_{i} \in Z(m)} \lnot e_{i}
-\end{align*}
-\begin{example}
-
-Consider the automaton in Figure \ref{fig:oficial_acc}.
-The set $Z_{0} = \{t_3, t_2\},\, Z_{1} = \{t_1\}$, where $t_{1}, t_{2}, t_{3} \in \delta$.
-Thus the encoded $\varphi_{O} = e_1 \land (\lnot e_2 \land \lnot e_3)$.
-Therefore if variable $e_{1}=\true$, $e_{2}=\false$ and $e_{3}=\false$, acceptance formula is $\true$.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-\node[cstate, initial] (g) {};
-\node[flabel, above] at (0, 1.5) {$\Fin\taccsq{0}{0}\land\Inf\tacc{1}{1}$};
-%
-\path[->]
-    (g) edge[loop, in=330, out=30, looseness=6]node[right, pos=0.4, scale=0.8] {$1$} node[accset,collacc1, pos=0.6] {1}(g)
-	(g) edge[loop, in=80, out=140, looseness=6] node[above, pos= 0.4, scale=0.8] {$2$} node[accsquare, collacc0, pos=0.6]{0}(g)
-	(g) edge[loop, in=220, out=280, looseness=6] node[below, pos=0.6, scale=0.8] {$3$} node[accsquare, collacc0, pos=0.4] {0} (g)
-	;
-\end{tikzpicture}
-    \caption{TELA with edges labeled with their numbers}
-    \label{fig:oficial_acc}
-\end{figure}
-
-\end{example}
-
-\subsection{New acceptance formula}
-\label{subsection:new}
-The new acceptance formula $\varphi_{N}$ determines the shape of the future acceptance condition of automaton, and the assignment of variables used in $\varphi_{N}$ determines new positions of acceptance marks on the transitions.
-
-The formula $\varphi_{N}$ is divided into two parts.
-The first part of the formula (4.1) expresses a simple constrain that the acceptance set that occurs in one clause cannot be in $\Inf$ and simultaneously in $\Fin$ form.
-
-The second part (4.2) of $\varphi_{N}$ is divided into two subformulas.
-The first subformula expresses that if the term $\Inf(k)$ is present in clause $c$, then at least one edge of the \emph{selection of edges} needs to be labeled with the acceptance mark $k$.
-
-Similarly, the second subformula implies that if the term $\Fin(k)$ is present in clause $c$, then none of the \emph{selected edges} can be labeled with the acceptance mark $k$.
-
-Using the variables presented in previous sections, we encode $\varphi_{N}$:
-\begin{align}
-    \varphi_{N} = & \bigwedge\limits_{1 \leq c \leq C}\;\bigwedge\limits_{1 \leq k \leq K} (\lnot p_{c,k} \lor \lnot n_{c,k}) \; \land \\
-    &\bigvee\limits_{1 \leq c \leq C} \; \bigwedge\limits_{1 \leq k \leq K} \Bigg[\big(  p_{c, k} \implies \bigvee\limits_{t_{i} \in \delta}(f_{i,k} \land e_{i})  \big) \land \\& \;\;\;\quad\quad\quad\quad\quad\quad\big(  n_{c, k} \implies \bigwedge\limits_{t_{i} \in \delta}(\lnot f_{i,k} \lor \lnot e_{i})  \big)\Bigg]
-\end{align}
-
-Observe the alternation of $\lor \land$ from the second part of the formula (4.2). This alternation determines that the formula $\psi'$ is in DNF.
-
-\subsection{Connection formula}
-\label{subsection:connection}
-The connection formula distinguishes the remaining two levels of simplification. Previously we denoted it as $\varphi_{C}$.
-
-Formula $\varphi_{C}$ defines requirements on the edges that are together in \emph{selection of edges}.
-These requirements are based on some properties of these edges.
-The more restrictive these requirements are, the more properties of \emph{relevant sets} are possessed by the \emph{selection of edges}.
-In other words, formula $\varphi_{C}$ holds if the \emph{selection of edges} has the required property.
-If not, the formula $\varphi_{C}$ is $\false$, and then the whole formula $\varphi$ is trivially satisfied.
-Formula $\varphi_{C}$ consists of two parts $\varphi_{C_{1}}$ and $\varphi_{C_{2}}$.
-
-The first part expresses the requirement that the \emph{selection of edges} is \emph{continuous}.
-Loosely said, this means that if a transition from \emph{selection of edges} enters some state $q$ then another transition that belongs to the \emph{selection of edges} exits the state $q$.
-This holds in the opposite direction as well.
-Thus if a transition from \emph{selection of edges} exits some state $q$ then another transition that belongs to the \emph{selection of edges} enters the state $q$.
-This is naturally expressed by equivalence.
-Note that if the selected transition is a self-loop, the destination and the source of the edge is the same state.
-
-The Boolean formula for the described statement:
-\begin{center}
-    $\varphi_{C_{1}}=\bigwedge\limits_{q \in Q} \Bigg( ( \bigvee\limits_{t_{i} \in \delta \cap \{ q \} \times  \Sigma \times  2^M \times Q} e_i ) \Longleftrightarrow( \bigvee\limits_{t_{j} \in \delta  \cap \ Q \times  \Sigma \times 2^M \times \{q\}} e_{j} ) \Bigg)$
-\end{center}
-
-\begin{example}
-   Consider the state $q$ in Figure \ref{fig:inNout} this part of formula is $(e_1 \lor e_2) \Longleftrightarrow (e_3 \lor e_4 \lor e_5)$.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-        \node[state] (q) at (0, 0) {$q$};
-
-	    \path[->]
-		(q) edge node[above, pos=0.4, scale=0.8] {$4$} ++(1.5, 0)
-		(q) edge node[above, pos=0.4, scale=0.8] {$5$} ++(1.3, 1)
-		(q) edge node[above, pos=0.4, scale=0.8] {$3$} ++(1.3, -1)
-	;
-	\path[<-]
-	(q) edge node[above, pos=0.6, scale=0.8] {$2$} ++(-1.3, -0.5)
-	(q) edge node[above, pos=0.6, scale=0.8] {$1$} ++(-1.3, 0.5)
-	;
-    \end{tikzpicture}
-    \caption{State $q$ and its entering and exiting edges}
-    \label{fig:inNout}
-\end{figure}
-\end{example}
-
-Even though we ensured that the selected transitions create continuous paths, the paths might not be connected.
-
-\begin{example}
-Consider the automaton in Figure \ref{fig:continuousNotConnected}.
-The transitions $t_1$ and $t_2$ are loops (the simplest cycles), but are not connected (each is located in different SCC).
-Thus $\{ t_{1}, t_{2}\}$ is not a \emph{relevant set}.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-			\node[cstate, initial] (i) at (0, 0) {};
-			\node[cstate] (j) at (2.5, 1) {};
-			\node[cstate] (k) at (2.5, -1) {};
-			\path[->]
-
-			(i) edge  (k)
-			(i) edge (j)
-			(j) edge[loop right, draw=red] node[right] {$1$} (j)
-			(k) edge[loop right, draw=red] node[right] {$2$} (k)
-
-		;
-    \end{tikzpicture}
-    \caption{Not connected cycles}
-    \label{fig:continuousNotConnected}
-\end{figure}
-\end{example}
-
-The second part of $\varphi_{C}$ specifies the properties of \emph{relevant set} that the edges that belong together to \emph{selected edges} need to have.
-We divided these properties into two levels Level 2 and Level 3, all described in ensuing sections.
-
-\subsubsection{SCC accuracy}
-The SCC accuracy corresponds with the Level of simplification 2. At this level, the formula specifies that the edges from one \emph{selection of edges} belong to the same SCC.
- Thus straightforwardly translated into QBF formula:
-\begin{center}
-    $\varphi_{C_{2}} = \bigvee\limits_{\delta_{S_{i}} \subseteq \delta}\Bigg( ( \bigvee\limits_{t_{i} \in \delta_{S_{i}}} e_i ) \land  ( \bigwedge\limits_{t_{j} \in \delta \setminus \delta_{S_{i}}} \lnot e_j  ) \Bigg)$
-\end{center}
-
-Finally, the connection formula is the conjunction of the two parts:
-\begin{center}
-    $\varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}$.
-\end{center}
-
-\begin{example}
-This example works as a motivation for simplification Level 3.
-For simplification Level 2, edges number $1$ and number $3$ in Figure \ref{fig:connected_motivation} can be in the same \emph{selection of edges} because they are placed in the same SCC.
-Obviously, since they are not connected, they cannot form \emph{relevant set}.
-Therefore we can specify a new level of simplification -- Level 3, which forces all edges from the \emph{selection of edges} to be cycles.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-    \node[state, initial] (g) {$0$};
-    \node[state, right=2cm of g](q){$1$};
-    %
-    \path[->, auto]
-    (g) edge[bend right] node[below, scale=0.8] {$2$}(q)
-    (q) edge[bend right]node[above, scale=0.8] {$4$} (g)
-    (q) edge[loop, in=65, out=115, looseness=10, draw=red] node[ scale=0.8] {$3$}(q)
-    (g) edge[loop, in=65, out=115, looseness=10, draw=red] node[scale=0.8] {$1$}(g)
-	;
-
-\end{tikzpicture}
-
-    \caption{Motivation for simplification L3}
-    \label{fig:connected_motivation}
-\end{figure}
-
-\end{example}
-
-\subsubsection{CYCLES accuracy}
-The simplification Level 2 ensures that \emph{selection of edges} is placed in one SCC, but we can go further and specify that \emph{selection of edges} is in fact a \emph{relevant set}.
-
-The intuition by simplification Level 3 is that the states of an automaton are  divided into two groups -- the states that belong to the group of \emph{selected states} and the states do not belong to the group of \emph{selected states}.
-Then we require that the states that belong to the same group are connected.
-We consider three possible scenarios that can occur while the states are divided into these two groups:
-\begin{enumerate}
-    \item The source of edge $t_{i}$ is in different group than the destination of edge $t_{i}$.
-    \item Source and destination of edge $t_{i}$ are in the \emph{selected states}.
-    \item Source and destination of edge $t_{i}$ are not in \emph{selected states}.
-\end{enumerate}
-Cases two and three do not violate the requirement that all connected states need to be in one group.
-Case one is a little more complicated.
-If an edge has a source in \emph{selected states} and destination not in the \emph{selected states}, we need to secure that the sequence of the edges returns to \emph{selected states}.
-In other words, there exists another edge that has a source in the not \emph{selected states} and destination in the \emph{selected states}.
-(Dually the other way around.)
-
-In order to express this in the QBF-formula we introduce a new set of quantified variables $\{ w_{1}, \ldots , w_{|Q|}$. The variables have the following semantics:
-\begin{align*}
-    w_{q} = \begin{cases}
-    1& \text{state } q \in Q \text{ is in the \emph{selected group}}\\
-    0& \text{else}
-    \end{cases}
-\end{align*}
-Since we want to ensure that for every \emph{selection of edges} (for every evaluation of variables that represent the edges) there exist such a groups we update the quantified prefix  $\overrightarrow{Q}$ in following way:
-\begin{align*}
-    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
-\end{align*}
-
-\smallskip
-\todo{neoptimalni}
-Since we want the statement to hold for every \emph{selection of edges}, we place the existentially quantified variables $\{w_{1}, \ldots, w_{|Q|} \}$ behind the universally quantified variables $\{e_{1}, \ldots, e_{|\delta|} \} $.
-At first we update the quantified prefix:
-\begin{align*}
-    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
-\end{align*}
-
-And then we can create the $\varphi_{C}$ formula:
-\begin{align*}
-    \varphi_{C} =  \Bigg[ &\bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (w_{q} \land w_{r}) \Bigg) \lor \\
-    &  \bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma  \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (\lnot w_{q} \land \lnot w_{r}) \Bigg) \lor \\ \Bigg( &\bigvee \limits_{\substack{t_{i} \in \{q\} \times \Sigma \times 2^M  \times \{ r \}, \\ q \in Q, r \in Q}} ( e_{i} \land w_{q} \land \lnot w_{r} )  \land \bigvee \limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} (e_{i} \land \lnot  w_{q} \land  w_{r}) \Bigg) \Bigg]
-\end{align*}
-
-\begin{example}
-Figure \ref{fig:cycles_sensitivity} displays an example of an automaton reduced by Level 4.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-
-    \node[state, initial] (g) {$0$};
-
-    \node[flabel] at (1.5, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{1}{1} \land \Inf\tacc{2}{2}$};
-    \node[state, right=2cm of g](q){$1$};
-
-    \path[->, auto]
-    (g) edge[bend right] node [accset,collacc0, pos=0.5]{0}node [below, pos=0.2, scale=0.8] {$2$}(q)
-    (q) edge[bend right] node[above, pos=0.8, scale=0.8] {$4$}(g)
-    (q) edge[loop, in=75, out=105, looseness=8]node[left, pos=0.1, scale=0.8]{5} (q)
-    (q) edge[loop, in=70, out=110, looseness=12] node[accset,collacc1] {1}node[left, pos=0.2, scale=0.8]{6} (q)
-    (g) edge[loop, in=75, out=105, looseness=8] node[accset,collacc2] {2} node[left, pos=0.1, scale=0.8]{1}(g)
-    (g) edge[loop, in=70, out=110, looseness=12] node[accset,collacc2, pos=0.4] {2} node[accset,collacc1, pos=0.7] {1} node[left, pos=0.2, scale=0.8]{3}(g)
-	;
-
-	\node[state, initial, right=4cm of g] (o){$0$};
-
-    \node[flabel] at (6.7, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{3}{3}$};
-    \node[state, right=2cm of o](p){$1$};
-
-    \path[->, auto]
-    (o) edge[bend right] node [accset,collacc0, pos=0.5]{0}(p)
-    (p) edge[bend right](o)
-    (p) edge[loop, in=75, out=105, looseness=8] (p)
-    (p) edge[loop, in=70, out=110, looseness=12] (p)
-    (o) edge[loop, in=75, out=105, looseness=8] (o)
-    (o) edge[loop, in=70, out=110, looseness=12] node[accset,collacc3] {3}(o)
-
-	;
-
-\end{tikzpicture}
-
-    \caption{TELA reduced with the CYCLES sensitivity}
-    \label{fig:cycles_sensitivity}
-\end{figure}
-\end{example}
-
-Similarly as in the previous section the connection formula is the conjunction of the two parts:
-\begin{align*}
-    \varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}.
-\end{align*}
-
-\subsection{Simplification procedure}
-\label{subsection:procedure}
-
-Now we describe our simplification algorithm taking as input an arbitrary TELA $\mathcal{A}$ with an acceptance condition $\psi$, required level of simplification $L$ and timeout period $T$.
-\begin{enumerate}
-    \item Convert acceptance condition $ \psi$ to DNF.
-    \item Apply the algorithm for simplification Level 1 described in Section \ref{sec:standard} on $\mathcal{A}$.
-    \item If $L = 1$ return automaton $\mathcal{A}$.
-     \item The initial value of $C$ is the number of clauses in $\psi$. Similarly, the initial value of $K$ is the number of acceptance sets of $\psi$ decreased by 1.
-    \item Encode $\mathcal{A}$ into QBF formula $\varphi$ according to the required level of simplification $L$.
-    \item Run a QBF-solver on $\varphi$ for $T$ seconds.
-    \begin{enumerate}
-        \item If the QBF-solver evaluates $\varphi$ as UNSAT, return automaton $\mathcal{A}'$.
-        \item If the timeout $T$ expires before the QBF-solver evaluates $\varphi$, return automaton $\mathcal{A}'$.
-        \item If the QBF-solver evaluates $\varphi$ as SAT. Decode the assignment into next TELA $\mathcal{A}'$.
-        \begin{enumerate}
-            \item If $K=1$, try to reduce $\varphi$ to $\true$ or $\false$, the simplest form of acceptance condition and return $\mathcal{A}$.
-            \item Decrease $K$ by 1 and go to (b).
-
-    \end{enumerate}
-
-    \end{enumerate}
-\end{enumerate}
-
-\todo{kde je optimalni misto, kam napsat:}
-For simplicity, we do not display letters on the edges of the automata in Figures. We can afford to omit them because they are not important for our simplification techniques.
-
-\begin{itemize}
-\item obecny popis konstrukce formule
-\item (level 3)
-\item level 4
-\end{itemize}
-
-%}}}
-%{{{ implementation and experimental evaluation
-
-\section{Implementation and experimental evaluation}\label{sec:experiments}
-
-%}}}
-%{{{ conclusion
-
-\section{Conclusions}\label{sec:conclusion}
-
-chceme sem napsat, ze QBF-based reduction se da pouzit napr. k
-nalezeni akceptacni formule s $c$ klauzulemi a $m$ znackami v CNF,
-pripadne k nalezeni ekvivalentniho paritniho automat s danou maximalni
-paritou $p$ a stejnou strukturou.
-
-%}}}
-
-%\bibliographystyle{plain}
-\bibliography{mc}
-
-\end{document}
 

### 2022-02-23 paper/main.tex
Preliminaries polished

@@
 \newcommand{\todo}[1]{\textcolor{red}{#1}}
-\def\Inf{\ensuremath{\mathsf{Inf}}}
-\def\Fin{\ensuremath{\mathsf{Fin}}}
+%\def\Inf{\ensuremath{\mathsf{Inf}}}
+%\def\Fin{\ensuremath{\mathsf{Fin}}}
+\DeclareMathOperator{\Inf}{\mathsf{Inf}}
+\DeclareMathOperator{\Fin}{\mathsf{Fin}}
 \def\false{\mathit{false}}
@@
 \emph{quantified Boolean formulas (QBF)}. For a given TELA and
-parameters $c,m$, it produces a QBF which is satisfiable if and only
-if there exists an automaton with the same structure, $m$ acceptance
-marks, an acceptance formula in disjunctive normal form with $c$
-clauses, and the same set of accepting runs as the original
+parameters $c,m$, it produces a QBF formula which is satisfiable if
+and only if there exists an automaton with the same structure, $m$
+acceptance marks, an acceptance formula in disjunctive normal form
+with $c$ clauses, and the same set of accepting runs as the original
 automaton. The placement of the marks on transitions and the
@@
 experimental results. Finally, Section~\ref{sec:conclusion} suggests
-other applications of our QBF-reduction technique and closes the
+other applications of our QBF-based reduction technique and closes the
 paper.
@@
     ranges over $M$.
-    $$\varphi ::= \true \mid \false \mid \Inf(m) \mid \Fin(m) \mid \left( \varphi \land \varphi \right) \mid \left( \varphi \lor \varphi \right)$$
+    $$\varphi ::= \true \mid \false \mid \Inf m \mid \Fin m \mid \left( \varphi \land \varphi \right) \mid \left( \varphi \lor \varphi \right)$$
 \end{itemize}
@@
 \emph{accepting} iff it satisfies the formula $\varphi$, where $\pi$
-satisfies $\Inf(m)$ iff $m\in M(\pi)$ and $\pi$ satisfies $\Fin(m)$
+satisfies $\Inf m$ iff $m\in M(\pi)$ and $\pi$ satisfies $\Fin m$
 iff $m\not\in M(\pi)$. The \emph{language} of automaton $\mA$ is the
@@
 
-\todo{Jan is here}
-
 An acceptance formula $\varphi$ is in \emph{disjunctive normal formal
   (DNF)} if it is a disjunction of clauses, where each clause is a
-conjunction of some terms of the form $\Fin(m)$ or $\Inf(m)$.  Formula
+conjunction of some terms of the form $\Fin m$ or $\Inf m$. Formula
 $\false$ corresponds to the disjunction of zero clauses and formula
-$\true$ is the clause with zero terms. Each acceptance formula can be
-transformed into an equivalent formula in DNF.
+$\true$ corresponds to one clause with zero terms. Each acceptance
+formula can be transformed into an equivalent formula in DNF.
 
@@
 marked with $m$, i.e., $\delta_m=\{(p,a,M',q)\in\delta\mid m\in
-M'\}$. The sets $\delta_m$ are called \emph{acceptance sets}.
-Finally, given an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we
-denote the set $\delta_{S,m}=\delta_S\cap\delta_m$.
-
-We often use acceptance marks $\tacc{1}{1},\tacc{2}{2},\ldots\in M$
-and we write $\Inf\tacc{4}{j}$ and $\Fin\tacc{8}{k}$ instead of
-$\Inf(\tacc{4}{j})$ and $\Fin(\tacc{8}{k})$, respectively. In our
-graphical representation of automata,
+M'\}$. The sets $\delta_m$ are called \emph{acceptance sets}.  Given
+an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we denote the set
+$\delta_{S,m}=\delta_S\cap\delta_m$. Finally, a set of transitions
+$D\subseteq\delta$ is called a \emph{cycle} is there exists a path
+from a state $p$ to the same state containing all transitions of $D$
+and no transition outside $D$.
+
+In graphical representation of automata, we often use acceptance marks
+$\tacc{1}{1},\tacc{2}{2},\ldots\in M$ and an edge 
 \begin{tikzpicture}[smallautomaton,baseline=-2pt]
@@
 
-% Finaly, \emph{loop} is a path from a state $p$ to the same state.
-
 \bigskip
@@
 \emph{Quantified Boolean formulas (QBFs)} are Boolean formulas
-extended with universal or existential quantification over
+extended with universal and existential quantification over
 propositional variables. We assume that subformulas of the form
@@
 
-A QBF $\varphi$ is in the \emph{prenex normal form} if it has the form
-$\varphi = Q_{1}x_{1}...Q_{n}x_{n} . \psi$, where
-$Q_1,\ldots,Q_n \in \{\forall, \exists\}$ and $\psi$ is a Boolean
-formula.
+% A QBF $\varphi$ is in the \emph{prenex normal form} if it has the form
+% $\varphi = Q_{1}x_{1}...Q_{n}x_{n} . \psi$, where
+% $Q_1,\ldots,Q_n \in \{\forall, \exists\}$ and $\psi$ is a Boolean
+% formula.
 
 %}}}
-%{{{ QBF-based reduction
-
-\section{QBF-based reduction}\label{sec:qbf}
+%{{{ construction of QBF formulas
+
+\section{Construction of QBF formulas}\label{sec:qbf}
 
@@
 pripadne k nalezeni ekvivalentniho paritniho automat s danou maximalni
-paritou $p$ a stejnou strukturou.
+paritou $p$ a stejnou strukturou. Nebo zkusit nalezt $\Fin$-less
+acceptance formuli.
 

### 2022-02-23 paper/main.tex
small progress in Section 3

@@
 \emph{quantified Boolean formulas (QBF)}. For a given TELA and
-parameters $c,m$, it produces a QBF formula which is satisfiable if
-and only if there exists an automaton with the same structure, $m$
+parameters $C,K$, it produces a QBF formula which is satisfiable if
+and only if there exists an automaton with the same structure, $K$
 acceptance marks, an acceptance formula in disjunctive normal form
-with $c$ clauses, and the same set of accepting runs as the original
+with $C$ clauses, and the same set of accepting runs as the original
 automaton. The placement of the marks on transitions and the
@@
 two simpler formulas such that their satisfiability implies the
-existence of an automaton with the same structure, $m$ acceptance
+existence of an automaton with the same structure, $K$ acceptance
 marks, and the same set of accepting runs, but not vice versa.
@@
 
-\emph{Quantified Boolean formulas (QBFs)} are Boolean formulas
+\emph{Quantified Boolean formulas (QBF)} are Boolean formulas
 extended with universal and existential quantification over
@@
 
+Recall that we aim to reduce the number of acceptance marks in a given
+TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ without altering its
+structure and language. In other words, we change only the acceptance
+marks on individual transitions and we replace the acceptance formula
+by a new one with less acceptance marks. To ensure that the modified
+automaton is equivalent to the original one, we actually construct the
+new acceptance formula such that the set of acceptance runs (modulo
+changed acceptance marks on transitions) is preserved.
+
+We start with the following simple observation. Definovat $inf(\varphi)$?
+
+\textbf{Jan is here!}
+
+Recall that our goal is to reduce the number of acceptance marks in a
+given TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ without altering its
+structure and language. Hence, our reduction replaces the set $M$ of
+marks by a smaller set, the acceptance formula $\varphi$ by a new one,
+and changes the sets of acceptance marks on individual transitions.
+
+Let us assume that we are given a TELA $\mA$ and two integer
+parameters $C,K\ge 0$, where $K$ is the desired number of acceptance
+marks after the reduction and $C$ is the desired number of clauses in
+the new acceptance formula in DNF. This section describes the
+construction of a QBF formula for $\mA,C,K$ such that satisfiability
+of this formula implies existence of the reduced automata with
+
+This section explains the construction of QBF formulas for a given
+TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ and two integer parameters
+$C,K\ge 0$ such that satisfiability of such a formula implies that
+there exists an acceptance formula in DNF with $C$ clauses and $K$
+acceptance marks and a placement of these marks on automata
+transitions such that $\mA$ with this acceptance formula and
+acceptance marks and acceptance formula
+
+such that their model describes an acceptance formula in DNF
+with $C$ clauses and $K$ acceptance marks and placement of these marks
+on automata transitions such that
+
 The aim of this section is to create a QBF-formula for given $n$ that guarantees the existence of equivalent acceptance condition of a given automaton with $n$ acceptance marks. 
@@
 
+Pridat do experimentu nejaky graf, ktery ukaze, jak rychle se nektere
+automaty redukovaly na minimum? Nebo spis ukazat vliv timeoutu na
+redukce?  (zvysovat timeout od 1s do 50s a vzdy ukazat miru
+redukce). Ty redukce ukazovat nejak jinak nez jen celkovym cislem
+(boxplot? Scatterplot?).
+
+Zkusit upravit redukci tak, aby vysledkem bylo TGBA a zkusit dale
+redukovat akceptacni znacky TGBA automatu redukovanych Spotem?
+
+Chceme ukazovat, ze redukce znacek muze nekdy umoznit dalsi redukci stavu?
+
 %}}}
@@
 acceptance formuli.
+
+Redukce netrva vecne, da se prakicky pouzit a zkracenim timeoutu na
+QBF-solver se da zrychlit (tradeoff s ucinnosti). Prinejmensim u
+Rabinizeru 4 je velmi dobra sance, ze se to vyplati. Pridat do
+experimentu nejaky graf, ktery ukaze, jak rychle se nektere automaty
+redukovaly na minimum?
 

### 2022-02-24 paper/main.tex
small progress in section 3

@@
 \DeclareMathOperator{\Fin}{\mathsf{Fin}}
+\newcommand{\minf}{\mathit{inf}}
 \def\false{\mathit{false}}
@@
 \emph{quantified Boolean formulas (QBF)}. For a given TELA and
-parameters $C,K$, it produces a QBF formula which is satisfiable if
-and only if there exists an automaton with the same structure, $K$
+parameters $k,n$, it produces a QBF formula which is satisfiable if
+and only if there exists an automaton with the same structure, $n$
 acceptance marks, an acceptance formula in disjunctive normal form
-with $C$ clauses, and the same set of accepting runs as the original
+with $k$ clauses, and the same set of accepting runs as the original
 automaton. The placement of the marks on transitions and the
@@
 two simpler formulas such that their satisfiability implies the
-existence of an automaton with the same structure, $K$ acceptance
+existence of an automaton with the same structure, $n$ acceptance
 marks, and the same set of accepting runs, but not vice versa.
@@
 
-%A transition $(p,a,M',q)\in\delta$ is a transition leading from state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$
+A transition $t=(p,a,M',q)\in\delta$ is the transition leading from
+state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$. We
+use $M(t)$ to refer to its set of acceptance marks $M'$. We extend
+this notation to a set $T\subseteq\delta$ of transitions such that
+$M(T)=\bigcup_{t\in T}M(t)$ denotes the set of acceptance marks that
+appear on transitions in $T$.
+
+\todo{Rozmyslet si, jak dal.}
+
 A run $\pi$ of $\mA$ over an infinite word
@@
 $\pi=(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots\in\delta^{\omega}$ where
-$q_0=q_I$. Let $M(\pi)$ denote set of all acceptance marks that occur
-in infinitely many transitions of run $\pi$. Run $\pi$ is
-\emph{accepting} iff it satisfies the formula $\varphi$, where $\pi$
-satisfies $\Inf m$ iff $m\in M(\pi)$ and $\pi$ satisfies $\Fin m$
-iff $m\not\in M(\pi)$. The \emph{language} of automaton $\mA$ is the
-set
+$q_0=q_I$. Let $\minf(\pi)$ denote the set of transitions that appear
+infinitely many times in $\pi$. Further, let $M(\pi)$ denote the set
+of acceptance marks that occur in infinitely many transitions of run
+$\pi$, i.e., $M(\pi)=M(\minf(\pi))$. Run $\pi$ is \emph{accepting} iff
+$M(\pi)$ satisfies the formula $\varphi$, where $M(\pi)$ satisfies
+$\Inf m$ iff $m\in M(\pi)$ and $M(\pi)$ satisfies $\Fin m$ iff
+$m\not\in M(\pi)$. The \emph{language} of automaton $\mA$ is the set
 $L(\mA)=\{u\in \Sigma^{\omega}\mid\textrm{there is an accepting run of
@@
 $(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
-such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
-is called a nontrivial \emph{strongly connected component (SCC)} if
-for each $p,q\in S$ there is a path from $p$ to $q$. An SCC S is
-\emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
-the rest of this paper, SCC always refers to a maximal SCC. Given an
-SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
-denotes the set of all transitions between states in $S$. Further, for
-each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
-marked with $m$, i.e., $\delta_m=\{(p,a,M',q)\in\delta\mid m\in
-M'\}$. The sets $\delta_m$ are called \emph{acceptance sets}.  Given
-an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we denote the set
+such that $p=q_0$ and $q=q_n$. We assume that each state $q$ is
+reachable from the initial state $q_I$ (i.e., $q=q_I$ or there is a
+path from $q_I$ to $q$) as states that are not reachable from $q_I$
+can be eliminated without any impact on $L(\mA)$. A nonempty set of
+states $S\subseteq Q$ is called a nontrivial \emph{strongly connected
+  component (SCC)} if for each $p,q\in S$ there is a path from $p$ to
+$q$. An SCC S is \emph{maximal} if there is no SCC $S'$ satisfying
+$S\subsetneq S'$. In the rest of this paper, SCC always refers to a
+maximal SCC. Given an SCC $S$,
+$\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$ denotes the
+set of all transitions between states in $S$. Further, for each mark
+$m\in M$, let $\delta_m$ denotes the set of all transitions marked
+with $m$, i.e., $\delta_m=\{(p,a,M',q)\in\delta\mid m\in M'\}$. The
+sets $\delta_m$ are called \emph{acceptance sets}.  Given an SCC $S$
+and a mark $m\in M$, by $\delta_{S,m}$ we denote the set
 $\delta_{S,m}=\delta_S\cap\delta_m$. Finally, a set of transitions
-$D\subseteq\delta$ is called a \emph{cycle} is there exists a path
-from a state $p$ to the same state containing all transitions of $D$
-and no transition outside $D$.
+$C\subseteq\delta$ is called a \emph{cycle} is there exists a path
+from a state $p$ to the same state containing all transitions of $C$
+and no transition outside $C$.
 
@@
 automaton is equivalent to the original one, we actually construct the
-new acceptance formula such that the set of acceptance runs (modulo
-changed acceptance marks on transitions) is preserved.
-
-We start with the following simple observation. Definovat $inf(\varphi)$?
+new acceptance formula such that the set of acceptance runs stays the
+same (modulo changed acceptance marks on transitions).
+
+We start with the several simple observations. Acceptance of a run
+$\pi$ is fully determined by the set $M(\pi)$ of acceptance marks that
+appear infinitely often on $\pi$, which is in turn fully determined by
+the set $\inf(\pi)$ of transitions appearing infinitely often in
+$\pi$.
+% \[
+%   M(\pi)=\bigcup_{(q_i,u_i,M_i,q_{i+1})\in\minf(\pi)}M_i
+% \]
+Further, each set of the form $\inf(\pi)$ is a cycle and vice versa.
+Hence, our reduction methods is looking for
+
+following simple observation. Definovat $inf(\varphi)$?
 

### 2022-02-24 paper/main.tex
improved intro.

@@
 Rabinizer~4~\cite{kretinsky.18.cav}, or
-SPOT~\cite{duret.16.atva2}. There are also some algorithms processing
+Spot~\cite{duret.16.atva2}. There are also some algorithms processing
 these automata, for example the emptiness check~\cite{baier.19.atva}
@@
 
-These observations motivate our study of a technique reducing the
-number of acceptance marks without altering the structure of the
-automaton. This paper presents a technique heavily based on
-\emph{quantified Boolean formulas (QBF)}. For a given TELA and
-parameters $k,n$, it produces a QBF formula which is satisfiable if
-and only if there exists an automaton with the same structure, $n$
-acceptance marks, an acceptance formula in disjunctive normal form
-with $k$ clauses, and the same set of accepting runs as the original
-automaton. The placement of the marks on transitions and the
-acceptance formula can be obtained from a model of the
-formula. Besides this formula, we describe also the construction of
-two simpler formulas such that their satisfiability implies the
-existence of an automaton with the same structure, $n$ acceptance
-marks, and the same set of accepting runs, but not vice versa.
+The number of acceptance marks can be algorithmically reduced to one
+as every TELA can be transformed to an equivalent Büchi automaton
+(this can be easily done for example by Spot~\cite{duret.16.atva2}),
+but this reduction is paid by dramatic changes of state-space: the
+number of states can increase exponentially and some important
+structural properties like determinism can be lost. This motivates our
+study of a technique reducing the number of acceptance marks without
+altering the structure of the automaton.
+
+We present a technique heavily based on \emph{quantified Boolean
+  formulas (QBF)}. For a given TELA and parameters $k,n$, it produces
+a QBF formula which is satisfiable if and only if there exists an
+automaton with the same structure, $n$ acceptance marks, an acceptance
+formula in disjunctive normal form with $k$ clauses, and the same set
+of accepting runs as the original automaton. The placement of the
+marks on transitions and the acceptance formula can be obtained from a
+model of the formula. Besides this formula, we describe also the
+construction of two simpler formulas such that their satisfiability
+implies the existence of an automaton with the same structure, $n$
+acceptance marks, and the same set of accepting runs, but not vice
+versa.
 
@@
 \paragraph*{Related results}
-We are not aware about any existing research primarily aimed at
-simplification of acceptance formulas of TELA or reduction of the
-number of its acceptance marks without increasing the number of
-states. There exists only a SAT-based approach that transforms a
-deterministic TELA automaton to an equivalent automaton with a given
-acceptance condition and a given number of
-states~\cite{baarir.15.lpar} (if such an automaton exists). Further,
-there are some SAT-based approaches aimed to reduce the number of
-states of automata over infinite words. More preceisely, there are
-reductions designed for nondeterministic Büchi
+There is a simple technique~\cite{babiak.13.spin} reducing the number
+of acceptance marks in \emph{transition-based genealised Büchi
+  automata (TGBA)} without changing its structure. We are not aware
+about any existing research aimed at simplification of acceptance
+formulas of TELA or reduction of the number of its acceptance marks
+without increasing the number of states. There exists only a SAT-based
+approach that transforms a deterministic TELA automaton to an
+equivalent automaton with a given acceptance condition and a given
+number of states~\cite{baarir.15.lpar} (if such an automaton
+exists). Further, there are some SAT-based approaches aimed to reduce
+the number of states of automata over infinite words. More preceisely,
+there are reductions designed for nondeterministic Büchi
 automata~\cite{ehlers.10.spin}, deterministic Büchi
@@
 
-A transition $t=(p,a,M',q)\in\delta$ is the transition leading from
+A tuple $t=(p,a,M',q)\in\delta$ is the \emph{transition} leading from
 state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$. We
-use $M(t)$ to refer to its set of acceptance marks $M'$. We extend
-this notation to a set $T\subseteq\delta$ of transitions such that
-$M(T)=\bigcup_{t\in T}M(t)$ denotes the set of acceptance marks that
-appear on transitions in $T$.
-
-\todo{Rozmyslet si, jak dal.}
+use $M(t)$ to refer to the set $M'$ of acceptance marks. Further, for
+a set $T\subseteq\delta$ of transitions, $M(T)=\bigcup_{t\in T}M(t)$
+denotes the set of acceptance marks that appear on transitions in $T$.
 
@@
 In graphical representation of automata, we often use acceptance marks
-$\tacc{1}{1},\tacc{2}{2},\ldots\in M$ and an edge 
+$\tacc{1}{1},\tacc{2}{2},\ldots\in M$. An edge 
 \begin{tikzpicture}[smallautomaton,baseline=-2pt]

### 2022-02-25 qbf/formula.py
qbf level1 fix

@@ -56,7 +56,7 @@ class Limboole_f_ctor(FormulaCreator):
 
         # edges that are true create continuous cycle of edges aka laso
 
-        if self.mode == 4:
+        if self.mode == 3:
             laso = laso2(aut, self.inner_edges, self.scc_edg)
             in_out = in_n_out(self.scc_state_info)
             laso.add_subf(in_out)
@@ -355,10 +355,6 @@ def laso_f(aut, inner_edges_nums, scc_state_info, scc_edg, inner_edges, mode):
     laso.add_subf(one_scc)
 
     # this part makes sure, that the edges are connected tzn QBF CONNECTED
-    if mode == 3:
-        neg = negate_part(aut, inner_edges, edges_translator)
-        laso.add_subf(neg)
-    return laso
 
 
 def equiv_f(old, new):

### 2022-02-25 qbf/q_main.py

@@ -5,8 +5,13 @@ from qbf.formula import *
 from qbf.optimization import *
 from qbf.z3_formula import *
 
+def add_or_append(dict, key, value):
+    if key in dict:
+        dict[key].append(value)
+    else:
+        dict[key] = [value]
 
-def edge_dictionary(aut):
+def edge_dictionary(aut, mode):
     """
 
     Args:
@@ -15,14 +20,36 @@ def edge_dictionary(aut):
     Returns: Dictionary {acceptance set number : [numbers of edges in this set]}
 
     """
-    list_of_all_edges = aut.edges()
+    #list_of_all_edges = aut.edges()
     edge_dict = {}
-    set_nums = aut.get_acceptance().used_sets()
+    scc_equiv_edges = []
+    representants = []
+    #set_nums = aut.get_acceptance().used_sets()
 
-    for acc_set in set_nums.sets():
-        edge_dict[acc_set] = list(map(aut.edge_number, list(
-            filter(lambda edg: edg.acc.has(acc_set), list_of_all_edges))))
-    return edge_dict
+    si = spot.scc_info(aut)
+    for scc in range(si.scc_count()):
+        marks_edges = {}
+        edges = si.inner_edges_of(scc)
+        for e in edges:
+            # edge_dict
+            if mode != 1:
+                for m in e.acc.sets():
+                    add_or_append(edge_dict, m, aut.edge_number(e))
+            if tuple( e.acc.sets()) in marks_edges:
+                marks_edges[tuple( e.acc.sets())].append(e)
+                # one num representant is enough to create formula
+            else:
+                marks_edges[tuple( e.acc.sets())] = [e]
+                representants.append(aut.edge_number(e))
+                if mode == 1:
+                    for m in e.acc.sets():
+                        add_or_append(edge_dict, m, aut.edge_number(e))
+
+        scc_equiv_edges.append(marks_edges)
+    #print(scc_equiv_edges)
+    return edge_dict, scc_equiv_edges, representants
+
+
 
 
 def get_edges(aut):
@@ -93,7 +120,7 @@ def create_formula(
         C,
         K,
         inner_edges,
-        mode, qbf_solver):
+        mode, qbf_solver, representants):
     """
 
     :param aut:
@@ -115,10 +142,14 @@ def create_formula(
         return None
     else:
 
+
         f_creator = Z3_f_ctor(edge_dict, scc_edg, scc_state_info
                       ,inner_edges_nums,C, K, inner_edges
                       , mode, qbf_solver)
[…]

### 2022-02-25 paper/main.tex
preliminaries polished

@@
 \newcommand{\minf}{\mathit{inf}}
+\newcommand{\mks}{\mathit{mks}}
 \def\false{\mathit{false}}
@@
 but this reduction is paid by dramatic changes of state-space: the
-number of states can increase exponentially and some important
-structural properties like determinism can be lost. This motivates our
-study of a technique reducing the number of acceptance marks without
-altering the structure of the automaton.
-
-We present a technique heavily based on \emph{quantified Boolean
+number of states can increase exponentially in the number of
+acceptance marks and some important structural properties like
+determinism can be lost. This motivates our study of a technique
+reducing the number of acceptance marks without altering the structure
+of the automaton.
+
+We present such a technique heavily based on \emph{quantified Boolean
   formulas (QBF)}. For a given TELA and parameters $k,n$, it produces
@@
 A tuple $t=(p,a,M',q)\in\delta$ is the \emph{transition} leading from
-state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$. We
-use $M(t)$ to refer to the set $M'$ of acceptance marks. Further, for
-a set $T\subseteq\delta$ of transitions, $M(T)=\bigcup_{t\in T}M(t)$
-denotes the set of acceptance marks that appear on transitions in $T$.
-
-A run $\pi$ of $\mA$ over an infinite word
+state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$.  A
+run $\pi$ of $\mA$ over an infinite word
 $u = u_{0}u_{1}u_{2}\dots \in \Sigma^{\omega}$ is an infinite sequence
@@
 $q_0=q_I$. Let $\minf(\pi)$ denote the set of transitions that appear
-infinitely many times in $\pi$. Further, let $M(\pi)$ denote the set
-of acceptance marks that occur in infinitely many transitions of run
-$\pi$, i.e., $M(\pi)=M(\minf(\pi))$. Run $\pi$ is \emph{accepting} iff
-$M(\pi)$ satisfies the formula $\varphi$, where $M(\pi)$ satisfies
-$\Inf m$ iff $m\in M(\pi)$ and $M(\pi)$ satisfies $\Fin m$ iff
-$m\not\in M(\pi)$. The \emph{language} of automaton $\mA$ is the set
+infinitely many times in $\pi$.
+%
+We use $\mks(\cdot)$ to denote the set of acceptance
+marks of various objects:
+\begin{itemize}
+\item for a transitions $t=(p,a,M',q)\in\delta$, $\mks(t)$ denotes the set $M'$,
+\item for a set $T\subseteq\delta$ of transitions,
+  $\mks(T)=\bigcup_{t\in T}\mks(t)$ denotes the set of marks that appear on
+  transitions in $T$,
+\item for a run $\pi$, $\mks(\pi)=\mks(\minf(\pi))$ denotes the set of
+  marks that occur in infinitely many transitions of run $\pi$.
+\end{itemize}
+%
+Run $\pi$ is \emph{accepting} iff $\mks(\pi)$ satisfies the formula
+$\varphi$, where a set of marks $M'\subseteq M$ satisfies $\Inf m$ iff
+$m\in M'$ and it satisfies $\Fin m$ iff $m\not\in M'$. The
+\emph{language} of automaton $\mA$ is the set
 $L(\mA)=\{u\in \Sigma^{\omega}\mid\textrm{there is an accepting run of
@@
 appear infinitely often on $\pi$, which is in turn fully determined by
-the set $\inf(\pi)$ of transitions appearing infinitely often in
+the set $\minf(\pi)$ of transitions appearing infinitely often in
 $\pi$.
@@
 % \]
-Further, each set of the form $\inf(\pi)$ is a cycle and vice versa.
-Hence, our reduction methods is looking for
-
-following simple observation. Definovat $inf(\varphi)$?
+Further, each set of the form $\minf(\pi)$ is a cycle and vice versa.
+Hence, our reduction methods is looking for a new acceptance formula
+$\psi$ with less accceptance marks than the original acceptance
+formula $\varphi$ and for new placement of the marks on transitions
+such that for each cycle $C\subseteq\delta$ it holds that the marks $M(C)$
+satisfies $\varphi$ if and only if $M(C)$ satisfies $\psi$.
 

### 2022-02-25 paper/main.tex
progress in section 3 (and improved prelim again)

@@
 \newcommand{\mks}{\mathit{mks}}
+%\newcommand{\oM}{\widehat{M}}
+\newcommand{\oM}{N}
+\newcommand{\rem}{\mathit{nm}}
+\newcommand{\mcycle}{\mathit{cycle}}
+\newcommand{\msat}{\mathit{satisfies}}
 \def\false{\mathit{false}}
@@
 A tuple $t=(p,a,M',q)\in\delta$ is the \emph{transition} leading from
-state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$.  A
-run $\pi$ of $\mA$ over an infinite word
+state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$.
+The set $M'$ is also refered by $\mks(t)$. For a set of transitions
+$T\subseteq\delta$, let $\mks(T)=\bigcup_{t\in T}\mks(t)$ denote the
+set of marks that appear on transitions in $T$.
+% of various objects:
+% \begin{itemize}
+% \item for a transition $t=(p,a,M',q)\in\delta$, $\mks(t)$ denotes the set $M'$,
+% \item for a set $T\subseteq\delta$ of transitions,
+%   $\mks(T)=\bigcup_{t\in T}\mks(t)$ denotes the set of marks that appear on
+%   transitions in $T$,
+% \item for a run $\pi$, $\mks(\pi)=\mks(\minf(\pi))$ denotes the set of
+%   marks that occur in infinitely many transitions of run $\pi$.
+% \end{itemize}
+
+A run $\pi$ of $\mA$ over an infinite word
 $u = u_{0}u_{1}u_{2}\dots \in \Sigma^{\omega}$ is an infinite sequence
@@
 $q_0=q_I$. Let $\minf(\pi)$ denote the set of transitions that appear
-infinitely many times in $\pi$.
-%
-We use $\mks(\cdot)$ to denote the set of acceptance
-marks of various objects:
-\begin{itemize}
-\item for a transitions $t=(p,a,M',q)\in\delta$, $\mks(t)$ denotes the set $M'$,
-\item for a set $T\subseteq\delta$ of transitions,
-  $\mks(T)=\bigcup_{t\in T}\mks(t)$ denotes the set of marks that appear on
-  transitions in $T$,
-\item for a run $\pi$, $\mks(\pi)=\mks(\minf(\pi))$ denotes the set of
-  marks that occur in infinitely many transitions of run $\pi$.
-\end{itemize}
-%
-Run $\pi$ is \emph{accepting} iff $\mks(\pi)$ satisfies the formula
-$\varphi$, where a set of marks $M'\subseteq M$ satisfies $\Inf m$ iff
-$m\in M'$ and it satisfies $\Fin m$ iff $m\not\in M'$. The
-\emph{language} of automaton $\mA$ is the set
+infinitely many times in $\pi$. Run $\pi$ is \emph{accepting} iff
+$\minf(\pi)$ satisfies the formula $\varphi$, where a set of $T$ of
+transitions satisfies $\Inf m$ iff $m\in\mks(T)$ and it satisfies
+$\Fin m$ iff $m\not\in\mks(T)$. The \emph{language} of automaton $\mA$ is
+the set
 $L(\mA)=\{u\in \Sigma^{\omega}\mid\textrm{there is an accepting run of
@@
 $(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
-such that $p=q_0$ and $q=q_n$. We assume that each state $q$ is
-reachable from the initial state $q_I$ (i.e., $q=q_I$ or there is a
-path from $q_I$ to $q$) as states that are not reachable from $q_I$
-can be eliminated without any impact on $L(\mA)$. A nonempty set of
-states $S\subseteq Q$ is called a nontrivial \emph{strongly connected
-  component (SCC)} if for each $p,q\in S$ there is a path from $p$ to
-$q$. An SCC S is \emph{maximal} if there is no SCC $S'$ satisfying
-$S\subsetneq S'$. In the rest of this paper, SCC always refers to a
-maximal SCC. Given an SCC $S$,
-$\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$ denotes the
-set of all transitions between states in $S$. Further, for each mark
-$m\in M$, let $\delta_m$ denotes the set of all transitions marked
-with $m$, i.e., $\delta_m=\{(p,a,M',q)\in\delta\mid m\in M'\}$. The
-sets $\delta_m$ are called \emph{acceptance sets}.  Given an SCC $S$
-and a mark $m\in M$, by $\delta_{S,m}$ we denote the set
-$\delta_{S,m}=\delta_S\cap\delta_m$. Finally, a set of transitions
-$C\subseteq\delta$ is called a \emph{cycle} is there exists a path
-from a state $p$ to the same state containing all transitions of $C$
-and no transition outside $C$.
+such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
+is called a nontrivial \emph{strongly connected component (SCC)} if
+for each $p,q\in S$ there is a path from $p$ to $q$. An SCC S is
+\emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
+the rest of this paper, SCC always refers to a maximal SCC. Given an
+SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
+denotes the set of all transitions between states in $S$. Further, for
+each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
+marked with $m$, i.e., $\delta_m=\{t\in\delta\mid m\in
+\mks(t)\}$. \todo{The sets $\delta_m$ are called \emph{acceptance
+    sets}. Given an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we
+  denote the set $\delta_{S,m}=\delta_S\cap\delta_m$.} A set of
+transitions $T\subseteq\delta$ is called a \emph{cycle} is there
+exists a path from a state $p$ to the same state containing all
+transitions of $T$ and no transition outside $T$. Finally, we assume
+that each TELA $\mA$ contains only states $q$ that are reachable from
+the initial state $q_I$ (i.e., $q=q_I$ or there is a path from $q_I$
+to $q$) as states that are not reachable from $q_I$ can be eliminated
+without any impact on $L(\mA)$.
 
@@
 TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ without altering its
-structure and language. In other words, we change only the acceptance
-marks on individual transitions and we replace the acceptance formula
-by a new one with less acceptance marks. To ensure that the modified
-automaton is equivalent to the original one, we actually construct the
-new acceptance formula such that the set of acceptance runs stays the
-same (modulo changed acceptance marks on transitions).
-
-We start with the several simple observations. Acceptance of a run
-$\pi$ is fully determined by the set $M(\pi)$ of acceptance marks that
-appear infinitely often on $\pi$, which is in turn fully determined by
-the set $\minf(\pi)$ of transitions appearing infinitely often in
-$\pi$.
-% \[
-%   M(\pi)=\bigcup_{(q_i,u_i,M_i,q_{i+1})\in\minf(\pi)}M_i
-% \]
-Further, each set of the form $\minf(\pi)$ is a cycle and vice versa.
-Hence, our reduction methods is looking for a new acceptance formula
-$\psi$ with less accceptance marks than the original acceptance
-formula $\varphi$ and for new placement of the marks on transitions
-such that for each cycle $C\subseteq\delta$ it holds that the marks $M(C)$
-satisfies $\varphi$ if and only if $M(C)$ satisfies $\psi$.
+structure and language. In other words, we are looking for a set $\oM$
+of acceptance marks, an accepance formula $\psi$ over $\oM$, and a
+function
+$\rem:\delta\rightarrow Q \times \Sigma \times 2^{\oM}\times Q$
+assigning new marks to transitions (i.e., for each
+$t=(p,a,M',q)\in\delta$, we assume that $\rem(t)=(p,a,\oM',q)$ for
+some $\oM'\subseteq\oM$) such that $|\oM|<|M|$ and the automaton
+$\mB=(Q,\oM,\Sigma,q_I,\rem(\delta),\psi)$ is equivalent to $\mA$. To
+ensure that the modified automaton is equivalent to the original one,
+we actually look for $\psi$ and $\rem$ such that each run
+$\pi=t_0t_1t_2\ldots$ of $\mA$ is accepting if and only if the run
+$\rem(t_0)\rem(t_1)\rem(t_2)\ldots$ of $\mB$ is accepting.
+
+We start with several simple observations. Acceptance of a run $\pi$
+is fully determined by the set $\minf(\pi)$.
+% of acceptance marks that
+% appear infinitely often on $\pi$, which is in turn fully determined by
+% the set $\minf(\pi)$ of transitions appearing infinitely often in
+% $\pi$.
+Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and
+vice versa. Hence, our reduction methods is looking for a new
+acceptance formula $\psi$ and function $\rem$ such that for each cycle
+$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and
+only if $\rem(T))$ satisifes $\psi$. This can be formally written
+by the formula
+\[
+  \forall T\subseteq\delta~.~\mcycle(T)\implies(T~\msat~\varphi ~\iff~ \rem(T)~\msat~\psi).
+\]
 

### 2022-02-26 paper/main.tex
progress in section 3

@@
 % $\pi$.
-Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and
-vice versa. Hence, our reduction methods is looking for a new
-acceptance formula $\psi$ and function $\rem$ such that for each cycle
-$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and
-only if $\rem(T))$ satisifes $\psi$. This can be formally written
-by the formula
+Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and vice
+versa. Hence, our reduction methods is looking for a new acceptance
+formula $\psi$ and function $\rem$ such that for each cycle
+$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and only
+if $\rem(T)$ satisifes $\psi$. This can be roughly denoted by the
+formula
 \[
-  \forall T\subseteq\delta~.~\mcycle(T)\implies(T~\msat~\varphi ~\iff~ \rem(T)~\msat~\psi).
+  \forall T\subseteq\delta~.~\mcycle(T)\implies\big(\msat_\varphi(T) ~\iff~ \msat_{\psi,\rem}(T)\big).
 \]
+In fact, this corresponds to the shape of the QBF formulas we will
+construct. As we are looking for $\psi$ and $\rem$ such that the
+formula holds, the subformula $\msat_{\psi,\rem}(T)$ contains many
+free variables representing possible instances of $\psi$ and
+$\rem$. If the formula is satisfiable, then its models encode desired
+instances of $\psi$ and $\rem$.  In the following, we assume that we
+are looking for a new acceptance formula $\psi$ in DNF. Note that our
+reduction method can be easily adopted to look for $\psi$ of a
+different shape.
+
+Now we describe the construction of QBF formulas in detail. The
+construction is parametrized by two integers $C,K\ge 0$, where $K$ is
+the desired number of acceptance marks and $C$ is the number of
+clauses of $\psi$. For simplicity, we assume that the reduced
+automaton will use the acceptance marks $N_K=\{1,2,\ldots,K\}$. First
+we desribe Boolean variables used in the constructed QBF formulas.
+\begin{itemize}
+\item The fact whether a transition $t$ is in the set $T$ is
+  represented by a boolean variable $e_t$. By $\vec{e}$ we denote the vector
+  of length $|\delta|$ of all variables $e_t$.
+\item 
+\end{itemize}
+
+\[
+  \forall \vec{e}~.~\mcycle(\vec{e})\implies\big(\msat_\varphi(\vec{e}) ~\iff~ \msat(\vec{e},\vec{k},\vec{p})\big).
+\]
 
 \textbf{Jan is here!}
+
+\subsection{Lightweight version of $\mcycle(\vec{e})$ (level 1)}
+
+\subsection{Balanced version of $\mcycle(\vec{e})$ (level 2)}
+
+\subsection{Strict version of  $\mcycle(\vec{e})$ (level 3)}
 
@@
 
-This section explains the construction of QBF formulas for a given
-TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ and two integer parameters
+This section explains the construction of QBF formulas for a given3TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ and two integer parameters
 $C,K\ge 0$ such that satisfiability of such a formula implies that

### 2022-02-26 main.py
tweak

@@ -40,13 +40,19 @@ def main(argv):
     parser.add_argument(
         "-Q",
         "--qbf_solver",
-        help="Choose wchich solver is going to be used.", choices=['limboole', 'z3'], default='limboole')
+        help="Choose wchich solver is going to be used.", choices=['limboole', 'z3'], default='z3')
+    parser.add_argument(
+        "-B",
+        "--base_level",
+        help="Choose whether qbf level 1 or standard level 1(original TELAtko).", choices=['telatko', 'z3'], default='z3')
 
     args = parser.parse_args()
 
 
     if not args.autfile:
         print("No automata to process.", file=sys.stderr)
+    if args.level > 3:
+        args.level = 3
 
     aut = spot.automata(args.autfile)
     timeouted = [0]
@@ -58,30 +64,24 @@ def main(argv):
         print_aut(origin, "problem", "w")
         try:
             spot.cleanup_acceptance_here(a)
-
-            #a_tele = process_automaton(tmp_a)
-
             acc_sets_count = a.get_acceptance().used_sets().count()
             clauses_count = len(
                 a.get_acceptance().to_dnf().top_disjuncts())
-            if args.level > 3:
-                args.level = 3
 
-            if args.level >= 1 and args.level <= 4:
+            if args.base_level == "telatko":
+                a = process_automaton(a)
+                if args.level > 1 and acc_sets_count != 0:
 
+                    auto = play(
+                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut, args.qbf_solver, 2)
+                else:
+                    auto = a
+            else:
                 if acc_sets_count == 0:
                     auto = a
                 else:
-
                     auto = play(
-                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut, args.qbf_solver)
-                    """
-                    if a_tele.get_acceptance().used_sets().count() < auto.get_acceptance().used_sets().count():
-                        print_aut(origin, "better_tele_l1.hoa", "a")
-                    """
-
-            else:
-                auto = a
+                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut, args.qbf_solver, 1)
 
             if args.outfile:
                 print_aut(auto, args.outfile, "a")

### 2022-02-26 qbf/q_main.py

@@ -153,58 +153,6 @@ def create_formula(
 
         return z3_formula
 
-    # quantified edges #e_1 ... #e_n
-    quant_edges = quant_all(inner_edges_nums)
-    # quantified variables ?w_1 ?w_2 ... ?w_n
-
-    # quant_edges += w_quant(aut)
-    if mode > 2:
-        quant_edges += w_quant(aut)
-
-    # edges that are true create continuous cycle of edges aka laso
-
-    if mode == 3:
-        laso = laso2(aut, inner_edges, scc_edg)
-        in_out = in_n_out(scc_state_info)
-        laso.add_subf(in_out)
-    else:
-        laso = laso_f(
-            aut,
-            inner_edges_nums,
-            scc_state_info,
-            scc_edg,
-            inner_edges,
-            mode)
-
-    # reqiurements on old acceptance formula
-    old = old_formula(aut.get_acceptance(), edge_dict)
-
-    # assignment of variables to create new acceptance formula
-    new = new_formula(inner_edges_nums, C, K)
-    # old and new need to be equivalent
-    eq = SATformula('<->')
-    eq.add_subf(old)
-    eq.add_subf(new)
-
-    impl = SATformula("->")
-    impl.add_subf(laso)
-    impl.add_subf(eq)
-
-    # root of QBFformula
-    con = SATformula("&")
-    con.add_subf(impl)
-
-    # not (Inf1 & Fin1)
-    inf_not_fin = inf_is_not_fin_clause(C, K)
-
-    # (Fin1 | Inf1) - we need control over formula
-    inf_or_fin = inf_or_fin_f(C, K)
-    con.add_subf(inf_not_fin)
-    con.add_subf(inf_or_fin)
-    # prints our formula into text file
-    SAT_output("sat_file", quant_edges, con)
-
-
 def try_evaluate0(aut, orig):
     """
     Try evaluate with K = 0(len of acceptance formula == 0)
@@ -326,7 +274,7 @@ def resolve_formula_atributes(minimized_atribute, C, K):
 
 
 def play(aut, C, K, mode, timeout, timeouted,
-         optimized_scc, minimized_atribute, qbf_solver):
+         optimized_scc, minimized_atribute, qbf_solver, tmp_mode):
 
     spot.cleanup_acceptance_here(aut)
 
@@ -345,7 +293,7 @@ def play(aut, C, K, mode, timeout, timeouted,
     scc_state_info, scc_edg = scc_info(aut)
 
[…]

### 2022-02-26 qbf/z3_formula.py

@@ -281,7 +281,7 @@ class Z3_f_ctor(FormulaCreator):
         # true
         one_scc = self.one_scc_f()
 
-        laso = And(in_out, one_scc)s
+        laso = And(in_out, one_scc)
         return laso
 
 

### 2022-02-26 main.py
telatko: print() cleanup

@@ -89,12 +89,8 @@ def main(argv):
                 print_aut(auto, None, " ")
 
             if not spot.are_equivalent(origin, auto):
-                print("not equivalent")
                 print_aut(origin, "not_eq", "w")
-                print("timeouted:", timeouted[0])
                 return
-            else:
-                print("equivalent")
 
             acc_sets_count2 = auto.get_acceptance().used_sets().count()
 
@@ -106,7 +102,6 @@ def main(argv):
                 "Automaton has too many acceptance sets, 32 is the limit.",
                 file=sys.stderr)
 
-    print("timeouted:", timeouted[0])
 
 
 if __name__ == "__main__":

### 2022-02-26 qbf/q_main.py

@@ -46,7 +46,6 @@ def edge_dictionary(aut, mode):
                         add_or_append(edge_dict, m, aut.edge_number(e))
 
         scc_equiv_edges.append(marks_edges)
-    #print(scc_equiv_edges)
     return edge_dict, scc_equiv_edges, representants
 
 
@@ -90,9 +89,7 @@ def scc_info(aut):
     for i in range(si.scc_count()):
         states = si.states_of(i)
         if si.is_trivial(i):
-            #print("trivial before", i)
             continue
-        #print("i: ", i)
         state_dict = {}
         for s in states:
             state_dict[s] = [[], []]
@@ -180,7 +177,6 @@ def count_sets(sets):
 
 def scc_optimized_formula(aut, acc, scc_state_info, C, K, L):
     formula = SATformula('&')
-    #print("scc state info:" , scc_state_info)
     weak = spot.scc_info(aut).weak_sccs()
 
     si = spot.scc_info(aut)
@@ -193,7 +189,6 @@ def scc_optimized_formula(aut, acc, scc_state_info, C, K, L):
     for scc in (si):
 
         if si.is_trivial(scc_counter):
-            #print("trivial", si.states_of(scc_counter))
             scc_counter += 1
             continue
 
@@ -205,9 +200,7 @@ def scc_optimized_formula(aut, acc, scc_state_info, C, K, L):
         eq = SATformula('<->')
         #old = old_formula_scc(aut.get_acceptance(), mark_edg_dict, edges_translator, si.acc_sets_of(scc))
 
-        # print(si.used_acc_of(scc_counter))
         if si.is_rejecting_scc(scc_counter):
-            # print("rejecting")
             # old je False
             old = SATformula("&")
             old.add_subf("t")
@@ -218,7 +211,6 @@ def scc_optimized_formula(aut, acc, scc_state_info, C, K, L):
 
         elif weak[scc_counter] and si.is_accepting_scc(scc_counter):
 
-            #print("weak accepting", si.states_of(scc_counter))
             old = SATformula("|")
             old.add_subf("t")
             old.add_subf("!t")
@@ -265,7 +257,6 @@ def scc_optimized_formula(aut, acc, scc_state_info, C, K, L):
 
 def resolve_formula_atributes(minimized_atribute, C, K):
     if minimized_atribute == 'clauses':
-        #print("reducing clauses")
 
         return FormulaAtribute.C
     else:
[…]

### 2022-02-26 qbf/z3_formula.py

@@ -26,9 +26,7 @@ class Z3_f_ctor(FormulaCreator):
         s = Solver()
         s.add(f)
         if s.check() == sat:
-            print("--satisfiable--")
             s = s.model()
-            print(s)
 
     def prune_scc_edg(self, representants):
         tmp = []
@@ -81,21 +79,17 @@ class Z3_f_ctor(FormulaCreator):
         old = self.old_formula(ACC_DNF(aut.get_acceptance().to_dnf()))
 
         new = self.new_formula()
-        #print(new)
         old = simplify(old)
-        #print(new)
         new = simplify(new)
 
         laso = simplify(laso)
 
         eq = self.equivalence_f(old, new)
-        #print(eq)
         impl = Implies(laso, eq)
         inf_not_fin = self.inf_is_not_fin_clause()
         inf_or_fin = self.inf_or_fin_f()
         inf_not_fin = simplify(inf_not_fin)
         inf_or_fin = simplify(inf_or_fin)
-        #print(inf_not_fin, inf_or_fin)
         #assert(False)
 
         formula = And(impl, inf_not_fin, inf_or_fin)
@@ -136,7 +130,6 @@ class Z3_f_ctor(FormulaCreator):
         main_dis.append(And(con1))
         main_dis.append(And(con2))
         main_dis.append(And(con3))
-        #print("main dis: ", main_dis)
         one_scc = self.one_scc_f()
         laso = And(Or(main_dis), one_scc)
 

### 2022-02-26 qbf/q_main.py
name of aut fce

@@ -150,16 +150,19 @@ def create_formula(
 
         return z3_formula
 
-def try_evaluate0(aut, orig):
+def try_evaluate0(aut, orig, qbf_run_info):
     """
     Try evaluate with K = 0(len of acceptance formula == 0)
     :param aut: spot::automaton
     :return: potentionally zero acceptance spot::automaton
     """
     last_eq_aut = spot.automaton(aut.to_str())
+    aut.set_name(qbf_run_info + "F_0_S")
+    last_eq_aut.set_name(qbf_run_info + "F_0_U")
     clear_aut_edges(aut)
     aut.set_acceptance(0, spot.acc_code.t())
     if spot.are_equivalent(orig, aut):
+
         return aut
 
     aut.set_acceptance(0, spot.acc_code.f())
@@ -263,6 +266,14 @@ def resolve_formula_atributes(minimized_atribute, C, K):
 
         return FormulaAtribute.K
 
+def update_run_info(K, res, tmp_mode):
+    if res == z3.unknown:
+        str_res = "T"
+    if res == unsat:
+        str_res = "U"
+    assert(str_res != "")
+    return "L_{0}_{1}_{2} ".format(str(tmp_mode), str(K), str_res)
+
 
 def play(aut, C, K, mode, timeout, timeouted,
          optimized_scc, minimized_atribute, qbf_solver, tmp_mode):
@@ -284,6 +295,7 @@ def play(aut, C, K, mode, timeout, timeouted,
     scc_state_info, scc_edg = scc_info(aut)
 
     # K = K - 1  # we dont want to try what we already know
+    qbf_run_info = ""
 
 
     currently_reduced = resolve_formula_atributes(minimized_atribute, C, K)
@@ -305,6 +317,7 @@ def play(aut, C, K, mode, timeout, timeouted,
         # QBF formula is written into ./sat_file
         formula = None
         if (optimized_scc):
+            # tohle pujde dopryc ?
             acc = ACC_DNF(aut.get_acceptance().to_dnf())
             scc_optimized_formula(aut, acc, scc_state_info, C, K, tmp_mode)
         else:
@@ -327,16 +340,15 @@ def play(aut, C, K, mode, timeout, timeouted,
             solver.add(formula)
             solver.set("timeout",1000* timeout)
             solver.push()
+            res = solver.check()
 
-
-            if solver.check() == sat:
+            if  res == sat:
                 s = solver.model()
-
[…]

### 2022-02-26 paper/main.tex
progress in section 3 (we are no page 6!)

@@
 clauses of $\psi$. For simplicity, we assume that the reduced
-automaton will use the acceptance marks $N_K=\{1,2,\ldots,K\}$. First
-we desribe Boolean variables used in the constructed QBF formulas.
-\begin{itemize}
-\item The fact whether a transition $t$ is in the set $T$ is
-  represented by a boolean variable $e_t$. By $\vec{e}$ we denote the vector
-  of length $|\delta|$ of all variables $e_t$.
-\item 
-\end{itemize}
-
+automaton will use the acceptance marks $N_K=\{1,2,\ldots,K\}$.
+%Further, we consider only subsets $T\subseteq\bigcup_{\textrm{SCC}~S}\delta_S$
+First we describe Boolean variables used in the constructed QBF
+formulas.
+\begin{itemize}
+\item For each transition $t\in\delta$, variable $e_t$ says whether
+  $t$ is in the current set $T$ or not.
+  \[e_t=\left\{
+      \begin{array}{ll}
+        1 & \textrm{if }t\in T\\
+        0 & \textrm{otherwise}    
+      \end{array}\right.
+  \]
+%  In fact, we consider variables 
+\item For each transition $t\in\delta$ and acceptance mark $k\in N_K$,
+  variable $n_{t,k}$ says whether $k$ is on the transition
+  $\rem(t)$ or not.
+  \[n_{t,k}=\left\{
+      \begin{array}{ll}
+        1 & \textrm{if }k\in\mks(\rem(t))\\
+        0 & \textrm{otherwise}    
+      \end{array}\right.
+  \]
+\item For each $c\in\{1,2,\ldots,C\}$ and acceptance mark $k\in N_K$,
+  variables $i_{c,k}$ and $f_{c,k}$ say whether the $c^\textrm{th}$
+  clause of $\psi$ contains terms $\Inf k$ or $\Fin k$, respectively.
+  \[i_{c,k}=\left\{
+      \begin{array}{ll}
+        1 & \textrm{if $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\
+        0 & \textrm{otherwise}    
+      \end{array}\right.
+  \]
+  \[f_{c,k}=\left\{
+      \begin{array}{ll}
+        1 & \textrm{if $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\
+        0 & \textrm{otherwise}    
+      \end{array}\right.
+  \]
+
+\end{itemize}
+By $\vec{e},\vec{n},\vec{i},\vec{f}$ we denote the vectors of all
+variables of the form $e_t$, $n_{t,k}$, $i_{c,k}$, and $f_{c,k}$,
+respectively. The constructed QBF formulas have the form 
 \[
-  \forall \vec{e}~.~\mcycle(\vec{e})\implies\big(\msat_\varphi(\vec{e}) ~\iff~ \msat(\vec{e},\vec{k},\vec{p})\big).
+  \Phi_{C,K}(\vec{n},\vec{i},\vec{f})~~=~~\forall \vec{e}~.~\mcycle(\vec{e})\implies\big(\msat_\varphi(\vec{e}) \iff \msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})\big).
 \]
-
-\textbf{Jan is here!}
+It remains to describe the subformulas $\mcycle(\vec{e})$,
+$\msat_\varphi(\vec{e})$, and
+$\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$. The subformula
+$\mcycle(\vec{e})$ will be defined as the last one and we present
+three versions of it.
+
+The subformula $\msat_\varphi(\vec{e})$ says whether $T$ satisfies the
+original acceptance formula $\varphi$ and it is derived directly from
+$\varphi$. Recall that $T$ satisfies $\Inf m$ iff $m\in\mks(T)$, which
+means that $T$ contains some transition with mark $m$. As the
+transitions with mark $m$ form the set $\delta_m$, $\Inf m$ can be
+expresses by $\bigvee_{t\in\delta_m}e_t$. Similarly, $T$ satisfies
+$\Fin m$ iff $m\not\in\mks(T)$, which can be expresses by
+$\bigwedge_{t\in\delta_m}\neg e_t$. Hence, $\msat_\varphi(\vec{e})$
+arises from $\varphi$ by replacing
+\begin{itemize}
+\item all terms of the form $\Inf m$ by $\bigvee_{t\in\delta_m}e_t$ and
+\item all terms of the form $\Fin m$ by $\bigwedge_{t\in\delta_m}\neg e_t$.
+\end{itemize}
+
+Now we construct the subformula
+$\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ that evaluates to true
+iff $\rem(T)$ satisfies~$\psi$. The subformula reflects the basic
+structure of $\psi$. As we assume that $\psi$ is a disjunction of $C$
+clauses, we have
+\[
+  \msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})=\bigvee_{c\in\{1,2,\ldots,C\}}\xi_{c,K}(\vec{e},\vec{n},\vec{i},\vec{f})
+\]
+where each $\xi_{c,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ corresponds to
+one clause. Recall that the presence of terms $\Inf k$ and $\Fin k$ in
+the $c^\textrm{th}$ clause is given by variables $i_{c,k}$ and
+$f_{c,k}$, respecively. $\Inf k$ is satisfied by $\rem(T)$ iff $T$
+contains a transition $t$ such that $k\in\mks(\rem(t))$, which can be
+expressed as $\bigvee_{t\in\delta}(e_t\wedge n_{t,k})$. Similarly,
+$\Fin k$ is satisfied by $\rem(T)$ iff there is no transition $t\in T$
+such that $k\in\mks(\rem(t))$, which can be expressed as
+$\bigwedge_{t\in\delta}\neg(e_t\wedge n_{t,k})$. Hence, we set
+\[
+  \xi_{c,K}(\vec{e},\vec{n},\vec{i},\vec{f})=
+  \bigwedge_{k\in N_K}\Big(i_{c,k}\implies\bigvee_{t\in\delta}(e_t\wedge n_{t,k})\Big)\wedge\Big(f_{c,k}\implies\bigwedge_{t\in\delta}\neg(e_t\wedge n_{t,k})\Big).
+\]
+
+It remains to define the subformula $\mcycle(\vec{e})$ with the intended meaning
+\[
+  \vec{e}\textrm{ represents a cycle} \iff \mcycle(\vec{e}).
+\]
+In fact, our reduction methods remains correct whenever ``implies'' holds.
+If $\mcycle(\vec{e})$
+
+  $\vec{e}$ represents a cycle. In fact, $\mcycle(\vec{e})$ needs to be
+satisfied by all cycles, but our reduction technique remains correct
+even if it is satisfied also by some sets $T\subseteq\delta$ that are
+not cycles. The only consequence is that
+
+Note that our reduction method remains
+correct even if $\mcycle(\vec{e})$ isAs we will see, However, our reduction method remains
+correct.
+
+The intended meaning
+of this formula is We offer three definitions:
+
+\todo{Jan is here!}
 
@@
 \subsection{Strict version of  $\mcycle(\vec{e})$ (level 3)}
+
+\subsection{Optimizations}
+
+SCC-based approach
+
+Its first part $\xi_1(\vec{i},\vec{f})$ says that any clause does not
+contain both $\Inf k$ and $\Fin k$ for the same mark $k$ as a clause
+with both such terms is always unsatisfiable. Note that the subformula
+$\xi_1(\vec{i},\vec{f})$ can be safely omitted, but it speeds up the QBF solving.
+
+\todo{Zbytek je stary text!}
 

### 2022-02-26 paper/main.tex
progress in section 3

@@
       \begin{array}{ll}
-        1 & \textrm{if $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\
+        1 & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\
         0 & \textrm{otherwise}    
@@
       \begin{array}{ll}
-        1 & \textrm{if $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\
+        1 & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\
         0 & \textrm{otherwise}    
@@
 \]
-In fact, our reduction methods remains correct whenever ``implies'' holds.
-If $\mcycle(\vec{e})$
-
-  $\vec{e}$ represents a cycle. In fact, $\mcycle(\vec{e})$ needs to be
-satisfied by all cycles, but our reduction technique remains correct
-even if it is satisfied also by some sets $T\subseteq\delta$ that are
-not cycles. The only consequence is that
-
-Note that our reduction method remains
-correct even if $\mcycle(\vec{e})$ isAs we will see, However, our reduction method remains
-correct.
-
-The intended meaning
-of this formula is We offer three definitions:
-
-\todo{Jan is here!}
-
-\subsection{Lightweight version of $\mcycle(\vec{e})$ (level 1)}
-
-\subsection{Balanced version of $\mcycle(\vec{e})$ (level 2)}
-
-\subsection{Strict version of  $\mcycle(\vec{e})$ (level 3)}
+In fact, only the direction ``$\Longrightarrow$'' is required for
+correctness of our reduction method. If $\mcycle(\vec{e})$ holds for
+some valuation of $\vec{e}$ that does not represent a cycle, then the
+required equivalence
+$\msat_\varphi(\vec{e}) \iff
+\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ on this valuation of
+$\vec{e}$ adds a superfluous constraint on $\psi$ represented by
+$\vec{i},\vec{f}$ and $\rem$ represented by $\vec{n}$. These
+superfluous constraints can lead to loss of reduction opportunities,
+but not to incorrectness. This observation allows us to trade the
+precision of $\mcycle(\vec{e})$ for simplicity of the formula.
+
+In the following, we define three versions of $\mcycle(\vec{e})$:
+\begin{itemize}
+\item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
+  only minimal requirements on $\vec{e}$. It does not use the
+  information about automaton structure, but it comes with an
+  interesting simplification of the whole formula $\Phi_{C,K}$.
+\item $\mcycle_2(\vec{e})$ is a balanced version. It uses the
+  information about the automaton structure to identify many
+  valuations of $\vec{e}$ that are not cycles. It is relatively
+  simple, but it is still satisfied by some valuations of $\vec{e}$
+  not corresponding to cycles.
+\item $\mcycle_3(\vec{e})$ is a strict version satisfied only by
+  the valuations of $\vec{e}$ representing cycles. Unfortunately,
+  it uses additional quantified variables. 
+\end{itemize}
+
+\subsection{Lightweight version $\mcycle_1(\vec{e})$}
+The lightweight version is based on the fact that every cycle is nonempty.
+Indeed, the only restriction imposed by
+\[
+  \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
+\]
+is that $\vec{e}$ represents a nonempty set.
+
+The fact that this version do
+
+\subsection{Balanced version $\mcycle_2(\vec{e})$}
+
+\subsection{Strict version $\mcycle_3(\vec{e})$}
 

### 2022-02-27 paper/main.tex
some progress in section 3

@@
 \newcommand{\msat}{\mathit{satisfies}}
+\newcommand{\Te}{T_{\vec{e}}}
 \def\false{\mathit{false}}
@@
 
-It remains to define the subformula $\mcycle(\vec{e})$ with the intended meaning
+It remains to define the subformula $\mcycle(\vec{e})$. Let $\Te$
+denote the set of transitions represented by $\vec{e}$. The original
+intended meaning of $\mcycle(\vec{e})$ is
 \[
-  \vec{e}\textrm{ represents a cycle} \iff \mcycle(\vec{e}).
+  \mcycle(\vec{e}) \iff \Te\textrm{ is a cycle}.
 \]
-In fact, only the direction ``$\Longrightarrow$'' is required for
-correctness of our reduction method. If $\mcycle(\vec{e})$ holds for
-some valuation of $\vec{e}$ that does not represent a cycle, then the
-required equivalence
+In fact, only the direction ``$\Longleftarrow$'' is required for the
+correctness of our reduction method. If $\Te$ is not a cycle and
+$\mcycle(\vec{e})$ holds, then the required equivalence
 $\msat_\varphi(\vec{e}) \iff
@@
 \item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
-  only minimal requirements on $\vec{e}$. It does not use the
+  only minimal requirements on $\Te$. It does not use the
   information about automaton structure, but it comes with an
   interesting simplification of the whole formula $\Phi_{C,K}$.
-\item $\mcycle_2(\vec{e})$ is a balanced version. It uses the
-  information about the automaton structure to identify many
-  valuations of $\vec{e}$ that are not cycles. It is relatively
-  simple, but it is still satisfied by some valuations of $\vec{e}$
-  not corresponding to cycles.
-\item $\mcycle_3(\vec{e})$ is a strict version satisfied only by
-  the valuations of $\vec{e}$ representing cycles. Unfortunately,
-  it uses additional quantified variables. 
+\item $\mcycle_2(\vec{e})$ is a balanced version. It says that every
+  transition in $\Te$ has to have a preceding and a succeeding
+  transition in $\Te$, which is a necessary condition on cycles, but
+  not a sufficient one.
+\item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
+  cycle. Unfortunately, it uses additional universally quantified
+  variables corresponding to automata states. Transformation to prenex
+  normal form turns the quantifiers to existential ones and the
+  resulting formula thus contains quantifier alternation.
 \end{itemize}
@@
 \]
-is that $\vec{e}$ represents a nonempty set.
-
-The fact that this version do
+is that $\Te$ is nonempty.
+
+With this lightweight version of $\mcycle(\vec{e})$, the formula
+$\Phi_{C,K}$ actually says that
+for every nonempty set
+$T\subseteq\delta$, $T$ satisfies $\varphi$ if and only if $\rem(T)$ satisfies
+$\psi$. Note that the formula is not affected by the placement of
+transitions in the automaton. The only aspect of a transition $t$
+reflected by the formula is its set of marks $\mks(t)$. Hence, we do
+not have to distinguish between transitions with the same sets of
+marks.
+
+Formally, we define an equivalence on transitions such that for each
+$t_1,t_2\in\delta$ we write $t_1\sim t_2$ whenever
+$\mks(t_1)=\mks(t_2)$.
+%This relation is clearly an equivalence.
+We prove the following statement. 
+\begin{lemma}
+  Assume that there is a function $\rem$ and a formula $\psi$ such
+  that for every nonempty set $T\subseteq\delta$ it holds that $T$
+  satisfies $\varphi$ iff $\rem(T)$ satisfies $\psi$. Then we can
+  safely assume that $\rem$ respects the quivalence $\sim$, that is
+  $t_1\sim t_2\implies\rem(t_1)=\rem(t_2)$.
+\end{lemma}
+\begin{proof}
+  \todo{Jan is here!}
+\end{proof}
+
+Formally, for each pair of transitions $t_1,t_2\in\delta$, we write
+$t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$. This relation is clearly an equivalence.
+Further, let us say that
+
+We show that if this formula holds for some $\rem$ and $\psi$,
+then it also holds for some $\rem'$ and $\psi$ such that $\rem'$
+assigns the same sets of acceptance marks to transitions originally
+marked by the same sets of marks.
 

### 2022-02-27 main.py
optimization for SCC

@@ -89,8 +89,10 @@ def main(argv):
                 print_aut(auto, None, " ")
 
             if not spot.are_equivalent(origin, auto):
+                assert(False)
                 print_aut(origin, "not_eq", "w")
                 return
+
 
             acc_sets_count2 = auto.get_acceptance().used_sets().count()
 

### 2022-02-27 qbf/formula.py

@@ -38,13 +38,11 @@ class FormulaCreator:
 
 
 # how can I write this more pretty?
-class Limboole_f_ctor(FormulaCreator):
+class Limboole_f_ctor():
     def __init__(self,  edge_dict, scc_edg, scc_state_info
                   ,inner_edges_nums,C, K, inner_edges
-                  , mode, qbf_solver):
-        super().__init__( edge_dict, scc_edg, scc_state_info
-                      ,inner_edges_nums,C, K, inner_edges
-                      , mode, qbf_solver)
+                  , mode)
+
     def get_qbf_formula(self, aut):
         # quantified edges #e_1 ... #e_n
         quant_edges = quant_all(self.inner_edges_nums)

### 2022-02-27 qbf/z3_formula.py

@@ -1,19 +1,21 @@
 from telatko2.classes import *
 from z3 import *
-from qbf.q_main import FormulaCreator
+#from qbf.q_main import FormulaCreator
 
 
 
-class Z3_f_ctor(FormulaCreator):
-    def __init__(self,  edge_dict, scc_edg, scc_state_info
-                  ,inner_edges_nums,C, K, inner_edges
-                  , mode, qbf_solver):
-        super().__init__( edge_dict, scc_edg, scc_state_info
-                      ,inner_edges_nums,C, K, inner_edges
-                      , mode, qbf_solver)
-
+class Z3_f_ctor():
+    def __init__(self,  edge_dict, scc_edg, scc_state_info , inner_edges_nums ,C, K, inner_edges, mode):
         self.univ_vars = []
         self.exist_vars = []
+        self.edge_dict = edge_dict
+        self.scc_edg = scc_edg
+        self.scc_state_info = scc_state_info
+        self.inner_edges_nums = inner_edges_nums
+        self.C = C
+        self.K = K
+        self.inner_edges = inner_edges
+        self.mode = mode # level of simplification
 
 
     def foo(self):
@@ -26,7 +28,74 @@ class Z3_f_ctor(FormulaCreator):
         s = Solver()
         s.add(f)
         if s.check() == sat:
+
             s = s.model()
+
+
+    def add_inf_fin(self, impl):
+        inf_not_fin = self.inf_is_not_fin_clause()
+        inf_or_fin = self.inf_or_fin_f()
+        inf_not_fin = simplify(inf_not_fin)
+        inf_or_fin = simplify(inf_or_fin)
+        formula = And(impl, inf_not_fin, inf_or_fin)
+        return formula
+
+    def accepting_scc(self, representants):
+        self.inner_edges_nums = representants
+        self.quant_all()
+
+        self.prune_scc_edg(representants)
+        laso = self.one_scc_f()
+        new = self.new_formula()
+        #new = simplify(new)
+
+        impl = Implies(laso, (new))
+        formula = self.add_inf_fin(impl)
+
+        formula = ForAll(self.univ_vars, formula)
+
+        return formula
+
+    def rejecting_scc_l23(self, aut):
+        self.quant_all()
+        if self.mode > 2:
+            self.w_quant(aut)
+
+        if self.mode == 3:
+            laso = self.laso_cycles(aut)
+            laso = simplify(laso)
+            in_out = self.in_n_out()
[…]

### 2022-02-28 paper/main.tex
lemma in section 3.1 proven!

@@
   interesting simplification of the whole formula $\Phi_{C,K}$.
-\item $\mcycle_2(\vec{e})$ is a balanced version. It says that every
-  transition in $\Te$ has to have a preceding and a succeeding
-  transition in $\Te$, which is a necessary condition on cycles, but
-  not a sufficient one.
+\item $\mcycle_2(\vec{e})$ is a balanced version. It says that $\Te$
+  is nonempty and every transition in $\Te$ has to have a preceding
+  and a succeeding transition in $\Te$, which is a necessary condition
+  for being a cycle, but not a sufficient one.
 \item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
@@
 \subsection{Lightweight version $\mcycle_1(\vec{e})$}
-The lightweight version is based on the fact that every cycle is nonempty.
-Indeed, the only restriction imposed by
+The lightweight version is defined as
 \[
@@
 \]
-is that $\Te$ is nonempty.
+which means only that $\Te$ is nonempty. This condition is satisfied by
+every cycle.
 
@@
 
-Formally, we define an equivalence on transitions such that for each
-$t_1,t_2\in\delta$ we write $t_1\sim t_2$ whenever
-$\mks(t_1)=\mks(t_2)$.
-%This relation is clearly an equivalence.
-We prove the following statement. 
+Formally, we define an equivalence $\sim\subseteq\delta\times\delta$
+on transitions such that $t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$.
 \begin{lemma}
-  Assume that there is a function $\rem$ and a formula $\psi$ such
-  that for every nonempty set $T\subseteq\delta$ it holds that $T$
-  satisfies $\varphi$ iff $\rem(T)$ satisfies $\psi$. Then we can
-  safely assume that $\rem$ respects the quivalence $\sim$, that is
-  $t_1\sim t_2\implies\rem(t_1)=\rem(t_2)$.
+  Assume that there is a function $\rem$ and a formula $\psi$ such that 
+  \[
+    \textrm{for every nonempty set $T\subseteq\delta$ it holds }
+      (T~\msat~\varphi~\iff~\rem(T)~\msat~\psi).\tag{$1$}\label{eq:lem1} 
+  \]  
+  Then there exists a function $\rem'$ that respects % the equivalence
+  $\sim$ (i.e.,
+  if $t_1\sim t_2$ then $\rem'(t_1)=\rem'(t_2)$) and
+  \[
+    \textrm{for every nonempty set $T\subseteq\delta$ it holds }
+    (T~\msat~\varphi~\iff~\rem'(T)~\msat~\psi).\tag{$2$}\label{eq:lem2} 
+  \]  
 \end{lemma}
 \begin{proof}
-  \todo{Jan is here!}
+  Let $\rem$ be a function and $\psi$ a formula such that
+  \eqref{eq:lem1} holds. To construct the function $\rem'$, we first
+  select one transition from each equivalence class of $\sim$. For
+  every transition $t\in\delta$, by $\overline{t}$ we denote the
+  selected transition equivalent to $t$. We define the function
+  $\rem'$ as $\rem'(t)=\rem(\overline{t})$. Clearly, $\rem'$ respects
+  the equivalence $\sim$. It remains to show that \eqref{eq:lem2}
+  holds for $\rem'$ and $\psi$.
+
+  Let $T\subseteq\delta$ be an arbitrary nonempty set. We construct
+  the set $\overline{T}=\{\overline{t}\mid t\in T\}$. As
+  $\mks(t)=\mks(\overline{t})$ holds for all transitions, we get
+  $\mks(T)=\mks(\overline{T})$ and thus
+  \[
+    T~\msat~\varphi~\iff~\overline{T}~\msat~\varphi.
+  \]
+  Now we apply $\eqref{eq:lem1}$ on $\overline{T}$ to get that
+  \[
+    \overline{T}~\msat~\varphi~\iff~\rem(\overline{T})~\msat~\psi.
+  \]
+  Finally, the definition of $\rem'$ implies that
+  $\rem'(T)=\rem(\overline{T})$ and thus
+  \[
+    \rem(\overline{T})~\msat~\psi~\iff~\rem'(T)~\msat~\psi.
+  \]
+  Altogether, we obtain
+  \[
+    T~\msat~\varphi~\iff~\overline{T}~\msat~\varphi~\iff~\rem(\overline{T})~\msat~\psi~\iff~\rem'(T)~\msat~\psi
+  \]
+  which proves that \eqref{eq:lem2} holds for $\rem'$ and $\psi$. 
 \end{proof}
 
-Formally, for each pair of transitions $t_1,t_2\in\delta$, we write
-$t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$. This relation is clearly an equivalence.
-Further, let us say that
-
-We show that if this formula holds for some $\rem$ and $\psi$,
-then it also holds for some $\rem'$ and $\psi$ such that $\rem'$
-assigns the same sets of acceptance marks to transitions originally
-marked by the same sets of marks.
+The previous lemma and its proof suggest implies that it is sufficient to
 

### 2022-02-28 paper/main.tex
section 3.1 completed

@@
 
-The previous lemma and its proof suggest implies that it is sufficient to
+The lemma suggests the following simplification of the whole formula
+$\Phi_{C,K}$ built with $\mcycle_1(\vec{e})$. Before we build the
+formula, we compute the equivalence $\sim$ on automata transitions and
+remove all transitions except one of each class. Then we build the
+formula for this pruned automaton as described above. The more
+transitions we removed, the shorter formula with less $e_t$ variables
+we obtain. If the formula $\Phi_{C,K}$ for the pruned automaton is
+satisfiable and we derive $\rem$ and $\psi$ from its model, we can use
+safely them also on the original automaton. We only need to extend
+$\rem$ to all transitions such that it changes the acceptance marks on
+all equivalent transitions in the same way.
 

### 2022-03-01 paper/main.tex
some figures and more text in sec. 3

@@
 In graphical representation of automata, we often use acceptance marks
-$\tacc{1}{1},\tacc{2}{2},\ldots\in M$. An edge 
+$\tacc{1}{1},\tacc{2}{2},\ldots\in M$. Further, an edge 
 \begin{tikzpicture}[smallautomaton,baseline=-2pt]
@@
 \begin{itemize}
-\item $\mcycle_1(\vec{e})$ is a lightweight version, which puts
-  only minimal requirements on $\Te$. It does not use the
-  information about automaton structure, but it comes with an
-  interesting simplification of the whole formula $\Phi_{C,K}$.
-\item $\mcycle_2(\vec{e})$ is a balanced version. It says that $\Te$
+\item $\mcycle_1(\vec{e})$ is a lightweight version, which only says
+  that $\Te$ is nonempty. It does not use the information about
+  automaton structure, but it comes with an interesting simplification
+  of the whole formula $\Phi_{C,K}$.
+\item $\mcycle_2(\vec{e})$ is an intermediate version. It says that $\Te$
   is nonempty and every transition in $\Te$ has to have a preceding
@@
   cycle. Unfortunately, it uses additional universally quantified
-  variables corresponding to automata states. Transformation to prenex
-  normal form turns the quantifiers to existential ones and the
-  resulting formula thus contains quantifier alternation.
+  variables corresponding to automata states. Transformation of
+  $\Phi_{C,K}$ to prenex normal form turns the quantifiers to
+  existential ones and the resulting formula thus contains quantifier
+  alternation.
 \end{itemize}
+We write $\Phi_{i,C,K}$ when we want to emphasize that the particular
+formula $\Phi_{C,K}$ contains the version $\mcycle_i(\vec{e})$.
 
@@
 
-\subsection{Balanced version $\mcycle_2(\vec{e})$}
+\subsection{Intermediate version $\mcycle_2(\vec{e})$}
+The intermediate version says that $\Te$ is nonempty and for each state
+$q\in Q$ it holds that $\Te$ contains a transition leading to $q$ if
+and only if it contains a transition leading from $q$.
+Formally,
+\[
+  \mcycle_2(\vec{e})=\bigvee_{t\in\delta}e_t~\wedge~
+  \bigwedge_{q\in Q}\Big(\bigvee_{t'\in\delta~\cap~Q\times\Sigma\times 2^M\times\{q\}}\hspace{-1cm}e_{t'}\hspace{1cm}\iff
+  \bigvee_{t''\in\delta~\cap~\{q\}\times\Sigma\times 2^M\times Q}\hspace{-1cm}e_{t''}\hspace{6mm}\Big).
+\]
+This condition is satisfied by every cycle, but also by some sets of
+transitions that are not cycles. Some examples of such sets are provided in Figure~\ref{fig:inter}.
+
+\begin{figure}[t]
+  \centering
+  \begin{tikzpicture}[smallautomaton]
+    \begin{scope}[initial angle=90,xshift=-5cm]
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (1.8,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10] (p)
+      (p) edge[out=30,in=150] (q)
+      (q) edge[out=30,in=-30,looseness=10] (q)
+      (q) edge[out=-150,in=-30] (p);
+    \end{scope}
+
+    \begin{scope}[initial angle=90,densely dotted,black!20]
+      \node[state,initial,solid,draw=black] (p) at (0,0) {};
+      \node[state,solid,draw=black] (q) at (1.8,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10,black,solid] (p)
+      (p) edge[out=30,in=150,black,solid] (q)
+      (q) edge[out=30,in=-30,looseness=10,black,solid] (q)
+      (q) edge[out=-150,in=-30] (p);
+    \end{scope}
+
+    \begin{scope}[initial angle=90,densely dotted,black!20,xshift=5cm]
+      \node[state,initial,solid,draw=black] (p) at (0,0) {};
+      \node[state,solid,draw=black] (q) at (1.8,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10,black,solid] (p)
+      (p) edge[out=30,in=150] (q)
+      (q) edge[out=30,in=-30,looseness=10,black,solid] (q)
+      (q) edge[out=-150,in=-30] (p);
+    \end{scope}
+  \end{tikzpicture}
+  \caption{An automaton structure where transition labels and
+    acceptance marks are not depicted (left) and two sets $\Te$ that
+    are not cycles even if $\mcycle_2(\vec{e})$ holds (middle and
+    right)}
+  \label{fig:inter}
+\end{figure}
 
 \subsection{Strict version $\mcycle_3(\vec{e})$}
+Before we give a definition of $\mcycle_3(\vec{e})$, we prove that
+cycles can be characterised in the following way.
+\begin{lemma}
+  A set $T\subseteq\delta$ is not a cycle if and only if there exists a set of
+  states $S\subseteq Q$ such that 
+  \begin{enumerate}
+  \item some transition $t\in T$ leads from a state in $S$ to a state
+    outside $S$ and
+  \item there is no transition in $T$ leading from a state outside $S$
+    to a state in $S$.
+  \end{enumerate}
+  \todo{Nebo neprazdna cast $T$ vede z $S$ do $S$ a jina neprazdna
+    cast $T$ vede z doplňu $S$ do doplňku $S$ a žádná hrana to
+    nekrosí.}
+\end{lemma}
+
+\begin{figure}[t]
+  \centering
+  \begin{tikzpicture}[smallautomaton]
+    \begin{scope}[initial angle=90]
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (2,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$1$} (p)
+      (p) edge[out=30,in=150] pic {acc={1}{1}} pic {l=$2$} (q)
+      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={2}{2}} pic[pos=.67] {acc={3}{3}} pic {l=$3$} (q)
+      (q) edge[out=-150,in=-30] pic {acc={3}{3}} pic {l=$4$} (p);
+
+      \node[acclabel] at (1,-1.5)
+      {$(\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{3}{3})$};
+    \end{scope}
+
+% \begin{scope}[initial angle=90]
+% \node[state,initial] (phi) at (0,0) {$\varphi\vphantom{\G}$};
+% \node[state] (Fb) at (1.5,-1.5) {$\F b$};
+% \path[->,auto]
+% (phi) edge[loop left]
+%   pic {acc={3}{1}} pic {l=$\true$} (phi)
+% (phi) edge[loop right]
+%   pic {acc={1}{3}} pic {l=$a$} (phi)
+% (phi) edge[out=-135,in=-105,looseness=10]
+%   pic[pos=.12] {acc={2}{2}} pic[pos=.3,left] {l=$b$} (phi)
+% %(phi) edge[out=-45,in=45] (Fb)
+% (phi) edge[out=-45,in=105]
+%   node[above right,inner sep=.3,pos=.45] {$\bar{b}$} (Fb)
+% (phi) edge[out=-45,in=-75,looseness=10]
+%   pic[pos=.12] {acc={2}{2}} (phi)
+% (Fb) edge[loop right]
+%   pic {acc={3}{1}} pic {l=$\bar{b}$} (phi)
+% (Fb) edge node {$b$} + (0,-1.1)
+% ;
+% \node[acclabel] at (.75,-3.2)
+%   {$\Fin\tacc{3}{1}\wedge(\Fin\tacc{2}{2}\vee\Fin\tacc{1}{3})$};
+% \end{scope}
+
+% \begin{scope}[xshift=4.8cm,initial angle=90]
+% \node[state,initial] (phi) at (0,-.7) {$\varphi\vphantom{\G}$};
+% \path[->,auto]
+% (phi) edge[loop left] 
+%   pic {acc={3}{1}} pic {l=$\true$} (phi)
+% (phi) edge[loop right]
+%   pic {acc={1}{3}} pic {l=$a$} (phi)
+% (phi) edge[out=-135,in=-105,looseness=10]
+%   pic[pos=.75] {accsq={0}{5}} pic[pos=.15] {acc={2}{2}}
+%   pic[left] {l=$b$} (phi)
+% (phi) edge[out=-45,in=-75,looseness=10]
+%   pic[pos=.15] {acc={2}{2}} pic[pos=.75] {acc={4}{4}}
+%   pic[right] {l=$\bar{b}$} (phi)
+% ;
+% \node[acclabel,align=center] at (0,-2.8)
+%   {$\Fin\tacc{3}{1}\wedge{}$\\$(\Fin\tacc{2}{2}\vee\Fin\tacc{1}{3})
+%    \wedge{}$\\$(\Fin\tacc{4}{4}\vee\Inf\taccsq{0}{5})$};
+% \end{scope}
+
+  \end{tikzpicture}
+  \caption{An automaton showing different reduction result for various reduction levels. \todo{Overit a dodelat!}}
+  \label{fig:levels}
+\end{figure}
 

### 2022-03-01 main.py
incremental solving

@@ -45,7 +45,10 @@ def main(argv):
         "-B",
         "--base_level",
         help="Choose whether qbf level 1 or standard level 1(original TELAtko).", choices=['telatko', 'z3'], default='z3')
-
+    parser.add_argument(
+        "-I",
+        "--incremental",
+        help="Use incremental solving.", action='store_true')
     args = parser.parse_args()
 
 
@@ -73,7 +76,7 @@ def main(argv):
                 if args.level > 1 and acc_sets_count != 0:
 
                     auto = play(
-                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut, args.qbf_solver, 2)
+                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut,  2, args.incremental)
                 else:
                     auto = a
             else:
@@ -81,7 +84,7 @@ def main(argv):
                     auto = a
                 else:
                     auto = play(
-                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut, args.qbf_solver, 1)
+                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut, 1, args.incremental)
 
             if args.outfile:
                 print_aut(auto, args.outfile, "a")
@@ -90,7 +93,7 @@ def main(argv):
 
             if not spot.are_equivalent(origin, auto):
                 assert(False)
-                print_aut(origin, "not_eq", "w")
+                
                 return
 
 

### 2022-03-01 telatko2/classes.py

@@ -457,7 +457,7 @@ class ACC_CNF(ACC):
                     marks_on_some_edges,
                     marks__on_all_edges)
                 if val is None:
-                    
+
                     clean_con.append(dis)
 
                 elif val == True:
@@ -532,6 +532,7 @@ def parse_dnf_acc(acc):
             for c in con:
                 if c.isdigit():
                     num.append(c)
+            
             new_dis.append(ACCMark(mtype, int(''.join(num))))
         formula.append(new_dis)
     return formula

### 2022-03-01 paper/main.tex
example showing reduction progress

@@
 \subsection{Strict version $\mcycle_3(\vec{e})$}
-Before we give a definition of $\mcycle_3(\vec{e})$, we prove that
+Before we give the definition of $\mcycle_3(\vec{e})$, we prove that
 cycles can be characterised in the following way.
 \begin{lemma}
-  A set $T\subseteq\delta$ is not a cycle if and only if there exists a set of
-  states $S\subseteq Q$ such that 
-  \begin{enumerate}
-  \item some transition $t\in T$ leads from a state in $S$ to a state
-    outside $S$ and
-  \item there is no transition in $T$ leading from a state outside $S$
-    to a state in $S$.
-  \end{enumerate}
-  \todo{Nebo neprazdna cast $T$ vede z $S$ do $S$ a jina neprazdna
-    cast $T$ vede z doplňu $S$ do doplňku $S$ a žádná hrana to
-    nekrosí.}
+  A set $T\subseteq\delta$ is a cycle if for each set of
+  states $S\subseteq Q$ it holds that
+  \[
+    \begin{split}
+    \textrm{if $T$ contains a transition leading from a state in $S$ to a state outside $S$,~~~~}\\[-.5ex]
+    \textrm{then it also contains a transition leading from a state outside $S$ to a state in $S$.}
+    \end{split}
+    \tag{$3$}\label{eq:lem3} 
+  \]  
 \end{lemma}
-
-\begin{figure}[t]
-  \centering
-  \begin{tikzpicture}[smallautomaton]
-    \begin{scope}[initial angle=90]
-      \node[state,initial] (p) at (0,0) {};
-      \node[state] (q) at (2,0) {};
-      \path[->,auto]
-      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$1$} (p)
-      (p) edge[out=30,in=150] pic {acc={1}{1}} pic {l=$2$} (q)
-      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={2}{2}} pic[pos=.67] {acc={3}{3}} pic {l=$3$} (q)
-      (q) edge[out=-150,in=-30] pic {acc={3}{3}} pic {l=$4$} (p);
-
-      \node[acclabel] at (1,-1.5)
-      {$(\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{3}{3})$};
-    \end{scope}
-
-% \begin{scope}[initial angle=90]
-% \node[state,initial] (phi) at (0,0) {$\varphi\vphantom{\G}$};
-% \node[state] (Fb) at (1.5,-1.5) {$\F b$};
-% \path[->,auto]
-% (phi) edge[loop left]
-%   pic {acc={3}{1}} pic {l=$\true$} (phi)
-% (phi) edge[loop right]
-%   pic {acc={1}{3}} pic {l=$a$} (phi)
-% (phi) edge[out=-135,in=-105,looseness=10]
-%   pic[pos=.12] {acc={2}{2}} pic[pos=.3,left] {l=$b$} (phi)
-% %(phi) edge[out=-45,in=45] (Fb)
-% (phi) edge[out=-45,in=105]
-%   node[above right,inner sep=.3,pos=.45] {$\bar{b}$} (Fb)
-% (phi) edge[out=-45,in=-75,looseness=10]
-%   pic[pos=.12] {acc={2}{2}} (phi)
-% (Fb) edge[loop right]
-%   pic {acc={3}{1}} pic {l=$\bar{b}$} (phi)
-% (Fb) edge node {$b$} + (0,-1.1)
-% ;
-% \node[acclabel] at (.75,-3.2)
-%   {$\Fin\tacc{3}{1}\wedge(\Fin\tacc{2}{2}\vee\Fin\tacc{1}{3})$};
-% \end{scope}
-
-% \begin{scope}[xshift=4.8cm,initial angle=90]
-% \node[state,initial] (phi) at (0,-.7) {$\varphi\vphantom{\G}$};
-% \path[->,auto]
-% (phi) edge[loop left] 
-%   pic {acc={3}{1}} pic {l=$\true$} (phi)
-% (phi) edge[loop right]
-%   pic {acc={1}{3}} pic {l=$a$} (phi)
-% (phi) edge[out=-135,in=-105,looseness=10]
-%   pic[pos=.75] {accsq={0}{5}} pic[pos=.15] {acc={2}{2}}
-%   pic[left] {l=$b$} (phi)
-% (phi) edge[out=-45,in=-75,looseness=10]
-%   pic[pos=.15] {acc={2}{2}} pic[pos=.75] {acc={4}{4}}
-%   pic[right] {l=$\bar{b}$} (phi)
-% ;
-% \node[acclabel,align=center] at (0,-2.8)
-%   {$\Fin\tacc{3}{1}\wedge{}$\\$(\Fin\tacc{2}{2}\vee\Fin\tacc{1}{3})
-%    \wedge{}$\\$(\Fin\tacc{4}{4}\vee\Inf\taccsq{0}{5})$};
-% \end{scope}
-
-  \end{tikzpicture}
-  \caption{An automaton showing different reduction result for various reduction levels. \todo{Overit a dodelat!}}
-  \label{fig:levels}
-\end{figure}
+\begin{proof}
+  \todo{Ted to musim dokazat.}
+\end{proof}
 
@@
 
-\todo{Zbytek je stary text!}
-
-Recall that our goal is to reduce the number of acceptance marks in a
-given TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ without altering its
-structure and language. Hence, our reduction replaces the set $M$ of
-marks by a smaller set, the acceptance formula $\varphi$ by a new one,
-and changes the sets of acceptance marks on individual transitions.
-
-Let us assume that we are given a TELA $\mA$ and two integer
-parameters $C,K\ge 0$, where $K$ is the desired number of acceptance
-marks after the reduction and $C$ is the desired number of clauses in
-the new acceptance formula in DNF. This section describes the
-construction of a QBF formula for $\mA,C,K$ such that satisfiability
-of this formula implies existence of the reduced automata with
-
-This section explains the construction of QBF formulas for a given3TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ and two integer parameters
-$C,K\ge 0$ such that satisfiability of such a formula implies that
-there exists an acceptance formula in DNF with $C$ clauses and $K$
-acceptance marks and a placement of these marks on automata
-transitions such that $\mA$ with this acceptance formula and
-acceptance marks and acceptance formula
-
-such that their model describes an acceptance formula in DNF
-with $C$ clauses and $K$ acceptance marks and placement of these marks
-on automata transitions such that
-
-The aim of this section is to create a QBF-formula for given $n$ that guarantees the existence of equivalent acceptance condition of a given automaton with $n$ acceptance marks. 
-\smallskip
-
-Let $\mathcal{A}$ with the acceptance condition $\psi$ be an original automaton on the input.
-Let $\mathcal{A}'$ denote automaton with simplified acceptance condition $\psi'$ equivalent to the original automaton $\mathcal{A}$.
-We encode $\psi$ and the placement of acceptance marks into QBF  $\varphi$.
-
-Let $\varphi$ be a QBF formula in \emph{prenex normal form}. Formula $\varphi$
-is composed of quantified prefix $\overrightarrow{Q}$ and three subformulae $\varphi_{O}, \varphi_{N} $ and $\varphi_{C}$.
-The exact shape of the formula is $\varphi = \overrightarrow{Q} .
-\  \varphi_{C} \implies ( \varphi_{O} \Longleftrightarrow \varphi_{N})$.
-Subformula $\varphi_{O}$ expresses properties of $\psi$.
-Subformula $\varphi_{N}$ expresses properties of $\psi'$, namely the number of acceptance marks and a shape of $\psi'$.
-Before we describe the role of $\varphi_{C}$, we need to introduce \emph{relevant set}.
-
-\smallskip
-
-Evaluation of a run of $\mathcal{A}$ is dependent on set of edges that are contained infinitely within the run.
-The set of these edges is from now on referred to as \emph{relevant set}, since only these edges are relevant for the evaluation of $\psi$.
-The \emph{relevant set} is a subset of $\delta_{S_{i}}$ for some SCC $S_{i}$ and any two edges from the \emph{relevant set} are connected by a sequence of edges that also belong to the \emph{relevant set}.
-The purpose of subformula $\varphi_{C}$ is to describe some properties of \emph{relevant sets}.
-For each level of simplification, $\varphi_{C}$ describes these properties more specifically.
-Naturally, the more we specify these properties, the more complex the formula $\varphi_{C}$ is.
-
-\smallskip
-
-Afterwards the formula $\varphi$ is created, we use a QBF-solver to find a model of $\varphi$.
-If no such model exists ($\varphi$ is UNSAT), we know that automaton $\mathcal{A}'$ with $n$ acceptance marks and equivalent to $\mathcal{A}$ does not exist.
-Otherwise, we use the model of $\varphi$ for decoding the automaton $\mathcal{A}'$.
-The structure of $\mathcal{A}'$ remains the same as the structure of $\mathcal{A}$.
-The only thing we change is the acceptance condition $\psi'$ and the placement of acceptance marks on the edges.
-
-\subsection{Motivation}
-Based on the fundamentals given in the previous section, the simplification Level 1 is based on a relationship between acceptance sets of an automaton.
-To achieve finer simplification levels, the relation between the inner structure of the automaton and its acceptance condition needs to be taken into consideration.
-
-As the acceptance condition and the inner structure differ for every specific automaton, finding algorithmic ways of reducing acceptance marks would be complicated.
-Instead, we encode the given automaton into a QBF-formula and let the solver find the redundant marks for us.
-
-\subsection{Encoding}
- In this section we propose encoding of the relevant parts of a given automaton into QBF-formula.
-\subsubsection{Intuition}
-The goal of the procedure is to find acceptance condition for a particular automaton with less acceptance marks.
-Whilst doing that, we need to ensure that the new automaton stays equivalent to the original automaton because the algorithm not only changes the acceptance condition it also changes the placement of acceptance marks on the edges.
-To do so, we need to ensure that every run on $\mathcal{A}'$ is evaluated the same way as run on $\mathcal{A}$ (these two automata have the same structure).
-
-Since we work with transition-based automata, the evaluation of an acceptance condition is dependent on the edges.
-Intuitively we want formula $\varphi$ to say that for every \emph{relevant set}, the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
-Unfortunately, to identify the exact edges that belong to the \emph{relevant set} of a run is not trivial, since we need to take into consideration every infinite suffix of every run on $\mathcal{A}$.
-Instead we let the formula $\varphi$ say that for every \emph{selection of edges} the acceptance condition $\psi'$ is evaluated as $\true$ $\iff$ $\psi$ is evaluated as $\true$.
-For now the \emph{selection of edges} can contain any edges of the automaton.
-Later we restrict the \emph{selection of edges} as we force it to have some properties of \emph{relevant set}.
-
-To express this in the QBF formula, we need to introduce variables that represent edges of the automaton.
-
-Let us enumerate set of edges $\delta$.
-Every edge of the automaton is represented by variable $e_{i}$, where the index corresponds with the enumeration, thus $e_{i}$ represents the edge $t_{i} \in \delta$.
-If such a variable is evaluated as $\true$, it means that the edge is contained in the \emph{selection of edges}.
-Analogously if a variable $e_{i}$ is evaluated as $\false$, it means that the edge is not contained in the \emph{selection of edges}.
-
-\smallskip
-
-If we want to consider every possible \emph{selection of edges}, we need to consider every possible evaluation of every variable that represents an edge.
-
-\todo{tohle bych odstranila}
-\begin{example}
-   Let $t_{1}, t_{2} \in \delta$ be the edges of the automaton from the Figure \ref{fig:elval_of_edges}.
-Since the automaton has two edges, we get four possible evaluation of variables representing the edges.
-    \begin{enumerate}
-        \item $e_{1}$ is $\true \land e_{2}$ is $\true \Longleftrightarrow t_{1}, t_{2}$ are in the \emph{selection of edges}.
-        \item $e_{1}$ is $\true \land e_{2}$ is $\false \Longleftrightarrow t_{1}$ is in the \emph{selection of edges} and $t_{2}$ is not in the \emph{selection of edges}.
-        \item $e_{1}$ is $\false \land e_{2}$ is $\true \Longleftrightarrow t_{1}$ is not in the \emph{selection of edges} and $t_{2}$ is in the \emph{selection of edges}.
-        \item $e_{1}$ is $\false \land e_{2}$ is $\false \Longleftrightarrow t_{1}, t_{2}$ are not in the \emph{selection of edges}.
-    \end{enumerate}
-\end{example}
-
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-			\node[state, initial] (z) at (0,0) {};
-
-			\path[->]
-
-            	(z) edge[loop, in = 70, out = 110, looseness=6] node [above, scale=0.7] {1} (z)
-            	(z) edge[loop, in = 68, out = 112, looseness=10] node [above, scale=0.7] {2} (z)
-			;
-		\end{tikzpicture}
-    \caption{TELA with enumerated edges}
-    \label{fig:elval_of_edges}
-\end{figure}
-
-Hereby the expressive power of quantified Boolean formula comes to use.
-If we universally quantify every variable $e_{i}$ that represents an edge $t_{i} \in \delta$, then we consider the situation when $t_{i}$ is contained in the \emph{selection of edges} and the situation when $t_{i}$ is not contained in the \emph{selection of edges}.
-
-That being said, we can proceed to the encoding of some information about automaton into the QBF-formula $\varphi$.
-As mentioned before, the formula $\varphi$ expresses that for every \emph{selection of edges}, the acceptance condition $\psi$ is evaluated as $\true$ $\mathit{iff}$ the acceptance condition $\psi'$ is evaluated as true.
-
-The subformula $\varphi_{O}$ encodes the shape of the acceptance condition $\psi$ and information which edges are labeled with which acceptance marks.
-The evaluation of $\varphi_{O}$ depends purely on the evaluation of variables that represent edges of the automaton.
-It is encoded in such way that the subformula $\varphi_{O}$ is satisfied iff the edges of \emph{selection of edges} satisfy $\psi$.
-The details of $\varphi_{O}$ encoding are described in Subsection \ref{subsection:original}.
-
-\smallskip
-
-The subformula $\varphi_{N}$ contains information about the number of acceptance marks we want in the new acceptance formula $\psi'$.
-Further, it contains free variables.
-The model of these free variables gives us the information about the shape of $\psi'$ and the placement of acceptance marks on the edges of $\mathcal{A}'$.
-The details of $\varphi_{N}$ encoding are described in Subsection \ref{subsection:new}.
-
-The quantification of all variables that represent the edges of the automaton ensures that we consider every possible \emph{selection of edges}, as described in Section \ref{subsection:quant_pref}. 
-\begin{center}
-    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . ( \varphi_{O} \Longleftrightarrow \varphi_{N} ) $
-\end{center}
-However, one can clearly notice that some edges will never be visited infinitely often in the same run.
-(For example, when some of the edges occur in different strongly connected components than the others.) In other words, these edges will never belong to any \emph{relevant set} together.
-Therefore it is possible to eliminate every evaluation of QBF, where variables representing such edges are evaluated as $\true$.
-It is natural to express this constrain as an implication.
-Let $\varphi_{C}$ be a formula describing that the \emph{selection of edges} has a potential be a \emph{relevant set}.
-The QBF representation is:
-\begin{center}
-    $\forall \{ e_{1}, \ldots, e_{|\delta|}\} . \varphi_{C} \implies (\varphi_{O} \Longleftrightarrow \varphi_{N})$
-\end{center}
-The matter of eliminating various combinations of edges is described in more detail in Subection \ref{subsection:connection}.
-
-\subsubsection{Variables}
-In this section, we describe the syntax and intended semantics of variables, parameters, and auxiliary letters used in this section.
-Hereby is the overview of the parameters used when building the formula:
-\begin{itemize}
-    \item K -- The number of acceptance marks in $\psi'$.
-    \item C -- the number of clauses in formula $\psi'$.
-\end{itemize}
-
-These parameters are derived from the original automaton $\mathcal{A}$. As the algorithm iterates (see Subection \ref{subsection:procedure} )  the parameter $K$ decreases.
-
-\smallskip
-
-The auxiliary letters have the following function:
-\begin{itemize}
-    \item $k$ -- the number of particular acceptance mark
-    \item $c$ -- the number of particular clause
-    \item $i$ -- the number of particular edge 
-\end{itemize}
-
-Hereby is the overview of the variables used when building the QBF formula $\varphi$ :
-\begin{itemize}
-    \item $$f_{i,k} = \begin{cases}
-    1& \text{edge }t_{i} \text{ is labeled with the acceptance mark }k\\
-    0& \text{else}
-\end{cases}$$
-\item $$p_{c,k} = \begin{cases}
-    1& \Inf(k) \text{ is in the clause }c\\
-    0& \text{else}
-\end{cases}$$
-\item $$n_{c,k}= \begin{cases}
-    1& \Fin(k) \text{ is in the clause }c\\
-    0& \text{else}
-\end{cases}$$
-\item $$e_{i} \;= \begin{cases}
-    1& \text{The edge } t_{i} \text{ is in the \emph{selection of edges} }\\
-    0& \text{else}
-\end{cases}$$
-\end{itemize}
-
-\subsection{QBF decoding}
-After we build $\varphi$, we use QBF-solver to find satisfying assignment of variables.
-If such a valuation exists, we can build a new automaton $\mathcal{A}'$ with the same structure and fewer acceptance marks.
-This new automaton is equivalent to the original one $\mathcal{A}$ -- this is guaranteed by the QBF-formula $\varphi$.
-There are two possible outcomes from the QBF-solver:
-\begin{enumerate}
-    \item $\varphi$ is UNSAT -- a model of $\varphi$ does not exist.
-That means that the acceptance condition of $\mathcal{A}$ cannot be simplified for given C and K.
-Thus we return automaton $\mathcal{A}$ and the information that $\mathcal{A}$ is the automaton with the simplest acceptance condition possible.
-    \item $\varphi$ is SAT -- QBF-solver provides (returns) a model of unquantified variables.
-Namely to $p_{c, k},n_{c,k} \text{ and } f_{i,k}, \text{ for each }1 \leq c \leq C, 1 \leq k \leq K, 1 \leq i \leq |\delta|$.
-    Firstly, we decode the variables $f_{i,k}$ and use them to identify new placement of the acceptance marks.
-Specifically, we create a new automaton $\mathcal{A}'$ with the same structure as $\mathcal{A}$.
-Then  we add acceptance mark $\tacc{5}{k}$ on edge $t_{i}$ for every $f_{i,k}$ variable evaluated as $\true$.
-
-    The new acceptance condition of automaton $\mathcal{A}'$ is then build from variables $p_{c,k}$ and $n_{c,k}$.
-The second part of $\varphi_{N}$ was encoded to produce new acceptance formula in DNF.
-(See the alternation between Boolean operators  in order $\lor, \land$ in Section \ref{subsection:new}.) Thus the shape of the new acceptance condition is set as DNF skeleton, and we simply add terms according to the evaluation of variables.
-If $p_{c,k}$ is $\true$, we add term $\Inf\tacc{5}{k}$ to the disjunct $c$.
-Similarly, if $n_{c,k}$ is $\true$, then we add $\Fin\tacc{5}{k}$ to the disjunct $c$.
-This way, we process every variable that QBF-solver gives us and produce the new automaton $\mathcal{A}'$
-    equivalent to the original one.
-
-\end{enumerate}
-
-\subsection{Quantified prefix}
-\label{subsection:quant_pref}
-The aim of the quantified prefix in formula $\varphi$ is to consider every possible selection of edges.
-As mentioned in the previous sections, edge $t_{i} \in \delta$ belongs to the \emph{selection of edges} if the variable $e_{i}$ that represents the edge $t_{i}$ is evaluated as $\true$.
-Therefore if we want to consider every possible \emph{selection of edges}, we need to consider every evaluation of the variables representing edges.
-We use a universal quantifier to express this.
-The universal quantifier states that the formula has to hold for both evaluations of the quantified variable.
-
-The QBF prefix that expresses the universal quantification for every variable representing an edge has the following shape:
-\begin{align*}
-    \overrightarrow{Q} = \forall\{e_{1}, \ldots ,e_{|\delta|} \}
-\end{align*}
-
-The advantage of QBF is that it is able to express a statement in shorter formula then if we tried to express the statement in standard Boolean formula with the use of conjunction.
-
-\begin{example}
-Consider the QBF $\theta = \forall e_{1} \forall e_{2} .
-(e_1 \lor e_2) $.
-The QBF $\theta $ transformed to standard Boolean formula is $\theta' = (e_{1} \lor e_{2}) \land (  e_{1} \lor \lnot e_{2}) \land  (  \lnot e_{1} \lor e_{2}) \land  (  \lnot e_{1} \lor \lnot e_{2}) .$
-\end{example}
-
-\subsection{Original acceptance formula}
-\label{subsection:original}
-
-As mentioned before, the satisfiability of $\varphi_{O}$ is purely resultant of assessment of $e_{i}, 1 \leq i \leq |\delta|$ variables representing edges in automaton.
-
-Let us enumerate the edges of the original automaton $\mathcal{A}$ and let $Z(m), m \in M$ be the set of edges labeled with mark $m$. 
-The encoding of original acceptance condition terms into $\varphi_{O}$ is defined as follows:
-%%davat tam rovnitko nebo nejak sipku??
-\begin{align*}
-    &\varphi(\Inf(m))=\bigvee\limits_{t_{i} \in Z(m)} e_{i}\\
-    &\varphi(\Fin(m)) = \bigwedge\limits_{t_{i} \in Z(m)} \lnot e_{i}
-\end{align*}
-\begin{example}
-
-Consider the automaton in Figure \ref{fig:oficial_acc}.
-The set $Z_{0} = \{t_3, t_2\},\, Z_{1} = \{t_1\}$, where $t_{1}, t_{2}, t_{3} \in \delta$.
-Thus the encoded $\varphi_{O} = e_1 \land (\lnot e_2 \land \lnot e_3)$.
-Therefore if variable $e_{1}=\true$, $e_{2}=\false$ and $e_{3}=\false$, acceptance formula is $\true$.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-\node[cstate, initial] (g) {};
-\node[flabel, above] at (0, 1.5) {$\Fin\taccsq{0}{0}\land\Inf\tacc{1}{1}$};
-%
-\path[->]
-    (g) edge[loop, in=330, out=30, looseness=6]node[right, pos=0.4, scale=0.8] {$1$} node[accset,collacc1, pos=0.6] {1}(g)
-	(g) edge[loop, in=80, out=140, looseness=6] node[above, pos= 0.4, scale=0.8] {$2$} node[accsquare, collacc0, pos=0.6]{0}(g)
-	(g) edge[loop, in=220, out=280, looseness=6] node[below, pos=0.6, scale=0.8] {$3$} node[accsquare, collacc0, pos=0.4] {0} (g)
-	;
-\end{tikzpicture}
-    \caption{TELA with edges labeled with their numbers}
-    \label{fig:oficial_acc}
-\end{figure}
-
-\end{example}
-
-\subsection{New acceptance formula}
-\label{subsection:new}
-The new acceptance formula $\varphi_{N}$ determines the shape of the future acceptance condition of automaton, and the assignment of variables used in $\varphi_{N}$ determines new positions of acceptance marks on the transitions.
-
-The formula $\varphi_{N}$ is divided into two parts.
-The first part of the formula (4.1) expresses a simple constrain that the acceptance set that occurs in one clause cannot be in $\Inf$ and simultaneously in $\Fin$ form.
-
-The second part (4.2) of $\varphi_{N}$ is divided into two subformulas.
-The first subformula expresses that if the term $\Inf(k)$ is present in clause $c$, then at least one edge of the \emph{selection of edges} needs to be labeled with the acceptance mark $k$.
-
-Similarly, the second subformula implies that if the term $\Fin(k)$ is present in clause $c$, then none of the \emph{selected edges} can be labeled with the acceptance mark $k$.
-
-Using the variables presented in previous sections, we encode $\varphi_{N}$:
-\begin{align}
-    \varphi_{N} = & \bigwedge\limits_{1 \leq c \leq C}\;\bigwedge\limits_{1 \leq k \leq K} (\lnot p_{c,k} \lor \lnot n_{c,k}) \; \land \\
-    &\bigvee\limits_{1 \leq c \leq C} \; \bigwedge\limits_{1 \leq k \leq K} \Bigg[\big(  p_{c, k} \implies \bigvee\limits_{t_{i} \in \delta}(f_{i,k} \land e_{i})  \big) \land \\& \;\;\;\quad\quad\quad\quad\quad\quad\big(  n_{c, k} \implies \bigwedge\limits_{t_{i} \in \delta}(\lnot f_{i,k} \lor \lnot e_{i})  \big)\Bigg]
-\end{align}
-
-Observe the alternation of $\lor \land$ from the second part of the formula (4.2). This alternation determines that the formula $\psi'$ is in DNF.
-
-\subsection{Connection formula}
-\label{subsection:connection}
-The connection formula distinguishes the remaining two levels of simplification. Previously we denoted it as $\varphi_{C}$.
-
-Formula $\varphi_{C}$ defines requirements on the edges that are together in \emph{selection of edges}.
-These requirements are based on some properties of these edges.
-The more restrictive these requirements are, the more properties of \emph{relevant sets} are possessed by the \emph{selection of edges}.
-In other words, formula $\varphi_{C}$ holds if the \emph{selection of edges} has the required property.
-If not, the formula $\varphi_{C}$ is $\false$, and then the whole formula $\varphi$ is trivially satisfied.
-Formula $\varphi_{C}$ consists of two parts $\varphi_{C_{1}}$ and $\varphi_{C_{2}}$.
-
-The first part expresses the requirement that the \emph{selection of edges} is \emph{continuous}.
-Loosely said, this means that if a transition from \emph{selection of edges} enters some state $q$ then another transition that belongs to the \emph{selection of edges} exits the state $q$.
-This holds in the opposite direction as well.
-Thus if a transition from \emph{selection of edges} exits some state $q$ then another transition that belongs to the \emph{selection of edges} enters the state $q$.
-This is naturally expressed by equivalence.
-Note that if the selected transition is a self-loop, the destination and the source of the edge is the same state.
-
-The Boolean formula for the described statement:
-\begin{center}
-    $\varphi_{C_{1}}=\bigwedge\limits_{q \in Q} \Bigg( ( \bigvee\limits_{t_{i} \in \delta \cap \{ q \} \times  \Sigma \times  2^M \times Q} e_i ) \Longleftrightarrow( \bigvee\limits_{t_{j} \in \delta  \cap \ Q \times  \Sigma \times 2^M \times \{q\}} e_{j} ) \Bigg)$
-\end{center}
-
-\begin{example}
-   Consider the state $q$ in Figure \ref{fig:inNout} this part of formula is $(e_1 \lor e_2) \Longleftrightarrow (e_3 \lor e_4 \lor e_5)$.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-        \node[state] (q) at (0, 0) {$q$};
-
-	    \path[->]
-		(q) edge node[above, pos=0.4, scale=0.8] {$4$} ++(1.5, 0)
-		(q) edge node[above, pos=0.4, scale=0.8] {$5$} ++(1.3, 1)
-		(q) edge node[above, pos=0.4, scale=0.8] {$3$} ++(1.3, -1)
-	;
-	\path[<-]
-	(q) edge node[above, pos=0.6, scale=0.8] {$2$} ++(-1.3, -0.5)
-	(q) edge node[above, pos=0.6, scale=0.8] {$1$} ++(-1.3, 0.5)
-	;
-    \end{tikzpicture}
-    \caption{State $q$ and its entering and exiting edges}
-    \label{fig:inNout}
-\end{figure}
-\end{example}
-
-Even though we ensured that the selected transitions create continuous paths, the paths might not be connected.
-
-\begin{example}
-Consider the automaton in Figure \ref{fig:continuousNotConnected}.
-The transitions $t_1$ and $t_2$ are loops (the simplest cycles), but are not connected (each is located in different SCC).
-Thus $\{ t_{1}, t_{2}\}$ is not a \emph{relevant set}.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-			\node[cstate, initial] (i) at (0, 0) {};
-			\node[cstate] (j) at (2.5, 1) {};
-			\node[cstate] (k) at (2.5, -1) {};
-			\path[->]
-
-			(i) edge  (k)
-			(i) edge (j)
-			(j) edge[loop right, draw=red] node[right] {$1$} (j)
-			(k) edge[loop right, draw=red] node[right] {$2$} (k)
-
-		;
-    \end{tikzpicture}
-    \caption{Not connected cycles}
-    \label{fig:continuousNotConnected}
-\end{figure}
-\end{example}
-
-The second part of $\varphi_{C}$ specifies the properties of \emph{relevant set} that the edges that belong together to \emph{selected edges} need to have.
-We divided these properties into two levels Level 2 and Level 3, all described in ensuing sections.
-
-\subsubsection{SCC accuracy}
-The SCC accuracy corresponds with the Level of simplification 2. At this level, the formula specifies that the edges from one \emph{selection of edges} belong to the same SCC.
- Thus straightforwardly translated into QBF formula:
-\begin{center}
-    $\varphi_{C_{2}} = \bigvee\limits_{\delta_{S_{i}} \subseteq \delta}\Bigg( ( \bigvee\limits_{t_{i} \in \delta_{S_{i}}} e_i ) \land  ( \bigwedge\limits_{t_{j} \in \delta \setminus \delta_{S_{i}}} \lnot e_j  ) \Bigg)$
-\end{center}
-
-Finally, the connection formula is the conjunction of the two parts:
-\begin{center}
-    $\varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}$.
-\end{center}
-
-\begin{example}
-This example works as a motivation for simplification Level 3.
-For simplification Level 2, edges number $1$ and number $3$ in Figure \ref{fig:connected_motivation} can be in the same \emph{selection of edges} because they are placed in the same SCC.
-Obviously, since they are not connected, they cannot form \emph{relevant set}.
-Therefore we can specify a new level of simplification -- Level 3, which forces all edges from the \emph{selection of edges} to be cycles.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-    \node[state, initial] (g) {$0$};
-    \node[state, right=2cm of g](q){$1$};
-    %
-    \path[->, auto]
-    (g) edge[bend right] node[below, scale=0.8] {$2$}(q)
-    (q) edge[bend right]node[above, scale=0.8] {$4$} (g)
-    (q) edge[loop, in=65, out=115, looseness=10, draw=red] node[ scale=0.8] {$3$}(q)
-    (g) edge[loop, in=65, out=115, looseness=10, draw=red] node[scale=0.8] {$1$}(g)
-	;
-
-\end{tikzpicture}
-
-    \caption{Motivation for simplification L3}
-    \label{fig:connected_motivation}
-\end{figure}
-
-\end{example}
-
-\subsubsection{CYCLES accuracy}
-The simplification Level 2 ensures that \emph{selection of edges} is placed in one SCC, but we can go further and specify that \emph{selection of edges} is in fact a \emph{relevant set}.
-
-The intuition by simplification Level 3 is that the states of an automaton are  divided into two groups -- the states that belong to the group of \emph{selected states} and the states do not belong to the group of \emph{selected states}.
-Then we require that the states that belong to the same group are connected.
-We consider three possible scenarios that can occur while the states are divided into these two groups:
-\begin{enumerate}
-    \item The source of edge $t_{i}$ is in different group than the destination of edge $t_{i}$.
-    \item Source and destination of edge $t_{i}$ are in the \emph{selected states}.
-    \item Source and destination of edge $t_{i}$ are not in \emph{selected states}.
-\end{enumerate}
-Cases two and three do not violate the requirement that all connected states need to be in one group.
-Case one is a little more complicated.
-If an edge has a source in \emph{selected states} and destination not in the \emph{selected states}, we need to secure that the sequence of the edges returns to \emph{selected states}.
-In other words, there exists another edge that has a source in the not \emph{selected states} and destination in the \emph{selected states}.
-(Dually the other way around.)
-
-In order to express this in the QBF-formula we introduce a new set of quantified variables $\{ w_{1}, \ldots , w_{|Q|}$. The variables have the following semantics:
-\begin{align*}
-    w_{q} = \begin{cases}
-    1& \text{state } q \in Q \text{ is in the \emph{selected group}}\\
-    0& \text{else}
-    \end{cases}
-\end{align*}
-Since we want to ensure that for every \emph{selection of edges} (for every evaluation of variables that represent the edges) there exist such a groups we update the quantified prefix  $\overrightarrow{Q}$ in following way:
-\begin{align*}
-    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
-\end{align*}
-
-\smallskip
-\todo{neoptimalni}
-Since we want the statement to hold for every \emph{selection of edges}, we place the existentially quantified variables $\{w_{1}, \ldots, w_{|Q|} \}$ behind the universally quantified variables $\{e_{1}, \ldots, e_{|\delta|} \} $.
-At first we update the quantified prefix:
-\begin{align*}
-    \overrightarrow{Q} = \forall \{ e_{1}, \ldots,  e_{|\delta|}\} \exists \{ w_{1}, \ldots, w_{|Q|}\}
-\end{align*}
-
-And then we can create the $\varphi_{C}$ formula:
-\begin{align*}
-    \varphi_{C} =  \Bigg[ &\bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (w_{q} \land w_{r}) \Bigg) \lor \\
-    &  \bigwedge\limits_{\substack{t_{i} \in \{q\}  \times \Sigma  \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} \Bigg( e_{i} \implies (\lnot w_{q} \land \lnot w_{r}) \Bigg) \lor \\ \Bigg( &\bigvee \limits_{\substack{t_{i} \in \{q\} \times \Sigma \times 2^M  \times \{ r \}, \\ q \in Q, r \in Q}} ( e_{i} \land w_{q} \land \lnot w_{r} )  \land \bigvee \limits_{\substack{t_{i} \in \{q\}  \times \Sigma \times 2^M \times \{ r \}, \\ q \in Q, r \in Q}} (e_{i} \land \lnot  w_{q} \land  w_{r}) \Bigg) \Bigg]
-\end{align*}
-
-\begin{example}
-Figure \ref{fig:cycles_sensitivity} displays an example of an automaton reduced by Level 4.
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-
-    \node[state, initial] (g) {$0$};
-
-    \node[flabel] at (1.5, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{1}{1} \land \Inf\tacc{2}{2}$};
-    \node[state, right=2cm of g](q){$1$};
-
-    \path[->, auto]
-    (g) edge[bend right] node [accset,collacc0, pos=0.5]{0}node [below, pos=0.2, scale=0.8] {$2$}(q)
-    (q) edge[bend right] node[above, pos=0.8, scale=0.8] {$4$}(g)
-    (q) edge[loop, in=75, out=105, looseness=8]node[left, pos=0.1, scale=0.8]{5} (q)
-    (q) edge[loop, in=70, out=110, looseness=12] node[accset,collacc1] {1}node[left, pos=0.2, scale=0.8]{6} (q)
-    (g) edge[loop, in=75, out=105, looseness=8] node[accset,collacc2] {2} node[left, pos=0.1, scale=0.8]{1}(g)
-    (g) edge[loop, in=70, out=110, looseness=12] node[accset,collacc2, pos=0.4] {2} node[accset,collacc1, pos=0.7] {1} node[left, pos=0.2, scale=0.8]{3}(g)
-	;
-
-	\node[state, initial, right=4cm of g] (o){$0$};
-
-    \node[flabel] at (6.7, 2.5){$ \Fin\tacc{0}{0} \land \Inf\tacc{3}{3}$};
-    \node[state, right=2cm of o](p){$1$};
-
-    \path[->, auto]
-    (o) edge[bend right] node [accset,collacc0, pos=0.5]{0}(p)
-    (p) edge[bend right](o)
-    (p) edge[loop, in=75, out=105, looseness=8] (p)
-    (p) edge[loop, in=70, out=110, looseness=12] (p)
-    (o) edge[loop, in=75, out=105, looseness=8] (o)
-    (o) edge[loop, in=70, out=110, looseness=12] node[accset,collacc3] {3}(o)
-
-	;
-
-\end{tikzpicture}
-
-    \caption{TELA reduced with the CYCLES sensitivity}
-    \label{fig:cycles_sensitivity}
-\end{figure}
-\end{example}
-
-Similarly as in the previous section the connection formula is the conjunction of the two parts:
-\begin{align*}
-    \varphi_{C} = \varphi_{C_{1}} \land \varphi_{C_{2}}.
-\end{align*}
-
-\subsection{Simplification procedure}
-\label{subsection:procedure}
-
+%}}}
+%{{{ algorithm and implementation
+
+\section{Reduction algorithm and its implementation}\label{sec:algimpl}
+Jak se nastavuji inicialni parametry a jak postupne klesaji.
+Jak prechazim od nizsich levelu k vyssim.
+Vyuzivame incremental QBF solving?
+
+\todo{Nasleduje stary text, ale pozor, obrazek je novy!}
 Now we describe our simplification algorithm taking as input an arbitrary TELA $\mathcal{A}$ with an acceptance condition $\psi$, required level of simplification $L$ and timeout period $T$.
@@
 
-%}}}
-%{{{ algorithm and implementation
-
-\section{Reduction algorithm and its implementation}\label{sec:algimpl}
-Jak se nastavuji inicialni parametry a jak postupne klesaji.
-Jak prechazim od nizsich levelu k vyssim.
-Vyuzivame incremental QBF solving?
+\begin{figure}[t]
+  \centering
+  \begin{tikzpicture}[smallautomaton]
+    \begin{scope}[initial angle=90,xshift=-3.5cm]
+      \node (l) at (1,1.3) {input:};
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (2,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
+      (p) edge[out=30,in=150] pic {acc={1}{1}} pic {l=$b$} (q)
+      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={2}{2}} pic[pos=.67] {acc={3}{3}} pic {l=$c$} (q)
+      (q) edge[out=-150,in=-30] pic {acc={3}{3}} pic {l=$d$} (p);
+
+      \node[acclabel] at (1,-1.4)
+      {\small$(\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{3}{3})$};
+    \end{scope}
+    \begin{scope}[initial angle=90,xshift=3.5cm]
+      \node (l) at (1,1.3) {level 1:};
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (2,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
+      (p) edge[out=30,in=150] pic {acc={2}{2}} pic {l=$b$} (q)
+      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={0}{0}} pic[pos=.67] {acc={2}{2}} pic {l=$c$} (q)
+      (q) edge[out=-150,in=-30] pic {acc={1}{1}} pic {l=$d$} (p);
+
+      \node[acclabel] at (1,-1.4)
+      {\small$(\Fin\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Fin\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})$};
+    \end{scope}
+    \begin{scope}[initial angle=90,xshift=-3.5cm,yshift=-4cm]
+      \node (l) at (1,1.3) {level 2:};
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (2,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10] pic {l=$a$} (p)
+      (p) edge[out=30,in=150] pic {acc={0}{0}} pic {l=$b$} (q)
+      (q) edge[out=30,in=-30,looseness=10] pic {acc={0}{0}} pic {l=$c$} (q)
+      (q) edge[out=-150,in=-30] pic {acc={1}{1}} pic {l=$d$} (p);
+
+      \node[acclabel] at (1,-1.4)
+      {\small$(\Fin\tacc{0}{0}\wedge\Inf\tacc{1}{1})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1})$};
+    \end{scope}
+    \begin{scope}[initial angle=90,xshift=3.5cm,yshift=-4cm]
+      \node (l) at (1,1.3) {level 3:};
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (2,0) {};
+      \path[->,auto]
+      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
+      (p) edge[out=30,in=150] pic {acc={0}{0}} pic {l=$b$} (q)
+      (q) edge[out=30,in=-30,looseness=10] pic {l=$c$} (q)
+      (q) edge[out=-150,in=-30] pic {l=$d$} (p);
+
+      \node[acclabel] at (1,-1.4)
+      {\small$\Fin\tacc{0}{0}$};
+    \end{scope}
+  \end{tikzpicture}
+  \smallskip
+
+  \caption{An example illustrating the results of the three basic
+    reduction levels: an input automaton and the automata
+    obtained by reduction of level 1, level 2, and level 3.
+    \todo{Pozor, prvni klauzule v levelu 2 je trash!}}
+  \label{fig:example}
+\end{figure}
 

### 2022-03-01 paper/main.tex
first part of Lemma 3 proven

@@
 exists a path from a state $p$ to the same state containing all
-transitions of $T$ and no transition outside $T$. Finally, we assume
-that each TELA $\mA$ contains only states $q$ that are reachable from
-the initial state $q_I$ (i.e., $q=q_I$ or there is a path from $q_I$
-to $q$) as states that are not reachable from $q_I$ can be eliminated
-without any impact on $L(\mA)$.
+transitions of $T$ (potentially multiple times) and no transition
+outside $T$. Finally, we assume that each TELA $\mA$ contains only
+states $q$ that are reachable from the initial state $q_I$ (i.e.,
+$q=q_I$ or there is a path from $q_I$ to $q$) as states that are not
+reachable from $q_I$ can be eliminated without any impact on $L(\mA)$.
 
@@
 \begin{lemma}
-  A set $T\subseteq\delta$ is a cycle if for each set of
+  A set $T\subseteq\delta$ is a cycle if and only if for each set of
   states $S\subseteq Q$ it holds that
-  \[
-    \begin{split}
-    \textrm{if $T$ contains a transition leading from a state in $S$ to a state outside $S$,~~~~}\\[-.5ex]
-    \textrm{then it also contains a transition leading from a state outside $S$ to a state in $S$.}
-    \end{split}
+  \begin{equation}
+    \begin{minipage}{0.85\linewidth}
+      \centering
+      if $T$ contains a transition leading from a state in $S$ to a state outside $S$,\\
+      then it also contains a transition leading from a state outside $S$ to a state in $S$.
+    \end{minipage}
+    % \begin{split}
+    % \textrm{if $T$ contains a transition leading from a state in $S$ to a state outside $S$,}\\[-.5ex]
+    % \textrm{then it also contains a transition leading from a state outside $S$ to a state in $S$.}
+    % \end{split}
     \tag{$3$}\label{eq:lem3} 
-  \]  
+  \end{equation}
 \end{lemma}
 \begin{proof}
-  \todo{Ted to musim dokazat.}
+  We first prove direction ``$\Longrightarrow$''. Let $T$ be a cycle
+  and $S\subseteq Q$ be a set of states such that there exists a
+  transition $t\in T$ leading from a state $p\in S$ to a state
+  $q\not\in S$. The definition of a cycle says that there exists a
+  path $t_1t_2\ldots t_n\in T^+$ from a state $p'$ back to $p'$ using
+  all transitions of $T$. Hence, $t=t_i$ for some $1\le i\le n$. Then
+  $t_{i+1}\ldots t_nt_1t_2\ldots t_{i-1}\in T^+$ is a path leading
+  from the state $q\not\in S$ to the state $p\in S$. This path has to
+  contain a transition $t'\in T$ leading from a state outside $S$ to a
+  state in $S$.
+
+  Now we prove the opposite direction ``$\Longleftarrow$''.
+
 \end{proof}

### 2022-03-01 main.py
gradual increasing of levels func

@@ -49,6 +49,10 @@ def main(argv):
         "-I",
         "--incremental",
         help="Use incremental solving.", action='store_true')
+    parser.add_argument(
+        "-G",
+        "--gradual",
+        help="Levels will gradually grow from L1 up to specified level.", action='store_true')
     args = parser.parse_args()
 
 
@@ -58,7 +62,9 @@ def main(argv):
         args.level = 3
 
     aut = spot.automata(args.autfile)
-    timeouted = [0]
+    tmp_mode = args.level
+    if args.gradual:
+        tmp_mode = 1
 
     for a in aut:
 
@@ -74,9 +80,11 @@ def main(argv):
             if args.base_level == "telatko":
                 a = process_automaton(a)
                 if args.level > 1 and acc_sets_count != 0:
+                    if tmp_mode != args.level:
+                        tmp_mode = 2
 
                     auto = play(
-                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut,  2, args.incremental)
+                        a, clauses_count, acc_sets_count, args.level, args.timeout, args.scc, args.minimized_atribut,  tmp_mode, args.incremental)
                 else:
                     auto = a
             else:
@@ -84,7 +92,7 @@ def main(argv):
                     auto = a
                 else:
                     auto = play(
-                        a, clauses_count, acc_sets_count, args.level, args.timeout, timeouted, args.scc, args.minimized_atribut, 1, args.incremental)
+                        a, clauses_count, acc_sets_count, args.level, args.timeout, args.scc, args.minimized_atribut, tmp_mode, args.incremental)
 
             if args.outfile:
                 print_aut(auto, args.outfile, "a")
@@ -93,8 +101,7 @@ def main(argv):
 
             if not spot.are_equivalent(origin, auto):
                 assert(False)
-                
-                return
+
 
 
             acc_sets_count2 = auto.get_acceptance().used_sets().count()

### 2022-03-01 qbf/q_main.py

@@ -257,7 +257,7 @@ def update_run_info(K, res, tmp_mode):
     if res == unsat:
         str_res = "U"
     assert(str_res != "")
-    return "L_{0}_{1}_{2} ".format(str(tmp_mode), str(K), str_res)
+    return "L{0}_{1}_{2} ".format(str(tmp_mode), str(K), str_res)
 
 def get_formula(aut, scc_state_info, C, K, tmp_mode, edge_dict, scc_edg, representants, optimized_scc, inner_edges_nums, inner_edges):
     if (optimized_scc):
@@ -313,9 +313,10 @@ def restrict_pn_vars(C, K):
 
 def update_inc_solver(solver, currently_reduced, C, K, inner_edges_nums):
     if currently_reduced == FormulaAtribute.K:
+        solver.push()
         solver.add(restrict_f_vars(K, inner_edges_nums))
         solver.add(restrict_pn_vars(C, K))
-        solver.push()
+
         return solver
     else:
         print("TBD")
@@ -334,7 +335,81 @@ def model_assert(model, K):
         if str(t)[0] == 'f' and str(t)[-1] == str(K):
             assert(is_false(model[t]))
 
-def play(aut, C, K, mode, timeout, timeouted,
+def inc_loop(aut, inner_edges_nums, inner_edges, mode, tmp_mode, scc_state_info, C, K,  scc_edg, optimized_scc, timeout, minimized_atribute):
+    currently_reduced = resolve_formula_atributes(minimized_atribute, C, K)
+
+    original = spot.automaton(aut.to_str())
+    qbf_run_info = ""
+    # dictionary {acc_set_num : [edge_nums]}
+    edge_dict, scc_equiv_edges, representants = edge_dictionary(aut, tmp_mode)
+    formula = get_formula(aut, scc_state_info, C, K, tmp_mode, edge_dict, scc_edg, representants, optimized_scc, inner_edges_nums, inner_edges)
+    solver = solver_init(formula, timeout)
+    model = None
+    if currently_reduced == FormulaAtribute.K:
+        K += 1
+    else:
+        C += 1
+
+    while C > 0 and K > 0:
+
+        if aut.get_acceptance().used_sets().count(
+        ) < 1 or aut.prop_state_acc() == spot.trival.yes_value:
+            return aut
+
+        res = solver.check()
+
+        if res == sat:
+            model = solver.model()
+
+            if K == 1:
+                # does automaton with one mark exist? --the solver asks, answer was yes
[…]

### 2022-03-01 qbf/z3_formula.py

@@ -149,7 +149,7 @@ class Z3_f_ctor():
         old = self.old_formula(ACC_DNF(aut.get_acceptance().to_dnf()))
 
         new = self.new_formula()
-
+        
         new = simplify(new)
 
         laso = simplify(laso)

### 2022-03-02 paper/main.tex
proof of Lemma 3 unfinished (klizi se mi oci, az rano)

--- 2022-03-01 paper/main.tex
@@
 \newcommand{\Te}{T_{\vec{e}}}
+\newcommand{\pre}{S_\mathit{pre}}
+\newcommand{\post}{S_\mathit{post}}
 \def\false{\mathit{false}}
@@
 \emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
-the rest of this paper, SCC always refers to a maximal SCC. Given an
-SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
+the rest of this paper, SCC always refers to a maximal SCC. Given a
+set of states $S\subseteq Q$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
 denotes the set of all transitions between states in $S$. Further, for
@@
 marked with $m$, i.e., $\delta_m=\{t\in\delta\mid m\in
-\mks(t)\}$. \todo{The sets $\delta_m$ are called \emph{acceptance
-    sets}. Given an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we
-  denote the set $\delta_{S,m}=\delta_S\cap\delta_m$.} A set of
-transitions $T\subseteq\delta$ is called a \emph{cycle} is there
-exists a path from a state $p$ to the same state containing all
-transitions of $T$ and no transition outside $T$. Finally, we assume
-that each TELA $\mA$ contains only states $q$ that are reachable from
-the initial state $q_I$ (i.e., $q=q_I$ or there is a path from $q_I$
-to $q$) as states that are not reachable from $q_I$ can be eliminated
-without any impact on $L(\mA)$.
+\mks(t)\}$.
+% \todo{The sets $\delta_m$ are called \emph{acceptance sets}. Given
+% an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we denote the set
+% $\delta_{S,m}=\delta_S\cap\delta_m$.}
+A set of transitions $T\subseteq\delta$ is called a \emph{cycle} is
+there exists a path from a state $p$ to the same state containing all
+transitions of $T$ (potentially multiple times) and no transition
+outside $T$. Finally, we assume that each TELA $\mA$ contains only
+states $q$ that are reachable from the initial state $q_I$ (i.e.,
+$q=q_I$ or there is a path from $q_I$ to $q$) as states that are not
+reachable from $q_I$ can be eliminated without any impact on $L(\mA)$.
 
@@
 \begin{lemma}
-  A set $T\subseteq\delta$ is a cycle if for each set of
+  A nonempty set $T\subseteq\delta$ is a cycle if and only if for each set of
   states $S\subseteq Q$ it holds that
-  \[
-    \begin{split}
-    \textrm{if $T$ contains a transition leading from a state in $S$ to a state outside $S$,~~~~}\\[-.5ex]
-    \textrm{then it also contains a transition leading from a state outside $S$ to a state in $S$.}
-    \end{split}
-    \tag{$3$}\label{eq:lem3} 
-  \]  
+  \begin{enumerate}
+  \item\label{en:1} all transitions in $T$ lead from a state in $S$ to a state in $S$ (i.e., $T\subseteq\delta_S$) or
+  \item\label{en:2} all transitions in $T$ lead from a state outside $S$ to a state outside $S$ (i.e., $T\subseteq\delta_{Q\smallsetminus S}$) or
+  \item\label{en:3} $T$ contains a transition leading from a state in $S$ to a state outside $S$ and a transition leading from a state outside
+    $S$ to a state in $S$.
+  \end{enumerate}
 \end{lemma}
 \begin{proof}
-  \todo{Ted to musim dokazat.}
+  We first prove the direction ``$\Longrightarrow$''. Let $T$ be a
+  cycle and $S\subseteq Q$ be an arbitrary set of states. We show that
+  if (\ref{en:1}) and (\ref{en:2}) do not hold, then (\ref{en:3}) has
+  to hold. Hence, assume that $T\not\subseteq\delta_S$ and
+  $T\not\subseteq\delta_{Q\smallsetminus S}$. Then there are two
+  cases.
+  \begin{itemize}
+  \item $T$ contains a transtition $t\in\delta_S$ and a transition
+    $t'\in\delta_{Q\smallsetminus S}$. The definition of a cycle
+    implies that there exists a path $t_1t_2\ldots t_n\in T^+$ from a
+    state $p$ back to $p$ containing both $t$ and $t'$. However, this
+    implies that $T$ contains a transtion leading from a state in $S$
+    to a state outside $S$ and a transtion leading from a state outside $S$
+    to a state in $S$.
+  \item $T$ contains a transition $t$ leading from a state in $S$ to a
+    state outside $S$ (or vice versa). However, as $T$ is a cycle,
+    there exists a path $t_1t_2\ldots t_n\in T^+$ that leads from a
+    state $p$ to the same state and contains $t$. Hence, $T$ has to
+    contain also a transition leading from a state outside $S$ to a
+    state in $S$ (or vice versa).
+  \end{itemize}
+  In both cases, $\ref{en:3}$ holds.
+
+  Now we prove the opposite direction ``$\Longleftarrow$''. As $T$ is
+  nonempty, there exists a transition $t\in T$ leading from a state $p$.
+  We use this state to define the sets $\post,\pre\subseteq Q$ of states as 
+  \begin{align}\notag
+    \post&=\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }p\textrm{ to }q\}\\
+    \pre &=\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }q\textrm{ to }p\}
+  \end{align}
+  Note that each transition $t'\in T$ leading from a state in
+  $q\in\post$ leads also to a state in $\post$ as the path in $T^+$
+  leading form $p$ to $q$ can be prolonged by this transition. In
+  other words, there is no transtition in $T$ leading from a state
+  in $\post$ to a state outside $\post$. Similarly, one can show that
+  there is no transition in $T$ leadning from outside $\pre$ to $\pre$.
+
+  \todo{Jan is here}
+
+  us we
+  choose Let us fix
+  a transition $t\in T$ leading from a state $p$ to a state $q$. Let
+  $\post$ be the set of all states reachable from $q$ using
+  transitions in $T$ and $\pre$ be the set of all states from which
+
+  that $T\subseteq\delta$ is a nonemtpy set such that~\eqref{eq:lem3}
+  holds for each set $S\subset Q$.
 \end{proof}

### 2022-03-02 qbf/q_main.py
inc test ben

@@ -153,8 +153,8 @@ def try_evaluate0(aut, orig, qbf_run_info):
     :return: potentionally zero acceptance spot::automaton
     """
     last_eq_aut = spot.automaton(aut.to_str())
-    aut.set_name(qbf_run_info + "F_0_S")
-    last_eq_aut.set_name(qbf_run_info + "F_0_U")
+    aut.set_name(aut.get_name() + "F_0_S")
+    last_eq_aut.set_name(last_eq_aut.get_name() + "F_0_U")
     clear_aut_edges(aut)
     aut.set_acceptance(0, spot.acc_code.t())
     if spot.are_equivalent(orig, aut):
@@ -337,7 +337,8 @@ def model_assert(model, K):
 
 def inc_loop(aut, inner_edges_nums, inner_edges, mode, tmp_mode, scc_state_info, C, K,  scc_edg, optimized_scc, timeout, minimized_atribute):
     currently_reduced = resolve_formula_atributes(minimized_atribute, C, K)
-
+    K = aut.get_acceptance().used_sets().count()
+    K -= 1
     original = spot.automaton(aut.to_str())
     qbf_run_info = ""
     # dictionary {acc_set_num : [edge_nums]}
@@ -359,17 +360,24 @@ def inc_loop(aut, inner_edges_nums, inner_edges, mode, tmp_mode, scc_state_info,
         res = solver.check()
 
         if res == sat:
+            print("sat, ", K)
             model = solver.model()
+            if aut.get_acceptance().used_sets().count() <= 1:
+                print(aut.get_acceptance(), aut.get_acceptance().used_sets().count())
+                return try_evaluate0(aut, original, qbf_run_info)
 
+            """
             if K == 1:
+                continue
                 # does automaton with one mark exist? --the solver asks, answer was yes
                 # therefore my question is if 0 marks can be used
                 # aut with one acc mark is SAT -- rozmysli si tohle jeste
                 process_variables(aut, model, scc_equiv_edges, tmp_mode)
-                return try_evaluate0(aut, original, qbf_run_info)
+
+            """
 
 
-            if mode == 1:
+            if tmp_mode == 1:
                 solver  = update_inc_solver(solver, currently_reduced, C, K, representants)
             else:
                 solver  = update_inc_solver(solver, currently_reduced, C, K, inner_edges_nums)
@@ -379,11 +387,17 @@ def inc_loop(aut, inner_edges_nums, inner_edges, mode, tmp_mode, scc_state_info,
             else:
                 C = C - 1
         else:
-
[…]

### 2022-03-02 paper/main.tex
section 3.3 nearly finished!

@@
 \newcommand{\Te}{T_{\vec{e}}}
-\newcommand{\pre}{S_\mathit{pre}}
-\newcommand{\post}{S_\mathit{post}}
+% \newcommand{\pre}{{\leftarrow\!\!S}}
+% \newcommand{\post}{{S\!\!\rightarrow}}
+\newcommand{\pre}{{S_\mathit{pre}}}
+\newcommand{\post}{{S_\mathit{post}}}
 \def\false{\mathit{false}}
@@
 adjacent transitions of the form
-$(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
+$\rho=(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
 such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
@@
   states $S\subseteq Q$ it holds that
-  \begin{enumerate}
+  \begin{enumerate}[A.]
   \item\label{en:1} all transitions in $T$ lead from a state in $S$ to a state in $S$ (i.e., $T\subseteq\delta_S$) or
@@
   \end{itemize}
-  In both cases, $\ref{en:3}$ holds.
-
-  Now we prove the opposite direction ``$\Longleftarrow$''. As $T$ is
-  nonempty, there exists a transition $t\in T$ leading from a state $p$.
-  We use this state to define the sets $\post,\pre\subseteq Q$ of states as 
-  \begin{align}\notag
-    \post&=\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }p\textrm{ to }q\}\\
-    \pre &=\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }q\textrm{ to }p\}
-  \end{align}
-  Note that each transition $t'\in T$ leading from a state in
-  $q\in\post$ leads also to a state in $\post$ as the path in $T^+$
-  leading form $p$ to $q$ can be prolonged by this transition. In
-  other words, there is no transtition in $T$ leading from a state
-  in $\post$ to a state outside $\post$. Similarly, one can show that
-  there is no transition in $T$ leadning from outside $\pre$ to $\pre$.
-
-  \todo{Jan is here}
-
-  us we
-  choose Let us fix
-  a transition $t\in T$ leading from a state $p$ to a state $q$. Let
-  $\post$ be the set of all states reachable from $q$ using
-  transitions in $T$ and $\pre$ be the set of all states from which
-
-  that $T\subseteq\delta$ is a nonemtpy set such that~\eqref{eq:lem3}
-  holds for each set $S\subset Q$.
+  In both cases, (\ref{en:3}) holds.
+
+  Now we prove the opposite direction ``$\Longleftarrow$'' by
+  contraposition. Assume that a nonempty set $T$ is not a cycle. We
+  show that there is a set $S\subseteq Q$ such neither (\ref{en:1})
+  nor (\ref{en:2}) nor (\ref{en:3}) holds. Let $p$ be a state such
+  that some transition of $T$ leads from $p$. We define the set
+  \[
+    \post=\{p\}\cup\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }p\textrm{ to }q\}
+  \]
+  of states reachable from $p$ via transitions in $T$ and the set
+  \[
+    \pre=\{p\}\cup\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }q\textrm{ to }p\}
+  \]  
+  of states from which $p$ is reachable via transtitions of $T$. As
+  there is a transition of $T$ leading from $p$, we have that
+  $T\not\subseteq\delta_{Q\smallsetminus\post}$ and
+  $T\not\subseteq\delta_{Q\smallsetminus\pre}$, i.e., (\ref{en:2})
+  does not holds for $\post$ and $\pre$. Further, the definition of
+  $\post$ implies that there is no transition of $T$ leading from a
+  state in $\post$ to a state outside $\post$, which means that
+  (\ref{en:3}) does not hold for $\post$. Similarly, $T$ contains no
+  transition leading from a state outside $\pre$ to a state in $\pre$,
+  which means that (\ref{en:3}) does not hold for $\pre$. Now we prove
+  by contradition that (\ref{en:1}) does not hold for at least one of
+  $\post,\pre$. Hence, let us assume that $T\subseteq\delta_\post$ and
+  $T\subseteq\delta_\pre$. Then for each $t_i\in T$ leading from $p_i$
+  to $q_i$ we have that $p_i\in\post$ and $q_i\in\pre$, which implies
+  that there is a path $\rho_i\in T^+$ containing $t_i$ and leading
+  from $p$ back to $p$. If we concatenate all these paths, we get the
+  path $\rho_1\rho_2\ldots\rho_{|T|}\in T^+$ that contains all transitions of
+  $T$ and leads from $p$ back to $p$, which means that $T$ is a cycle.
+  This is a contradiction.
 \end{proof}
+
+The formula $\mcycle_3(\vec{e})$ says that $\Te$ is nonempty and each
+set $S\subseteq Q$ satisfies (\ref{en:1}) or (\ref{en:2}) or
+(\ref{en:3}).  For each state $q\in Q$, variable $s_q$ says whether
+$q$ is in the current set $S$ or not.
+\[s_q=\left\{
+    \begin{array}{ll}
+      1 & \textrm{if }q\in S\\
+      0 & \textrm{otherwise}    
+    \end{array}\right.
+\]
+By $\vec{s}$ we denote the vectors of all variables of the form $s_q$.
+The formula $\mcycle_3(\vec{e})$ is defined as follows.
+\begin{align*}
+  \mcycle_3(\vec{e})
+  &=\bigvee_{t\in\delta}e_t~\wedge~\forall\vec{s}~.~\zeta_{\ref{en:1}}(\vec{e},\vec{s})\,\vee\,
+    \zeta_{\ref{en:2}}(\vec{e},\vec{s})\,\vee\,
+    \zeta_{\ref{en:3}}(\vec{e},\vec{s})\\ %\qquad\textrm{where}\\
+  \zeta_{\ref{en:1}}(\vec{e},\vec{s}) &= \bigwedge_{t=(p,a,M',q)\in\delta}\hspace{-5mm}\big(e_t\implies (s_p\wedge s_q)\big)\\
+  \zeta_{\ref{en:2}}(\vec{e},\vec{s}) &= \bigwedge_{t=(p,a,M',q)\in\delta}\hspace{-5mm}\big(e_t\implies (\neg s_p\wedge \neg s_q)\big)\\
+  \zeta_{\ref{en:3}}(\vec{e},\vec{s}) &= \big(\bigvee_{t=(p,a,M',q)\in\delta}\hspace{-5mm}e_t\wedge s_p\wedge \neg s_q\big) ~\wedge~
+  \big(\bigvee_{t=(p,a,M',q)\in\delta}\hspace{-5mm}e_t\wedge \neg s_p\wedge s_q\big)
+\end{align*}
+\todo{Zminit, ze z univerzalniho kvantifikatoru se stane existencni.}
 

### 2022-03-02 paper/main.tex
paper: some Impl & Experiments text + TeX for tables

@@
 
+\usepackage{multirow}
+\usepackage{booktabs}
+\usepackage{xspace}
+\usepackage{makecell}
+\usepackage[table]{xcolor}
+%\usepackage{arydshln}
+%\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
+
 \usepackage{tikz}
@@
 \usetikzlibrary{extshapes}
+
+\usepackage{pgfplots}
+\pgfplotsset{compat=newest}
 
@@
 \newcommand{\mB}{\mathcal{B}}
+
+\newcommand{\telatko}{\texttt{telatko}\xspace}
 
@@
 We have implemented our reduction technique in a tool called
-\texttt{telatko}. We show that the tool can reduce acceptance marks in
+\telatko. We show that the tool can reduce acceptance marks in
 automata produced by Delag~\cite{muller.17.gandalf},
@@
 \end{figure}
+%}}}
+%{{{ algorithm and implementation
+
+\section{Reduction algorithm and its implementation}\label{sec:algimpl}
+
+The presented reduction algorithms have been implemented in a tool called
+\telatko. It uses the Spot library~\cite{duret.16.atva2} for automata parsing
+and manipulation, and the Z3 Theorem Prover~\cite{demoura.08.tacas} to solve
+the satisfiability of QBF formulas. Moreover, in the experimental evaluation
+we use several tools from the Spot library, i.e. \texttt{genltl},
+\texttt{ltlfilt}, \texttt{ltl2tgba}, \texttt{autcross} and \texttt{autfilt},
+to generate, simplify and translate LTL formulas and to compare and simplify
+automata. \telatko is implemented in Python.
+
+Jak se nastavuji inicialni parametry a jak postupne klesaji.
+Jak prechazim od nizsich levelu k vyssim.
+Vyuzivame incremental QBF solving?
 
@@
 \section{Experimental evaluation}\label{sec:experiments}
+
+We have used the tool \texttt{genltl} from Spot to generate LTL formulas
+from literature and from scalable patterns. Using the values from 1 to 5 for
+the parametrized formulas, we have generated 407 unique LTL formulas. These
+have been translated to deterministic or nondeterministic TELA using four
+state-of-the-art translators \texttt{ltl2tgba} (from the Spot library),
+\texttt{ltl3tela}, \texttt{ltl2dela} (known as Delag), and \texttt{ltl2dgra}
+(known as Rabinizer 4). The generated automata sets contain less than 407
+automata as not all of the formulas are successfully translated due to timeouts.
+Table \ref{tab:ltltranslators} shows the exact versions, the homepage of the
+used tools and the number of input automata produced by the given tool from
+the LTL formulas.
+
+Spot version 2.10.4 and Z3 version 4.8.13 have been used in the experimental
+evaluation.
+
+\begin{table}
+\caption{LTL to automata translators used for the experimental evaluation
+and the number of successfully translated formulas.}
+\label{tab:ltltranslators}
+\centering
+\begin{tabular}{lr@{~~~}lr}
+\toprule
+tool        & version&homepage & \# formulas \\
+\midrule
+\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 404\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} \todo{RELEASE} & 376\\
+Delag & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 397\\
+Rabinizer & \todo ? & & 386\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{Test: heatmap}
+\centering
+\begin{tabular}{c|ccccc|}
+\multicolumn{1}{c}{} & 3 & 4 & 5 & 6 & \multicolumn{1}{c}{7} \\
+%\specialrule{2pt}{0pt}{0pt}
+\cline{2-6}
+3 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \cellcolor{blue!25}\makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+4 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+5 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+6 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+7 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+\cline{2-6}
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{Test: Data}
+\centering
+\begin{tabular}{cccccc}
+\toprule
+& & marks & \% red. & time & timeouts \\
+\midrule
+\multirow{4}{*}{\rotatebox[origin=c]{90}{\makecell{Delag \\ (420 marks)}}} & level 1 & 240 & 12.3 & 133.7 & 42 \\
+& level 2 & 240 & 12.3 & 133.7 & 42 \\
+& level 3 & 240 & 12.3 & 133.7 & 42 \\
+& multilevel & 240 & 12.3 & 133.7 & 42 \\
+\midrule
+\multirow{4}{*}{\rotatebox[origin=c]{90}{\makecell{Rabinizer 4 \\ (420 marks)}}} & level 1 & 240 & 12.3 & 133.7 & 42 \\
+& level 2 & 240 & 12.3 & 133.7 & 42 \\
+& level 3 & 240 & 12.3 & 133.7 & 42 \\
+& multilevel & 240 & 12.3 & 133.7 & 42 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{tikzpicture}
+\begin{axis}[
+	ymode=log,
+	height=5cm,
+	width=10cm,
+	grid=major,
+]
+\addplot coordinates {
+	(1, 0.3)
+	(2, 0.8)
+	(3, 1.5)
+	(4, 2.7)
+	(5, 3.9)
+	(6, 4.1)
+};
+\addlegendentry{Delag}
+\end{axis}
+\end{tikzpicture}
 

### 2022-03-03 paper/main.tex
smal changes in structure of the paper

@@
 construction of three mentioned QBF formulas. The reduction algorithm
-based on these formulas is described together with its implementation
-in Section~\ref{sec:algimpl}. Section~\ref{sec:experiments} shows the
-experimental results. Finally, Section~\ref{sec:conclusion} suggests
-other applications of our QBF-based reduction technique and closes the
-paper.
+based on these formulas is presented in
+Section~\ref{sec:algo}. Section~\ref{sec:implementation} describes our
+tool \telatko implemening the reduction technique.
+Experimental results are shown in
+Section~\ref{sec:experiments}. Finally, Section~\ref{sec:conclusion}
+suggests other applications of our QBF-based reduction technique and
+closes the paper.
 
@@
 %}}}
-%{{{ algorithm and implementation
-
-\section{Reduction algorithm and its implementation}\label{sec:algimpl}
+%{{{ algorithm
+
+\section{Reduction algorithm}\label{sec:algo}
 Jak se nastavuji inicialni parametry a jak postupne klesaji.
@@
 \end{figure}
+
 %}}}
-%{{{ algorithm and implementation
-
-\section{Reduction algorithm and its implementation}\label{sec:algimpl}
+%{{{ implementation
+
+\section{Implementation}\label{sec:implementation}
 
@@
 
-Jak se nastavuji inicialni parametry a jak postupne klesaji.
-Jak prechazim od nizsich levelu k vyssim.
-Vyuzivame incremental QBF solving?
+\todo{Jake parametry ten tool ma? Co potrebuje k behu? Kde je dostupny? Pod jakou licenci?}
 

### 2022-03-03 paper/main.tex
small changes in section 3.

@@
 
-With this lightweight version of $\mcycle(\vec{e})$, the formula
-$\Phi_{C,K}$ actually says that
-for every nonempty set
-$T\subseteq\delta$, $T$ satisfies $\varphi$ if and only if $\rem(T)$ satisfies
-$\psi$. Note that the formula is not affected by the placement of
-transitions in the automaton. The only aspect of a transition $t$
-reflected by the formula is its set of marks $\mks(t)$. Hence, we do
-not have to distinguish between transitions with the same sets of
-marks.
+The formula $\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$ actually
+says that for every nonempty set $T\subseteq\delta$, $T$ satisfies
+$\varphi$ if and only if $\rem(T)$ satisfies $\psi$. Note that the
+formula is not affected by the placement of transitions in the
+automaton. The only aspect of a transition $t$ reflected by the
+formula is its set of marks $\mks(t)$. Hence, we do not have to
+distinguish between transitions with the same sets of marks.
 
@@
 The lemma suggests the following simplification of the whole formula
-$\Phi_{C,K}$ built with $\mcycle_1(\vec{e})$. Before we build the
+$\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$. Before we build the
 formula, we compute the equivalence $\sim$ on automata transitions and
@@
 transitions we removed, the shorter formula with less $e_t$ variables
-we obtain. If the formula $\Phi_{C,K}$ for the pruned automaton is
+we obtain. If the formula $\Phi_{1,C,K}$ for the pruned automaton is
 satisfiable and we derive $\rem$ and $\psi$ from its model, we can use
@@
 \end{align*}
-\todo{Zminit, ze z univerzalniho kvantifikatoru se stane existencni.}
 
 \subsection{Optimizations}
+
+\todo{Honza je tu!}
 

### 2022-03-03 paper/main.tex
add Table with cummulative numbers

@@
 
-\begin{table}
+For each automata set (that is, the output of individual LTL translator), we
+have run \telatko with various reduction levels, always with the incremental
+solving turned on. We have again filtered out the automata for which any run
+of \telatko timed out after 120 seconds. We do this to make sure the numbers
+of acceptance marks can be meaningfully compared between different reduction
+levels of the same automata set.
+
+Table~\ref{tab:cumm-data} shows the number of marks in the input sets and after
+each reduction level, together with the reduction ratio and time spent by the
+individual reduction level. The columns \textit{solver response} and
+\textit{solver timeout} show the number of automata for which the last query to
+QBF solver has been successfully answered, or the solver timed out within a 30
+seconds limit to a single query, respectively. An unanswered QBF formula (i.e.
+the solver timed out) is considered unsatisfiable.
+
+\begin{table}[h]
 \caption{LTL to automata translators used for the experimental evaluation
@@
 \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} \todo{RELEASE} & 376\\
-Delag & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 397\\
-Rabinizer & \todo ? & & 386\\
+\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 397\\
+\texttt{ltl2dgra} & 21.0 & & 386\\
 \bottomrule
@@
 
-\begin{table}
-\caption{Test: heatmap}
+%\begin{table}[t]
+%\centering
+%\begin{tabular}{c|ccccc|}
+%\multicolumn{1}{c}{} & 3 & 4 & 5 & 6 & \multicolumn{1}{c}{7} \\
+%%\specialrule{2pt}{0pt}{0pt}
+%\cline{2-6}
+%3 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \cellcolor{blue!25}\makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+%4 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+%5 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+%6 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+%7 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
+%\cline{2-6}
+%\end{tabular}
+%\end{table}
+
+\begin{table}[h]
+\caption{The cummulative number of acceptance marks before and after reduction
+for various automata sets. \textit{X/Y solved aut.} means that out of
+\textit{Y} successfully translated LTL formulas, every run of \telatko ended
+within 120 seconds time limit on \textit{X} automata. We sum up the acceptance
+marks only on these automata. The columns \textit{solver response} and
+\textit{solver timeout} refer to last call to the QBF solver (that is, the
+numbers of its successful responses within 30 seconds limit and timeouts).
+}
+\label{tab:cumm-data}
 \centering
-\begin{tabular}{c|ccccc|}
-\multicolumn{1}{c}{} & 3 & 4 & 5 & 6 & \multicolumn{1}{c}{7} \\
-%\specialrule{2pt}{0pt}{0pt}
-\cline{2-6}
-3 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \cellcolor{blue!25}\makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-4 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-5 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-6 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-7 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-\cline{2-6}
-\end{tabular}
+\input{table-data.tex}
 \end{table}
 
-\begin{table}
-\caption{Test: Data}
-\centering
-\begin{tabular}{cccccc}
-\toprule
-& & marks & \% red. & time & timeouts \\
-\midrule
-\multirow{4}{*}{\rotatebox[origin=c]{90}{\makecell{Delag \\ (420 marks)}}} & level 1 & 240 & 12.3 & 133.7 & 42 \\
-& level 2 & 240 & 12.3 & 133.7 & 42 \\
-& level 3 & 240 & 12.3 & 133.7 & 42 \\
-& multilevel & 240 & 12.3 & 133.7 & 42 \\
-\midrule
-\multirow{4}{*}{\rotatebox[origin=c]{90}{\makecell{Rabinizer 4 \\ (420 marks)}}} & level 1 & 240 & 12.3 & 133.7 & 42 \\
-& level 2 & 240 & 12.3 & 133.7 & 42 \\
-& level 3 & 240 & 12.3 & 133.7 & 42 \\
-& multilevel & 240 & 12.3 & 133.7 & 42 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{tikzpicture}
-\begin{axis}[
-	ymode=log,
-	height=5cm,
-	width=10cm,
-	grid=major,
-]
-\addplot coordinates {
-	(1, 0.3)
-	(2, 0.8)
-	(3, 1.5)
-	(4, 2.7)
-	(5, 3.9)
-	(6, 4.1)
-};
-\addlegendentry{Delag}
-\end{axis}
-\end{tikzpicture}
+%\begin{tikzpicture}
+%\begin{axis}[
+%	ymode=log,
+%	height=5cm,
+%	width=10cm,
+%	grid=major,
+%]
+%\addplot coordinates {
+%	(1, 0.3)
+%	(2, 0.8)
+%	(3, 1.5)
+%	(4, 2.7)
+%	(5, 3.9)
+%	(6, 4.1)
+%};
+%\addlegendentry{Delag}
+%\end{axis}
+%\end{tikzpicture}
 

### 2022-03-03 paper/main.tex
some progress in section 3.4

--- 2022-03-03 paper/main.tex
@@
   these tools. In the case of Rabinizer~4, the cumulative number of
-  acceptance marks was reduced to less than one half.
+  acceptance marks was reduced to less than one half. \todo{Sazet ltl3tela a jine tooly vsude jako \texttt{ltl3tela} apod?}
 \end{abstract}
@@
 
-\subsection{Optimizations}
+\subsection{SCC-based optimization}
+
+Now we present a formula optimization based on the fact that every
+cycle is completely included in the transition set $\delta_S$ of some
+SCC $S$. Note that we cannot reduce the number of acceptance marks in
+each SCC separately as all SCCs have to share one acceptance formula.
+However, for each SCC we can build a $\Phi_{C,K}$ formula separately and
+then we can put all these formulas into one conjunction such that they
+share the variables $\vec{i},\vec{f}$.
 
 \todo{Honza je tu!}
+
+For a given automaton $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, we first
+compute all SCCs. Each SCC $S$ induces an automaton $(S,M,\Sigma,q',\delta_S,\varphi)$, where $q'$ is an arbitrary state
+of $S$.
+
+the optimization saying that So far, we have not use the fact that each cycle is A
 

### 2022-03-03 paper/table-data.tex

\begin{tabular}{ccccccc}                                                        
\toprule                                                                        
automata set & level & marks & reduction [\%] & time [s] & \makecell{solver \\ response} & \makecell{solver \\ timeout} \\

\midrule
\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\389/404 solved aut.\\538 acc. marks}}
&
level 1 &
421 & 21.75 & 419.2 & 389 & 0
\\
&
level 2 &
421 & 21.75 & 14.13 & 389 & 0
\\
&
level 3 &
412 & 23.42 & 232.98 & 382 & 7
\\
&
multi &
412 & 23.42 & 686.96 & 383 & 6
\\
\midrule
\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\371/376 solved aut.\\562 acc. marks}}
&
level 1 &
546 & 2.85 & 783.19 & 353 & 18
\\
&
level 2 &
548 & 2.49 & 80.53 & 351 & 20
\\
&
level 3 &
540 & 3.91 & 163.15 & 346 & 25
\\
&
multi &
537 & 4.45 & 2295.88 & 346 & 25
\\
\midrule
\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\388/397 solved aut.\\766 acc. marks}}
&
level 1 &
553 & 27.81 & 898.88 & 370 & 18
\\
&
level 2 &
563 & 26.5 & 99.36 & 367 & 21
\\
&
level 3 &
569 & 25.72 & 355.55 & 357 & 31
\\
&
multi &
541 & 29.37 & 2538.7 & 361 & 27
\\
\midrule
\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\378/386 solved aut.\\1084 acc. marks}}
&
level 1 &
616 & 43.17 & 828.84 & 362 & 16
\\
&
level 2 &
627 & 42.16 & 132.48 & 358 & 20
\\
&
level 3 &
627 & 42.16 & 194.59 & 352 & 26
\\
&
multi &
606 & 44.1 & 2339.83 & 352 & 26
\\

\bottomrule                                                                     
\end{tabular}

### 2022-03-04 paper/main.tex
add quantile chart
clarify number of input automata

@@
 automata as not all of the formulas are successfully translated due to timeouts.
+We have also removed automata with 0 or 1 acceptance marks as there is little
+point reducing these.
 Table \ref{tab:ltltranslators} shows the exact versions, the homepage of the
-used tools and the number of input automata produced by the given tool from
-the LTL formulas.
+used tools and the number of automata produced by the given tool from
+the LTL formulas after filtering.
 
@@
 \toprule
-tool        & version&homepage & \# formulas \\
+tool        & version&homepage & automata \\
 \midrule
-\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 404\\
-\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} \todo{RELEASE} & 376\\
-\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 397\\
-\texttt{ltl2dgra} & 21.0 & & 386\\
+\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} \todo{RELEASE} & 108\\
+\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\
+\texttt{ltl2dgra} & 21.0 & & 274\\
 \bottomrule
@@
 
-%\begin{tikzpicture}
-%\begin{axis}[
-%	ymode=log,
-%	height=5cm,
-%	width=10cm,
-%	grid=major,
-%]
-%\addplot coordinates {
-%	(1, 0.3)
-%	(2, 0.8)
-%	(3, 1.5)
-%	(4, 2.7)
-%	(5, 3.9)
-%	(6, 4.1)
-%};
-%\addlegendentry{Delag}
-%\end{axis}
-%\end{tikzpicture}
+Looking at Figure~\ref{fig:quantile}, we ascertain that for every set, the
+reduction of most automata takes under a second.
+
+\begin{figure}[h]
+\caption{Performance of \telatko on different automata sets. The $x$ axis
+represents the number of reduced automata, the $y$ axis shows the time spent
+reducing (in seconds).}
+\label{fig:quantile}
+\centering
+\input{quantile.tex}
+\end{figure}
 

### 2022-03-04 paper/main.tex
section 3 finished!

--- 2022-03-03 paper/main.tex
@@
 \end{itemize}
-We write $\Phi_{i,C,K}$ when we want to emphasize that the particular
-formula $\Phi_{C,K}$ contains the version $\mcycle_i(\vec{e})$.
+We write $\Phi_{j,C,K}$ when we want to emphasize that the particular
+formula $\Phi_{C,K}$ contains the version $\mcycle_j(\vec{e})$.
 
@@
 
-Now we present a formula optimization based on the fact that every
-cycle is completely included in the transition set $\delta_S$ of some
-SCC $S$. Note that we cannot reduce the number of acceptance marks in
-each SCC separately as all SCCs have to share one acceptance formula.
-However, for each SCC we can build a $\Phi_{C,K}$ formula separately and
-then we can put all these formulas into one conjunction such that they
-share the variables $\vec{i},\vec{f}$.
-
-\todo{Honza je tu!}
-
-For a given automaton $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, we first
-compute all SCCs. Each SCC $S$ induces an automaton $(S,M,\Sigma,q',\delta_S,\varphi)$, where $q'$ is an arbitrary state
-of $S$.
-
-the optimization saying that So far, we have not use the fact that each cycle is A
-
-SCC-based approach
-
-Its first part $\xi_1(\vec{i},\vec{f})$ says that any clause does not
-contain both $\Inf k$ and $\Fin k$ for the same mark $k$ as a clause
-with both such terms is always unsatisfiable. Note that the subformula
-$\xi_1(\vec{i},\vec{f})$ can be safely omitted, but it speeds up the QBF solving.
+Now we present an optimized formula construction based on the fact
+that every cycle is completely included in the transition set
+$\delta_S$ of some SCC $S$. Note that we cannot reduce the number of
+acceptance marks in each SCC separately as all SCCs have to share the
+same acceptance formula after the reduction. However, for each SCC we
+can build the formula $\Phi_{C,K}$ separately and then we can put all
+these formulas into one conjunction such that they share the variables
+$\vec{i},\vec{f}$.
+
+Now we describe the construction of optimized QBF formulas
+$\Phi^\mathit{opt}_{j,C,K}$ formally. Given an automaton
+$\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$ and its SCC $S$, by $\mA^S$ we
+denote the automaton corresponding to the part of $\mA$ delimited by
+$S$, i.e., $\mA^S=(S,M,\Sigma,q',\delta_S,\varphi)$ where $q'$ is an
+arbitrary state of $S$. Further, let $\vec{n}_S$ denote the vector of
+variables $n_{t,k}$ where $t\in\delta_S$ and $k\in N_k$. By
+$\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})$ we denote the formula
+$\Phi_{j,C,K}$ constructed for $\mA^S$ as presented above.
+Finally, we define the optimized QBF formula as
+\[
+  \Phi^\mathit{opt}_{j,C,K}(\vec{n}',\vec{i},\vec{f})=\bigwedge_{\textrm{SCC }S}\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})
+\]
+where $\vec{n}'$ is the vector of all varibles contained in vectors
+$\vec{n}_S$ for SCCs $S$.
+
+One effect of this optimization is that the formula
+$\Phi^\mathit{opt}_{j,C,K}$ contains no variables $n_{t,k}$ %and $e_t$
+for transitions $t$ that do not lead between states of the same SCC.
+The acceptance marks on such a transition $t$ do not affect the
+acceptance of any run as $t$ appears at most once on each run. For
+these transitions $t$, we can define $\rem(t)$ such that
+$\mks(\rem(t))=\emptyset$.
 

### 2022-03-04 paper/main.tex
section 4 completed

@@
 %\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
+\usepackage[noend]{algorithm2e}
 
@@
 \section{Reduction algorithm}\label{sec:algo}
-Jak se nastavuji inicialni parametry a jak postupne klesaji.
-Jak prechazim od nizsich levelu k vyssim.
-Vyuzivame incremental QBF solving?
-
-\todo{Nasleduje stary text, ale pozor, obrazek je novy!}
-Now we describe our simplification algorithm taking as input an arbitrary TELA $\mathcal{A}$ with an acceptance condition $\psi$, required level of simplification $L$ and timeout period $T$.
-\begin{enumerate}
-    \item Convert acceptance condition $ \psi$ to DNF.
-    \item Apply the algorithm for simplification Level 1 described in Section \ref{sec:standard} on $\mathcal{A}$.
-    \item If $L = 1$ return automaton $\mathcal{A}$.
-     \item The initial value of $C$ is the number of clauses in $\psi$. Similarly, the initial value of $K$ is the number of acceptance sets of $\psi$ decreased by 1.
-    \item Encode $\mathcal{A}$ into QBF formula $\varphi$ according to the required level of simplification $L$.
-    \item Run a QBF-solver on $\varphi$ for $T$ seconds.
-    \begin{enumerate}
-        \item If the QBF-solver evaluates $\varphi$ as UNSAT, return automaton $\mathcal{A}'$.
-        \item If the timeout $T$ expires before the QBF-solver evaluates $\varphi$, return automaton $\mathcal{A}'$.
-        \item If the QBF-solver evaluates $\varphi$ as SAT. Decode the assignment into next TELA $\mathcal{A}'$.
-        \begin{enumerate}
-            \item If $K=1$, try to reduce $\varphi$ to $\true$ or $\false$, the simplest form of acceptance condition and return $\mathcal{A}$.
-            \item Decrease $K$ by 1 and go to (b).
-
-    \end{enumerate}
-
-    \end{enumerate}
-\end{enumerate}
-
-\todo{kde je optimalni misto, kam napsat:}
-For simplicity, we do not display letters on the edges of the automata in Figures. We can afford to omit them because they are not important for our simplification techniques.
-
-\begin{itemize}
-\item obecny popis konstrukce formule
-\item (level 3)
-\item level 4
-\end{itemize}
+
+This section explains how we use the QBF formulas constructed in the
+previous section to reduce the number of acceptance marks in TELA
+automata. First we describe a \emph{single-level} reduction, which
+uses only a single kind of QBF formulas. More precisely, we talk about
+\emph{level~1} reduction if $\Phi^\mathit{opt}_{1,C,K}$ is
+used. Analogously, we talk about \emph{level~2} and \emph{level~3}
+reduction when $\Phi^\mathit{opt}_{2,C,K}$ or
+$\Phi^\mathit{opt}_{3,C,K}$ is used, respectively.
+
+\begin{algorithm}[t]
+  \DontPrintSemicolon
+  \SetKwFunction{FextendBackward}{extendBackward}
+  \SetKwProg{PI}{Procedure}{}{}
+  \PI{$\mathit{SingleLevelReduction}(\mA,j,\mathit{reduceC})$}{
+    \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, level $j\in\{1,2,3\}$, $\mathit{reduceC}\in\{\true,\false\}$}
+    \KwOut{an equivalent TELA with the same structure as $\mA$ and with at most as many acceptance marks as in $\mA$} %with the same or lower number of acceptance~marks}
+    \;
+    \If{$L(\mA)=\emptyset$}{
+%      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'=\{(p,a,\emptyset,q)\mid(p,a,M',q)\in\delta\}$
+      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all acceptance marks removed
+    }
+    $C_\mA\gets$ the number of clauses in the acceptance formula of $\mA$ transformed to DNF\;
+    $K_\mA\gets$ the number of acceptance marks in $\mA$\;
+    $C\gets C_\mA$\;
+    $K\gets K_\mA$\;
+    $\mathit{stop}\gets\false$\;
+    \While{$\neg\mathit{stop}~\wedge~K>0$}{
+      $\Phi\gets\Phi^\mathit{opt}_{j,C,K{-}1}$ for $\mA$\;
+      \leIf{$\mathit{satisfiable}(\Phi)$}{$K\gets K{-}1$}{$\mathit{stop}\gets\true$}
+    }
+    \If{$\mathit{reduceC}$}{
+      $\mathit{stop}\gets\false$\;
+      \While{$\neg\mathit{stop}~\wedge~C>1$}{
+        $\Phi\gets\Phi^\mathit{opt}_{j,C-1,K}$ for $\mA$\;
+        \leIf{$\mathit{satisfiable}(\Phi)$}{$C\gets C{-}1$}{$\mathit{stop}\gets\true$}
+      }
+    }
+    \eIf{$K=K_\mA~\wedge~C=C_\mA$}{
+      \KwRet $\mA$
+    }{
+      compute $\rem$ and $\psi$ from a model of $\Phi^\mathit{opt}_{j,C,K}$\;
+      \KwRet $(Q,N_K,\Sigma,q_I,\rem(\delta),\psi)$
+    }
+  }\;
+  \caption{The single-level reduction procedure}
+  \label{alg:single}
+\end{algorithm}
+
+The reduction procedure is given in Algorithm~\ref{alg:single}. Besides
+the reduction of acceptance marks, the algorithm also reduces the
+number of clauses in acceptance formula if the last argument
+$\mathit{reduceC}$ is set to $\true$. In the first step, the procedure
+checks whether the input automaton represents the empty language and
+if this is the case, it returns the same automaton with acceptance
+conditon $\false$ and with all acceptance marks removed. Otherwise it
+gradually decreases the number of marks until 0 is reached or the QBF
+solver behind the function $\mathit{satisfiable}(\Phi)$ fails to
+reduce the number of marks, i.e., it claims unsatisfiability of $\Phi$
+or it runs out of resources. In the next step, if $\mathit{reduceC}$
+is set to $\true$ then the procedure gradually reduces the number of
+clauses. It stops if no equivalent automaton with less clauses is
+found or the number of clauses is reduced to one. Note that the
+acceptance condition with 0 clauses is $\false$ and this case is
+treated at the beginning of the procedure. Finally, if the procedure
+succeeds to reduce the number of marks or clauses, if constructs the
+modified automaton. Otherwise, it returns the original automaton.
+\todo{do sekce o implementaci napsat, ze namisto opakovaneho budovani
+  $\Phi_{j,C,K}$ puzivame inkrementalni pristup.}
 
@@
     \begin{scope}[initial angle=90,xshift=-3.5cm]
-      \node (l) at (1,1.3) {input:};
+      \node (l) at (1,1.3) {automaton $\mA$};
       \node[state,initial] (p) at (0,0) {};
@@
       \node[acclabel] at (1,-1.4)
-      {\small$(\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{3}{3})$};
+      {$(\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{3}{3})$};
     \end{scope}
     \begin{scope}[initial angle=90,xshift=3.5cm]
-      \node (l) at (1,1.3) {level 1:};
+      \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,1,\true)$}; %{level 1:};
       \node[state,initial] (p) at (0,0) {};
@@
       \node[acclabel] at (1,-1.4)
-      {\small$(\Fin\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Fin\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})$};
+      {$\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2}$};
     \end{scope}
     \begin{scope}[initial angle=90,xshift=-3.5cm,yshift=-4cm]
-      \node (l) at (1,1.3) {level 2:};
+      \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,2,\true)$}; %{level 2:};
       \node[state,initial] (p) at (0,0) {};
@@
       \node[acclabel] at (1,-1.4)
-      {\small$(\Fin\tacc{0}{0}\wedge\Inf\tacc{1}{1})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1})$};
+      {$\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}$};
     \end{scope}
     \begin{scope}[initial angle=90,xshift=3.5cm,yshift=-4cm]
-      \node (l) at (1,1.3) {level 3:};
+      \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,3,\true)$}; %{level 3:};
       \node[state,initial] (p) at (0,0) {};
@@
       \node[acclabel] at (1,-1.4)
-      {\small$\Fin\tacc{0}{0}$};
+      {$\Fin\tacc{0}{0}$};
     \end{scope}
@@
     reduction levels: an input automaton and the automata
-    obtained by reduction of level 1, level 2, and level 3.
-    \todo{Pozor, prvni klauzule v levelu 2 je trash!}}
+    obtained by reduction of level 1, level 2, and level 3.}
   \label{fig:example}
 \end{figure}
+
+Figure~\ref{fig:example} shows a very simple automaton $\mA$ and the
+three automata produced by calls of
+$\mathit{SingleLevelReduction}(\mA,j,\true)$ for $j\in\{1,2,3\}$. The
+figure clearly illustrates that the higher level of reduction we use,
+the more acceptance marks can be reduced. On the other side, lower
+levels are typically faster. The best results can be often achieved by
+combining reductions of all levels. We call this aproach
+\emph{multi-level} reduction. It is a straightforward sequential
+application of the three levels, see Algorithm~\ref{alg:multi}.
+
+\begin{algorithm}[t]
+  \DontPrintSemicolon
+  \SetKwFunction{FextendBackward}{extendBackward}
+  \SetKwProg{PI}{Procedure}{}{}
+  \PI{$\mathit{MultiLevelReduction}(\mA,\mathit{reduceC})$}{
+    \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$ and $\mathit{reduceC}\in\{\true,\false\}$}
+    \KwOut{an equivalent TELA with the same structure as $\mA$ and with at most as many acceptance marks as in $\mA$} %with the same or lower number of acceptance~marks}
+    \;
+    $\mA\gets\mathit{SingleLevelReduction}(\mA,1,\mathit{reduceC})$\;
+    $\mA\gets\mathit{SingleLevelReduction}(\mA,2,\mathit{reduceC})$\;
+    $\mA\gets\mathit{SingleLevelReduction}(\mA,3,\mathit{reduceC})$\;
+    \KwRet $\mA$
+  }\;
+  \caption{The multi-level reduction procedure}
+  \label{alg:multi}
+\end{algorithm}
 

### 2022-03-04 paper/main.tex
min tweak
experiments: progress

@@
 are looking for a new acceptance formula $\psi$ in DNF. Note that our
-reduction method can be easily adopted to look for $\psi$ of a
+reduction method can be easily adapted to look for $\psi$ of a
 different shape.
@@
 
-%\begin{table}[t]
-%\centering
-%\begin{tabular}{c|ccccc|}
-%\multicolumn{1}{c}{} & 3 & 4 & 5 & 6 & \multicolumn{1}{c}{7} \\
-%%\specialrule{2pt}{0pt}{0pt}
-%\cline{2-6}
-%3 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \cellcolor{blue!25}\makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-%4 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-%5 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-%6 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-%7 & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} & \makecell{\tiny{12} \\ \tiny{34}} \\
-%\cline{2-6}
-%\end{tabular}
-%\end{table}
-
 \begin{table}[h]
@@
 \input{table-data.tex}
+\end{table}
+
+Table~\ref{tab:heatmap} shows the number of automata coming from all four sets
+for every input and output number of acceptance marks. The results of the
+multi-level reduction technique are presented. The table indicates that in most
+cases only 1 or 2 marks can be saved. However, some of the automata with
+more than 7 acceptance marks could be, in theory, probably reduced more as
+the query to the QBF solver mostly remains unanswered in the time limit.
+Moreover, some of the results are promising, e.g. in 17 cases, we have
+successfully reduced 8 or more acceptance marks to only 4.
+
+\begin{table}[t]
+\caption{The effect of reduction.
+A cell on coordinates $(x,y)$ contains the number of automata that have been
+reduced from $x$ to $y$ acceptance marks. If the cell contains a sum of two
+numbers, the latter represents number of automata where the last query to QBF
+solver has not been answered in the time limit.}
+\label{tab:heatmap}
+\setlength{\tabcolsep}{4pt}
+\centering
+\input{table-heatmap.tex}
 \end{table}

### 2022-03-04 paper/main.tex
TELA automata

@@
 conditions. Further, algorithms handling TELA can automatically handle
-all automata with traditional acceptance conditions. TELA automata can
+all automata with traditional acceptance conditions. TELA can
 be obtained for example by translating LTL formulas by

### 2022-03-04 paper/main.tex
incremental approach in section 4

--- 2022-03-04 paper/main.tex
@@
 modified automaton. Otherwise, it returns the original automaton.
-\todo{do sekce o implementaci napsat, ze namisto opakovaneho budovani
-  $\Phi_{j,C,K}$ puzivame inkrementalni pristup.}
+
+Note that the algorithm can be also reformulated to use an incremental
+approach instead of building a new formula $\Phi$ in each iteration.
+The incremental version of the first \textbf{while} loop builds the formula
+$\Phi=\Phi^\mathit{opt}_{j,C,K{-}1}$ only in the first iteration. In
+each subsequent iteration, it extends the formula with a condition
+saying that one more mark is not used in the automaton, i.e., the mark
+is neither on edges, nor in the acceptance formula. For example, if we
+want to say that mark $k\in N_K$ is not used in the automaton, we replace
+$\Phi$ by
+\[
+  \Phi~\wedge~\bigwedge_{t\in\delta}\neg n_{t,k}~\wedge\bigwedge_{c\in\{1,2,\ldots C\}}(\neg i_{c,k}\wedge\neg f_{c,k}).
+\]
+The second \emph{while} can transformed to an incremental version in a
+similar way. The advantage of the incremental approach comes from the
+fact that some QBF solvers can decide an extended formula faster as
+they profit from the information computed when solving the original
+formula.
 
@@
 
-  \caption{An example illustrating the results of the three basic
-    reduction levels: an input automaton and the automata
-    obtained by reduction of level 1, level 2, and level 3.}
+  \caption{An example illustrating the results of the three
+    single-level reductions: an input automaton $\mA$ and the automata
+    obtained by reducing it with level 1, level 2, and level 3.}
   \label{fig:example}

### 2022-03-04 paper/table-data.tex
table-data.tex fix times
remove automata with 0-1 acc. marks

@@
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\389/404 solved aut.\\538 acc. marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\79/89 solved aut.\\236 acc.~marks}}
 &
 level 1 &
-421 & 21.75 & 419.2 & 389 & 0
+236 & 0.0 & 269.4 & 79 & 0
 \\
@@
 level 2 &
-421 & 21.75 & 14.13 & 389 & 0
+236 & 0.0 & 284.4 & 79 & 0
 \\
@@
 level 3 &
-412 & 23.42 & 232.98 & 382 & 7
+227 & 3.8 & 517.4 & 72 & 7
 \\
@@
 multi &
-412 & 23.42 & 686.96 & 383 & 6
+227 & 3.8 & 537.4 & 73 & 6
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\371/376 solved aut.\\562 acc. marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\103/108 solved aut.\\419 acc.~marks}}
 &
 level 1 &
-546 & 2.85 & 783.19 & 353 & 18
+403 & 3.8 & 658.2 & 85 & 18
 \\
@@
 level 2 &
-548 & 2.49 & 80.53 & 351 & 20
+405 & 3.3 & 738.9 & 83 & 20
 \\
@@
 level 3 &
-540 & 3.91 & 163.15 & 346 & 25
+397 & 5.3 & 901.5 & 78 & 25
 \\
@@
 multi &
-537 & 4.45 & 2295.88 & 346 & 25
+394 & 6.0 & 2170.6 & 78 & 25
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\388/397 solved aut.\\766 acc. marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\153/162 solved aut.\\618 acc.~marks}}
 &
 level 1 &
-553 & 27.81 & 898.88 & 370 & 18
+435 & 29.6 & 788.4 & 135 & 18
 \\
@@
 level 2 &
-563 & 26.5 & 99.36 & 367 & 21
+445 & 28.0 & 887.7 & 132 & 21
 \\
@@
 level 3 &
-569 & 25.72 & 355.55 & 357 & 31
+451 & 27.0 & 1243.3 & 122 & 31
 \\
@@
 multi &
-541 & 29.37 & 2538.7 & 361 & 27
+423 & 31.6 & 2427.8 & 126 & 27
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\378/386 solved aut.\\1084 acc. marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\266/274 solved aut.\\972 acc.~marks}}
 &
 level 1 &
-616 & 43.17 & 828.84 & 362 & 16
+606 & 37.7 & 774.2 & 250 & 16
 \\
@@
 level 2 &
-627 & 42.16 & 132.48 & 358 & 20
+617 & 36.5 & 906.6 & 246 & 20
 \\
@@
 level 3 &
-627 & 42.16 & 194.59 & 352 & 26
+617 & 36.5 & 1101.0 & 240 & 26
 \\
@@
 multi &
-606 & 44.1 & 2339.83 & 352 & 26
+596 & 38.7 & 2285.3 & 240 & 26
 \\

### 2022-03-04 paper/main.tex
implement

--- 2022-03-04 paper/main.tex
@@
 modified automaton. Otherwise, it returns the original automaton.
-
-Note that the algorithm can be also reformulated to use an incremental
-approach instead of building a new formula $\Phi$ in each iteration.
-The incremental version of the first \textbf{while} loop builds the formula
-$\Phi=\Phi^\mathit{opt}_{j,C,K{-}1}$ only in the first iteration. In
-each subsequent iteration, it extends the formula with a condition
-saying that one more mark is not used in the automaton, i.e., the mark
-is neither on edges, nor in the acceptance formula. For example, if we
-want to say that mark $k\in N_K$ is not used in the automaton, we replace
-$\Phi$ by
-\[
-  \Phi~\wedge~\bigwedge_{t\in\delta}\neg n_{t,k}~\wedge\bigwedge_{c\in\{1,2,\ldots C\}}(\neg i_{c,k}\wedge\neg f_{c,k}).
-\]
-The second \emph{while} can transformed to an incremental version in a
-similar way. The advantage of the incremental approach comes from the
-fact that some QBF solvers can decide an extended formula faster as
-they profit from the information computed when solving the original
-formula.
+\todo{do sekce o implementaci napsat, ze namisto opakovaneho budovani
+  $\Phi_{j,C,K}$ puzivame inkrementalni pristup.}
 
@@
 
-  \caption{An example illustrating the results of the three
-    single-level reductions: an input automaton $\mA$ and the automata
-    obtained by reducing it with level 1, level 2, and level 3.}
+  \caption{An example illustrating the results of the three basic
+    reduction levels: an input automaton and the automata
+    obtained by reduction of level 1, level 2, and level 3.}
   \label{fig:example}
@@
 to generate, simplify and translate LTL formulas and to compare and simplify
-automata. \telatko is implemented in Python.
+automata.
+\telatko is implemented in Python and has a CLI interface which allows users to specify details about a run.
+The tool expects file containing automata in HOA format ~\cite{babiak.15.cav} and processes the input file, the result is saved in the specified output file again in HOA format.
+If there is no output file given, \telatko prints the result to $\mathit{stdout}$.
+The switch \texttt{-L} specifies the level of simplification, \texttt{-I} turns on the incremental solving and \texttt{-T} specifies the timeout of Z3-solver in seconds.
+The default settings is the incremental version of \emph{multi-level} (see Algorithm~\ref{alg:multi}).
 

### 2022-03-04 paper/main.tex
experiments: polished

--- 2022-03-04 paper/main.tex
@@
 and manipulation, and the Z3 Theorem Prover~\cite{demoura.08.tacas} to solve
-the satisfiability of QBF formulas. Moreover, in the experimental evaluation
-we use several tools from the Spot library, i.e. \texttt{genltl},
-\texttt{ltlfilt}, \texttt{ltl2tgba}, \texttt{autcross} and \texttt{autfilt},
-to generate, simplify and translate LTL formulas and to compare and simplify
-automata. \telatko is implemented in Python.
+the satisfiability of QBF formulas. \telatko is implemented in Python.
 
@@
 the LTL formulas after filtering.
+
+We have used multiple other helper tools in the experimental evaluation.
+Alongside \texttt{genltl}, we have used \texttt{ltlfilt} to simplify the
+formulas and to remove duplicates, and \texttt{autcross} to compare automata and
+measure the running times.
 
@@
 
-Looking at Figure~\ref{fig:quantile}, we ascertain that for every set, the
-reduction of most automata takes under a second.
+Figure~\ref{fig:quantile} presents the time spent by reducing the individual
+automata sets. The chart shows a pleasing finding that for every set, most
+automata are reduced in under a second and the high cummulative running times
+are caused by few complicated automata.
 

### 2022-03-04 paper/main.tex

--- 2022-03-04 paper/main.tex
@@
 and manipulation, and the Z3 Theorem Prover~\cite{demoura.08.tacas} to solve
-the satisfiability of QBF formulas. Moreover, in the experimental evaluation
-we use several tools from the Spot library, i.e. \texttt{genltl},
-\texttt{ltlfilt}, \texttt{ltl2tgba}, \texttt{autcross} and \texttt{autfilt},
-to generate, simplify and translate LTL formulas and to compare and simplify
-automata.
-\telatko is implemented in Python and has a CLI interface which allows users to specify details about a run.
+the satisfiability of QBF formulas. \telatko is implemented in Python
+and has a CLI interface which allows users to specify details about a run.
 The tool expects file containing automata in HOA format ~\cite{babiak.15.cav} and processes the input file, the result is saved in the specified output file again in HOA format.

### 2022-03-04 paper/main.tex
conclusion started
slight polish

@@
 \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
-\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} \todo{RELEASE} & 108\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 108\\
 \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\
@@
 \section{Conclusions}\label{sec:conclusion}
+
+We have presented a QBF-based method that reduces the number of
+acceptance marks in transition-based Emerson-Lei automata. The method
+has ben implemented and our experimental results show that
 

### 2022-03-04 paper/main.tex
section 5 polished

@@
 
-The presented reduction algorithms have been implemented in a tool called
-\telatko. It uses the Spot library~\cite{duret.16.atva2} for automata parsing
-and manipulation, and the Z3 Theorem Prover~\cite{demoura.08.tacas} to solve
-the satisfiability of QBF formulas. \telatko is implemented in Python
-and has a CLI interface which allows users to specify details about a run.
-The tool expects file containing automata in HOA format ~\cite{babiak.15.cav} and processes the input file, the result is saved in the specified output file again in HOA format.
-If there is no output file given, \telatko prints the result to $\mathit{stdout}$.
-The switch \texttt{-L} specifies the level of simplification, \texttt{-I} turns on the incremental solving and \texttt{-T} specifies the timeout of Z3-solver in seconds.
-The default settings is the incremental version of \emph{multi-level} (see Algorithm~\ref{alg:multi}).
-
-\todo{Jake parametry ten tool ma? Co potrebuje k behu? Kde je dostupny? Pod jakou licenci?}
+The presented reduction algorithms have been implemented in a tool
+called \telatko. The tool is implemented in Python 3 and uses the Spot
+library~\cite{duret.16.atva2} for automata parsing and manipulation
+and the theorem prover Z3~\cite{demoura.08.tacas} to solve the
+satisfiability of QBF formulas. \telatko is available at
+\url{https://gitlab.fi.muni.cz/xschwar3/telatko} under the GNU GPLv3
+license. The tool can be executed by the command 
+\begin{center}
+  \texttt{telatko -F <input.hoa> [-L }$j$\texttt{] [-C] [-I] [-T }$t$\texttt{] [-O <output.hoa>]}
+\end{center}
+where
+\begin{description}
+\item[\texttt{-F <input.hoa>}] specifies the file with the input automaton in HOA format~\cite{babiak.15.cav}, 
+\item[\texttt{-L }$j$] specifies the reduction level; if omitted, the multi-level reduction is used,
+\item[\texttt{-C}] switches on the reduction of the number of clauses
+  after the number of marks is reduced (it corresponds to
+  $\mathit{reduceC}=\true$ in Algortihms~\ref{alg:single}
+  and~\ref{alg:multi}),
+\item[\texttt{-I}] switches on the incremental version,
+\item[\texttt{-T }$t$] sets the timeout for QBF query to $t$ seconds (the default value is 50 seconds), 
+\item[\texttt{-O <output.hoa>}] specifies the output file; if omitted,
+  the produced automaton is sent to \textit{stdout} in the HOA format.
+\end{description}
 
@@
 \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
-\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 108\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} \todo{RELEASE} & 108\\
 \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\

### 2022-03-04 paper/main.tex
abstract polished

@@
   % acceptance marks.
-  In fact, we present three formula constructions
-  trading speed for higher reduction
-  efficiency. We evaluated our reduction technique on TELA produced by
-  state-of-the-art tools Delag, ltl3tela, Rabinizer~4, and Spot. The
-  technique reduced some acceptance marks in automata produced by all
-  these tools. In the case of Rabinizer~4, the cumulative number of
-  acceptance marks was reduced to less than one half. \todo{Sazet ltl3tela a jine tooly vsude jako \texttt{ltl3tela} apod?}
+  In fact, we present three formula constructions trading speed for
+  higher reduction efficiency. We evaluated our reduction technique on
+  TELA produced by state-of-the-art tools of the libraries Owl and
+  Spot and by the tool \texttt{ltl3tela}. The technique reduced some
+  acceptance marks in automata produced by all the tools. The achieved
+  reduction on automata produced by tools of the Owl library was
+  between 30 and 40\%.
 \end{abstract}

### 2022-03-04 paper/main.tex
progress

@@
 be obtained for example by translating LTL formulas by
-Delag~\cite{muller.17.gandalf}, ltl3tela~\cite{major.19.atva},
-Rabinizer~4~\cite{kretinsky.18.cav}, or
-Spot~\cite{duret.16.atva2}. There are also some algorithms processing
+\texttt{ltl2dela} (known as Delag)~\cite{muller.17.gandalf}),
+\texttt{ltl3tela}~\cite{major.19.atva},
+\texttt{ltl2dgra} (known as Rabinizer~4)~\cite{kretinsky.18.cav},
+or \texttt{ltl2tgba} which is a part of the Spot library~\cite{duret.16.atva2}.
+There are also some algorithms processing
 these automata, for example the emptiness check~\cite{baier.19.atva}
@@
 and Spot~\cite{duret.16.atva2}.  While the reduction is relatively
-modest for TELA produced by ltl3tela and Spot, it is substantial for
-automata produced by Rabinizer~4. \todo{A jak vyjde Delag? Fixnout i
-  Abstract.}
+modest for TELA produced by \texttt{ltl3tela} and Spot,
+it is substantial for automata produced by tools of the Owl library.
 
@@
 \midrule
+\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\
+\texttt{ltl2dgra} & 21.0 & & 274\\
 \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
 \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 108\\
-\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\
-\texttt{ltl2dgra} & 21.0 & & 274\\
 \bottomrule

### 2022-03-04 paper/main.tex
intro: order of tools

@@
 automata produced by Delag~\cite{muller.17.gandalf},
-ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav},
-and Spot~\cite{duret.16.atva2}.  While the reduction is relatively
+Rabinizer~4~\cite{kretinsky.18.cav} (both included in the Owl library),
+Spot~\cite{duret.16.atva2}
+and \texttt{ltl3tela}~\cite{major.19.atva}.  While the reduction is relatively
 modest for TELA produced by \texttt{ltl3tela} and Spot,

### 2022-03-04 paper/main.tex
conclusion

@@
 acceptance marks in transition-based Emerson-Lei automata. The method
-has ben implemented and our experimental results show that
-
-chceme sem napsat, ze QBF-based reduction se da pouzit napr. k
-nalezeni akceptacni formule s $c$ klauzulemi a $m$ znackami v CNF,
-pripadne k nalezeni ekvivalentniho paritniho automat s danou maximalni
-paritou $p$ a stejnou strukturou. Nebo zkusit nalezt $\Fin$-less
-acceptance formuli.
-
-Redukce netrva vecne, da se prakicky pouzit a zkracenim timeoutu na
-QBF-solver se da zrychlit (tradeoff s ucinnosti). Prinejmensim u
-Rabinizeru 4 je velmi dobra sance, ze se to vyplati. Pridat do
-experimentu nejaky graf, ktery ukaze, jak rychle se nektere automaty
-redukovaly na minimum?
+has been implemented and our experimental results show that it is possible
+to use the QBF solvers to find equivalent acceptance conditions of TELA
+without changes to their structure.
+The reduction of acceptance marks is not the only purpose of the approach,
+one could possibly look for e.g. equivalent parity automata or automata without
+$\Fin$-marks, depending on the subsequent use of the automata.
+Even though the QBF queries can be time-consuming, in practice it can be well
+used and one can find a good trade-off between speed and efficiency by setting
+a reasonable timeout.
+
+In the future work, it could be examined how the reductions of the acceptance
+condition can contribute to further reduction of the state space by enabling
+various optimizations, e.g. simulations.
 

### 2022-03-04 paper/main.tex
section 6 polished

--- 2022-03-04 paper/main.tex
@@
 be obtained for example by translating LTL formulas by
-\texttt{ltl2dela} (known as Delag)~\cite{muller.17.gandalf}),
-\texttt{ltl3tela}~\cite{major.19.atva},
-\texttt{ltl2dgra} (known as Rabinizer~4)~\cite{kretinsky.18.cav},
-or \texttt{ltl2tgba} which is a part of the Spot library~\cite{duret.16.atva2}.
-There are also some algorithms processing
+Delag~\cite{muller.17.gandalf}, ltl3tela~\cite{major.19.atva},
+Rabinizer~4~\cite{kretinsky.18.cav}, or
+Spot~\cite{duret.16.atva2}. There are also some algorithms processing
 these automata, for example the emptiness check~\cite{baier.19.atva}
@@
 and Spot~\cite{duret.16.atva2}.  While the reduction is relatively
-modest for TELA produced by \texttt{ltl3tela} and Spot,
-it is substantial for automata produced by tools of the Owl library.
+modest for TELA produced by ltl3tela and Spot, it is substantial for
+automata produced by Rabinizer~4. \todo{A jak vyjde Delag? Fixnout i
+  Abstract.}
 
@@
 automata~\cite{ehlers.10.sat}, and deterministic generalized Büchi
-automata~\cite{baarir.14.forte}. Note that all these techniques are
+automata~\cite{baarir.14.forte}. Note that these techniques are
 usually very slow and their authors typically suggest to use them only
@@
 
-We have used the tool \texttt{genltl} from Spot to generate LTL formulas
-from literature and from scalable patterns. Using the values from 1 to 5 for
-the parametrized formulas, we have generated 407 unique LTL formulas. These
-have been translated to deterministic or nondeterministic TELA using four
-state-of-the-art translators \texttt{ltl2tgba} (from the Spot library),
-\texttt{ltl3tela}, \texttt{ltl2dela} (known as Delag), and \texttt{ltl2dgra}
-(known as Rabinizer 4). The generated automata sets contain less than 407
-automata as not all of the formulas are successfully translated due to timeouts.
-We have also removed automata with 0 or 1 acceptance marks as there is little
-point reducing these.
-Table \ref{tab:ltltranslators} shows the exact versions, the homepage of the
-used tools and the number of automata produced by the given tool from
-the LTL formulas after filtering.
-
-We have used multiple other helper tools in the experimental evaluation.
-Alongside \texttt{genltl}, we have used \texttt{ltlfilt} to simplify the
-formulas and to remove duplicates, and \texttt{autcross} to compare automata and
-measure the running times.
-
-Spot version 2.10.4 and Z3 version 4.8.13 have been used in the experimental
-evaluation.
-
-For each automata set (that is, the output of individual LTL translator), we
-have run \telatko with various reduction levels, always with the incremental
-solving turned on. We have again filtered out the automata for which any run
-of \telatko timed out after 120 seconds. We do this to make sure the numbers
-of acceptance marks can be meaningfully compared between different reduction
-levels of the same automata set.
-
-Table~\ref{tab:cumm-data} shows the number of marks in the input sets and after
-each reduction level, together with the reduction ratio and time spent by the
-individual reduction level. The columns \textit{solver response} and
-\textit{solver timeout} show the number of automata for which the last query to
-QBF solver has been successfully answered, or the solver timed out within a 30
-seconds limit to a single query, respectively. An unanswered QBF formula (i.e.
-the solver timed out) is considered unsatisfiable.
-
-\begin{table}[h]
+We have used the tool \texttt{genltl} of the Spot
+library~\cite{duret.16.atva2} to generate LTL formulas from
+literature. The parametrized formulas were generated for all
+combinations of parameter values from 1 to 5. Altogether, we have
+generated 407 unique LTL formulas. These have been translated to
+deterministic TELA by two state-of-the-art translators from the Owl
+library~\cite{kretinsky.18.atva} and to nondeterministic TELA by
+another two state-of-the-art translators, namely \texttt{ltl2tgba}
+from the Spot library~\cite{duret.16.atva2}) and
+\texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
+produced by individual tools contain less than 407 automata as not all
+of the formulas are successfully translated due to timeout of 60
+seconds. Further, we have removed automata with 0 or 1 acceptance
+marks as there is little point reducing these. Table
+\ref{tab:ltltranslators} shows the exact versions of used translators,
+the homepage of the used tools and the number of automata produced by
+the given tool from the LTL formulas after filtering.
+
+We have used multiple other helper tools in the experimental
+evaluation.  Alongside \texttt{genltl}, we have used \texttt{ltlfilt}
+to simplify the formulas and to remove duplicates, and
+\texttt{autcross} to compare automata and measure the running times.
+
+Spot version 2.10.4 and Z3 version 4.8.13 have been used in the
+experimental evaluation.
+
+For each automata set (that is, the output of an individual LTL
+translator), we have applied all single-level reductions and the
+multi-level reduction, always with incremental approach and without
+reduction of the number of clauses. The timeout for each QBF query was
+set to 30 seconds. We have again filtered out the automata for which
+any of the four \telatko runs timed out after 120 seconds. We do this
+to make sure the numbers of acceptance marks can be meaningfully
+compared between different reduction levels of the same automata set.
+
+Table~\ref{tab:cumm-data} shows the cummulative numbers of marks in
+the input automata sets and after each reduction, together with the
+reduction ratio and total time spent by the considered reduction. The
+columns \textit{solver response} and \textit{solver timeout} show the
+number of automata for which the last query to QBF solver has been
+successfully answered, or the solver timed out within a 30 seconds
+limit to a single query, respectively. The timeout of the last QBF
+query implies that the automaton could be potentially further reduced
+by the reduction with a longer time limit.
+
+\begin{table}[b]
 \caption{LTL to automata translators used for the experimental evaluation
@@
 \midrule
+\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 108\\
 \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\
 \texttt{ltl2dgra} & 21.0 & & 274\\
-\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
-\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 108\\
 \bottomrule
@@
 
-Table~\ref{tab:heatmap} shows the number of automata coming from all four sets
-for every input and output number of acceptance marks. The results of the
-multi-level reduction technique are presented. The table indicates that in most
-cases only 1 or 2 marks can be saved. However, some of the automata with
-more than 7 acceptance marks could be, in theory, probably reduced more as
-the query to the QBF solver mostly remains unanswered in the time limit.
-Moreover, some of the results are promising, e.g. in 17 cases, we have
-successfully reduced 8 or more acceptance marks to only 4.
+Table~\ref{tab:heatmap} shows the number of automata coming from all
+four sets that have a given number of original acceptance marks and a
+given number of acceptance marks after multi-level reduction. The
+table indicates that in many cases only 1 or 2 marks can be
+saved. However, the achieved reduction is substantial for some
+automata with a higher number of original acceptance marks. For
+example, in 17 cases, we have successfully reduced 8 or more
+acceptance marks to only 4. 
+% Moreover, there were many timeouts of the last QBF query for manof these automata.
+% Let us note that some that some reductions are the results are promising, e.g. in 17 cases, we have successfully
+% reduced 8 or more acceptance marks to only 4.
 
@@
 
-Figure~\ref{fig:quantile} presents the time spent by reducing the individual
-automata sets. The chart shows a pleasing finding that for every set, most
-automata are reduced in under a second and the high cummulative running times
-are caused by few complicated automata.
+Figure~\ref{fig:quantile} presents the time spent by multi-level
+reduction of automata in individual automata sets. The chart shows a
+pleasing finding that for every set, most automata are reduced in
+under a second and the high cummulative running times are caused by
+few complicated automata.
 
@@
 
-Pridat do experimentu nejaky graf, ktery ukaze, jak rychle se nektere
-automaty redukovaly na minimum? Nebo spis ukazat vliv timeoutu na
-redukce?  (zvysovat timeout od 1s do 50s a vzdy ukazat miru
-redukce). Ty redukce ukazovat nejak jinak nez jen celkovym cislem
-(boxplot? Scatterplot?).
-
-Zkusit upravit redukci tak, aby vysledkem bylo TGBA a zkusit dale
-redukovat akceptacni znacky TGBA automatu redukovanych Spotem?
-
-Chceme ukazovat, ze redukce znacek muze nekdy umoznit dalsi redukci stavu?
+% Pridat do experimentu nejaky graf, ktery ukaze, jak rychle se nektere
+% automaty redukovaly na minimum? Nebo spis ukazat vliv timeoutu na
+% redukce?  (zvysovat timeout od 1s do 50s a vzdy ukazat miru
+% redukce). Ty redukce ukazovat nejak jinak nez jen celkovym cislem
+% (boxplot? Scatterplot?).
+
+% Zkusit upravit redukci tak, aby vysledkem bylo TGBA a zkusit dale
+% redukovat akceptacni znacky TGBA automatu redukovanych Spotem?
+
+% Chceme ukazovat, ze redukce znacek muze nekdy umoznit dalsi redukci stavu?
 

### 2022-03-04 paper/main.tex

@@
 automata produced by Delag~\cite{muller.17.gandalf},
-ltl3tela~\cite{major.19.atva}, Rabinizer~4~\cite{kretinsky.18.cav},
-and Spot~\cite{duret.16.atva2}.  While the reduction is relatively
-modest for TELA produced by ltl3tela and Spot, it is substantial for
-automata produced by Rabinizer~4. \todo{A jak vyjde Delag? Fixnout i
-  Abstract.}
+Rabinizer~4~\cite{kretinsky.18.cav} (both included in the Owl library),
+Spot~\cite{duret.16.atva2}
+and \texttt{ltl3tela}~\cite{major.19.atva}.  While the reduction is relatively
+modest for TELA produced by \texttt{ltl3tela} and Spot,
+it is substantial for automata produced by tools of the Owl library.
 

### 2022-03-04 paper/main.tex
almost submitted for SAT 2022 (+some todo notes)

@@
 
-We have presented a QBF-based method that reduces the number of
+We have presented an application of QBF solving to reduction of the number of
 acceptance marks in transition-based Emerson-Lei automata. The method
@@
 to use the QBF solvers to find equivalent acceptance conditions of TELA
-without changes to their structure.
-The reduction of acceptance marks is not the only purpose of the approach,
-one could possibly look for e.g. equivalent parity automata or automata without
-$\Fin$-marks, depending on the subsequent use of the automata.
-Even though the QBF queries can be time-consuming, in practice it can be well
-used and one can find a good trade-off between speed and efficiency by setting
-a reasonable timeout.
+with less acceptance marks and without changes to automata structure.
+
+The reduction of acceptance marks is not the only application of the
+presented approach: one could possibly look for, e.g., equivalent
+parity automata or automata without $\Fin$-marks, depending on the
+subsequent use of the automata. Even though the QBF queries can be
+time-consuming, in practice it can be well used and one can find a
+good trade-off between speed and efficiency by choosing precision of
+the reduction and setting a reasonable timeout.
 
@@
 condition can contribute to further reduction of the state space by enabling
-various optimizations, e.g. simulations.
+various optimizations, e.g., simulations.
 
 %}}}
+
+\section{TODO}
+
+Do submission:
+\begin{itemize}
+\item do experimentu napsat, jaky hardware pouzivame (merit cas bez toho nedava smysl)
+\item spustit experimenty bez timeoutu na telatko (?a delsi timeout na QBF?)
+\item prelestit text v poslednich dvou kapitolach
+\end{itemize}
+
+\noindent
+Do budoucna:
+\begin{itemize}
+\item level 1 bez SCC optimalizace
+\item nahradit weak SCC jednou hranou pro neakceptujici a jinou pro akceptujici
+\item sladit konstrukci QBF formuli s implementaci
+\item podivat se znovu na uzitecnost a potrebnost optimalizaci typu ``klauzule neni false'' a ``alespon jedna klauzule je neprazdna''
+\item sladit popis algoritmu s implementaci
+\item v experimentech nejak sledovat vztah usetreneho casu a usetrenych znacek?
+\item zkusit aplikovat na jine typy automatu (treba TGBA nebo Rabin)?
+\item zkusit, zda/jak redukce znacek umozni redukovat stavovy prostor
+\end{itemize}
 

### 2022-04-21 paper/main.tex
automaton nameparagraph

@@
 
+The parameters of the last query of the QBF solver are encoded in the name of the automaton.
+
+The name of the automaton is in the following format:
+\begin{center}
+  \texttt{level\_marks\_solver}
+\end{center}
+where 
+\begin{itemize}
+    \item \texttt{level} $\in \{L1, L2, L3, T, F\}$ and it gives us the information about which level of simplification was applied (or in case of $T$ and $F$ if the acceptance formula can be $\true$ or $\false$).
+    \item \texttt{marks} $\in \mathbf{N}$ representing the number of marks we built the formula for.
+    \item \texttt{solver} $\in \{U, T, S\}$ where $U$ means that the query was unsatisfiable, $T$ means the solver timeouted and $S$ means that the simplification is satisfiable.
+\end{itemize}
+
+Note that we are only interested in the result of the last query, therefore we get the information that the simplification was satisfiable only if the result automaton has zero acceptance marks.
+Meaning that the acceptance condition is either $\true$ or $\false$.
+
 %}}}

### 2022-04-22 paper/table-data.tex
experiments (autcross -T 300) - worse numbers

@@
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\79/89 solved aut.\\236 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\158/162 solved aut.\\667 acc.~marks}}
 &
 level 1 &
-236 & 0.0 & 269.4 & 79 & 0
+487 & 27.0 & 1148.9 & 133 & 25
 \\
@@
 level 2 &
-236 & 0.0 & 284.4 & 79 & 0
+492 & 26.2 & 1262.6 & 132 & 26
 \\
@@
 level 3 &
-227 & 3.8 & 517.4 & 72 & 7
+499 & 25.2 & 1701.5 & 120 & 38
 \\
@@
 multi &
-227 & 3.8 & 537.4 & 73 & 6
+479 & 28.2 & 3521.4 & 124 & 34
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\103/108 solved aut.\\419 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\274/274 solved aut.\\1084 acc.~marks}}
 &
 level 1 &
-403 & 3.8 & 658.2 & 85 & 18
+706 & 34.9 & 1213.1 & 249 & 25
 \\
@@
 level 2 &
-405 & 3.3 & 738.9 & 83 & 20
+716 & 33.9 & 1416.9 & 246 & 28
 \\
@@
 level 3 &
-397 & 5.3 & 901.5 & 78 & 25
+716 & 33.9 & 1766.9 & 240 & 34
 \\
@@
 multi &
-394 & 6.0 & 2170.6 & 78 & 25
+695 & 35.9 & 3707.9 & 241 & 33
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\153/162 solved aut.\\618 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\81/89 solved aut.\\246 acc.~marks}}
 &
 level 1 &
-435 & 29.6 & 788.4 & 135 & 18
+246 & 0.0 & 640.7 & 81 & 0
 \\
@@
 level 2 &
-445 & 28.0 & 887.7 & 132 & 21
+246 & 0.0 & 691.7 & 80 & 1
 \\
@@
 level 3 &
-451 & 27.0 & 1243.3 & 122 & 31
+237 & 3.7 & 1101.5 & 73 & 8
 \\
@@
 multi &
-423 & 31.6 & 2427.8 & 126 & 27
+237 & 3.7 & 1170.3 & 74 & 7
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\266/274 solved aut.\\972 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\107/108 solved aut.\\463 acc.~marks}}
 &
 level 1 &
-606 & 37.7 & 774.2 & 250 & 16
+447 & 3.5 & 932.8 & 85 & 22
 \\
@@
 level 2 &
-617 & 36.5 & 906.6 & 246 & 20
+449 & 3.0 & 1017.0 & 83 & 24
 \\
@@
 level 3 &
-617 & 36.5 & 1101.0 & 240 & 26
+441 & 4.8 & 1183.9 & 78 & 29
 \\
@@
 multi &
-596 & 38.7 & 2285.3 & 240 & 26
+438 & 5.4 & 2973.9 & 78 & 29
 \\

### 2022-04-24 paper/table-data.tex
new data (RANGE=1..4, no timeout)

--- 2022-03-04 paper/table-data.tex
@@
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\79/89 solved aut.\\236 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\129/129 solved aut.\\523 acc.~marks}}
 &
 level 1 &
-236 & 0.0 & 269.4 & 79 & 0
+387 & 26.0 & 796.9 & 111 & 18
 \\
@@
 level 2 &
-236 & 0.0 & 284.4 & 79 & 0
+395 & 24.5 & 1057.1 & 108 & 21
 \\
@@
 level 3 &
-227 & 3.8 & 517.4 & 72 & 7
+393 & 24.9 & 9366.6 & 104 & 25
 \\
@@
 multi &
-227 & 3.8 & 537.4 & 73 & 6
+379 & 27.5 & 11486.6 & 104 & 25
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\103/108 solved aut.\\419 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\234/234 solved aut.\\882 acc.~marks}}
 &
 level 1 &
-403 & 3.8 & 658.2 & 85 & 18
+552 & 37.4 & 786.3 & 220 & 14
 \\
@@
 level 2 &
-405 & 3.3 & 738.9 & 83 & 20
+569 & 35.5 & 956.2 & 216 & 18
 \\
@@
 level 3 &
-397 & 5.3 & 901.5 & 78 & 25
+562 & 36.3 & 1276.4 & 211 & 23
 \\
@@
 multi &
-394 & 6.0 & 2170.6 & 78 & 25
+543 & 38.4 & 2525.3 & 211 & 23
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\153/162 solved aut.\\618 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\70/70 solved aut.\\198 acc.~marks}}
 &
 level 1 &
-435 & 29.6 & 788.4 & 135 & 18
+198 & 0.0 & 590.9 & 70 & 0
 \\
@@
 level 2 &
-445 & 28.0 & 887.7 & 132 & 21
+198 & 0.0 & 630.8 & 70 & 0
 \\
@@
 level 3 &
-451 & 27.0 & 1243.3 & 122 & 31
+189 & 4.5 & 2127.0 & 63 & 7
 \\
@@
 multi &
-423 & 31.6 & 2427.8 & 126 & 27
+189 & 4.5 & 2180.6 & 64 & 6
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\266/274 solved aut.\\972 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\90/91 solved aut.\\343 acc.~marks}}
 &
 level 1 &
-606 & 37.7 & 774.2 & 250 & 16
+328 & 4.4 & 486.5 & 78 & 12
 \\
@@
 level 2 &
-617 & 36.5 & 906.6 & 246 & 20
+330 & 3.8 & 567.0 & 75 & 15
 \\
@@
 level 3 &
-617 & 36.5 & 1101.0 & 240 & 26
+322 & 6.1 & 717.0 & 71 & 19
 \\
@@
 multi &
-596 & 38.7 & 2285.3 & 240 & 26
+319 & 7.0 & 1622.5 & 72 & 18
 \\

### 2022-04-24 paper/main.tex
progress in Experiments

@@
 literature. The parametrized formulas were generated for all
-combinations of parameter values from 1 to 5. Altogether, we have
-generated 407 unique LTL formulas. These have been translated to
+combinations of parameter values from 1 to 4. Altogether, we have
+generated 348 unique LTL formulas. These have been translated to
 deterministic TELA by two state-of-the-art translators from the Owl
@@
 another two state-of-the-art translators, namely \texttt{ltl2tgba}
-from the Spot library~\cite{duret.16.atva2}) and
+from the Spot library~\cite{duret.16.atva2} and
 \texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
-produced by individual tools contain less than 407 automata as not all
+produced by individual tools contain less than 348 automata as not all
 of the formulas are successfully translated due to timeout of 60
@@
 reduction of the number of clauses. The timeout for each QBF query was
-set to 30 seconds. We have again filtered out the automata for which
-any of the four \telatko runs timed out after 120 seconds. We do this
-to make sure the numbers of acceptance marks can be meaningfully
-compared between different reduction levels of the same automata set.
+set to 30 seconds.
+%We have again filtered out the automata for which
+%any of the four \telatko runs timed out after 120 seconds. We do this
+%to make sure the numbers of acceptance marks can be meaningfully
+%compared between different reduction levels of the same automata set.
+Additionally, we have filtered one automaton from the \texttt{ltl3tela} set
+that cannot be processed by \telatko due to an extensive use of system
+resources.
 
@@
 \midrule
-\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\
-\texttt{ltl2dgra} & 21.0 & & 274\\
-\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
-\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 108\\
+\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
+\texttt{ltl2dgra} & 21.0 & & 234\\
+\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 90\\
 \bottomrule
@@
 \caption{The cummulative number of acceptance marks before and after reduction
-for various automata sets. \textit{X/Y solved aut.} means that out of
-\textit{Y} successfully translated LTL formulas, every run of \telatko ended
-within 120 seconds time limit on \textit{X} automata. We sum up the acceptance
-marks only on these automata. The columns \textit{solver response} and
+for various automata sets.
+%\textit{X/Y solved aut.} means that out of
+%\textit{Y} successfully translated LTL formulas, every run of \telatko ended
+%within 120 seconds time limit on \textit{X} automata. We sum up the acceptance
+%marks only on these automata.
+The columns \textit{solver response} and
 \textit{solver timeout} refer to last call to the QBF solver (that is, the
@@
 automata with a higher number of original acceptance marks. For
-example, in 17 cases, we have successfully reduced 8 or more
+example, in 22 cases, we have successfully reduced 7 or more
 acceptance marks to only 4.

### 2022-04-25 paper/table-data.tex

@@
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\129/129 solved aut.\\523 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\129 automata\\523 acc.~marks}}
 &
@@
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\234/234 solved aut.\\882 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\234 automata\\882 acc.~marks}}
 &
@@
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\70/70 solved aut.\\198 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\70 automata\\198 acc.~marks}}
 &
@@
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\90/91 solved aut.\\343 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\90 automata\\343 acc.~marks}}
 &

### 2022-04-25 paper/main.tex
asi nejake male zmeny (netusim jake :-))
polishing of sections 1--3

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
+\documentclass[a4paper,UKenglish,cleveref,autoref,thm-restate]{lipics-v2021}
 %\documentclass[a4paper,UKenglish,cleveref, autoref, anonymous, thm-restate]{lipics-v2021}
@@
 \newcommand{\Nset}{{\mathbb{N}_0}}
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
+\newcommand{\todo}[1]{\marginpar{\textcolor{red}{#1}}}
 %\def\Inf{\ensuremath{\mathsf{Inf}}}
@@
 \begin{abstract}
-  \emph{Transition-based Emerson-Lei automata (TELA)} become a popular
+  \emph{Transition-based Emerson-Lei automata (TELA)} became a popular
   formalism that generalizes many traditional kinds of automata over
@@
   % acceptance marks.
-  In fact, we present three formula constructions trading speed for
-  higher reduction efficiency. We evaluated our reduction technique on
-  TELA produced by state-of-the-art tools of the libraries Owl and
-  Spot and by the tool \texttt{ltl3tela}. The technique reduced some
-  acceptance marks in automata produced by all the tools. The achieved
-  reduction on automata produced by tools of the Owl library was
-  between 30 and 40\%.
+  In fact, we present three formula constructions
+  % trading speed for higher reduction efficiency.
+  with increasing reduction potential and formula complexity, and thus
+  also decreasing speed of their satisfiability solving. We evaluated
+  our reduction technique on TELA produced by state-of-the-art tools
+  of the libraries Owl and Spot and by the tool \texttt{ltl3tela}. The
+  technique reduced some acceptance marks in automata produced by all
+  the tools. On automata with more than one acceptance mark produced
+  by tools Delag and Rabinizer~4 of the Owl library, our technique reduced
+  reduced 27.5\% and 38.4\% of acceptance marks, respectively. \todo{zkontrolovat!}
+  % The achieved
+  % reduction of acceptance marks on automata produced by tools of the Owl library was
+  % between 25 and 40\%.
 \end{abstract}
@@
 
-In 1987, Emerson and Lei~\cite{emerson.87.scp} introduced automata
-over infinite words where acceptance conditions are arbitrary
-combinations of acceptance primitives saying that a certain set of
-states should be visited finitely often or infinitely often. In 2015,
-the same kind of acceptance condition was described in the \emph{Hanoi
-  omega-automata format (HOAF)}~\cite{babiak.15.cav}. The only
-difference is that the acceptance primitives talk about finitely or
-infinitely often visited acceptance marks rather than sets of states.
-Acceptance marks are placed on transitions and each mark identifies
-the set of transitions containing this mark. Hence, these automata are
-called \emph{transition-based Emerson-Lei automata (TELA)} and they
+Automata over infinite words like Büchi, Rabin, Streett, or parity
+automata play a crucial role in many algorithms related to the
+concurrency theory. In particular, they are used in specification,
+verification, analysis, monitoring, and synthesis of various systems
+with infinite behaviour. In 1987, Emerson and
+Lei~\cite{emerson.87.scp} introduced automata over infinite words
+where acceptance conditions are arbitrary combinations of acceptance
+primitives saying that a certain set of states should be visited
+finitely often or infinitely often. In 2015, the same kind of
+acceptance condition was described in the \emph{Hanoi omega-automata
+  format (HOAF)}~\cite{babiak.15.cav}. The only difference is that the
+acceptance primitives talk about finitely or infinitely often visited
+acceptance marks rather than sets of states.  Acceptance marks are
+placed on transitions and each mark identifies the set of transitions
+containing this mark. Hence, these automata are called
+\emph{transition-based Emerson-Lei automata (TELA)} and they
 generalize many traditional kinds of automata over infinite words
@@
 conditions. Further, algorithms handling TELA can automatically handle
-all automata with traditional acceptance conditions. TELA can
-be obtained for example by translating LTL formulas by
-\texttt{ltl2dela} (known as Delag)~\cite{muller.17.gandalf}),
-\texttt{ltl3tela}~\cite{major.19.atva},
-\texttt{ltl2dgra} (known as Rabinizer~4)~\cite{kretinsky.18.cav},
-or \texttt{ltl2tgba} which is a part of the Spot library~\cite{duret.16.atva2}.
-There are also some algorithms processing
-these automata, for example the emptiness check~\cite{baier.19.atva}
-or translation of TELA to parity
+all automata with traditional acceptance conditions. TELA can be
+obtained for example by translating LTL formulas by \texttt{ltl2dela}
+(known as Delag)~\cite{muller.17.gandalf},
+\texttt{ltl3tela}~\cite{major.19.atva}, \texttt{ltl2dgra} (known as
+Rabinizer~4)~\cite{kretinsky.18.cav}, or \texttt{ltl2tgba} which is a
+part of the Spot library~\cite{duret.16.atva2}.  There are also
+algorithms processing these automata, for example the emptiness
+check~\cite{baier.19.atva} or translation of TELA to parity
 automata~\cite{renkin.20.atva,casares.22.tacas}.
@@
 and \texttt{ltl3tela}~\cite{major.19.atva}.  While the reduction is relatively
-modest for TELA produced by \texttt{ltl3tela} and Spot,
-it is substantial for automata produced by tools of the Owl library.
+modest on TELA produced by \texttt{ltl3tela} and Spot,
+it is substantial on automata produced by the tools of the Owl library.
 
@@
 
-A run $\pi$ of $\mA$ over an infinite word
-$u = u_{0}u_{1}u_{2}\dots \in \Sigma^{\omega}$ is an infinite sequence
+A \emph{run} $\pi$ of $\mA$ over an infinite word
+$u = u_{0}u_{1}u_{2}\ldots \in \Sigma^{\omega}$ is an infinite sequence
 of adjacent transitions
-$\pi=(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots\in\delta^{\omega}$ where
+$\pi=(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\ldots\in\delta^{\omega}$ where
 $q_0=q_I$. Let $\minf(\pi)$ denote the set of transitions that appear
 infinitely many times in $\pi$. Run $\pi$ is \emph{accepting} iff
-$\minf(\pi)$ satisfies the formula $\varphi$, where a set of $T$ of
+$\minf(\pi)$ satisfies the formula $\varphi$, where a set $T$ of
 transitions satisfies $\Inf m$ iff $m\in\mks(T)$ and it satisfies
@@
   (DNF)} if it is a disjunction of clauses, where each clause is a
-conjunction of some terms of the form $\Fin m$ or $\Inf m$. Formula
-$\false$ corresponds to the disjunction of zero clauses and formula
-$\true$ corresponds to one clause with zero terms. Each acceptance
-formula can be transformed into an equivalent formula in DNF.
+conjunction of terms of the form $\Fin m$ or $\Inf m$.  Each
+acceptance formula can be transformed into an equivalent formula
+in~DNF. Formula $\false$ corresponds to the disjunction of zero
+clauses and formula $\true$ corresponds to one clause with zero terms.
 
@@
 adjacent transitions of the form
-$\rho=(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
+$\rho=(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\ldots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
 such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
@@
 the rest of this paper, SCC always refers to a maximal SCC. Given a
-set of states $S\subseteq Q$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
-denotes the set of all transitions between states in $S$. Further, for
-each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
-marked with $m$, i.e., $\delta_m=\{t\in\delta\mid m\in
-\mks(t)\}$.
+set of states $S\subseteq Q$, let $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
+denote the set of all transitions between states in $S$. Further, for
+each mark $m\in M$, let $\delta_m=\{t\in\delta\mid m\in
+\mks(t)\}$ denote the set of all transitions
+marked with $m$.
 % \todo{The sets $\delta_m$ are called \emph{acceptance sets}. Given
@@
 A set of transitions $T\subseteq\delta$ is called a \emph{cycle} is
-there exists a path from a state $p$ to the same state containing all
-transitions of $T$ (potentially multiple times) and no transition
-outside $T$. Finally, we assume that each TELA $\mA$ contains only
-states $q$ that are reachable from the initial state $q_I$ (i.e.,
-$q=q_I$ or there is a path from $q_I$ to $q$) as states that are not
-reachable from $q_I$ can be eliminated without any impact on $L(\mA)$.
+there exists a path from a state $p$ to the same state containing each
+transition of $T$ at least once and no transition outside
+$T$. Finally, we assume that each TELA $\mA$ contains only states $q$
+that are reachable from the initial state $q_I$ (i.e., $q=q_I$ or
+there is a path from $q_I$ to $q$) as states that are not reachable
+from $q_I$ can be eliminated without any impact on $L(\mA)$.
 
@@
 is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
-assignment of the equivalent Boolean formula. A QBF formula is
+assignment of an equivalent Boolean formula. A QBF formula is
 \emph{satisfiable} iff it has a model.
@@
 free variables representing possible instances of $\psi$ and
-$\rem$. If the formula is satisfiable, then its models encode desired
-instances of $\psi$ and $\rem$.  In the following, we assume that we
-are looking for a new acceptance formula $\psi$ in DNF. Note that our
-reduction method can be easily adapted to look for $\psi$ of a
-different shape.
+$\rem$. If the formula is satisfiable, then each of its models encodes
+a desired instance of $\psi$ and $\rem$.  In the following, we assume
+that we are looking for a new acceptance formula $\psi$ in DNF. Note
+that our reduction method can be easily adapted to look for $\psi$ of
+a different shape.
 
@@
 %Further, we consider only subsets $T\subseteq\bigcup_{\textrm{SCC}~S}\delta_S$
-First we describe Boolean variables used in the constructed QBF
-formulas.
+We start with a description of Boolean variables used in the
+constructed QBF formulas.
 \begin{itemize}
@@
 transitions with mark $m$ form the set $\delta_m$, $\Inf m$ can be
-expresses by $\bigvee_{t\in\delta_m}e_t$. Similarly, $T$ satisfies
-$\Fin m$ iff $m\not\in\mks(T)$, which can be expresses by
+expressed by $\bigvee_{t\in\delta_m}e_t$. Similarly, $T$ satisfies
+$\Fin m$ iff $m\not\in\mks(T)$, which can be expressed by
 $\bigwedge_{t\in\delta_m}\neg e_t$. Hence, $\msat_\varphi(\vec{e})$
@@
 \]
-In fact, only the direction ``$\Longleftarrow$'' is required for the
-correctness of our reduction method. If $\Te$ is not a cycle and
-$\mcycle(\vec{e})$ holds, then the required equivalence
+In fact, only the direction ``$\Longleftarrow$'' is needed for the
+correctness of our reduction method. If there are some valuations of
+$\vec{e}$ such that $\mcycle(\vec{e})$ holds and $\Te$ is not a cycle,
+then we will superfluously require the equivalence
 $\msat_\varphi(\vec{e}) \iff
-\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ on this valuation of
-$\vec{e}$ adds a superfluous constraint on $\psi$ represented by
-$\vec{i},\vec{f}$ and $\rem$ represented by $\vec{n}$. These
-superfluous constraints can lead to loss of reduction opportunities,
-but not to incorrectness. This observation allows us to trade the
-precision of $\mcycle(\vec{e})$ for simplicity of the formula.
-
-In the following, we define three versions of $\mcycle(\vec{e})$:
+\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ on these valuations.
+% adds a superfluous constraint on $\psi$ (represented by
+% $\vec{i},\vec{f}$) and $\rem$ (represented by $\vec{n}$).
+These superfluous constraints can lead to loss of reduction
+opportunities, but not to incorrectness. This observation allows us to
+trade the precision of $\mcycle(\vec{e})$ for its simplicity.
+formula.
+%
+We define three versions of $\mcycle(\vec{e})$:
 \begin{itemize}
@@
 \item $\mcycle_2(\vec{e})$ is an intermediate version. It says that $\Te$
-  is nonempty and every transition in $\Te$ has to have a preceding
+  is nonempty and every transition in $\Te$ has % to have
+  a preceding
   and a succeeding transition in $\Te$, which is a necessary condition
@@
 \end{itemize}
-We write $\Phi_{j,C,K}$ when we want to emphasize that the particular
+We write $\Phi_{j,C,K}$ %when we want
+to emphasize that a particular
 formula $\Phi_{C,K}$ contains the version $\mcycle_j(\vec{e})$.
@@
 
-Formally, we define an equivalence $\sim\subseteq\delta\times\delta$
+Formally, we define an equivalence ${\sim}\subseteq\delta\times\delta$
 on transitions such that $t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$.
@@
   the set $\overline{T}=\{\overline{t}\mid t\in T\}$. As
-  $\mks(t)=\mks(\overline{t})$ holds for all transitions, we get
+  $\mks(t)=\mks(\overline{t})$ for all transitions, we get
   $\mks(T)=\mks(\overline{T})$ and thus
@@
 remove all transitions except one of each class. Then we build the
-formula for this pruned automaton as described above. The more
-transitions we removed, the shorter formula with less $e_t$ variables
-we obtain. If the formula $\Phi_{1,C,K}$ for the pruned automaton is
-satisfiable and we derive $\rem$ and $\psi$ from its model, we can use
-safely them also on the original automaton. We only need to extend
-$\rem$ to all transitions such that it changes the acceptance marks on
-all equivalent transitions in the same way.
+formula $\Phi_{1,C,K}$ for this pruned automaton. The more transitions
+we removed, the shorter formula with less $e_t$ variables we
+obtain. If the formula $\Phi_{1,C,K}$ for the pruned automaton is
+satisfiable, we derive $\rem$ and $\psi$ from its model and use them
+with the original automaton. We only need to extend $\rem$ to all
+transitions such that it changes the acceptance marks on all
+equivalent transitions in the same way. \todo{je to OK?}
 
@@
   to $q_i$ we have that $p_i\in\post$ and $q_i\in\pre$, which implies
-  that there is a path $\rho_i\in T^+$ containing $t_i$ and leading
-  from $p$ back to $p$. If we concatenate all these paths, we get the
-  path $\rho_1\rho_2\ldots\rho_{|T|}\in T^+$ that contains all transitions of
-  $T$ and leads from $p$ back to $p$, which means that $T$ is a cycle.
-  This is a contradiction.
+  that
+  \begin{itemize}
+  \item $p_i=p$ (we set $\rho'_i=\varepsilon$ in this case) or
+    there is a path $\rho'_i\in T^+$ leading from $p$ to $p_i$, and
+  \item $q_i=p$ (we set $\rho''_i=\varepsilon$ in this case) or
+    there is a path $\rho''_i\in T^+$ leading from $q_i$ to $p$.
+  \end{itemize}
+  Then there is a path $\rho_i=\rho'_it_i\rho''_i\in T^+$ leading from
+  $p$ back to $p$ and containing $t_i$. If we concatenate all these
+  paths, we get the path $\rho_1\rho_2\ldots\rho_{|T|}\in T^+$ that
+  contains all transitions of $T$ and leads from $p$ back to $p$,
+  which means that $T$ is a cycle.  This is a contradiction.
 \end{proof}
@@
 arbitrary state of $S$. Further, let $\vec{n}_S$ denote the vector of
-variables $n_{t,k}$ where $t\in\delta_S$ and $k\in N_k$. By
+variables $n_{t,k}$ where $t\in\delta_S$ and $k\in N_K$. By
 $\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})$ we denote the formula
@@
 \begin{itemize}
+\item upravit nazev a abstrakt, zduraznit uzitecnost TELA pro concurrency (analysis of probabilistic systems...)
 \item do experimentu napsat, jaky hardware pouzivame (merit cas bez toho nedava smysl)

### 2022-04-26 paper/main.tex
new numbers

--- 2022-04-24 paper/main.tex
@@
 Spot version 2.10.4 and Z3 version 4.8.13 have been used in the
-experimental evaluation.
+experimental evaluation. The experiments have been run on a computer
+with Intel® Core™ i7-8550U processor and 16 GB of memory running
+Debian 11.
 
@@
 %compared between different reduction levels of the same automata set.
-Additionally, we have filtered one automaton from the \texttt{ltl3tela} set
-that cannot be processed by \telatko due to an extensive use of system
-resources.
+%Additionally, we have filtered one automaton from the \texttt{ltl3tela} set
+%that cannot be processed by \telatko due to an extensive use of system
+%resources.
 
@@
 \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
-\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 90\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
 \bottomrule

### 2022-04-26 paper/table-data.tex
new numbers

@@
 level 1 &
-387 & 26.0 & 796.9 & 111 & 18
+387 & 26.0 & 736.5 & 111 & 18
 \\
@@
 level 2 &
-395 & 24.5 & 1057.1 & 108 & 21
+395 & 24.5 & 954.6 & 108 & 21
 \\
@@
 level 3 &
-393 & 24.9 & 9366.6 & 104 & 25
+397 & 24.1 & 7089.0 & 103 & 26
 \\
@@
 multi &
-379 & 27.5 & 11486.6 & 104 & 25
+379 & 27.5 & 9128.6 & 105 & 24
 \\
@@
 level 1 &
-552 & 37.4 & 786.3 & 220 & 14
+552 & 37.4 & 718.7 & 220 & 14
 \\
@@
 level 2 &
-569 & 35.5 & 956.2 & 216 & 18
+569 & 35.5 & 866.0 & 216 & 18
 \\
@@
 level 3 &
-562 & 36.3 & 1276.4 & 211 & 23
+561 & 36.4 & 1148.5 & 213 & 21
 \\
@@
 multi &
-543 & 38.4 & 2525.3 & 211 & 23
+543 & 38.4 & 2310.9 & 211 & 23
 \\
@@
 level 1 &
-198 & 0.0 & 590.9 & 70 & 0
+198 & 0.0 & 34.6 & 70 & 0
 \\
@@
 level 2 &
-198 & 0.0 & 630.8 & 70 & 0
+198 & 0.0 & 66.4 & 70 & 0
 \\
@@
 level 3 &
-189 & 4.5 & 2127.0 & 63 & 7
+189 & 4.5 & 1256.5 & 63 & 7
 \\
@@
 multi &
-189 & 4.5 & 2180.6 & 64 & 6
+189 & 4.5 & 1288.2 & 64 & 6
 \\
 \midrule
-\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\90 automata\\343 acc.~marks}}
+\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\91 automata\\348 acc.~marks}}
 &
 level 1 &
-328 & 4.4 & 486.5 & 78 & 12
+332 & 4.6 & 454.3 & 79 & 12
 \\
@@
 level 2 &
-330 & 3.8 & 567.0 & 75 & 15
+334 & 4.0 & 537.8 & 76 & 15
 \\
@@
 level 3 &
-322 & 6.1 & 717.0 & 71 & 19
+326 & 6.3 & 680.7 & 72 & 19
 \\
@@
 multi &
-319 & 7.0 & 1622.5 & 72 & 18
+323 & 7.2 & 1548.2 & 73 & 18
 \\

### 2022-04-26 paper/main.tex
section 4 polished.

@@
 formula, we compute the equivalence $\sim$ on automata transitions and
-remove all transitions except one of each class. Then we build the
-formula $\Phi_{1,C,K}$ for this pruned automaton. The more transitions
-we removed, the shorter formula with less $e_t$ variables we
-obtain. If the formula $\Phi_{1,C,K}$ for the pruned automaton is
-satisfiable, we derive $\rem$ and $\psi$ from its model and use them
-with the original automaton. We only need to extend $\rem$ to all
-transitions such that it changes the acceptance marks on all
-equivalent transitions in the same way. \todo{je to OK?}
+temporarily remove all transitions except one of each equivalence class. Then we
+build the formula $\Phi_{1,C,K}$ for the pruned automaton. The more
+transitions we removed, the shorter formula with less $e_t$ variables
+we obtain. If the formula $\Phi_{1,C,K}$ for the pruned automaton is
+satisfiable, we derive $\rem$ and $\psi$ from its model and extend % use them
+% with the original automaton. We only need to extend
+$\rem$ to all transitions of the original automaton such that
+it changes the acceptance marks on all equivalent transitions in the
+same way. In the following, we use this simplification whenever
+$\Phi_{1,C,K}$ is employed.
+\todo{Terko, je to OK?}
 
@@
 uses only a single kind of QBF formulas. More precisely, we talk about
-\emph{level~1} reduction if $\Phi^\mathit{opt}_{1,C,K}$ is
-used. Analogously, we talk about \emph{level~2} and \emph{level~3}
-reduction when $\Phi^\mathit{opt}_{2,C,K}$ or
+\emph{level~1}, \emph{level~2}, or \emph{level~3} reduction when
+$\Phi^\mathit{opt}_{1,C,K}$, $\Phi^\mathit{opt}_{2,C,K}$, or
 $\Phi^\mathit{opt}_{3,C,K}$ is used, respectively.
+
+% \begin{algorithm}[t]
+%   \DontPrintSemicolon
+%   \SetKwFunction{FextendBackward}{extendBackward}
+%   \SetKwProg{PI}{Procedure}{}{}
+%   \PI{$\mathit{SingleLevelReduction}(\mA,j,\mathit{reduceC})$}{
+%     \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, level $j\in\{1,2,3\}$, $\mathit{reduceC}\in\{\true,\false\}$}
+%     \KwOut{an equivalent TELA with the same structure as $\mA$ and with at most as many acceptance marks as in $\mA$} %with the same or lower number of acceptance~marks}
+%     \;
+%   %   \If{$L(\mA)=\emptyset$}{
+%   % %      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'=\{(p,a,\emptyset,q)\mid(p,a,M',q)\in\delta\}$
+%   %       \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all acceptance marks removed
+%   %     }
+%     $C_\mA\gets$ the number of clauses in the acceptance formula of $\mA$ transformed to DNF\;
+%     $K_\mA\gets$ the number of acceptance marks in $\mA$\;
+%     $C\gets C_\mA$\;
+%     $K\gets K_\mA$\;
+%     $\mathit{stop}\gets\false$\;
+%     \While(\tcp*[f]{reduction of the number of marks}){$\neg\mathit{stop}~\wedge~K>0$}{
+%       $\Phi\gets\Phi^\mathit{opt}_{j,C,K{-}1}$ for $\mA$\;
+%       \leIf{$\mathit{satisfiable}(\Phi)$}{$K\gets K{-}1$}{$\mathit{stop}\gets\true$}
+%     }
+%     \If(\tcp*[f]{check whether the condition $\false$ fits}){$K=1~\wedge~\mathit{stop}$}{
+%       $\Phi\gets\Phi^\mathit{opt}_{j,0,0}$ for $\mA$\; 
+%       \If{$\mathit{satisfiable}(\Phi)$}{
+%         \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all marks removed
+%       }
+%     }  
+%     \If(\tcp*[f]{reduction of the number of clauses}){$\mathit{reduceC}$}{
+%       $\mathit{stop}\gets\false$\;
+%       \While{$\neg\mathit{stop}~\wedge~C>1$}{
+%         $\Phi\gets\Phi^\mathit{opt}_{j,C-1,K}$ for $\mA$\;
+%         \leIf{$\mathit{satisfiable}(\Phi)$}{$C\gets C{-}1$}{$\mathit{stop}\gets\true$}
+%       }
+%     }
+%     \eIf{$K=K_\mA~\wedge~C=C_\mA$}{
+%       \KwRet $\mA$
+%     }{
+%       compute $\rem$ and $\psi$ from a model of $\Phi^\mathit{opt}_{j,C,K}$\;
+%       \KwRet $(Q,N_K,\Sigma,q_I,\rem(\delta),\psi)$
+%     }
+%   }\;
+%   \caption{The single-level reduction procedure}
+%   \label{alg:single}
+% \end{algorithm}
 
@@
     \;
-    \If{$L(\mA)=\emptyset$}{
-%      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'=\{(p,a,\emptyset,q)\mid(p,a,M',q)\in\delta\}$
-      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all acceptance marks removed
-    }
+  %   \If{$L(\mA)=\emptyset$}{
+  % %      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'=\{(p,a,\emptyset,q)\mid(p,a,M',q)\in\delta\}$
+  %       \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all acceptance marks removed
+  %     }
     $C_\mA\gets$ the number of clauses in the acceptance formula of $\mA$ transformed to DNF\;
@@
     $K\gets K_\mA$\;
-    $\mathit{stop}\gets\false$\;
-    \While{$\neg\mathit{stop}~\wedge~K>0$}{
-      $\Phi\gets\Phi^\mathit{opt}_{j,C,K{-}1}$ for $\mA$\;
-      \leIf{$\mathit{satisfiable}(\Phi)$}{$K\gets K{-}1$}{$\mathit{stop}\gets\true$}
+%    \While(\tcp*[f]{reduction of the number of marks}){$K>1~\wedge~\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$}{$K\gets K{-}1$}
+    \lWhile{$K>1~\wedge~\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$}{$K\gets K{-}1$}
+    \If{$K=1$}{
+      \If(\tcp*[f]{check the condition $\true$}){all cycles in $\mA$ are accepting}{ %$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,1,0})$}{
+        \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\true)$ where $\delta'$ is $\delta$ with all marks removed
       }
-    \If{$\mathit{reduceC}$}{
-      $\mathit{stop}\gets\false$\;
-      \While{$\neg\mathit{stop}~\wedge~C>1$}{
-        $\Phi\gets\Phi^\mathit{opt}_{j,C-1,K}$ for $\mA$\;
-        \leIf{$\mathit{satisfiable}(\Phi)$}{$C\gets C{-}1$}{$\mathit{stop}\gets\true$}
+      \If(\tcp*[f]{check the condition $\false$}){all cycles in $\mA$ are rejecting}{ %$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,0,0})$}{
+        \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all marks removed
       }
     }  
-    \eIf{$K=K_\mA~\wedge~C=C_\mA$}{
-      \KwRet $\mA$
-    }{
+    \If(\tcp*[f]{reduction of the number of clauses}){$\mathit{reduceC}$}{
+      \lWhile{$C>1~\wedge~\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C-1,K})$}{$C\gets C{-}1$}
+    }
+    \If{$K<K_\mA~\vee~C<C_\mA$}{
       compute $\rem$ and $\psi$ from a model of $\Phi^\mathit{opt}_{j,C,K}$\;
@@
     }
+    \KwRet $\mA$
   }\;
@@
 
-The reduction procedure is given in Algorithm~\ref{alg:single}. Besides
-the reduction of acceptance marks, the algorithm also reduces the
-number of clauses in acceptance formula if the last argument
-$\mathit{reduceC}$ is set to $\true$. In the first step, the procedure
-checks whether the input automaton represents the empty language and
-if this is the case, it returns the same automaton with acceptance
-conditon $\false$ and with all acceptance marks removed. Otherwise it
-gradually decreases the number of marks until 0 is reached or the QBF
-solver behind the function $\mathit{satisfiable}(\Phi)$ fails to
-reduce the number of marks, i.e., it claims unsatisfiability of $\Phi$
-or it runs out of resources. In the next step, if $\mathit{reduceC}$
-is set to $\true$ then the procedure gradually reduces the number of
-clauses. It stops if no equivalent automaton with less clauses is
-found or the number of clauses is reduced to one. Note that the
-acceptance condition with 0 clauses is $\false$ and this case is
-treated at the beginning of the procedure. Finally, if the procedure
-succeeds to reduce the number of marks or clauses, if constructs the
-modified automaton. Otherwise, it returns the original automaton.
-
-Note that the algorithm can be also reformulated to use an incremental
-approach instead of building a new formula $\Phi$ in each iteration.
-The incremental version of the first \textbf{while} loop builds the formula
+The reduction procedure called $\mathit{SingleLevelReduction}$ is
+given in Algorithm~\ref{alg:single}. Besides the reduction of
+acceptance marks, the algorithm also reduces the number of clauses in
+acceptance formula if the last argument $\mathit{reduceC}$ is set to
+$\true$.
+% In the first step, the procedure checks whether the input
+% automaton represents the empty language and if this is the case, it
+% returns the same automaton with acceptance conditon $\false$ and with
+% all acceptance marks removed. Otherwise it
+The first \textbf{while} loop gradually decreases the number of marks
+until $K=1$ is reached or the QBF solver behind the function
+$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$ fails to reduce
+the number of marks, i.e., it claims unsatisfiability of the formula
+or it runs out of resources. If the loop ends with $K=1$, we check
+whether an acceptance condition without any mark (i.e., $\true$ or
+$\false$) can be used. These checks are based on an inspection of the
+automaton rather than on QBF solving. If some of the checks succeeds,
+we return the correspodning automaton without any acceptance
+mark. Otherwise, if $\mathit{reduceC}$ is set to $\true$ then the
+procedure gradually reduces the number of clauses in the second
+\textbf{while} loop. Note that the loop never checks for acceptance
+condition with 0 clauses as it is equivalent to $\false$ and this case
+was treated above. Finally, if the procedure succeeds to reduce the
+number of marks or clauses, it constructs the modified
+automaton. Otherwise, it returns the original automaton.
+
+The algorithm can be reformulated to use an incremental approach
+instead of building a new formula in each iteration of the
+\textbf{while} loops. The incremental version of the first
+\textbf{while} loop builds the formula
 $\Phi=\Phi^\mathit{opt}_{j,C,K{-}1}$ only in the first iteration. In
-each subsequent iteration, it extends the formula with a condition
+each subsequent iteration, it extends this formula with a condition
 saying that one more mark is not used in the automaton, i.e., the mark
 is neither on edges, nor in the acceptance formula. For example, if we
-want to say that mark $k\in N_K$ is not used in the automaton, we replace
-$\Phi$ by
+want to say that the mark $k\in N_K$ is not used, we replace $\Phi$ by
 \[
@@
 \]
-The second \emph{while} can transformed to an incremental version in a
-similar way. The advantage of the incremental approach comes from the
-fact that some QBF solvers can decide an extended formula faster as
-they profit from the information computed when solving the original
-formula.
+The second \textbf{while} can transformed to an incremental version
+similarly. The incremental approach benefits from the fact that some
+QBF solvers can decide an extended formula faster as they reuse the
+information computed when solving the original formula.
 
@@
     \end{scope}
-    \begin{scope}[initial angle=90,xshift=-3.5cm,yshift=-4cm]
+    \begin{scope}[initial angle=90,xshift=-3.5cm,yshift=-4.1cm]
       \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,2,\true)$}; %{level 2:};
@@
     \end{scope}
-    \begin{scope}[initial angle=90,xshift=3.5cm,yshift=-4cm]
+    \begin{scope}[initial angle=90,xshift=3.5cm,yshift=-4.1cm]
       \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,3,\true)$}; %{level 3:};
@@
     \;
-    $\mA\gets\mathit{SingleLevelReduction}(\mA,1,\mathit{reduceC})$\;
-    $\mA\gets\mathit{SingleLevelReduction}(\mA,2,\mathit{reduceC})$\;
+    $\mA\gets\mathit{SingleLevelReduction}(\mA,1,\false)$\;
+    $\mA\gets\mathit{SingleLevelReduction}(\mA,2,\false)$\;
     $\mA\gets\mathit{SingleLevelReduction}(\mA,3,\mathit{reduceC})$\;

### 2022-04-26 paper/main.tex
section 5 polished

@@
 
-The parameters of the last query of the QBF solver are encoded in the name of the automaton.
-
-The name of the automaton is in the following format:
-\begin{center}
-  \texttt{level\_marks\_solver}
-\end{center}
-where 
-\begin{itemize}
-    \item \texttt{level} $\in \{L1, L2, L3, T, F\}$ and it gives us the information about which level of simplification was applied (or in case of $T$ and $F$ if the acceptance formula can be $\true$ or $\false$).
-    \item \texttt{marks} $\in \mathbf{N}$ representing the number of marks we built the formula for.
-    \item \texttt{solver} $\in \{U, T, S\}$ where $U$ means that the query was unsatisfiable, $T$ means the solver timeouted and $S$ means that the simplification is satisfiable.
-\end{itemize}
-
-Note that we are only interested in the result of the last query, therefore we get the information that the simplification was satisfiable only if the result automaton has zero acceptance marks.
+If some call of the function
+$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$ in the first
+\textbf{while} loop of Algorithm~\ref{alg:single} does not return
+$\true$, then the name of the output automaton encodes the reason for
+it. In case of a single level reduction, the name has the form
+\texttt{L$j$\_$k$\_$X$}, where $j$ is the considered level, $k=K-1$ is
+the number of acceptance mark considered by the formula, $X$ is either
+\texttt{U} if the formula is unsatisfiable or \texttt{T} if the solver
+ran out of resources. If $X$ is \texttt{T}, we know that additional
+resources (typically longer timeout) may lead to further reductions.
+If the multi-level reduction is used, the automaton name contains the
+information from all levels. For example, the name `\texttt{L1\_5\_U
+  L2\_3\_U L3\_1\_T}' means that level 1 reduced the number of marks
+to 6 (reduction to 5 is impossible on this level), level 2 reduced it
+to 4, and level 3 to 2 and the automaton may be potentially reduced
+further.
+
+% that does not . 
+% The parameters of the last query of the QBF solver are encoded in the name of the automaton.
+
+% The name of the automaton is in the following format:
+% \begin{center}
+%   \texttt{level\_marks\_solver}
+% \end{center}
+% where 
+% \begin{itemize}
+%     \item \texttt{level} $\in \{L1, L2, L3, T, F\}$ and it gives us the information about which level of simplification was applied (or in case of $T$ and $F$ if the acceptance formula can be $\true$ or $\false$).
+%     \item \texttt{marks} $\in \mathbf{N}$ representing the number of marks we built the formula for.
+%     \item \texttt{solver} $\in \{U, T, S\}$ where $U$ means that the query was unsatisfiable, $T$ means the solver timeouted and $S$ means that the simplification is satisfiable.
+% \end{itemize}
+
+% Note that we are only interested in the result of the last query, therefore we get the information that the simplification was satisfiable only if the result automaton has zero acceptance marks.
 Meaning that the acceptance condition is either $\true$ or $\false$.

### 2022-04-26 paper/main.tex
HW info

--- 2022-04-26 paper/main.tex
@@
 
-Spot version 2.10.4 and Z3 version 4.8.13 have been used in the
+Spot version 2.10.3.dev and Z3 version 4.8.15 have been used in the
 experimental evaluation. The experiments have been run on a computer
-with Intel® Core™ i7-8550U processor and 16 GB of memory running
-Debian 11.
+with Intel® Core™ i7-8700 processor and 32 GB of memory running
+Ubuntu 20.04.4.
 

### 2022-04-26 paper/table-data.tex
data od Terky

@@
 level 1 &
-387 & 26.0 & 736.5 & 111 & 18
+387 & 26.0 & 755.4 & 111 & 18
 \\
@@
 level 2 &
-395 & 24.5 & 954.6 & 108 & 21
+390 & 25.4 & 1024.5 & 109 & 20
 \\
@@
 level 3 &
-397 & 24.1 & 7089.0 & 103 & 26
+393 & 24.9 & 7273.3 & 104 & 25
 \\
@@
 multi &
-379 & 27.5 & 9128.6 & 105 & 24
+371 & 29.1 & 9111.4 & 108 & 21
 \\
@@
 level 1 &
-552 & 37.4 & 718.7 & 220 & 14
+552 & 37.4 & 738.7 & 220 & 14
 \\
@@
 level 2 &
-569 & 35.5 & 866.0 & 216 & 18
+569 & 35.5 & 907.3 & 216 & 18
 \\
@@
 level 3 &
-561 & 36.4 & 1148.5 & 213 & 21
+561 & 36.4 & 1163.4 & 213 & 21
 \\
@@
 multi &
-543 & 38.4 & 2310.9 & 211 & 23
+542 & 38.5 & 2320.7 & 213 & 21
 \\
@@
 level 1 &
-198 & 0.0 & 34.6 & 70 & 0
+198 & 0.0 & 49.3 & 70 & 0
 \\
@@
 level 2 &
-198 & 0.0 & 66.4 & 70 & 0
+198 & 0.0 & 78.9 & 70 & 0
 \\
@@
 level 3 &
-189 & 4.5 & 1256.5 & 63 & 7
+189 & 4.5 & 1222.8 & 63 & 7
 \\
@@
 multi &
-189 & 4.5 & 1288.2 & 64 & 6
+189 & 4.5 & 1259.1 & 64 & 6
 \\
@@
 level 1 &
-332 & 4.6 & 454.3 & 79 & 12
+332 & 4.6 & 465.9 & 79 & 12
 \\
@@
 level 2 &
-334 & 4.0 & 537.8 & 76 & 15
+334 & 4.0 & 553.9 & 76 & 15
 \\
@@
 level 3 &
-326 & 6.3 & 680.7 & 72 & 19
+326 & 6.3 & 691.8 & 73 & 18
 \\
@@
 multi &
-323 & 7.2 & 1548.2 & 73 & 18
+323 & 7.2 & 1545.9 & 73 & 18
 \\

### 2022-04-26 paper/main.tex
fix typos

@@
 previous section to reduce the number of acceptance marks in TELA
-automata. First we describe a \emph{single-level} reduction, which
+automata. First, we describe a \emph{single-level} reduction, which
 uses only a single kind of QBF formulas. More precisely, we talk about
@@
 given in Algorithm~\ref{alg:single}. Besides the reduction of
-acceptance marks, the algorithm also reduces the number of clauses in
-acceptance formula if the last argument $\mathit{reduceC}$ is set to
+acceptance marks, the algorithm also reduces the number of clauses in the acceptance formula if the last argument $\mathit{reduceC}$ is set to
 $\true$.
@@
 automaton rather than on QBF solving. If some of the checks succeeds,
-we return the correspodning automaton without any acceptance
+we return the corresponding automaton without any acceptance
 mark. Otherwise, if $\mathit{reduceC}$ is set to $\true$ then the
@@
 \]
-The second \textbf{while} can transformed to an incremental version
+The second \textbf{while} can be transformed to an incremental version
 similarly. The incremental approach benefits from the fact that some
@@
 levels are typically faster. The best results can be often achieved by
-combining reductions of all levels. We call this aproach
+combining reductions of all levels. We call this approach
 \emph{multi-level} reduction. It is a straightforward sequential
@@
 $\true$, then the name of the output automaton encodes the reason for
-it. In case of a single level reduction, the name has the form
+it. In the case of a single level reduction, the name has the form
 \texttt{L$j$\_$k$\_$X$}, where $j$ is the considered level, $k=K-1$ is
-the number of acceptance mark considered by the formula, $X$ is either
+the number of acceptance marks considered by the formula, $X$ is either
 \texttt{U} if the formula is unsatisfiable or \texttt{T} if the solver
@@
 \texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
-produced by individual tools contain less than 348 automata as not all
-of the formulas are successfully translated due to timeout of 60
+produced by individual tools contains less than 348 automata as not all
+of the formulas are successfully translated due to a timeout of 60
 seconds. Further, we have removed automata with 0 or 1 acceptance
-marks as there is little point reducing these. Table
+marks as there is little point in reducing these. Table
 \ref{tab:ltltranslators} shows the exact versions of used translators,
-the homepage of the used tools and the number of automata produced by
+the homepage of the used tools, and the number of automata produced by
 the given tool from the LTL formulas after filtering.
@@
 translator), we have applied all single-level reductions and the
-multi-level reduction, always with incremental approach and without
+multi-level reduction, always with an incremental approach and without
 reduction of the number of clauses. The timeout for each QBF query was
@@
 
-Table~\ref{tab:cumm-data} shows the cummulative numbers of marks in
+Table~\ref{tab:cumm-data} shows the cumulative numbers of marks in
 the input automata sets and after each reduction, together with the
@@
 \begin{table}[h]
-\caption{The cummulative number of acceptance marks before and after reduction
+\caption{The cumulative number of acceptance marks before and after reduction
 for various automata sets.
@@
 pleasing finding that for every set, most automata are reduced in
-under a second and the high cummulative running times are caused by
-few complicated automata.
+under a second and the high cumulative running times are caused by
+a few complicated automata.
 
@@
 to use the QBF solvers to find equivalent acceptance conditions of TELA
-with less acceptance marks and without changes to automata structure.
+with fewer acceptance marks and without changes to automata structure.
 
@@
 time-consuming, in practice it can be well used and one can find a
-good trade-off between speed and efficiency by choosing precision of
+good trade-off between speed and efficiency by choosing the precision of
 the reduction and setting a reasonable timeout.

### 2022-04-26 paper/main.tex
progress in section 6

--- 2022-04-26 paper/main.tex
@@
 \texttt{U} if the formula is unsatisfiable or \texttt{T} if the solver
-ran out of resources. If $X$ is \texttt{T}, we know that additional
-resources (typically longer timeout) may lead to further reductions.
-If the multi-level reduction is used, the automaton name contains the
-information from all levels. For example, the name `\texttt{L1\_5\_U
-  L2\_3\_U L3\_1\_T}' means that level 1 reduced the number of marks
-to 6 (reduction to 5 is impossible on this level), level 2 reduced it
-to 4, and level 3 to 2 and the automaton may be potentially reduced
-further.
+did not decided within the time limit. If $X$ is \texttt{T}, we longer
+timeout may lead to further reductions. a If the multi-level reduction
+is used, the automaton name contains the information from all
+levels. For example, the name `\texttt{L1\_5\_U L2\_3\_U L3\_1\_T}'
+means that level 1 reduced the number of marks to 6 (reduction to 5 is
+impossible on this level), level 2 reduced it to 4, and level 3 to 2
+as the QBF solver did not finished in the time limit when trying to
+reduce the number of marks to
+1.%when tryint the automaton may be potentially reduced further.
 
@@
 literature. The parametrized formulas were generated for all
-combinations of parameter values from 1 to 4. Altogether, we have
-generated 348 unique LTL formulas. These have been translated to
-deterministic TELA by two state-of-the-art translators from the Owl
+combinations of parameter values from 1 to 4. We used the tool
+\texttt{ltlfilt} of the Spot library to simplify the formulas and
+remove duplicates. Altogether, we have generated 348 unique LTL
+formulas. These have been translated to deterministic TELA by two
+state-of-the-art translators from the Owl
 library~\cite{kretinsky.18.atva} and to nondeterministic TELA by
@@
 seconds. Further, we have removed automata with 0 or 1 acceptance
-marks as there is little point reducing these. Table
-\ref{tab:ltltranslators} shows the exact versions of used translators,
-the homepage of the used tools and the number of automata produced by
-the given tool from the LTL formulas after filtering.
-
-We have used multiple other helper tools in the experimental
-evaluation.  Alongside \texttt{genltl}, we have used \texttt{ltlfilt}
-to simplify the formulas and to remove duplicates, and
-\texttt{autcross} to compare automata and measure the running times.
-
-Spot version 2.10.3.dev and Z3 version 4.8.15 have been used in the
-experimental evaluation. The experiments have been run on a computer
-with Intel® Core™ i7-8700 processor and 32 GB of memory running
-Ubuntu 20.04.4.
-
-For each automata set (that is, the output of an individual LTL
-translator), we have applied all single-level reductions and the
-multi-level reduction, always with incremental approach and without
-reduction of the number of clauses. The timeout for each QBF query was
-set to 30 seconds.
-%We have again filtered out the automata for which
-%any of the four \telatko runs timed out after 120 seconds. We do this
-%to make sure the numbers of acceptance marks can be meaningfully
-%compared between different reduction levels of the same automata set.
-%Additionally, we have filtered one automaton from the \texttt{ltl3tela} set
-%that cannot be processed by \telatko due to an extensive use of system
-%resources.
-
-Table~\ref{tab:cumm-data} shows the cummulative numbers of marks in
-the input automata sets and after each reduction, together with the
-reduction ratio and total time spent by the considered reduction. The
-columns \textit{solver response} and \textit{solver timeout} show the
-number of automata for which the last query to QBF solver has been
-successfully answered, or the solver timed out within a 30 seconds
-limit to a single query, respectively. The timeout of the last QBF
-query implies that the automaton could be potentially further reduced
-by the reduction with a longer time limit.
+marks as there is a little point reducing these. Table
+\ref{tab:ltltranslators} shows the exact versions and homepages of
+used translators, and the number of automata produced by each
+translator after removing the automata with less than two marks.
 
@@
 
-\begin{table}[h]
-\caption{The cummulative number of acceptance marks before and after reduction
-for various automata sets.
+For each automata set (that is, the automata produced by an individual
+LTL translator), we have applied all single-level reductions and the
+multi-level reduction, always with incremental approach and without
+reduction of the number of clauses. The timeout for each QBF query was
+set to 30 seconds.
+
+For the experiments we used the version of \texttt{telatko} build with
+Spot library version 2.10.4 and Z3 version 4.8.15. The experiments
+have been run on a computer with Intel® Core™ i7-8700 processor and 32
+GB of memory running Ubuntu 20.04.4. We used the tool
+\texttt{autcross} of the Spot library to get the statistics of the
+reduced automata and the running times.
+
+Table~\ref{tab:cumm-data} shows the cummulative numbers of marks in
+the input automata sets and after each reduction, together with the
+reduction ratio and total time spent by the considered reduction. The
+column \textit{solver timeout} shows the number of automata for which
+the last query to QBF solver did not finished within the 30 seconds
+limit. The timeout of the last QBF query implies that the automaton
+may be potentially further reduced by the reduction with a longer time
+limit.
+
+\begin{table}[t]
+\caption{The cummulative numbers of acceptance marks before and after reduction
+for various reductions and automata sets.
 %\textit{X/Y solved aut.} means that out of
@@
 %marks only on these automata.
-The columns \textit{solver response} and
-\textit{solver timeout} refer to last call to the QBF solver (that is, the
-numbers of its successful responses within 30 seconds limit and timeouts).
-}
+The column \textit{solver timeout} indicates the number of instances
+when the last call to the QBF solver timed out.  }
 \label{tab:cumm-data}
 \centering
+\setlength{\tabcolsep}{10pt}
 \input{table-data.tex}
@@
 automata with a higher number of original acceptance marks. For
-example, in 22 cases, we have successfully reduced 7 or more
+example, in 24 cases, we have successfully reduced 7 or more
 acceptance marks to only 4.
@@
 
-\begin{figure}[h]
+\begin{figure}[b]
 \caption{Performance of \telatko on different automata sets. The $x$ axis
@@
 
+\todo{dodelat popis os u quantile plotu}
+
 % Pridat do experimentu nejaky graf, ktery ukaze, jak rychle se nektere
@@
 
-\section{TODO}
-
-Do submission:
-\begin{itemize}
-\item upravit nazev a abstrakt, zduraznit uzitecnost TELA pro concurrency (analysis of probabilistic systems...)
-\item do experimentu napsat, jaky hardware pouzivame (merit cas bez toho nedava smysl)
-\item spustit experimenty bez timeoutu na telatko (?a delsi timeout na QBF?)
-\item prelestit text v poslednich dvou kapitolach
-\end{itemize}
-
-\noindent
-Do budoucna:
-\begin{itemize}
-\item level 1 bez SCC optimalizace
-\item nahradit weak SCC jednou hranou pro neakceptujici a jinou pro akceptujici
-\item sladit konstrukci QBF formuli s implementaci
-\item podivat se znovu na uzitecnost a potrebnost optimalizaci typu ``klauzule neni false'' a ``alespon jedna klauzule je neprazdna''
-\item sladit popis algoritmu s implementaci
-\item v experimentech nejak sledovat vztah usetreneho casu a usetrenych znacek?
-\item zkusit aplikovat na jine typy automatu (treba TGBA nebo Rabin)?
-\item zkusit, zda/jak redukce znacek umozni redukovat stavovy prostor
-\end{itemize}
+% \section{TODO}
+
+% \noindent
+% Do budoucna:
+% \begin{itemize}
+% \item level 1 bez SCC optimalizace
+% \item nahradit weak SCC jednou hranou pro neakceptujici a jinou pro akceptujici
+% \item sladit konstrukci QBF formuli s implementaci
+% \item podivat se znovu na uzitecnost a potrebnost optimalizaci typu ``klauzule neni false'' a ``alespon jedna klauzule je neprazdna''
+% \item sladit popis algoritmu s implementaci
+% \item v experimentech nejak sledovat vztah usetreneho casu a usetrenych znacek?
+% \item zkusit aplikovat na jine typy automatu (treba TGBA nebo Rabin)?
+% \item zkusit, zda/jak redukce znacek umozni redukovat stavovy prostor
+% \end{itemize}
 

### 2022-04-26 paper/main.tex
intro percentage

--- 2022-04-26 paper/main.tex
@@
   by tools Delag and Rabinizer~4 of the Owl library, our technique reduced
-  reduced 27.5\% and 38.4\% of acceptance marks, respectively. \todo{zkontrolovat!}
+  reduced 29.1\% and 38.5\% of acceptance marks, respectively.
   % The achieved

### 2022-04-26 paper/main.tex

--- 2022-04-26 paper/main.tex
@@
 \texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
-produced by individual tools contain less than 348 automata as not all
-of the formulas are successfully translated due to timeout of 60
+produced by individual tools contains less than 348 automata as not all
+of the formulas are successfully translated due to a timeout of 60
 seconds. Further, we have removed automata with 0 or 1 acceptance

### 2022-04-26 paper/table-data.tex

--- 2022-04-26 paper/table-data.tex
@@
-\begin{tabular}{ccccccc}                                                        
+\begin{tabular}{ccrrrr}
 \toprule                                                                        
-automata set & level & marks & reduction [\%] & time [s] & \makecell{solver \\ response} & \makecell{solver \\ timeout} \\
+automata set & level & marks & reduction [\%] & time [s] & \makecell{solver \\[-1ex] timeout} \\
 
@@
 level 1 &
-387 & 26.0 & 736.5 & 111 & 18
+387 & 26.0 & 755.4 &  18
 \\
@@
 level 2 &
-395 & 24.5 & 954.6 & 108 & 21
+390 & 25.4 & 1024.5 &  20
 \\
@@
 level 3 &
-397 & 24.1 & 7089.0 & 103 & 26
+393 & 24.9 & 7273.3 &  25
 \\
@@
 multi &
-379 & 27.5 & 9128.6 & 105 & 24
+371 & 29.1 & 9111.4 &  21
 \\
@@
 level 1 &
-552 & 37.4 & 718.7 & 220 & 14
+552 & 37.4 & 738.7 & 14
 \\
@@
 level 2 &
-569 & 35.5 & 866.0 & 216 & 18
+569 & 35.5 & 907.3 &  18
 \\
@@
 level 3 &
-561 & 36.4 & 1148.5 & 213 & 21
+561 & 36.4 & 1163.4 &  21
 \\
@@
 multi &
-543 & 38.4 & 2310.9 & 211 & 23
+542 & 38.5 & 2320.7 &  21
 \\
@@
 level 1 &
-198 & 0.0 & 34.6 & 70 & 0
+198 & 0.0 & 49.3 &  0
 \\
@@
 level 2 &
-198 & 0.0 & 66.4 & 70 & 0
+198 & 0.0 & 78.9 & 0
 \\
@@
 level 3 &
-189 & 4.5 & 1256.5 & 63 & 7
+189 & 4.5 & 1222.8 & 7
 \\
@@
 multi &
-189 & 4.5 & 1288.2 & 64 & 6
+189 & 4.5 & 1259.1 & 6
 \\
@@
 level 1 &
-332 & 4.6 & 454.3 & 79 & 12
+332 & 4.6 & 465.9 &  12
 \\
@@
 level 2 &
-334 & 4.0 & 537.8 & 76 & 15
+334 & 4.0 & 553.9 &  15
 \\
@@
 level 3 &
-326 & 6.3 & 680.7 & 72 & 19
+326 & 6.3 & 691.8 &  18
 \\
@@
 multi &
-323 & 7.2 & 1548.2 & 73 & 18
+323 & 7.2 & 1545.9 & 18
 \\

### 2022-04-26 paper/main.tex
je to ok :)
upravy v experimentech
intro typos

@@
   number of %different
-  acceptance marks. We introduce a new technique reducing the number
+  acceptance marks. We introduce a new technique for reducing the number
   of %different
-  acceptance marks in TELA with use of \emph{quantified Boolean
+  acceptance marks in TELA with the use of \emph{quantified Boolean
     formulas (QBF)}.
@@
   by tools Delag and Rabinizer~4 of the Owl library, our technique reduced
-  reduced 29.1\% and 38.5\% of acceptance marks, respectively.
+  29.1\% and 38.5\% of acceptance marks, respectively.
   % The achieved
@@
 $\Phi_{1,C,K}$ is employed.
-\todo{Terko, je to OK?}
 

### 2022-04-26 paper/main.tex
section 6

@@
 impossible on this level), level 2 reduced it to 4, and level 3 to 2
-as the QBF solver did not finished in the time limit when trying to
+as the QBF solver did not finish in the time limit when trying to
 reduce the number of marks to
@@
 seconds. Further, we have removed automata with 0 or 1 acceptance
-marks as there is a little point reducing these. Table
+marks as there is a little point in reducing these. Table
 \ref{tab:ltltranslators} shows the exact versions and homepages of
@@
 
-For the experiments we used the version of \texttt{telatko} build with
+For the experiments, we used the version of \texttt{telatko} build with
 Spot library version 2.10.4 and Z3 version 4.8.15. The experiments
@@
 \begin{table}[t]
-\caption{The cummulative numbers of acceptance marks before and after reduction
+\caption{The cumulative numbers of acceptance marks before and after reduction
 for various reductions and automata sets.
@@
 
-\begin{figure}[b]
+\begin{figure}[b!]
 \caption{Performance of \telatko on different automata sets. The $x$ axis
@@
 \end{figure}
-
-\todo{dodelat popis os u quantile plotu}
 

### 2022-04-26 paper/main.tex
final polishing

--- 2022-04-26 paper/main.tex
@@
 \begin{table}[t]
-\caption{The cummulative numbers of acceptance marks before and after reduction
-for various reductions and automata sets.
+  \caption{The cummulative numbers of acceptance marks before and
+    after reduction for various reductions and automata sets. The
+    column \emph{reduction} shows the cummulative percentage of saved
+    acceptance marks and \emph{time} reports the cummulative reduction
+    time.
 %\textit{X/Y solved aut.} means that out of
@@
 
-Table~\ref{tab:heatmap} shows the number of automata coming from all
+Table~\ref{tab:heatmap} shows the number of automata from all
 four sets that have a given number of original acceptance marks and a
@@
 automata with a higher number of original acceptance marks. For
-example, in 24 cases, we have successfully reduced 7 or more
+example, in 24 cases, we have reduced 7 or more
 acceptance marks to only 4.
@@
 reduced from $x$ to $y$ acceptance marks. If the cell contains a sum of two
-numbers, the latter represents number of automata where the last query to QBF
-solver has not been answered in the time limit.}
+numbers, the latter represents number of automata where the last call to QBF
+solver reached the time limit.}
 \label{tab:heatmap}
@@
 
-\todo{dodelat popis os u quantile plotu}
-
 % Pridat do experimentu nejaky graf, ktery ukaze, jak rychle se nektere
@@
 
-We have presented an application of QBF solving to reduction of the number of
-acceptance marks in transition-based Emerson-Lei automata. The method
-has been implemented and our experimental results show that it is possible
-to use the QBF solvers to find equivalent acceptance conditions of TELA
-with fewer acceptance marks and without changes to automata structure.
+We have presented a method reducing the number of acceptance marks in
+transition-based Emerson-Lei automata with use of QBF solving and
+without altering automata structure. We have implemented the method in
+a tool called \texttt{telatko}. The presented experimental results
+show that the tool can indeed reduce the number of acceptance marks in
+automata produced by state-of-the-art LTL to automata translators. The
+reduction was significant on automata produced by the translators of
+the Owl library.
 
 The reduction of acceptance marks is not the only application of the
-presented approach: one could possibly look for, e.g., equivalent
-parity automata or automata without $\Fin$-marks, depending on the
-subsequent use of the automata. Even though the QBF queries can be
-time-consuming, in practice it can be well used and one can find a
-good trade-off between speed and efficiency by choosing the precision of
-the reduction and setting a reasonable timeout.
-
-In the future work, it could be examined how the reductions of the acceptance
-condition can contribute to further reduction of the state space by enabling
-various optimizations, e.g., simulations.
+presented approach. For example, it can be easily adopted to look for
+an equivalent automaton with the same structure and an acceptance
+formula of a specific form (e.g., without any $\Fin m$ terms). Even
+though the QBF queries can be time-consuming, in practice one can
+often find a good trade-off between speed and efficiency by adjusting 
+the formula precision and choosing a reasonable timeout.
+
+In future, we plan to improve the reduction process (e.g., use the
+bisection method to reduce the number of acceptance marks more
+efficiently) and study whether the reduction of acceptance condition
+can enable further reductions or simplifications of automata
+structures.
 

### 2022-04-26 paper/table-data.tex
final polishing

@@
 level 1 &
-387 & 26.0 & 755.4 &  18
+387~ & 26.0~~~~~~ & 755.4~ &  18~~~
 \\
@@
 level 2 &
-390 & 25.4 & 1024.5 &  20
+390~ & 25.4~~~~~~ & 1024.5~ &  20~~~
 \\
@@
 level 3 &
-393 & 24.9 & 7273.3 &  25
+393~ & 24.9~~~~~~ & 7273.3~ &  25~~~
 \\
@@
 multi &
-371 & 29.1 & 9111.4 &  21
+371~ & 29.1~~~~~~ & 9111.4~ &  21~~~
 \\
@@
 level 1 &
-552 & 37.4 & 738.7 & 14
+552~ & 37.4~~~~~~ & 738.7~ & 14~~~
 \\
@@
 level 2 &
-569 & 35.5 & 907.3 &  18
+569~ & 35.5~~~~~~ & 907.3~ &  18~~~
 \\
@@
 level 3 &
-561 & 36.4 & 1163.4 &  21
+561~ & 36.4~~~~~~ & 1163.4~ &  21~~~
 \\
@@
 multi &
-542 & 38.5 & 2320.7 &  21
+542~ & 38.5~~~~~~ & 2320.7~ &  21~~~
 \\
@@
 level 1 &
-198 & 0.0 & 49.3 &  0
+198~ & 0.0~~~~~~ & 49.3~ &  0~~~
 \\
@@
 level 2 &
-198 & 0.0 & 78.9 & 0
+198~ & 0.0~~~~~~ & 78.9~ & 0~~~
 \\
@@
 level 3 &
-189 & 4.5 & 1222.8 & 7
+189~ & 4.5~~~~~~ & 1222.8~ & 7~~~
 \\
@@
 multi &
-189 & 4.5 & 1259.1 & 6
+189~ & 4.5~~~~~~ & 1259.1~ & 6~~~
 \\
@@
 level 1 &
-332 & 4.6 & 465.9 &  12
+332~ & 4.6~~~~~~ & 465.9~ &  12~~~
 \\
@@
 level 2 &
-334 & 4.0 & 553.9 &  15
+334~ & 4.0~~~~~~ & 553.9~ &  15~~~
 \\
@@
 level 3 &
-326 & 6.3 & 691.8 &  18
+326~ & 6.3~~~~~~ & 691.8~ &  18~~~
 \\
@@
 multi &
-323 & 7.2 & 1545.9 & 18
+323~ & 7.2~~~~~~ & 1545.9~ & 18~~~
 \\

### 2022-04-26 paper/main.tex

@@
 \begin{table}[t]
-  \caption{The cummulative numbers of acceptance marks before and
-    after reduction for various reductions and automata sets. The
-    column \emph{reduction} shows the cummulative percentage of saved
-    acceptance marks and \emph{time} reports the cummulative reduction
-    time.
-%\textit{X/Y solved aut.} means that out of
-%\textit{Y} successfully translated LTL formulas, every run of \telatko ended
-%within 120 seconds time limit on \textit{X} automata. We sum up the acceptance
-%marks only on these automata.
-The column \textit{solver timeout} indicates the number of instances
-when the last call to the QBF solver timed out.  }
+  \caption{The cumulative numbers of acceptance marks before and after
+    reduction for various reductions and automata sets. The column
+    \emph{reduction} shows the cumulative percentage of saved
+    acceptance marks and \emph{time} reports the cumulative reduction
+    time. The column \textit{solver timeout} indicates the number of
+    instances when the last call to the QBF solver timed out.  }
 \label{tab:cumm-data}

### 2022-10-14 paper/main.tex
polished, but not resubmited (easychair is locked)
polishing the paper before FOSSACS 2023 submission

@@
-\documentclass[a4paper,UKenglish,cleveref,autoref,thm-restate]{lipics-v2021}
-%\documentclass[a4paper,UKenglish,cleveref, autoref, anonymous, thm-restate]{lipics-v2021}
-%
-%This is a template for producing LIPIcs articles. 
-%See lipics-v2021-authors-guidelines.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling autoref support, use "autoref"
-%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
-%for enabling thm-restate support, use "thm-restate"
-%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
-%for producing a PDF according the PDF/A standard, add "pdfa"
-
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g.
-when preparing a pre-final version to be uploaded to arXiv or another public repository% \usepackage[numbers]{natbib}
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
+% vim: tw=78 ai spell spelllang=en
+
+\documentclass[runningheads]{llncs}
+
+%\usepackage[paperheight=235mm, paperwidth=155mm,textwidth=12.2cm,textheight=19.3cm]{geometry}
 \usepackage{multirow}
@@
 \usepackage[noend]{algorithm2e}
+\usepackage{amsmath,amssymb}
+\usepackage{enumerate}
 
@@
 
+% setting the color of comments in algorithm
+\definecolor{algoColorComment}{RGB}{204, 51, 0}
+\definecolor{algoColorComment2}{RGB}{204, 153, 102}
+\newcommand{\decmt}[1]{{\color{algoColorComment2}\texttt{#1}}}
+\SetCommentSty{decmt}
+
 %}}}
+%{{{ hyperref and ORCID ref
+
+\makeatletter
+\RequirePackage[bookmarks,unicode,colorlinks=true]{hyperref}%
+   \def\@citecolor{blue}%
+   \def\@urlcolor{blue}%
+   \def\@linkcolor{blue}%
+\def\UrlFont{\rmfamily}
+\def\orcidID#1{\smash{\href{http://orcid.org/#1}{\protect\raisebox{-1.25pt}{\protect\includegraphics{orcid_color.eps}}}}}
+\makeatother
+
+\usepackage{academicons}
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+
+%}}}
+
+\begin{document}
 
@@
 \title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving}
-\author{Tereza Schwarzov\'{a}}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
-\author{Jan Strej\v{c}ek}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{strejcek@fi.muni.cz}{https://orcid.org/0000-0001-5873-403X}{}
-\author{Juraj Major}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{}{}
+\author{Tereza Schwarzov\'{a} \and
+        Jan Strej\v{c}ek\orcidID{0000-0001-5873-403X}\thanks{Corresponding author: \email{strejcek@fi.muni.cz}} \and
+        Juraj Major}
 \authorrunning{T. Schwarzov\'{a}, J. Strej\v{c}ek, and J. Major}
-\Copyright{Tereza Schwarzov\'{a}, Jan Strej\v{c}ek, and Juraj Major}
-\ccsdesc[500]{Theory of computation~Logic}
-\ccsdesc[500]{Theory of computation~Automata over infinite objects}
-\keywords{Emerson-Lei automata, TELA, automata reduction, QBF, telatko}
-\funding{This work has been supported by the Czech Science Foundation grant GA19-24397S.}
+\titlerunning{Reducing Acceptance Marks in TELA by QBF-solving}
+\institute{%Faculty of Informatics,
+  Masaryk University, Brno, Czech Republic
+}
 
 %}}}
-
-\begin{document}
 
@@
   number of %different
-  acceptance marks. We introduce a new technique for reducing the number
-  of %different
+  acceptance marks. We introduce a new technique for reducing the
+  number of %different
   acceptance marks in TELA with the use of \emph{quantified Boolean
@@
   % acceptance marks.
-  In fact, we present three formula constructions
-  % trading speed for higher reduction efficiency.
-  with increasing reduction potential and formula complexity, and thus
-  also decreasing speed of their satisfiability solving. We evaluated
+  %
+  % In fact, we present three formula constructions
+  % % trading speed for higher reduction efficiency.
+  % with increasing reduction potential and formula complexity, and thus
+  % also decreasing speed of their satisfiability solving.
+  We evaluated
   our reduction technique on TELA produced by state-of-the-art tools
@@
 Automata over infinite words like Büchi, Rabin, Streett, or parity
-automata play a crucial role in many algorithms related to the
-concurrency theory. In particular, they are used in specification,
-verification, analysis, monitoring, and synthesis of various systems
-with infinite behaviour. In 1987, Emerson and
-Lei~\cite{emerson.87.scp} introduced automata over infinite words
-where acceptance conditions are arbitrary combinations of acceptance
-primitives saying that a certain set of states should be visited
-finitely often or infinitely often. In 2015, the same kind of
-acceptance condition was described in the \emph{Hanoi omega-automata
-  format (HOAF)}~\cite{babiak.15.cav}. The only difference is that the
-acceptance primitives talk about finitely or infinitely often visited
-acceptance marks rather than sets of states.  Acceptance marks are
-placed on transitions and each mark identifies the set of transitions
-containing this mark. Hence, these automata are called
-\emph{transition-based Emerson-Lei automata (TELA)} and they
+automata play a crucial role in many algorithms related to 
+concurrency theory, game theory, and formal methods in general. In
+particular, they are used in specification, verification, analysis,
+monitoring, and synthesis of various systems with infinite
+behaviour. In 1987, Emerson and Lei~\cite{emerson.87.scp} introduced
+automata over infinite words where acceptance conditions are arbitrary
+combinations of acceptance primitives saying that a certain set of
+states should be visited finitely often or infinitely often. In 2015,
+the same kind of acceptance condition was described in the \emph{Hanoi
+  omega-automata format (HOAF)}~\cite{babiak.15.cav}. The only
+difference is that the acceptance primitives talk about finitely or
+infinitely often visited acceptance marks rather than sets of states.
+Acceptance marks are placed on transitions and each mark identifies
+the set of transitions containing this mark. Hence, these automata are
+called \emph{transition-based Emerson-Lei automata (TELA)} and they
 generalize many traditional kinds of automata over infinite words
@@
 (this can be easily done for example by Spot~\cite{duret.16.atva2}),
-but this reduction is paid by dramatic changes of state-space: the
+but this reduction is paid by dramatic changes of state space: the
 number of states can increase exponentially in the number of
@@
 We present such a technique heavily based on \emph{quantified Boolean
-  formulas (QBF)}. For a given TELA and parameters $k,n$, it produces
+  formulas (QBF)}. For a given TELA and parameters $c,k$, it produces
 a QBF formula which is satisfiable if and only if there exists an
-automaton with the same structure, $n$ acceptance marks, an acceptance
-formula in disjunctive normal form with $k$ clauses, and the same set
+automaton with the same structure, $k$ acceptance marks, an acceptance
+formula in disjunctive normal form with $c$ clauses, and the same set
 of accepting runs as the original automaton. The placement of the
@@
 construction of two simpler formulas such that their satisfiability
-implies the existence of an automaton with the same structure, $n$
+implies the existence of an automaton with the same structure, $k$
 acceptance marks, and the same set of accepting runs, but not vice
@@
 automata produced by Delag~\cite{muller.17.gandalf},
-Rabinizer~4~\cite{kretinsky.18.cav} (both included in the Owl library),
-Spot~\cite{duret.16.atva2}
-and \texttt{ltl3tela}~\cite{major.19.atva}.  While the reduction is relatively
-modest on TELA produced by \texttt{ltl3tela} and Spot,
-it is substantial on automata produced by the tools of the Owl library.
+Rabinizer~4~\cite{kretinsky.18.cav} (both included in the Owl
+library), Spot~\cite{duret.16.atva2} and
+\texttt{ltl3tela}~\cite{major.19.atva}.  While the reduction is
+relatively modest on TELA produced by \texttt{ltl3tela} and Spot, it
+is substantial on automata produced by the tools of the Owl library.
 
@@
   automata (TGBA)} without changing its structure. We are not aware
-about any existing research aimed at simplification of acceptance
+of any existing research aimed at simplification of acceptance
 formulas of TELA or reduction of the number of its acceptance marks
@@
   A \emph{transition-based Emerson-Lei automaton (TELA)} is a tuple
-  $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$, where
+  $\mA =(Q,M,\Sigma,\delta,q_I,\varphi)$, where
   \begin{itemize}
@@
   \item $\Sigma$ is a finite \emph{alphabet},
-  \item $q_I\in Q$ is an \emph{initial state},
   \item $\delta \subseteq Q \times \Sigma \times 2^{M} \times Q$ is a
-    \emph{transition relation}, and
+    \emph{transition relation},
+  \item $q_I\in Q$ is an \emph{initial state}, and
   \item $\varphi$ is the \emph{acceptance condition} constructed
@@
 state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$.
-The set $M'$ is also refered by $\mks(t)$. For a set of transitions
+The set $M'$ is also refered to by $\mks(t)$. For a set of transitions
 $T\subseteq\delta$, let $\mks(T)=\bigcup_{t\in T}\mks(t)$ denote the
@@
 transitions satisfies $\Inf m$ iff $m\in\mks(T)$ and it satisfies
-$\Fin m$ iff $m\not\in\mks(T)$. The \emph{language} of automaton $\mA$ is
+$\Fin m$ iff $m\not\in\mks(T)$. The \emph{language} of $\mA$ is
 the set
@@
 % $\delta_{S,m}=\delta_S\cap\delta_m$.}
-A set of transitions $T\subseteq\delta$ is called a \emph{cycle} is
+A set of transitions $T\subseteq\delta$ is called a \emph{cycle} if
 there exists a path from a state $p$ to the same state containing each
@@
 
-In graphical representation of automata, we often use acceptance marks
+In graphical representation, we often use acceptance marks
 $\tacc{1}{1},\tacc{2}{2},\ldots\in M$. Further, an edge 
@@
 Recall that we aim to reduce the number of acceptance marks in a given
-TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ without altering its
-structure and language. In other words, we are looking for a set $\oM$
-of acceptance marks, an accepance formula $\psi$ over $\oM$, and a
-function
+TELA $\mA =(Q,M,\Sigma,\delta,q_I,\varphi)$ without altering its
+structure and language. In other words, we are looking for
+\begin{itemize}
+\item a set $\oM$ of acceptance marks,
+\item an acceptance formula $\psi$ over $\oM$, and 
+\item a function
   $\rem:\delta\rightarrow Q \times \Sigma \times 2^{\oM}\times Q$
@@
   $t=(p,a,M',q)\in\delta$, we assume that $\rem(t)=(p,a,\oM',q)$ for
-some $\oM'\subseteq\oM$) such that $|\oM|<|M|$ and the automaton
-$\mB=(Q,\oM,\Sigma,q_I,\rem(\delta),\psi)$ is equivalent to $\mA$. To
+  some $\oM'\subseteq\oM$)
+\end{itemize}
+such that $|\oM|<|M|$ and the automaton
+$\mB=(Q,\oM,\Sigma,\rem(\delta),q_I,\psi)$ is equivalent to $\mA$. To
 ensure that the modified automaton is equivalent to the original one,
@@
 Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and vice
-versa. Hence, our reduction methods is looking for a new acceptance
+versa. Hence, our reduction method is looking for a new acceptance
 formula $\psi$ and function $\rem$ such that for each cycle
@@
       \begin{array}{ll}
-        1 & \textrm{if }t\in T\\
+        1~~ & \textrm{if }t\in T\\[.5ex]
         0 & \textrm{otherwise}    
@@
       \begin{array}{ll}
-        1 & \textrm{if }k\in\mks(\rem(t))\\
+        1~~ & \textrm{if }k\in\mks(\rem(t))\\[.5ex]
         0 & \textrm{otherwise}    
@@
       \begin{array}{ll}
-        1 & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\
+        1~~ & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\[.5ex]
         0 & \textrm{otherwise}    
@@
       \begin{array}{ll}
-        1 & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\
+        1~~ & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\[.5ex]
         0 & \textrm{otherwise}    
@@
 \[
-  \Phi_{C,K}(\vec{n},\vec{i},\vec{f})~~=~~\forall \vec{e}~.~\mcycle(\vec{e})\implies\big(\msat_\varphi(\vec{e}) \iff \msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})\big).
-\]
-It remains to describe the subformulas $\mcycle(\vec{e})$,
-$\msat_\varphi(\vec{e})$, and
-$\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$. The subformula
-$\mcycle(\vec{e})$ will be defined as the last one and we present
-three versions of it.
+  \Phi_{C,K}(\vec{n},\vec{i},\vec{f})~~=~~\forall \vec{e}~.~\mcycle(\vec{e})\implies\big(\msat_\varphi(\vec{e}) \iff \msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})\big),
+\]
+where the subformulas $\mcycle(\vec{e})$, $\msat_\varphi(\vec{e})$,
+and $\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ are defined
+as follows. The subformula $\mcycle(\vec{e})$ will be defined as the last
+one and we present three versions of it.
 
@@
 
-Now we construct the subformula
+Next, we construct the subformula
 $\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ that evaluates to true
@@
 trade the precision of $\mcycle(\vec{e})$ for its simplicity.
-formula.
-%
+
 We define three versions of $\mcycle(\vec{e})$:
@@
 \item $\mcycle_1(\vec{e})$ is a lightweight version, which only says
-  that $\Te$ is nonempty. It does not use the information about
-  automaton structure, but it comes with an interesting simplification
-  of the whole formula $\Phi_{C,K}$.
-\item $\mcycle_2(\vec{e})$ is an intermediate version. It says that $\Te$
-  is nonempty and every transition in $\Te$ has % to have
-  a preceding
-  and a succeeding transition in $\Te$, which is a necessary condition
-  for being a cycle, but not a sufficient one.
+  that $\Te$ is nonempty and $\Te\subseteq\delta_S$ for some SCC $S$.
+  Except for SCCs, it does not use the information about automaton
+  structure, but it comes with an interesting simplification of the
+  whole formula $\Phi_{C,K}$.
+\item $\mcycle_2(\vec{e})$ is an intermediate version. It says that
+  $\Te$ is nonempty, $\Te\subseteq\delta_S$ for some SCC $S$, and
+  every transition in $\Te$ has % to have
+  a preceding and a succeeding transition in $\Te$, which is a
+  necessary condition for being a cycle, but not a sufficient one.
 \item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
@@
 \end{itemize}
-We write $\Phi_{j,C,K}$ %when we want
-to emphasize that a particular
-formula $\Phi_{C,K}$ contains the version $\mcycle_j(\vec{e})$.
+We write $\Phi_{j,C,K}$ % where $j\in\{1,2,3\}$
+when we want to emphasize
+that a particular formula $\Phi_{C,K}$ contains the version
+$\mcycle_j(\vec{e})$.
 
@@
 The lightweight version is defined as
+% \[
+%   \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
+% \]
+% which means only that $\Te$ is nonempty.
 \[
-  \mcycle_1(\vec{e})=\bigvee_{t\in\delta}e_t
-\]
-which means only that $\Te$ is nonempty. This condition is satisfied by
-every cycle.
+  \mcycle_1(\vec{e})=\bigvee_{\text{SCC } S}~~\big(\bigvee_{t\in\delta_S}e_t ~~\wedge \bigwedge_{t'\in\delta\smallsetminus\delta_S}\neg e_{t'}\big)
+\]
+which means only that $\Te$ is nonempty and $\Te\subseteq\delta_S$ for
+some SCC $S$.
+This condition is satisfied by every cycle.
 
 The formula $\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$ actually
-says that for every nonempty set $T\subseteq\delta$, $T$ satisfies
-$\varphi$ if and only if $\rem(T)$ satisfies $\psi$. Note that the
-formula is not affected by the placement of transitions in the
-automaton. The only aspect of a transition $t$ reflected by the
-formula is its set of marks $\mks(t)$. Hence, we do not have to
-distinguish between transitions with the same sets of marks.
+says that for every nonempty set $T\subseteq\delta_S$ whenere $S$ is
+an SCC, $T$ satisfies $\varphi$ if and only if $\rem(T)$ satisfies
+$\psi$. Note that the formula is not affected by the placement of
+transitions in the automaton. \todo{Tady to predelat, aby to
+  reflektovalo $T\subseteq\delta_S$.}  The only aspect of a transition
+$t$ reflected by the formula is its set of marks $\mks(t)$. Hence, we
+do not have to distinguish between transitions with the same sets of
+marks.
 
@@
   \[
-    \textrm{for every nonempty set $T\subseteq\delta$ it holds }
+    \textrm{for every set $\emptyset\neq T\subseteq\delta$ it holds }
       (T~\msat~\varphi~\iff~\rem(T)~\msat~\psi).\tag{$1$}\label{eq:lem1} 
@@
   \[
-    \textrm{for every nonempty set $T\subseteq\delta$ it holds }
+    \textrm{for every set $\emptyset\neq T\subseteq\delta$ it holds }
     (T~\msat~\varphi~\iff~\rem'(T)~\msat~\psi).\tag{$2$}\label{eq:lem2} 
@@
   \[
-    T~\msat~\varphi~\iff~\overline{T}~\msat~\varphi~\iff~\rem(\overline{T})~\msat~\psi~\iff~\rem'(T)~\msat~\psi
-  \]
-  which proves that \eqref{eq:lem2} holds for $\rem'$ and $\psi$. 
+    \setlength{\arraycolsep}{0pt}
+    \begin{array}{rl}
+      T~\msat~\varphi~ 
+      & \iff~\overline{T}~\msat~\varphi \\
+      & \iff~\rem(\overline{T})~\msat~\psi\\
+      & \iff~\rem'(T)~\msat~\psi
+    \end{array}
+  \]
+  which proves that \eqref{eq:lem2} holds for $\rem'$ and $\psi$.\qed
 \end{proof}
@@
 \subsection{Intermediate version $\mcycle_2(\vec{e})$}
-The intermediate version says that $\Te$ is nonempty and for each state
-$q\in Q$ it holds that $\Te$ contains a transition leading to $q$ if
-and only if it contains a transition leading from $q$.
-Formally,
+The intermediate version says that $\Te$ is nonempty,
+$\Te\subseteq\delta_S$ for some SCC $S$, and for each state $q\in Q$
+it holds that $\Te$ contains a transition leading to $q$ if and only
+if it contains a transition leading from $q$. Formally,
 \[
-  \mcycle_2(\vec{e})=\bigvee_{t\in\delta}e_t~\wedge~
+%  \mcycle_2(\vec{e})=\bigvee_{t\in\delta}e_t~\wedge~
+  \mcycle_2(\vec{e})=\mcycle_1(\vec{e})~\wedge~
   \bigwedge_{q\in Q}\Big(\bigvee_{t'\in\delta~\cap~Q\times\Sigma\times 2^M\times\{q\}}\hspace{-1cm}e_{t'}\hspace{1cm}\iff
@@
   \begin{tikzpicture}[smallautomaton]
-    \begin{scope}[initial angle=90,xshift=-5cm]
+    \begin{scope}[initial angle=90,xshift=-4cm]
       \node[state,initial] (p) at (0,0) {};
@@
 
-    \begin{scope}[initial angle=90,densely dotted,black!20,xshift=5cm]
+    \begin{scope}[initial angle=90,densely dotted,black!20,xshift=4cm]
       \node[state,initial,solid,draw=black] (p) at (0,0) {};
@@
   contraposition. Assume that a nonempty set $T$ is not a cycle. We
-  show that there is a set $S\subseteq Q$ such neither (\ref{en:1})
-  nor (\ref{en:2}) nor (\ref{en:3}) holds. Let $p$ be a state such
-  that some transition of $T$ leads from $p$. We define the set
+  show that there is a set $S\subseteq Q$ such that neither
+  (\ref{en:1}) nor (\ref{en:2}) nor (\ref{en:3}) holds. Let $p$ be a
+  state such that some transition of $T$ leads from $p$. We define the
+  set $\post$ of states reachable from $p$ via transitions in $T$ and the set
+  $\pre$ of states from which $p$ is reachable via transtitions of $T$.
   \[
-    \post=\{p\}\cup\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }p\textrm{ to }q\}
-  \]
-  of states reachable from $p$ via transitions in $T$ and the set
-  \[
+    \begin{array}{r}
+    \post=\{p\}\cup\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }p\textrm{ to }q\}\\[.5ex]
     \pre=\{p\}\cup\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }q\textrm{ to }p\}
-  \]  
-  of states from which $p$ is reachable via transtitions of $T$. As
-  there is a transition of $T$ leading from $p$, we have that
+    \end{array}
+  \]  
+  As there is a transition of $T$ leading from $p$, we have that
   $T\not\subseteq\delta_{Q\smallsetminus\post}$ and
@@
   contains all transitions of $T$ and leads from $p$ back to $p$,
-  which means that $T$ is a cycle.  This is a contradiction.
+  which means that $T$ is a cycle.  This is a contradiction.\qed
 \end{proof}
@@
     \begin{array}{ll}
-      1 & \textrm{if }q\in S\\
+      1~~ & \textrm{if }q\in S\\[.5ex]
       0 & \textrm{otherwise}    
@@
 
-\subsection{SCC-based optimization}
-
-Now we present an optimized formula construction based on the fact
-that every cycle is completely included in the transition set
-$\delta_S$ of some SCC $S$. Note that we cannot reduce the number of
-acceptance marks in each SCC separately as all SCCs have to share the
-same acceptance formula after the reduction. However, for each SCC we
-can build the formula $\Phi_{C,K}$ separately and then we can put all
-these formulas into one conjunction such that they share the variables
-$\vec{i},\vec{f}$.
-
-Now we describe the construction of optimized QBF formulas
-$\Phi^\mathit{opt}_{j,C,K}$ formally. Given an automaton
-$\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$ and its SCC $S$, by $\mA^S$ we
-denote the automaton corresponding to the part of $\mA$ delimited by
-$S$, i.e., $\mA^S=(S,M,\Sigma,q',\delta_S,\varphi)$ where $q'$ is an
-arbitrary state of $S$. Further, let $\vec{n}_S$ denote the vector of
-variables $n_{t,k}$ where $t\in\delta_S$ and $k\in N_K$. By
-$\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})$ we denote the formula
-$\Phi_{j,C,K}$ constructed for $\mA^S$ as presented above.
-Finally, we define the optimized QBF formula as
-\[
-  \Phi^\mathit{opt}_{j,C,K}(\vec{n}',\vec{i},\vec{f})=\bigwedge_{\textrm{SCC }S}\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})
-\]
-where $\vec{n}'$ is the vector of all varibles contained in vectors
-$\vec{n}_S$ for SCCs $S$.
-
-One effect of this optimization is that the formula
-$\Phi^\mathit{opt}_{j,C,K}$ contains no variables $n_{t,k}$ %and $e_t$
-for transitions $t$ that do not lead between states of the same SCC.
-The acceptance marks on such a transition $t$ do not affect the
-acceptance of any run as $t$ appears at most once on each run. For
-these transitions $t$, we can define $\rem(t)$ such that
+\subsection{Optimizations}
+
+Finally, we mention two simple optimizations of the formula
+construction, which are always applied in the rest of the paper.
+
+The first optimization is based on the fact that every cycle is
+completely included in the transition set $\delta_S$ of some SCC
+$S$. Hence, all transitions $t$ that do not lead between states of the
+same SCC can be completely ignored during the formula
+construction. The acceptance marks on such a transition $t$ do not
+affect the acceptance of any run as $t$ appears at most once on each
+run. For these transitions $t$, we can define $\rem(t)$ such that
 $\mks(\rem(t))=\emptyset$.
+
+The second optimization is specific for $\Phi_{3,C,K}$. In
+construction of $\mcycle_3(\vec{e})$, we replace the subformula
+$\bigvee_{t\in\delta}e_t$ enforcing the nonemptiness of $\Te$ by
+$\mcycle_2(\vec{e})$. This modification does not change the overall
+semantics of $\mcycle_3(\vec{e})$, but our prelimnary experiments
+showed that QBF solvers can often solve the modified formula
+$\Phi_{3,C,K}$ faster.
+
+% \subsection{SCC-based optimization}
+
+% Now we present an optimized formula construction based on the fact
+% that every cycle is completely included in the transition set
+% $\delta_S$ of some SCC $S$. Note that we cannot reduce the number of
+% acceptance marks in each SCC separately as all SCCs have to share the
+% same acceptance formula after the reduction. However, for each SCC we
+% can build the formula $\Phi_{C,K}$ separately and then we can put all
+% these formulas into one conjunction such that they share the variables
+% $\vec{i},\vec{f}$.
+
+% Now we describe the construction of optimized QBF formulas
+% $\Phi^\mathit{opt}_{j,C,K}$ formally. Given an automaton
+% $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$ and its SCC $S$, by $\mA^S$ we
+% denote the automaton corresponding to the part of $\mA$ delimited by
+% $S$, i.e., $\mA^S=(S,M,\Sigma,q',\delta_S,\varphi)$ where $q'$ is an
+% arbitrary state of $S$. Further, let $\vec{n}_S$ denote the vector of
+% variables $n_{t,k}$ where $t\in\delta_S$ and $k\in N_K$. By
+% $\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})$ we denote the formula
+% $\Phi_{j,C,K}$ constructed for $\mA^S$ as presented above.
+% Finally, we define the optimized QBF formula as
+% \[
+%   \Phi^\mathit{opt}_{j,C,K}(\vec{n}',\vec{i},\vec{f})=\bigwedge_{\textrm{SCC }S}\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})
+% \]
+% where $\vec{n}'$ is the vector of all varibles contained in vectors
+% $\vec{n}_S$ for SCCs $S$.
+
+% One effect of this optimization is that the formula
+% $\Phi^\mathit{opt}_{j,C,K}$ contains no variables $n_{t,k}$ %and $e_t$
+% for transitions $t$ that do not lead between states of the same SCC.
+% The acceptance marks on such a transition $t$ do not affect the
+% acceptance of any run as $t$ appears at most once on each run. For
+% these transitions $t$, we can define $\rem(t)$ such that
+% $\mks(\rem(t))=\emptyset$.
 
@@
 \emph{level~1}, \emph{level~2}, or \emph{level~3} reduction when
-$\Phi^\mathit{opt}_{1,C,K}$, $\Phi^\mathit{opt}_{2,C,K}$, or
-$\Phi^\mathit{opt}_{3,C,K}$ is used, respectively.
+$\Phi_{1,C,K}$, $\Phi_{2,C,K}$, or
+$\Phi_{3,C,K}$ is used, respectively.
 
@@
 %       \While{$\neg\mathit{stop}~\wedge~C>1$}{
-%         $\Phi\gets\Phi^\mathit{opt}_{j,C-1,K}$ for $\mA$\;
+%         $\Phi\gets\Phi_{j,C-1,K}$ for $\mA$\;
 %         \leIf{$\mathit{satisfiable}(\Phi)$}{$C\gets C{-}1$}{$\mathit{stop}\gets\true$}
@@
 %     }{
-%       compute $\rem$ and $\psi$ from a model of $\Phi^\mathit{opt}_{j,C,K}$\;
+%       compute $\rem$ and $\psi$ from a model of $\Phi_{j,C,K}$\;
 %       \KwRet $(Q,N_K,\Sigma,q_I,\rem(\delta),\psi)$
@@
   \PI{$\mathit{SingleLevelReduction}(\mA,j,\mathit{reduceC})$}{
-    \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, level $j\in\{1,2,3\}$, $\mathit{reduceC}\in\{\true,\false\}$}
+    \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, $j\in\{1,2,3\}$, $\mathit{reduceC}\in\{\true,\false\}$}
     \KwOut{an equivalent TELA with the same structure as $\mA$ and with at most as many acceptance marks as in $\mA$} %with the same or lower number of acceptance~marks}
@@
   %     }
-    $C_\mA\gets$ the number of clauses in the acceptance formula of $\mA$ transformed to DNF\;
+    $C_\mA\gets$ the number of clauses in the formula $\varphi$ transformed to~DNF\;
     $K_\mA\gets$ the number of acceptance marks in $\mA$\;
@@
     $K\gets K_\mA$\;
-%    \While(\tcp*[f]{reduction of the number of marks}){$K>1~\wedge~\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$}{$K\gets K{-}1$}
-    \lWhile{$K>1~\wedge~\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$}{$K\gets K{-}1$}
+%    \While(\tcp*[f]{reduction of the number of marks}){$K>1~\wedge~\mathit{satisfiable}(\Phi_{j,C,K{-}1})$}{$K\gets K{-}1$}
+    \lWhile{$K>1~\wedge~\mathit{satisfiable}(\Phi_{j,C,K{-}1})$}{$K\gets K{-}1$}
     \If{$K=1$}{
-      \If(\tcp*[f]{check the condition $\true$}){all cycles in $\mA$ are accepting}{ %$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,1,0})$}{
+      \If(\tcp*[f]{check the condition $\true$}){all cycles in $\mA$ are accepting}{ %$\mathit{satisfiable}(\Phi_{j,1,0})$}{
         \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\true)$ where $\delta'$ is $\delta$ with all marks removed
       }
-      \If(\tcp*[f]{check the condition $\false$}){all cycles in $\mA$ are rejecting}{ %$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,0,0})$}{
+      \If(\tcp*[f]{check the condition $\false$}){all cycles in $\mA$ are rejecting}{ %$\mathit{satisfiable}(\Phi_{j,0,0})$}{
         \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all marks removed
@@
     \If(\tcp*[f]{reduction of the number of clauses}){$\mathit{reduceC}$}{
-      \lWhile{$C>1~\wedge~\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C-1,K})$}{$C\gets C{-}1$}
+      \lWhile{$C>1~\wedge~\mathit{satisfiable}(\Phi_{j,C-1,K})$}{$C\gets C{-}1$}
     }
     \If{$K<K_\mA~\vee~C<C_\mA$}{
-      compute $\rem$ and $\psi$ from a model of $\Phi^\mathit{opt}_{j,C,K}$\;
-      \KwRet $(Q,N_K,\Sigma,q_I,\rem(\delta),\psi)$
+      compute $\rem$ and $\psi$ from a model of $\Phi_{j,C,K}$\;
+      \KwRet $(Q,N_K,\Sigma,\rem(\delta),q_I,\psi)$
     }
@@
 until $K=1$ is reached or the QBF solver behind the function
-$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$ fails to reduce
+$\mathit{satisfiable}(\Phi_{j,C,K{-}1})$ fails to reduce
 the number of marks, i.e., it claims unsatisfiability of the formula
@@
 \textbf{while} loop builds the formula
-$\Phi=\Phi^\mathit{opt}_{j,C,K{-}1}$ only in the first iteration. In
+$\Phi=\Phi_{j,C,K{-}1}$ only in the first iteration. In
 each subsequent iteration, it extends this formula with a condition
@@
 The presented reduction algorithms have been implemented in a tool
-called \telatko. The tool is implemented in Python 3 and uses the Spot
-library~\cite{duret.16.atva2} for automata parsing and manipulation
+called \telatko. It is implemented in Python 3 and uses the Spot
+library~\cite{duret.16.atva2} for automata parsing and manipulation,
 and the theorem prover Z3~\cite{demoura.08.tacas} to solve the
-satisfiability of QBF formulas. \telatko is available at
-\url{https://gitlab.fi.muni.cz/xschwar3/telatko} under the GNU GPLv3
+satisfiability of QBF formulas. Our tool is available at
+\begin{center}
+\url{https://gitlab.fi.muni.cz/xschwar3/telatko}
+\end{center}
+under the GNU GPLv3
 license. The tool can be executed by the command 
@@
 If some call of the function
-$\mathit{satisfiable}(\Phi^\mathit{opt}_{j,C,K{-}1})$ in the first
+$\mathit{satisfiable}(\Phi_{j,C,K{-}1})$ in the first
 \textbf{while} loop of Algorithm~\ref{alg:single} does not return
@@
 \texttt{U} if the formula is unsatisfiable or \texttt{T} if the solver
-did not decided within the time limit. If $X$ is \texttt{T}, we longer
-timeout may lead to further reductions. a If the multi-level reduction
+did not decided within the time limit. If $X$ is \texttt{T}, a longer
+timeout may lead to further reductions. If the multi-level reduction
 is used, the automaton name contains the information from all
@@
 state-of-the-art translators from the Owl
-library~\cite{kretinsky.18.atva} and to nondeterministic TELA by
+library~\cite{kretinsky.18.atva}, namely \texttt{ltl2dela} (known as
+Delag)~\cite{muller.17.gandalf}, and \texttt{ltl2dgra} (known as
+Rabinzer~4)~\cite{kretinsky.18.cav}, and to nondeterministic TELA by
 another two state-of-the-art translators, namely \texttt{ltl2tgba}
@@
 \texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
-produced by individual tools contains less than 348 automata as not all
-of the formulas are successfully translated due to a timeout of 60
+produced by individual tools contains less than 348 automata as not
+all of the formulas are successfully translated due to a timeout of 60
 seconds. Further, we have removed automata with 0 or 1 acceptance
@@
 \centering
-\begin{tabular}{lr@{~~~}lr}
+\begin{tabular}{lr@{~~~~~~}lr}
 \toprule
-tool        & version&homepage & automata \\
+tool        & version&homepage & ~~~automata \\
 \midrule
-\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
+\texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
 \texttt{ltl2dgra} & 21.0 & & 234\\
@@
 multi-level reduction, always with incremental approach and without
-reduction of the number of clauses. The timeout for each QBF query was
+reducing of the number of clauses. The timeout for each QBF query was
 set to 30 seconds.
-
-For the experiments, we used the version of \texttt{telatko} build with
-Spot library version 2.10.4 and Z3 version 4.8.15. The experiments
-have been run on a computer with Intel® Core™ i7-8700 processor and 32
-GB of memory running Ubuntu 20.04.4. We used the tool
+%
+All reductions have been performed by the tool \texttt{telatko}
+build with Spot library version 2.10.4 and Z3 version 4.8.15. The
+experiments have been run on a computer with Intel$^\text{®}$ Core™ i7-8700
+processor and 32 GB of memory running Ubuntu 20.04.4. We used the tool
 \texttt{autcross} of the Spot library to get the statistics of the
@@
 \label{tab:heatmap}
-\setlength{\tabcolsep}{4pt}
+\setlength{\tabcolsep}{2.5pt}
 \centering
@@
 \label{fig:quantile}
+\bigskip
 \centering
@@
 without altering automata structure. We have implemented the method in
-a tool called \texttt{telatko}. The presented experimental results
+a tool called \texttt{telatko}. The current applications of the tool
+are twofold. First, it can reduce the number of acceptance marks of a
+given TELA. Second, it discloses how tools producing TELA are
+economical with acceptance marks. The presented experimental results
 show that the tool can indeed reduce the number of acceptance marks in
-automata produced by state-of-the-art LTL to automata translators. The
-reduction was significant on automata produced by the translators of
-the Owl library.
+automata produced by all considered state-of-the-art LTL to automata
+translators on formulas from literature. Further, it clearly shows
+that the translators of the Owl library are significantly less
+economical with acceptance marks than the other two translators.
 
@@
 
-%\bibliographystyle{plain}
+\bibliographystyle{plain}
+%\bibliographystyle{abbrvnat}
 \bibliography{mc}

### 2022-10-14 paper/main.tex
typo fix

@@
 $T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and only
-if $\rem(T)$ satisifes $\psi$. This can be roughly denoted by the
+if $\rem(T)$ satisfies $\psi$. This can be roughly denoted by the
 formula
@@
   \begin{itemize}
-  \item $T$ contains a transtition $t\in\delta_S$ and a transition
+  \item $T$ contains a transition $t\in\delta_S$ and a transition
     $t'\in\delta_{Q\smallsetminus S}$. The definition of a cycle
@@
     state $p$ back to $p$ containing both $t$ and $t'$. However, this
-    implies that $T$ contains a transtion leading from a state in $S$
-    to a state outside $S$ and a transtion leading from a state outside $S$
+    implies that $T$ contains a transition leading from a state in $S$
+    to a state outside $S$ and a transition leading from a state outside $S$
     to a state in $S$.
@@
   set $\post$ of states reachable from $p$ via transitions in $T$ and the set
-  $\pre$ of states from which $p$ is reachable via transtitions of $T$.
+  $\pre$ of states from which $p$ is reachable via transitions of $T$.
   \[
@@
 
-The second optimization is specific for $\Phi_{3,C,K}$. In
+The second optimization is specific for $\Phi_{3,C,K}$. In the
 construction of $\mcycle_3(\vec{e})$, we replace the subformula
@@
 $\mcycle_2(\vec{e})$. This modification does not change the overall
-semantics of $\mcycle_3(\vec{e})$, but our prelimnary experiments
+semantics of $\mcycle_3(\vec{e})$, but our preliminary experiments
 showed that QBF solvers can often solve the modified formula
@@
 \texttt{U} if the formula is unsatisfiable or \texttt{T} if the solver
-did not decided within the time limit. If $X$ is \texttt{T}, a longer
+did not decide within the time limit. If $X$ is \texttt{T}, a longer
 timeout may lead to further reductions. If the multi-level reduction
@@
 column \textit{solver timeout} shows the number of automata for which
-the last query to QBF solver did not finished within the 30 seconds
+the last query to QBF solver did not finish within the 30 seconds
 limit. The timeout of the last QBF query implies that the automaton
@@
 
-In future, we plan to improve the reduction process (e.g., use the
+In the future, we plan to improve the reduction process (e.g., use the
 bisection method to reduce the number of acceptance marks more

### 2022-10-14 paper/main.tex
polishing of 3.1

--- 2022-10-14 paper/main.tex
@@
 
-The formula $\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$ actually
-says that for every nonempty set $T\subseteq\delta_S$ whenere $S$ is
-an SCC, $T$ satisfies $\varphi$ if and only if $\rem(T)$ satisfies
-$\psi$. Note that the formula is not affected by the placement of
-transitions in the automaton. \todo{Tady to predelat, aby to
-  reflektovalo $T\subseteq\delta_S$.}  The only aspect of a transition
-$t$ reflected by the formula is its set of marks $\mks(t)$. Hence, we
-do not have to distinguish between transitions with the same sets of
-marks.
-
-Formally, we define an equivalence ${\sim}\subseteq\delta\times\delta$
-on transitions such that $t_1\sim t_2$ whenever $\mks(t_1)=\mks(t_2)$.
-\begin{lemma}
-  Assume that there is a function $\rem$ and a formula $\psi$ such that 
+The formula $\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$ says that
+for every nonempty set $T\subseteq\delta_S$ whenere $S$ is an SCC, $T$
+satisfies $\varphi$ if and only if $\rem(T)$ satisfies $\psi$. Note
+that the only aspects of a transition $t$ reflected by the formula are
+its set of marks $\mks(t)$ and its affiliation to an SCC.
+% In partcular, it is not affected by its placement inside the
+% corresponding SCCs. The only aspect of a transition $t$ reflected by
+% the formula is its set of marks $\mks(t)$.
+Hence, we do not have to distinguish between transitions that are
+affiliated to the same SCC an have the same sets of marks.
+
+Let us now fix an SCC $S$. We define an equivalence
+${\sim_S}\subseteq\delta_S\times\delta_S$ on transitions % in $\delta_S$
+such that $t_1\sim_S t_2$ whenever $\mks(t_1)=\mks(t_2)$.
+\begin{lemma} Assume that there is a function $\rem$ and a formula
+  $\psi$ such that
   \[
-    \textrm{for every set $\emptyset\neq T\subseteq\delta$ it holds }
+    \textrm{for every set $\emptyset\neq T\subseteq\delta_S$ it holds }
       (T~\msat~\varphi~\iff~\rem(T)~\msat~\psi).\tag{$1$}\label{eq:lem1} 
@@
   Then there exists a function $\rem'$ that respects % the equivalence
-  $\sim$ (i.e.,
-  if $t_1\sim t_2$ then $\rem'(t_1)=\rem'(t_2)$) and
+  $\sim_S$ (i.e.,
+  if $t_1\sim_S t_2$ then $\rem'(t_1)=\rem'(t_2)$) and
   \[
-    \textrm{for every set $\emptyset\neq T\subseteq\delta$ it holds }
+    \textrm{for every set $\emptyset\neq T\subseteq\delta_S$ it holds }
     (T~\msat~\varphi~\iff~\rem'(T)~\msat~\psi).\tag{$2$}\label{eq:lem2} 
@@
   \eqref{eq:lem1} holds. To construct the function $\rem'$, we first
-  select one transition from each equivalence class of $\sim$. For
-  every transition $t\in\delta$, by $\overline{t}$ we denote the
-  selected transition equivalent to $t$. We define the function
-  $\rem'$ as $\rem'(t)=\rem(\overline{t})$. Clearly, $\rem'$ respects
-  the equivalence $\sim$. It remains to show that \eqref{eq:lem2}
-  holds for $\rem'$ and $\psi$.
-
-  Let $T\subseteq\delta$ be an arbitrary nonempty set. We construct
+  select one transition from each equivalence class of~$\sim_S$. For
+  every transition $t\in\delta_S$, by $\overline{t}$ we denote the
+  selected transition equivalent to $t$ and we define the function
+  $\rem'$ as $\rem'(t)=\rem(\overline{t})$. Note that we do not need
+  to discuss the value of $\rem'$ on transitions outside $\delta_S$ as
+  it is not relevant for the lemma. Clearly, $\rem'$ respects the
+  equivalence $\sim_S$. It remains to show that \eqref{eq:lem2} holds
+  for $\rem'$ and $\psi$.
+
+  Let $T\subseteq\delta_S$ be a % an arbitrary
+  nonempty set. We construct
   the set $\overline{T}=\{\overline{t}\mid t\in T\}$. As
-  $\mks(t)=\mks(\overline{t})$ for all transitions, we get
-  $\mks(T)=\mks(\overline{T})$ and thus
+  $\mks(t)=\mks(\overline{t})$ for all transitions of $\delta_S$, we
+  get $\mks(T)=\mks(\overline{T})$ and thus
   \[
@@
 $\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$. Before we build the
-formula, we compute the equivalence $\sim$ on automata transitions and
-temporarily remove all transitions except one of each equivalence class. Then we
-build the formula $\Phi_{1,C,K}$ for the pruned automaton. The more
-transitions we removed, the shorter formula with less $e_t$ variables
-we obtain. If the formula $\Phi_{1,C,K}$ for the pruned automaton is
-satisfiable, we derive $\rem$ and $\psi$ from its model and extend % use them
+formula, we compute the equivalences $\sim_S$ for all SCCs and
+temporarily remove all transitions affilited to SCCs except one of
+each equivalence class. Then we build the formula $\Phi_{1,C,K}$ for
+the pruned automaton. The more transitions we removed, the shorter
+formula with less $e_t$ variables we obtain. If the formula
+$\Phi_{1,C,K}$ for the pruned automaton is satisfiable, we derive
+$\rem$ and $\psi$ from its model and extend % use them
 % with the original automaton. We only need to extend
@@
   $p$ back to $p$ and containing $t_i$. If we concatenate all these
-  paths, we get the path $\rho_1\rho_2\ldots\rho_{|T|}\in T^+$ that
+/  paths, we get the path $\rho_1\rho_2\ldots\rho_{|T|}\in T^+$ that
   contains all transitions of $T$ and leads from $p$ back to $p$,
@@
 The second optimization is specific for $\Phi_{3,C,K}$. In
-construction of $\mcycle_3(\vec{e})$, we replace the subformula
+the construction of $\mcycle_3(\vec{e})$, we replace the subformula
 $\bigvee_{t\in\delta}e_t$ enforcing the nonemptiness of $\Te$ by

### 2022-10-14 paper/main.tex

@@
 
-The second optimization is specific for $\Phi_{3,C,K}$. In
-the construction of $\mcycle_3(\vec{e})$, we replace the subformula
+The second optimization is specific for $\Phi_{3,C,K}$. In the
+construction of $\mcycle_3(\vec{e})$, we replace the subformula
 $\bigvee_{t\in\delta}e_t$ enforcing the nonemptiness of $\Te$ by
 $\mcycle_2(\vec{e})$. This modification does not change the overall
-semantics of $\mcycle_3(\vec{e})$, but our prelimnary experiments
+semantics of $\mcycle_3(\vec{e})$, but our preliminary experiments
 showed that QBF solvers can often solve the modified formula

### 2022-10-14 paper/main.tex
fixed typo + TODO (kdyby snad nahodou)

@@
 formula, we compute the equivalences $\sim_S$ for all SCCs and
-temporarily remove all transitions affilited to SCCs except one of
+temporarily remove all transitions affiliated to SCCs except one of
 each equivalence class. Then we build the formula $\Phi_{1,C,K}$ for

### 2023-03-08 paper/main.tex
minor typo
main in LIPICS style

@@
 
-\documentclass[runningheads]{llncs}
-
-%\usepackage[paperheight=235mm, paperwidth=155mm,textwidth=12.2cm,textheight=19.3cm]{geometry}
+\documentclass[a4paper,UKenglish,cleveref,autoref,thm-restate]{lipics-v2021}
+
+\bibliographystyle{plainurl}% the mandatory bibstyle
+
 \usepackage{multirow}
@@
 
-\makeatletter
-\RequirePackage[bookmarks,unicode,colorlinks=true]{hyperref}%
-   \def\@citecolor{blue}%
-   \def\@urlcolor{blue}%
-   \def\@linkcolor{blue}%
-\def\UrlFont{\rmfamily}
-\def\orcidID#1{\smash{\href{http://orcid.org/#1}{\protect\raisebox{-1.25pt}{\protect\includegraphics{orcid_color.eps}}}}}
-\makeatother
-
-\usepackage{academicons}
-\definecolor{orcidlogocol}{HTML}{A6CE39}
+% \makeatletter
+% \RequirePackage[bookmarks,unicode,colorlinks=true]{hyperref}%
+%    \def\@citecolor{blue}%
+%    \def\@urlcolor{blue}%
+%    \def\@linkcolor{blue}%
+% \def\UrlFont{\rmfamily}
+% \def\orcidID#1{\smash{\href{http://orcid.org/#1}{\protect\raisebox{-1.25pt}{\protect\includegraphics{orcid_color.eps}}}}}
+% \makeatother
+
+% \usepackage{academicons}
+% \definecolor{orcidlogocol}{HTML}{A6CE39}
 
@@
 
+%{{{ title and authors
+
+\title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving}
+\author{Tereza Schwarzov\'{a}}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
+\author{Jan Strej\v{c}ek}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{strejcek@fi.muni.cz}{https://orcid.org/0000-0001-5873-403X}{}
+\author{Juraj Major}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{}{}
+\authorrunning{T. Schwarzov\'{a}, J. Strej\v{c}ek, and J. Major}
+\Copyright{Tereza Schwarzov\'{a}, Jan Strej\v{c}ek, and Juraj Major}
+\ccsdesc[500]{Theory of computation~Logic}
+\ccsdesc[500]{Theory of computation~Automata over infinite objects}
+\keywords{Emerson-Lei automata, TELA, automata reduction, QBF, telatko}
+%\funding{This work has been supported by the Czech Science Foundation grant GA19-24397S.}
+
+%}}}
+
 \begin{document}
-
-%{{{ title and authors
-
-\title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving}
-\author{Tereza Schwarzov\'{a} \and
-        Jan Strej\v{c}ek\orcidID{0000-0001-5873-403X}\thanks{Corresponding author: \email{strejcek@fi.muni.cz}} \and
-        Juraj Major}
-\authorrunning{T. Schwarzov\'{a}, J. Strej\v{c}ek, and J. Major}
-\titlerunning{Reducing Acceptance Marks in TELA by QBF-solving}
-\institute{%Faculty of Informatics,
-  Masaryk University, Brno, Czech Republic
-}
-
-%}}}
 
@@
 Hence, we do not have to distinguish between transitions that are
-affiliated to the same SCC an have the same sets of marks.
+affiliated to the same SCC and have the same sets of marks.
 

### 2023-03-09 paper/main.tex
fixes to the transformation to LIPIcs

@@
 
-\title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving}
-\author{Tereza Schwarzov\'{a}}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
-\author{Jan Strej\v{c}ek}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{strejcek@fi.muni.cz}{https://orcid.org/0000-0001-5873-403X}{}
-\author{Juraj Major}{Faculty of Informatics, Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{}{}
+\title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF-Solving}
+\author{Tereza Schwarzov\'{a}}{%Faculty of Informatics,
+  Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
+\author{Jan Strej\v{c}ek}{%Faculty of Informatics,
+  Masaryk University, Brno, Czech Republic}{strejcek@fi.muni.cz}{https://orcid.org/0000-0001-5873-403X}{}
+\author{Juraj Major}{%Faculty of Informatics,
+  Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{}{}
 \authorrunning{T. Schwarzov\'{a}, J. Strej\v{c}ek, and J. Major}
@@
   \]
-  which proves that \eqref{eq:lem2} holds for $\rem'$ and $\psi$.\qed
+  which proves that \eqref{eq:lem2} holds for $\rem'$ and $\psi$.%\qed
 \end{proof}
@@
   contains all transitions of $T$ and leads from $p$ back to $p$,
-  which means that $T$ is a cycle.  This is a contradiction.\qed
+  which means that $T$ is a cycle.  This is a contradiction.%\qed
 \end{proof}

### 2023-03-09 paper/main.tex
Upravy na zaklade Martinovych pripominek (part I)

@@
   number of %different
-  acceptance marks in TELA with the use of \emph{quantified Boolean
+  acceptance marks in TELA based on satisfiability of \emph{quantified Boolean
     formulas (QBF)}.
@@
 Their popularity comes probably from the fact that these automata can
-often use less states than equivalent automata with simpler acceptance
+often use fewer states than equivalent automata with simpler acceptance
 conditions. Further, algorithms handling TELA can automatically handle
@@
 
-We present such a technique heavily based on \emph{quantified Boolean
-  formulas (QBF)}. For a given TELA and parameters $c,k$, it produces
+Our reduction technique is heavily based on \emph{quantified Boolean
+  formulas (QBF)}. For a given TELA and parameters $C,K$, it produces
 a QBF formula which is satisfiable if and only if there exists an
-automaton with the same structure, $k$ acceptance marks, an acceptance
-formula in disjunctive normal form with $c$ clauses, and the same set
-of accepting runs as the original automaton. The placement of the
-marks on transitions and the acceptance formula can be obtained from a
-model of the formula. Besides this formula, we describe also the
-construction of two simpler formulas such that their satisfiability
-implies the existence of an automaton with the same structure, $k$
-acceptance marks, and the same set of accepting runs, but not vice
-versa.
+automaton with the same structure, $K$ acceptance marks, an acceptance
+formula in disjunctive normal form with $C$ cubes (i.e., conjunctive
+clauses), and the same set of accepting runs as the original
+automaton. The placement of the marks on transitions and the
+acceptance formula can be obtained from a model of the
+formula. Besides this formula, we describe also the construction of
+two simpler formulas whose satisfiability implies the existence of an
+automaton with the same structure, $K$ acceptance marks, and the same
+set of accepting runs, but not vice versa.
 
@@
 There is a simple technique~\cite{babiak.13.spin} reducing the number
-of acceptance marks in \emph{transition-based genealised Büchi
+of acceptance marks in \emph{transition-based generalised Büchi
   automata (TGBA)} without changing its structure. We are not aware
@@
 
-In this section we recall the basic terms around TELA and QBF.
+In this section we recall the basic terms related to TELA and QBF.
 
@@
 An acceptance formula $\varphi$ is in \emph{disjunctive normal formal
-  (DNF)} if it is a disjunction of clauses, where each clause is a
+  (DNF)} if it is a disjunction of cubes, where each cube is a
 conjunction of terms of the form $\Fin m$ or $\Inf m$.  Each
@@
 in~DNF. Formula $\false$ corresponds to the disjunction of zero
-clauses and formula $\true$ corresponds to one clause with zero terms.
+cubes and formula $\true$ corresponds to one cube with zero terms.
 
@@
 that we are looking for a new acceptance formula $\psi$ in DNF. Note
-that our reduction method can be easily adapted to look for $\psi$ of
-a different shape.
+that our reduction method can be easily adapted to look for $\psi$ in
+CNF or in a different shape.
 
@@
 construction is parametrized by two integers $C,K\ge 0$, where $K$ is
-the desired number of acceptance marks and $C$ is the number of
-clauses of $\psi$. For simplicity, we assume that the reduced
+the desired number of acceptance marks and $C$ is the number of cubes
+of $\psi$. Without loss of generality, we assume that the reduced
 automaton will use the acceptance marks $N_K=\{1,2,\ldots,K\}$.
@@
   variables $i_{c,k}$ and $f_{c,k}$ say whether the $c^\textrm{th}$
-  clause of $\psi$ contains terms $\Inf k$ or $\Fin k$, respectively.
+  cube of $\psi$ contains terms $\Inf k$ or $\Fin k$, respectively.
   \[i_{c,k}=\left\{
       \begin{array}{ll}
-        1~~ & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Inf k$}\\[.5ex]
+        1~~ & \textrm{if the $c^\textrm{th}$ cube of $\psi$ contains $\Inf k$}\\[.5ex]
         0 & \textrm{otherwise}    
@@
       \begin{array}{ll}
-        1~~ & \textrm{if the $c^\textrm{th}$ clause of $\psi$ contains $\Fin k$}\\[.5ex]
+        1~~ & \textrm{if the $c^\textrm{th}$ cube of $\psi$ contains $\Fin k$}\\[.5ex]
         0 & \textrm{otherwise}    
@@
 structure of $\psi$. As we assume that $\psi$ is a disjunction of $C$
-clauses, we have
+cubes, we have
 \[
@@
 where each $\xi_{c,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ corresponds to
-one clause. Recall that the presence of terms $\Inf k$ and $\Fin k$ in
-the $c^\textrm{th}$ clause is given by variables $i_{c,k}$ and
-$f_{c,k}$, respecively. $\Inf k$ is satisfied by $\rem(T)$ iff $T$
+one cube. Recall that the presence of terms $\Inf k$ and $\Fin k$ in
+the $c^\textrm{th}$ cube is given by variables $i_{c,k}$ and
+$f_{c,k}$, respectively. $\Inf k$ is satisfied by $\rem(T)$ iff $T$
 contains a transition $t$ such that $k\in\mks(\rem(t))$, which can be
@@
   that $\Te$ is nonempty and $\Te\subseteq\delta_S$ for some SCC $S$.
-  Except for SCCs, it does not use the information about automaton
+  Except for SCCs, it does not use the information about the automaton
   structure, but it comes with an interesting simplification of the
@@
   \]
-  Now we apply $\eqref{eq:lem1}$ on $\overline{T}$ to get that
+  Now we apply $\eqref{eq:lem1}$ to $\overline{T}$ and we get % that
   \[
@@
 
-    \begin{scope}[initial angle=90,densely dotted,black!20]
+    \begin{scope}[initial angle=90,densely dotted,black!35]
       \node[state,initial,solid,draw=black] (p) at (0,0) {};
@@
 
-    \begin{scope}[initial angle=90,densely dotted,black!20,xshift=4cm]
+    \begin{scope}[initial angle=90,densely dotted,black!35,xshift=4cm]
       \node[state,initial,solid,draw=black] (p) at (0,0) {};
@@
     are not cycles even if $\mcycle_2(\vec{e})$ holds (middle and
-    right)}
+    right).}
   \label{fig:inter}
@@
   $p$ back to $p$ and containing $t_i$. If we concatenate all these
-/  paths, we get the path $\rho_1\rho_2\ldots\rho_{|T|}\in T^+$ that
+  paths, we get the path $\rho_1\rho_2\ldots\rho_{|T|}\in T^+$ that
   contains all transitions of $T$ and leads from $p$ back to $p$,
@@
 %   %     }
-%     $C_\mA\gets$ the number of clauses in the acceptance formula of $\mA$ transformed to DNF\;
+%     $C_\mA\gets$ the number of cubes in the acceptance formula of $\mA$ transformed to DNF\;
 %     $K_\mA\gets$ the number of acceptance marks in $\mA$\;
@@
 %     }  
-%     \If(\tcp*[f]{reduction of the number of clauses}){$\mathit{reduceC}$}{
+%     \If(\tcp*[f]{reduction of the number of cubes}){$\mathit{reduceC}$}{
 %       $\mathit{stop}\gets\false$\;
@@
   %     }
-    $C_\mA\gets$ the number of clauses in the formula $\varphi$ transformed to~DNF\;
+    $C_\mA\gets$ the number of cubes in the formula $\varphi$ transformed to~DNF\;
     $K_\mA\gets$ the number of acceptance marks in $\mA$\;
@@
     }  
-    \If(\tcp*[f]{reduction of the number of clauses}){$\mathit{reduceC}$}{
+    \If(\tcp*[f]{reduction of the number of cubes}){$\mathit{reduceC}$}{
       \lWhile{$C>1~\wedge~\mathit{satisfiable}(\Phi_{j,C-1,K})$}{$C\gets C{-}1$}
@@
 given in Algorithm~\ref{alg:single}. Besides the reduction of
-acceptance marks, the algorithm also reduces the number of clauses in the acceptance formula if the last argument $\mathit{reduceC}$ is set to
+acceptance marks, the algorithm also reduces the number of cubes in the acceptance formula if the last argument $\mathit{reduceC}$ is set to
 $\true$.
@@
 mark. Otherwise, if $\mathit{reduceC}$ is set to $\true$ then the
-procedure gradually reduces the number of clauses in the second
+procedure gradually reduces the number of cubes in the second
 \textbf{while} loop. Note that the loop never checks for acceptance
-condition with 0 clauses as it is equivalent to $\false$ and this case
+condition with 0 cubes as it is equivalent to $\false$ and this case
 was treated above. Finally, if the procedure succeeds to reduce the
-number of marks or clauses, it constructs the modified
+number of marks or cubes, it constructs the modified
 automaton. Otherwise, it returns the original automaton.
@@
     single-level reductions: an input automaton $\mA$ and the automata
-    obtained by reducing it with level 1, level 2, and level 3.}
+    obtained by reducing it with level 1, level 2, and level~3.}
   \label{fig:example}
@@
 \item[\texttt{-L }$j$] specifies the reduction level; if omitted, the multi-level reduction is used,
-\item[\texttt{-C}] switches on the reduction of the number of clauses
+\item[\texttt{-C}] switches on the reduction of the number of cubes
   after the number of marks is reduced (it corresponds to
@@
 \item[\texttt{-I}] switches on the incremental version,
-\item[\texttt{-T }$t$] sets the timeout for QBF query to $t$ seconds (the default value is 50 seconds), 
+\item[\texttt{-T }$t$] sets the timeout for each QBF query to $t$ seconds (the default value is 50 seconds), 
 \item[\texttt{-O <output.hoa>}] specifies the output file; if omitted,
@@
 \texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
-produced by individual tools contains less than 348 automata as not
+produced by the individual tools contains less than 348 automata as not
 all of the formulas are successfully translated due to a timeout of 60
@@
 multi-level reduction, always with incremental approach and without
-reducing of the number of clauses. The timeout for each QBF query was
+reducing of the number of cubes. The timeout for each QBF query was
 set to 30 seconds.
@@
 All reductions have been performed by the tool \texttt{telatko}
-build with Spot library version 2.10.4 and Z3 version 4.8.15. The
+built with Spot library version 2.10.4 and Z3 version 4.8.15. The
 experiments have been run on a computer with Intel$^\text{®}$ Core™ i7-8700
@@
 column \textit{solver timeout} shows the number of automata for which
-the last query to QBF solver did not finish within the 30 seconds
+the last query to QBF solver did not finish within the 30~seconds
 limit. The timeout of the last QBF query implies that the automaton
@@
 given number of acceptance marks after multi-level reduction. The
-table indicates that in many cases only 1 or 2 marks can be
+table indicates that in many cases only 1 or 2~marks can be
 saved. However, the achieved reduction is substantial for some

### 2023-03-10 paper/main.tex
few more changes to the paper

@@
 Their popularity comes probably from the fact that these automata can
-often use fewer states than equivalent automata with simpler acceptance
-conditions. Further, algorithms handling TELA can automatically handle
-all automata with traditional acceptance conditions. TELA can be
-obtained for example by translating LTL formulas by \texttt{ltl2dela}
-(known as Delag)~\cite{muller.17.gandalf},
-\texttt{ltl3tela}~\cite{major.19.atva}, \texttt{ltl2dgra} (known as
-Rabinizer~4)~\cite{kretinsky.18.cav}, or \texttt{ltl2tgba} which is a
-part of the Spot library~\cite{duret.16.atva2}.  There are also
-algorithms processing these automata, for example the emptiness
+often use fewer states than equivalent automata with simpler
+acceptance conditions. Further, algorithms handling TELA can
+automatically handle all automata with traditional acceptance
+conditions. TELA can be obtained for example by translating LTL
+formulas with tools \texttt{ltl2dela} (known as
+Delag)~\cite{muller.17.gandalf} or \texttt{ltl2dgra} (known as
+Rabinizer~4)~\cite{kretinsky.18.cav} of the Owl library,
+\texttt{ltl2tgba} of the Spot library~\cite{duret.16.atva2}, or
+\texttt{ltl3tela}~\cite{major.19.atva}. There are also algorithms
+processing these automata, for example the emptiness
 check~\cite{baier.19.atva} or translation of TELA to parity
@@
 in the size of the original QBF. Let $V$ be the set of all
-propositional variables. A mapping $\mu:V\rightarrow\{\true,\false\}$
+propositional variables. A mapping $\mu:V\rightarrow\{0,1\}$
 is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
@@
 \centering
-\begin{tabular}{lr@{~~~~~~}lr}
+% \begin{tabular}{lr@{~~~~~~}lr}
+% \toprule
+% tool        & version&homepage & ~~~automata \\
+% \midrule
+% \texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
+% \texttt{ltl2dgra} & 21.0 & & 234\\
+% \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
+% \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
+% \bottomrule
+% \end{tabular}
+%
+\begin{tabular}{lrlrrr}
   \toprule
-tool        & version&homepage & ~~~automata \\
+  &&&& \multicolumn{2}{c}{automata with} \\ \cmidrule(lr){5-6}
+  tool &\!\!\!version&homepage & T/O & 0--1 mark & \textbf{2+ marks}\\
   \midrule
-\texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
-\texttt{ltl2dgra} & 21.0 & & 234\\
-\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
-\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
+  \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}}
+       & ? & ? & \textbf{129} \\
+  \texttt{ltl2dgra} & 21.0 &
+       & ? & ? & \textbf{234}\\
+  \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr}
+       & ? & ? & \textbf{70}\\
+  \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela}
+       & ? & ? & \textbf{91}\\
   \bottomrule

### 2023-03-14 paper/main.tex

@@
 
-\title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF-Solving}
+\title{Reducing Acceptance Marks in Emerson-Lei Automata by QBF Solving}
 \author{Tereza Schwarzov\'{a}}{%Faculty of Informatics,
@@
 \begin{abstract}
-  \emph{Transition-based Emerson-Lei automata (TELA)} became a popular
-  formalism that generalizes many traditional kinds of automata over
-  infinite words including Büchi, co-Büchi, Rabin, Streett, and parity
-  automata. Transitions in a TELA are labelled with acceptance marks
-  and its accepting formula is a positive Boolean combination of terms
-  saying that a particular mark has to be visited infinitely or
-  finitely often. % in an accepting run.
+  This paper presents a novel application of QBF solving to automata
+  reduction. We focus on \emph{Transition-based Emerson-Lei automata
+    (TELA)}, which is a popular formalism that generalizes many
+  traditional kinds of automata over infinite words including Büchi,
+  co-Büchi, Rabin, Streett, and parity automata. Transitions in a TELA
+  are labelled with acceptance marks and its accepting formula is a
+  positive Boolean combination of atoms saying that a particular mark
+  has to be visited infinitely or finitely
+  often. % in an accepting run.
   Algorithms processing these automata are often very sensitive to the
@@
   number of %different
-  acceptance marks in TELA based on satisfiability of \emph{quantified Boolean
-    formulas (QBF)}.
+  acceptance marks in TELA based on satisfiability of \emph{quantified
+    Boolean formulas (QBF)}.
   % : if a constructed QBF formula is satisfiable then
@@
   (DNF)} if it is a disjunction of cubes, where each cube is a
-conjunction of terms of the form $\Fin m$ or $\Inf m$.  Each
+conjunction of atoms of the form $\Fin m$ or $\Inf m$.  Each
 acceptance formula can be transformed into an equivalent formula
 in~DNF. Formula $\false$ corresponds to the disjunction of zero
-cubes and formula $\true$ corresponds to one cube with zero terms.
+cubes and formula $\true$ corresponds to the cube with zero atoms.
 
@@
   variables $i_{c,k}$ and $f_{c,k}$ say whether the $c^\textrm{th}$
-  cube of $\psi$ contains terms $\Inf k$ or $\Fin k$, respectively.
+  cube of $\psi$ contains atoms $\Inf k$ or $\Fin k$, respectively.
   \[i_{c,k}=\left\{
@@
 \begin{itemize}
-\item all terms of the form $\Inf m$ by $\bigvee_{t\in\delta_m}e_t$ and
-\item all terms of the form $\Fin m$ by $\bigwedge_{t\in\delta_m}\neg e_t$.
+\item all atoms of the form $\Inf m$ by $\bigvee_{t\in\delta_m}e_t$ and
+\item all atoms of the form $\Fin m$ by $\bigwedge_{t\in\delta_m}\neg e_t$.
 \end{itemize}
@@
 where each $\xi_{c,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ corresponds to
-one cube. Recall that the presence of terms $\Inf k$ and $\Fin k$ in
+one cube. Recall that the presence of atoms $\Inf k$ and $\Fin k$ in
 the $c^\textrm{th}$ cube is given by variables $i_{c,k}$ and
@@
 \caption{Performance of \telatko on different automata sets. The $x$ axis
-represents the number of reduced automata, the $y$ axis shows the time spent
+represents the number of processed automata, the $y$ axis shows the time spent
 reducing (in seconds).}
@@
 an equivalent automaton with the same structure and an acceptance
-formula of a specific form (e.g., without any $\Fin m$ terms). Even
+formula of a specific form (e.g., without any $\Fin m$ atoms). Even
 though the QBF queries can be time-consuming, in practice one can

### 2023-03-17 paper/main.tex
upravy v textu clanku

@@
 \emph{color appearance record}~\cite{renkin.20.atva} transforms a TELA
-with $s$ states and $m$ acceptance marks into a parity automaton with
-up to $s\cdot m!$ states. Further, the emptiness
-check~\cite{baier.19.atva} is exponential in the number of acceptance
-marks that appear in acceptance primitives saying that a mark has to
-be visited finitely often, while it is only polynomial in other
-measures of the input automaton.
+with $m$ acceptance marks and $s$ states into a parity automaton with
+up to $m!\cdot s$ states. Further, the emptiness-check
+algorithm~\cite{baier.19.atva} is exponential in the number of
+acceptance marks that appear in acceptance primitives saying that a
+mark has to be visited finitely often, while it is only polynomial in
+other measures of the input automaton.
 
@@
 terms used in the paper. Section~\ref{sec:qbf} explains the
-construction of three mentioned QBF formulas. The reduction algorithm
+construction of the three mentioned QBF formulas. The reduction algorithm
 based on these formulas is presented in
@@
 
-An acceptance formula $\varphi$ is in \emph{disjunctive normal formal
+An acceptance formula $\varphi$ is in \emph{disjunctive normal form
   (DNF)} if it is a disjunction of cubes, where each cube is a
@@
 a desired instance of $\psi$ and $\rem$. In the following, we assume
-that we are looking for a new acceptance formula $\psi$ in DNF. Note
-that our reduction method can be easily adapted to look for $\psi$ in
-CNF or in a different shape.
-
-Now we describe the construction of QBF formulas in detail. The
+that we are looking for a new acceptance formula $\psi$ in DNF. The
+choice of DNF is not fundamental, but inherited from our previous
+attempt to reduce acceptance formulas. The presented method can be
+easily adapted to look for $\psi$ in CNF or in a different shape.
+
+Now we describe the construction of the QBF formulas in detail. The
 construction is parametrized by two integers $C,K\ge 0$, where $K$ is
@@
 \]
-where the subformulas $\mcycle(\vec{e})$, $\msat_\varphi(\vec{e})$,
-and $\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ are defined
-as follows. The subformula $\mcycle(\vec{e})$ will be defined as the last
+where $\forall\vec{e}$ denotes the sequence composed of $\forall e_t$
+for all variables $e_t$ and the subformulas $\mcycle(\vec{e})$,
+$\msat_\varphi(\vec{e})$, and
+$\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ are defined as
+follows. The subformula $\mcycle(\vec{e})$ will be defined as the last
 one and we present three versions of it.
@@
   Altogether, we obtain
+  % \[
+  %   \setlength{\arraycolsep}{0pt}
+  %   \begin{array}{rl}
+  %     T~\msat~\varphi~ 
+  %     & \iff~\overline{T}~\msat~\varphi \\
+  %     & \iff~\rem(\overline{T})~\msat~\psi\\
+  %     & \iff~\rem'(T)~\msat~\psi
+  %   \end{array}
+  % \]
   \[
-    \setlength{\arraycolsep}{0pt}
-    \begin{array}{rl}
       T~\msat~\varphi~ 
-      & \iff~\overline{T}~\msat~\varphi \\
-      & \iff~\rem(\overline{T})~\msat~\psi\\
-      & \iff~\rem'(T)~\msat~\psi
-    \end{array}
+      \iff~\overline{T}~\msat~\varphi
+      \iff~\rem(\overline{T})~\msat~\psi
+      \iff~\rem'(T)~\msat~\psi
   \]  
@@
   $T\not\subseteq\delta_{Q\smallsetminus\pre}$, i.e., (\ref{en:2})
-  does not holds for $\post$ and $\pre$. Further, the definition of
+  does not hold for $\post$ and $\pre$. Further, the definition of
   $\post$ implies that there is no transition of $T$ leading from a
@@
 
+\todo{Sem nebo za tu dalsi subsection strcity analyzu slozitosti.}
+
 \subsection{Optimizations}
@@
 $\Phi_{3,C,K}$ faster.
+
+\todo{Zminit, ze nase QBF formule navic rikaji, ze v zadne
+  klauzulu/cude neni zaroven $\Fin m$ a $\Inf m$ a ze cela formule
+  ma alepson 1 $\Inf$ nebo $\Fin$? Je to tak, Terko?}
+
+\todo{Zminit, ze jsme zkouseli symmetry breaking, ale nefungoval?}
 
@@
 \]
-The second \textbf{while} can be transformed to an incremental version
-similarly. The incremental approach benefits from the fact that some
-QBF solvers can decide an extended formula faster as they reuse the
-information computed when solving the original formula.
+The second \textbf{while} loop can be transformed to an incremental
+version similarly. The incremental approach benefits from the fact
+that some QBF solvers can decide an extended formula faster as they
+reuse the information computed when solving the original formula.
 
@@
 
+\todo{Nekde zminit, ze inkrementalni pristup je implementovan jen pro prvni while loop? Je to tak, Terko?}
+
 If some call of the function
@@
 
+\todo{Upravit experimenty: pridat dalsi 2 sady automatu.}
 We have used the tool \texttt{genltl} of the Spot
@@
 multi-level reduction, always with incremental approach and without
-reducing of the number of cubes. The timeout for each QBF query was
-set to 30 seconds.
+reducing of the number of cubes.
+\todo{Vysvetlit, proc neredukujeme cubes?}
+The timeout for each QBF query was set to 30 seconds.
 %
@@
 
+\todo{popisek Fig. 3 neni hezky. Je to wall time nebo CPU time?}
 \begin{figure}[b!]
-\caption{Performance of \telatko on different automata sets. The $x$ axis
-represents the number of processed automata, the $y$ axis shows the time spent
-reducing (in seconds).}
+  \caption{Performance of \telatko on different automata sets. The $x$
+    axis represents the number of automata that are processed by
+    telatko in the time shown on the $y$ axis.}
 \label{fig:quantile}

### 2023-03-17 paper/main.tex
verze pro prvni submission na sat 2023

@@
 \newcommand{\todo}[1]{\marginpar{\textcolor{red}{#1}}}
+%\newcommand{\todo}[1]{}
 %\def\Inf{\ensuremath{\mathsf{Inf}}}
@@
 
-\todo{Sem nebo za tu dalsi subsection strcity analyzu slozitosti.}
+\todo{Sem nebo za tu dalsi subsection strcit analyzu slozitosti.}
 
@@
 \centering
-% \begin{tabular}{lr@{~~~~~~}lr}
+\begin{tabular}{lr@{~~~~~~}lr}
+\toprule
+tool        & version&homepage & ~~~automata \\
+\midrule
+\texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
+\texttt{ltl2dgra} & 21.0 & & 234\\
+\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
+\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
+\bottomrule
+\end{tabular}
+%
+% \begin{tabular}{lrlrrr}
 %   \toprule
-% tool        & version&homepage & ~~~automata \\
+%   &&&& \multicolumn{2}{c}{automata with} \\ \cmidrule(lr){5-6}
+%   tool &\!\!\!version&homepage & T/O & 0--1 mark & \textbf{2+ marks}\\
 %   \midrule
-% \texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
-% \texttt{ltl2dgra} & 21.0 & & 234\\
-% \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
-% \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
+%   \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}}
+%        & ? & ? & \textbf{129} \\
+%   \texttt{ltl2dgra} & 21.0 &
+%        & ? & ? & \textbf{234}\\
+%   \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr}
+%        & ? & ? & \textbf{70}\\
+%   \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela}
+%        & ? & ? & \textbf{91}\\
 %   \bottomrule
 % \end{tabular}
-%
-\begin{tabular}{lrlrrr}
-  \toprule
-  &&&& \multicolumn{2}{c}{automata with} \\ \cmidrule(lr){5-6}
-  tool &\!\!\!version&homepage & T/O & 0--1 mark & \textbf{2+ marks}\\
-  \midrule
-  \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}}
-       & ? & ? & \textbf{129} \\
-  \texttt{ltl2dgra} & 21.0 &
-       & ? & ? & \textbf{234}\\
-  \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr}
-       & ? & ? & \textbf{70}\\
-  \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela}
-       & ? & ? & \textbf{91}\\
-  \bottomrule
-\end{tabular}
 \end{table}
@@
 reducing of the number of cubes.
-\todo{Vysvetlit, proc neredukujeme cubes?}
+\todo{Vysvetlit, proc neredukujeme pocet cubes?}
 The timeout for each QBF query was set to 30 seconds.
@@
 
-\todo{popisek Fig. 3 neni hezky. Je to wall time nebo CPU time?}
+\todo{popisek Fig. 3 neni hezky. Je to wall time nebo CPU time? Je to
+  kumulovany cas nebo pro kazdy automat zvlast? Zmenit popisky os v
+  obrazku?}
 \begin{figure}[b!]
   \caption{Performance of \telatko on different automata sets. The $x$
-    axis represents the number of automata that are processed by
-    telatko in the time shown on the $y$ axis.}
+    axis represents the number of automata from each automata
+    set that are processed by \telatko in the time shown on the $y$
+    axis.}
 \label{fig:quantile}
@@
 
-\bibliographystyle{plain}
 %\bibliographystyle{abbrvnat}

### 2023-03-19 paper/main.tex
s/TELA automaton/TELA/

@@
 without increasing the number of states. There exists only a SAT-based
-approach that transforms a deterministic TELA automaton to an
+approach that transforms a deterministic TELA to an
 equivalent automaton with a given acceptance condition and a given
@@
 This section explains how we use the QBF formulas constructed in the
-previous section to reduce the number of acceptance marks in TELA
-automata. First, we describe a \emph{single-level} reduction, which
+previous section to reduce the number of acceptance marks in TELA.
+First, we describe a \emph{single-level} reduction, which
 uses only a single kind of QBF formulas. More precisely, we talk about

### 2023-03-20 paper/main.tex
complexity analysis added (and a bit more)

--- 2023-03-17 paper/main.tex
@@
 \usepackage{enumerate}
+\usepackage{tablefootnote}
 
@@
 \newcommand{\mB}{\mathcal{B}}
+\newcommand{\bigO}{\mathcal{O}}
 
@@
 there is a path from $q_I$ to $q$) as states that are not reachable
-from $q_I$ can be eliminated without any impact on $L(\mA)$.
+from $q_I$ can be eliminated without any impact on $L(\mA)$. We also
+assume that each TELA has at least one SCC as automata without any SCC
+trivially desribe an empty language.
 
@@
 
-\todo{Sem nebo za tu dalsi subsection strcity analyzu slozitosti.}
-
-\subsection{Optimizations}
+\subsection{Complexity of formulas}
+The constructed formulas $\Phi_{j,C,K}$ for $j\in\{1,2,3\}$ use
+$|\delta|$ universally quantified variables $e_t$, $|\delta|\cdot K$
+free variables $n_{t,k}$, and $C\cdot K$ free variables $i_{c,k}$ and
+$f_{c,k}$. The formula $\Phi_{3,C,K}$ additionally uses $|Q|$
+variables $s_q$ that are existentially quantified (when the formula is
+transformed to prenex normal form).
+
+To analyze the length of the formulas, we start with its subformulas.
+One can easily check that
+$|\msat_\varphi(\vec{e})|\in\bigO(|\varphi|\cdot|\delta|)$ and
+$|\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})|\in\bigO(C\cdot
+K\cdot|\delta|)$. Further,
+$|\mcycle_1(\vec{e})|,|\mcycle_2(\vec{e})|\in\bigO(S\cdot|\delta|)$, % and
+%$|\mcycle_2(\vec{e})|\in\bigO(S\cdot|\delta|+|\delta|)=\bigO(S\cdot|\delta|)$,
+where $S$ is the number of SCCs in the automaton.  Next,
+$|\mcycle_3(\vec{e})|\in\bigO(|\delta|+|Q|)$, which can be simplified
+to $|\mcycle_3(\vec{e})|\in\bigO(|\delta|)$ as $|Q|\le|\delta|$
+follows from the assumptions that all states are reachable and each
+automaton has at least one SCC. Altogether, we get
+$|\Phi_{1,C,K}|,|\Phi_{2,C,K}|\in\bigO(S\cdot|\delta|+|\varphi|\cdot|\delta|+C\cdot
+K\cdot|\delta|)=\bigO\big((S+|\varphi|+C\cdot K)\cdot|\delta|\big)$
+and
+$|\Phi_{3,C,K}|\in\bigO(|\delta|+|\varphi|\cdot|\delta|+C\cdot
+K\cdot|\delta|)=\bigO\big((|\varphi|+C\cdot K)\cdot|\delta|\big)$.
+
+\subsection{Optimizations of formulas}
 
@@
 $\bigvee_{t\in\delta}e_t$ enforcing the nonemptiness of $\Te$ by
-$\mcycle_2(\vec{e})$. This modification does not change the overall
-semantics of $\mcycle_3(\vec{e})$, but our preliminary experiments
-showed that QBF solvers can often solve the modified formula
-$\Phi_{3,C,K}$ faster.
+$\mcycle_2(\vec{e})$. This modification prolongs the formula, but it
+does not change the overall semantics of $\mcycle_3(\vec{e})$ and our
+preliminary experiments showed that QBF solvers can often solve the
+modified formula $\Phi_{3,C,K}$ faster.
+
+We have also made some experiments with breaking the symmetries in the
+formula models. In particular, we have ordered new acceptance marks by
+their placements on transitions and we have ordered the cubes by their
+content. As the effect of these modifications was inconclusive, we do
+not describe it here.
 
@@
   ma alepson 1 $\Inf$ nebo $\Fin$? Je to tak, Terko?}
-
-\todo{Zminit, ze jsme zkouseli symmetry breaking, ale nefungoval?}
 
@@
 
-\begin{table}[b]
+\begin{table}[t]
 \caption{LTL to automata translators used for the experimental evaluation
@@
 %
-\begin{tabular}{lrlrrr}
+\begin{tabular}{lrrrr}
   \toprule
-  &&&& \multicolumn{2}{c}{automata with} \\ \cmidrule(lr){5-6}
-  tool &\!\!\!version&homepage & T/O & 0--1 mark & \textbf{2+ marks}\\
+  &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
+  tool &\!\!\!version/homepage & timeout & 0 or 1 mark & \textbf{at least 2 marks}\\
   \midrule
-  \texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}}
+  \texttt{ltl2dela} & 21.0\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
        & ? & ? & \textbf{129} \\
-  \texttt{ltl2dgra} & 21.0 &
+  \texttt{ltl2dgra} & 21.0\footnotemark[1]
        & ? & ? & \textbf{234}\\
-  \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr}
+  \texttt{ltl2tgba} & 2.10.4\tablefootnote{\url{https://spot.lrde.epita.fr}}
        & ? & ? & \textbf{70}\\
-  \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela}
+  \texttt{ltl3tela} & 2.2.0\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
        & ? & ? & \textbf{91}\\
@@
 multi-level reduction, always with incremental approach and without
-reducing of the number of cubes.
-\todo{Vysvetlit, proc neredukujeme cubes?}
+reducing of the number of cubes. We do not reduce the number of cubes
+as our primary aim is to reduce the number of acceptance marks.
+%\todo{Vysvetlit, proc neredukujeme pocet cubes?}
 The timeout for each QBF query was set to 30 seconds.

### 2023-03-20 paper/main.tex
typo
changes in experiments (text)

@@
 assume that each TELA has at least one SCC as automata without any SCC
-trivially desribe an empty language.
+trivially describe an empty language.
 
@@
 K\cdot|\delta|)=\bigO\big((|\varphi|+C\cdot K)\cdot|\delta|\big)$.
+Note that the formula $\Phi_{3,C,K}$ is asymptotically shorter than
+$\Phi_{1,C,K}$ and $\Phi_{2,C,K}$, but it contains an additional quantifier
+alternation.
 
@@
 
-Finally, we mention two simple optimizations of the formula
+Finally, we mention several three simple optimizations of the formula
 construction, which are always applied in the rest of the paper.
@@
 
+The third optimization adds to each formula $\Phi_{j,C,K}$ a conjunct
+\[
+   \bigwedge_{c\in\{1,2,\ldots,C\}}\bigwedge_{k\in N_K} \neg
+   i_{c,k}~\vee~\neg f_{c,k}
+\]
+saying that any cube does not contain both $\Inf k$ and $\Fin k$ for any
+$k$ as such a cube cannot be satisfied by any run.
+
 We have also made some experiments with breaking the symmetries in the
@@
 
-\todo{Zminit, ze nase QBF formule navic rikaji, ze v zadne
-  klauzulu/cude neni zaroven $\Fin m$ a $\Inf m$ a ze cela formule
-  ma alepson 1 $\Inf$ nebo $\Fin$? Je to tak, Terko?}
+% \todo{Zminit, ze cela formule
+%   ma alepson 1 $\Inf$ nebo $\Fin$? Je to tak, Terko?}
 
@@
 
-\todo{Nekde zminit, ze inkrementalni pristup je implementovan jen pro prvni while loop? Je to tak, Terko?}
-
 If some call of the function
@@
 reduce the number of marks to
-1.%when tryint the automaton may be potentially reduced further.
+1.%when trying the automaton may be potentially reduced further.
 
@@
 
-\todo{Upravit experimenty: pridat dalsi 2 sady automatu.}
+%\todo{Upravit experimenty: pridat dalsi 2 sady automatu.}
 We have used the tool \texttt{genltl} of the Spot
@@
 library~\cite{kretinsky.18.atva}, namely \texttt{ltl2dela} (known as
-Delag)~\cite{muller.17.gandalf}, and \texttt{ltl2dgra} (known as
+Delag)~\cite{muller.17.gandalf} and \texttt{ltl2dgra} (known as
 Rabinzer~4)~\cite{kretinsky.18.cav}, and to nondeterministic TELA by
@@
 from the Spot library~\cite{duret.16.atva2} and
-\texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
-produced by the individual tools contains less than 348 automata as not
-all of the formulas are successfully translated due to a timeout of 60
-seconds. Further, we have removed automata with 0 or 1 acceptance
-marks as there is a little point in reducing these. Table
+\texttt{ltl3tela}~\cite{major.19.atva}. Each of the four considered
+tools failed on some formulas: it usually reached a timeout of 60
+seconds or it produced an automaton that cannot be parsed by the Spot
+library.
+% produced by the individual tools contains less than 348 automata as not
+% all of the formulas are successfully translated due to a timeout of 60
+% seconds.
+Further, we have removed automata with 0 or 1 acceptance marks as
+there is a little point in reducing these. Table
 \ref{tab:ltltranslators} shows the exact versions and homepages of
-used translators, and the number of automata produced by each
-translator after removing the automata with less than two marks.
+used translators. For each translator, it also provides the numbers of
+fails, produced automata with less than two marks, and automata with
+at least two marks that are actually used for the experimental
+evaluation of our reduction technique.
 
 \begin{table}[t]
-\caption{LTL to automata translators used for the experimental evaluation
-and the number of successfully translated formulas.}
+  \caption{LTL to automata translators used for the experimental
+    evaluation and the numbers of fails and successfully constructed
+    automata with less than two marks and with at least two marks.}
 \label{tab:ltltranslators}
@@
   &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
-  tool &\!\!\!version/homepage & timeout & 0 or 1 mark & \textbf{at least 2 marks}\\
+  tool &\!\!\!version/homepage & fails & 0 or 1 mark & \textbf{at least 2 marks}\\
   \midrule

### 2023-03-20 paper/main.tex
Table 1: add numbers

@@
   \texttt{ltl2dela} & 21.0\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
-       & ? & ? & \textbf{129} \\
+       & 5 & 214 & \textbf{129} \\
   \texttt{ltl2dgra} & 21.0\footnotemark[1]
-       & ? & ? & \textbf{234}\\
+       & 12 & 102 & \textbf{234}\\
   \texttt{ltl2tgba} & 2.10.4\tablefootnote{\url{https://spot.lrde.epita.fr}}
-       & ? & ? & \textbf{70}\\
+       & 0 & 278 & \textbf{70}\\
   \texttt{ltl3tela} & 2.2.0\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
-       & ? & ? & \textbf{91}\\
+       & 18 & 239 & \textbf{91}\\
   \bottomrule

### 2023-03-20 paper/main.tex
changes in text of experiments

--- 2023-03-20 paper/main.tex
@@
 \ref{tab:ltltranslators} shows the exact versions and homepages of
-used translators. For each translator, it also provides the numbers of
-fails, produced automata with less than two marks, and automata with
-at least two marks that are actually used for the experimental
-evaluation of our reduction technique.
+used translators. For each translator, it also provides the number of
+fails, the number of produced automata with less than two marks, and
+the number of automata with at least two marks. The last number is
+typeset in bold as these automata are actually used for the
+experimental evaluation of our reduction technique.
 
@@
 
-For each automata set (that is, the automata produced by an individual
-LTL translator), we have applied all single-level reductions and the
-multi-level reduction, always with incremental approach and without
-reducing of the number of cubes. We do not reduce the number of cubes
-as our primary aim is to reduce the number of acceptance marks.
+For each automata set (that is, the automata with at least two marks
+produced by an individual LTL translator), we have applied all
+single-level reductions and the multi-level reduction, always with
+incremental approach and without reducing of the number of cubes. We
+do not reduce the number of cubes as our primary aim is to reduce the
+number of acceptance marks.
 %\todo{Vysvetlit, proc neredukujeme pocet cubes?}

### 2023-03-20 paper/main.tex
minor typo

@@
 The reduction of acceptance marks is not the only application of the
-presented approach. For example, it can be easily adopted to look for
+presented approach. For example, it can be easily adapted to look for
 an equivalent automaton with the same structure and an acceptance

### 2023-03-20 paper/main.tex
typos

@@
 exists). Further, there are some SAT-based approaches aimed to reduce
-the number of states of automata over infinite words. More preceisely,
+the number of states of automata over infinite words. More precisely,
 there are reductions designed for nondeterministic Büchi
@@
 information about all accepting and non-accepting automata cycles. We
-expect that ACD could potentialy be used to reduce the number of
+expect that ACD could potentially be used to reduce the number of
 acceptance marks or to simplify the acceptance condition. However,
@@
 Section~\ref{sec:algo}. Section~\ref{sec:implementation} describes our
-tool \telatko implemening the reduction technique.
+tool \telatko implementing the reduction technique.
 Experimental results are shown in
@@
 state $p$ to state $q$ labelled with $a$ and acceptance marks $M'$.
-The set $M'$ is also refered to by $\mks(t)$. For a set of transitions
+The set $M'$ is also referred to by $\mks(t)$. For a set of transitions
 $T\subseteq\delta$, let $\mks(T)=\bigcup_{t\in T}\mks(t)$ denote the
@@
 The formula $\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$ says that
-for every nonempty set $T\subseteq\delta_S$ whenere $S$ is an SCC, $T$
+for every nonempty set $T\subseteq\delta_S$ where $S$ is an SCC, $T$
 satisfies $\varphi$ if and only if $\rem(T)$ satisfies $\psi$. Note
@@
   after the number of marks is reduced (it corresponds to
-  $\mathit{reduceC}=\true$ in Algortihms~\ref{alg:single}
+  $\mathit{reduceC}=\true$ in Algorithms~\ref{alg:single}
   and~\ref{alg:multi}),
@@
 Delag)~\cite{muller.17.gandalf} and \texttt{ltl2dgra} (known as
-Rabinzer~4)~\cite{kretinsky.18.cav}, and to nondeterministic TELA by
+Rabinizer~4)~\cite{kretinsky.18.cav}, and to nondeterministic TELA by
 another two state-of-the-art translators, namely \texttt{ltl2tgba}

### 2023-03-20 paper/main.tex
changes in Fig.3

--- 2023-03-20 paper/main.tex
@@
 from the Spot library~\cite{duret.16.atva2} and
-\texttt{ltl3tela}~\cite{major.19.atva}. Each of the four considered
-tools failed on some formulas: it usually reached a timeout of 60
-seconds or it produced an automaton that cannot be parsed by the Spot
-library.
+\texttt{ltl3tela}~\cite{major.19.atva}. Three of the considered
+translators failed on some formulas: it usually reached a timeout of
+60 seconds or it produced an automaton that cannot be parsed by the
+Spot library.
 % produced by the individual tools contains less than 348 automata as not
@@
   &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
-  tool &\!\!\!version/homepage & fails & 0 or 1 mark & \textbf{at least 2 marks}\\
+  tool &\!\!\!version/homepage & ~~~~fails & less than 2 marks & \textbf{at least 2 marks}\\
   \midrule
@@
 
-For each automata set (that is, the automata with at least two marks
+To each automata set (that is, the automata with at least two marks
 produced by an individual LTL translator), we have applied all
@@
 Figure~\ref{fig:quantile} presents the time spent by multi-level
-reduction of automata in individual automata sets. The chart shows a
-pleasing finding that for every set, most automata are reduced in
-under a second and the high cumulative running times are caused by
-a few complicated automata.
-
-\todo{popisek Fig. 3 neni hezky. Je to wall time nebo CPU time? Je to
-  kumulovany cas nebo pro kazdy automat zvlast? Zmenit popisky os v
-  obrazku?}
+reduction on individual automata of each automata sets. The chart
+shows a pleasing finding that for every set, most automata are reduced
+in under a second and the high cumulative running times are caused by 
+a relatively small number of complicated automata.
+
 \begin{figure}[b!]
-  \caption{Performance of \telatko on different automata sets. The $x$
-    axis represents the number of automata from each automata
-    set that are processed by \telatko in the time shown on the $y$
-    axis.}
+  \caption{Performance of \telatko on individual automata of each automata sets.
+    The graph shows the time ($y$ axis) needed by \telatko
+    to process the $x$-th automaton of the set, where automata of each set are ordered by their processing time.
+    % The $x$ axis represents the number of automata from each automata
+    % set that are processed by \telatko in the time shown on the $y$
+    % axis.
+  }
 \label{fig:quantile}

### 2023-03-20 paper/main.tex
kill the orphan

@@
 called \emph{alternating cycle decomposition
-  (ACD)}~\cite{casares.21.icalp} which compactly represents the
-information about all accepting and non-accepting automata cycles. We
-expect that ACD could potentially be used to reduce the number of
+  (ACD)}~\cite{casares.21.icalp} which compactly represents
+%the information about
+all accepting and non-accepting automata cycles. We
+expect that ACD could %potentially
+be used to reduce the number of
 acceptance marks or to simplify the acceptance condition. However,

### 2023-03-20 paper/main.tex
polished third optimization

--- 2023-03-20 paper/main.tex
@@
 
-The third optimization adds to each formula $\Phi_{j,C,K}$ a conjunct
+The third optimization extends $\Phi_{j,C,K}$ into the conjunction
 \[
-   \bigwedge_{c\in\{1,2,\ldots,C\}}\bigwedge_{k\in N_K} \neg
-   i_{c,k}~\vee~\neg f_{c,k}
-\]
-saying that any cube does not contain both $\Inf k$ and $\Fin k$ for any
-$k$ as such a cube cannot be satisfied by any run.
+   \Phi_{j,C,K}~\wedge \bigwedge_{c\in\{1,2,\ldots,C\}}\bigwedge_{k\in N_K} (\neg
+   i_{c,k}\vee\neg f_{c,k}).
+\]
+The added part says that any cube does not contain both $\Inf k$ and
+$\Fin k$ for any $k$. A cube with both $\Inf k$ and $\Fin k$ would be
+useless as it cannot be satisfied by any run.
 
@@
 
-\begin{table}[t]
+\begin{table}[b!]
 \caption{The effect of reduction.
@@
 
-\begin{figure}[b!]
+\begin{figure}[t]
   \caption{Performance of \telatko on individual automata of each automata sets.

### 2023-05-20 paper/main.tex
last update for SAT 2023 submission!
SAT23 final: enhancements & typos

@@
 conditions. TELA can be obtained for example by translating LTL
-formulas with tools \texttt{ltl2dela} (known as
-Delag)~\cite{muller.17.gandalf} or \texttt{ltl2dgra} (known as
-Rabinizer~4)~\cite{kretinsky.18.cav} of the Owl library,
+(linear temporal logic) formulas with tools \texttt{ltl2dela}
+(known as Delag)~\cite{muller.17.gandalf} or \texttt{ltl2dgra}
+(known as Rabinizer~4)~\cite{kretinsky.18.cav} of the Owl library,
 \texttt{ltl2tgba} of the Spot library~\cite{duret.16.atva2}, or
@@
 
+By choosing an appropriate format of the acceptance condition, one can
+easily represent many classical kinds of automata over infinite words,
+e.g. a B\"uchi automaton is a TELA with $M = \{\tacc{1}{1}\}$,
+$\varphi = \Inf\tacc{1}{1}$, and the acceptance mark \tacc{1}{1} being
+placed on all accepting transitions.
+Similarly, a Rabin automaton with $k$ acceptance
+pairs has $M = \{\tacc{5}{1}, \tacc{3}{1'} \ldots, \tacc{5}{k}, \tacc{3}{k'}\}$
+and $\varphi = (\Fin\tacc{5}{1} \land \Inf\tacc{3}{1'}) \lor \ldots
+\lor (\Fin\tacc{5}{k} \land \Inf\tacc{3}{k'})$.
+A state-based acceptance can be represented by TELA by placing the
+corresponding acceptance marks on all transitions leaving the accepting
+states.
+
 \bigskip
@@
 attempt to reduce acceptance formulas. The presented method can be
-easily adapted to look for $\psi$ in CNF or in a different shape.
+easily adapted to look for $\psi$ in conjunctive normal form (CNF)
+or in a different shape.
 
@@
 $\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ are defined as
-follows. The subformula $\mcycle(\vec{e})$ will be defined as the last
-one and we present three versions of it.
+follows. Three versions of the subformula $\mcycle(\vec{e})$ will
+be presented.
 
@@
 it holds that $\Te$ contains a transition leading to $q$ if and only
-if it contains a transition leading from $q$. Formally,
+if it contains a transition leaving $q$. Formally,
 \[
@@
   A nonempty set $T\subseteq\delta$ is a cycle if and only if for each set of
-  states $S\subseteq Q$ it holds that
+  states $S\subseteq Q$ one of the following conditions hold.
   \begin{enumerate}[A.]
-  \item\label{en:1} all transitions in $T$ lead from a state in $S$ to a state in $S$ (i.e., $T\subseteq\delta_S$) or
-  \item\label{en:2} all transitions in $T$ lead from a state outside $S$ to a state outside $S$ (i.e., $T\subseteq\delta_{Q\smallsetminus S}$) or
+  \item\label{en:1} All transitions in $T$ lead from a state in $S$ to a state in $S$ (i.e., $T\subseteq\delta_S$).
+  \item\label{en:2} All transitions in $T$ lead from a state outside $S$ to a state outside $S$ (i.e., $T\subseteq\delta_{Q\smallsetminus S}$).
   \item\label{en:3} $T$ contains a transition leading from a state in $S$ to a state outside $S$ and a transition leading from a state outside
@@
   which means that (\ref{en:3}) does not hold for $\pre$. Now we prove
-  by contradition that (\ref{en:1}) does not hold for at least one of
+  by contradiction that (\ref{en:1}) does not hold for at least one of
   $\post,\pre$. Hence, let us assume that $T\subseteq\delta_\post$ and
@@
 
-Finally, we mention several three simple optimizations of the formula
+Finally, we mention three simple optimizations of the formula
 construction, which are always applied in the rest of the paper.
@@
 \]
-The added part says that any cube does not contain both $\Inf k$ and
-$\Fin k$ for any $k$. A cube with both $\Inf k$ and $\Fin k$ would be
-useless as it cannot be satisfied by any run.
+The added part says that no cube contains both $\Inf k$ and $\Fin k$
+for any $k$. A cube with both $\Inf k$ and $\Fin k$ would be useless
+as it cannot be satisfied by any run.
 
@@
 \textbf{while} loop of Algorithm~\ref{alg:single} does not return
-$\true$, then the name of the output automaton encodes the reason for
+$\true$, then the name of the output automaton, included in the
+generated HOA, encodes the reason for
 it. In the case of a single level reduction, the name has the form
@@
 library~\cite{duret.16.atva2} to generate LTL formulas from
-literature. The parametrized formulas were generated for all
+literature, using all available formula sets implemented in
+\texttt{genltl} 2.10.4.
+The parametrized formulas were generated for all
 combinations of parameter values from 1 to 4. We used the tool
@@
 single-level reductions and the multi-level reduction, always with
-incremental approach and without reducing of the number of cubes. We
+incremental approach and without reducing the number of cubes. We
 do not reduce the number of cubes as our primary aim is to reduce the
@@
 may be potentially further reduced by the reduction with a longer time
-limit.
+limit. The reduction ratio drop in higher levels, e.g. between level~1
+and level~2 for \texttt{ltl2dgra}, happens due to the QBF solver timeouts
+occuring earlier as the formula being solved in higher level is more
+complex. For \texttt{ltl2dela} and \texttt{ltl2dgra}, the automata sets
+do not contain any automaton where level~2 or level~3 achieves a better
+result than level~1. However, all levels contribute to the reductions in
+the multilevel setting.
 
@@
     time. The column \textit{solver timeout} indicates the number of
-    instances when the last call to the QBF solver timed out.  }
+    instances where the last call to the QBF solver timed out.  }
 \label{tab:cumm-data}
@@
 Figure~\ref{fig:quantile} presents the time spent by multi-level
-reduction on individual automata of each automata sets. The chart
+reduction on individual automata of each automata set. The chart
 shows a pleasing finding that for every set, most automata are reduced

### 2023-05-29 paper/main.tex
SAT 2023: small enhancements

--- 2023-03-20 paper/main.tex
@@
 
-TELA have attracted a lot of attention during the last five years.
+TELA have attracted a lot of attention during the last five
+years~\cite{baier.19.atva,major.19.atva,muller.17.gandalf,renkin.20.atva}.
 Their popularity comes probably from the fact that these automata can
@@
   formulas (QBF)}. For a given TELA and parameters $C,K$, it produces
-a QBF formula which is satisfiable if and only if there exists an
+a QBF which is satisfiable if and only if there exists an
 automaton with the same structure, $K$ acceptance marks, an acceptance
-formula in disjunctive normal form with $C$ cubes (i.e., conjunctive
-clauses), and the same set of accepting runs as the original
+formula in disjunctive normal form with $C$ cubes (i.e., conjunction
+of literals), and the same set of accepting runs as the original
 automaton. The placement of the marks on transitions and the
@@
 terms used in the paper. Section~\ref{sec:qbf} explains the
-construction of the three mentioned QBF formulas. The reduction algorithm
+construction of the three mentioned QBF. The reduction algorithm
 based on these formulas is presented in
@@
 propositional variables. A mapping $\mu:V\rightarrow\{0,1\}$
-is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
-assignment of an equivalent Boolean formula. A QBF formula is
+is a \emph{model} of a QBF $\varphi$ iff it is a satisfying
+assignment of an equivalent Boolean formula. A QBF is
 \emph{satisfiable} iff it has a model.
@@
 
-\section{Construction of QBF formulas}\label{sec:qbf}
+\section{Construction of formulas}\label{sec:qbf}
 
@@
 TELA $\mA =(Q,M,\Sigma,\delta,q_I,\varphi)$ without altering its
-structure and language. In other words, we are looking for
+structure and language. In other words, we look for
 \begin{itemize}
-\item a set $\oM$ of acceptance marks,
+\item a set $\oM$ of acceptance marks such that $|\oM|<|M|$,
 \item an acceptance formula $\psi$ over $\oM$, and 
@@
   some $\oM'\subseteq\oM$)
+\item such that the automaton
+$\mB=(Q,\oM,\Sigma,\rem(\delta),q_I,\psi)$ is equivalent to $\mA$.
 \end{itemize}
-such that $|\oM|<|M|$ and the automaton
-$\mB=(Q,\oM,\Sigma,\rem(\delta),q_I,\psi)$ is equivalent to $\mA$. To
-ensure that the modified automaton is equivalent to the original one,
+To ensure that the modified automaton is equivalent to the original one,
 we actually look for $\psi$ and $\rem$ such that each run
@@
 \[
-  \forall T\subseteq\delta~.~\mcycle(T)\implies\big(\msat_\varphi(T) ~\iff~ \msat_{\psi,\rem}(T)\big).
-\]
-In fact, this corresponds to the shape of the QBF formulas we will
+  \forall T\subseteq\delta~.~\mcycle(T)\implies\big(\msat_\varphi(T) ~\iff~ \msat_{\psi}(\rem(T))\big).
+\]
+In fact, this corresponds to the shape of the QBF we will
 construct. As we are looking for $\psi$ and $\rem$ such that the
-formula holds, the subformula $\msat_{\psi,\rem}(T)$ contains many
+formula holds, the subformula $\msat_{\psi}(\rem(T))$ contains many
 free variables representing possible instances of $\psi$ and
@@
 
-Now we describe the construction of the QBF formulas in detail. The
+Now we describe the construction of the QBF in detail. The
 construction is parametrized by two integers $C,K\ge 0$, where $K$ is
@@
 We start with a description of Boolean variables used in the
-constructed QBF formulas.
+constructed QBF.
 \begin{itemize}
@@
 variables of the form $e_t$, $n_{t,k}$, $i_{c,k}$, and $f_{c,k}$,
-respectively. The constructed QBF formulas have the form 
+respectively. The constructed QBF have the form
 \[
@@
   \end{tikzpicture}
-  \caption{An automaton structure where transition labels and
-    acceptance marks are not depicted (left) and two sets $\Te$ that
+  \caption{An automaton structure where the transition set forms a cycle
+    (left) and two sets $\Te$ that
     are not cycles even if $\mcycle_2(\vec{e})$ holds (middle and
-    right).}
+    right). The transition labels and acceptance marks are not depicted.}
   \label{fig:inter}
@@
 
-This section explains how we use the QBF formulas constructed in the
+This section explains how we use the QBF constructed in the
 previous section to reduce the number of acceptance marks in TELA.
 First, we describe a \emph{single-level} reduction, which
-uses only a single kind of QBF formulas. More precisely, we talk about
+uses only a single kind of QBF. More precisely, we talk about
 \emph{level~1}, \emph{level~2}, or \emph{level~3} reduction when
@@
 and the theorem prover Z3~\cite{demoura.08.tacas} to solve the
-satisfiability of QBF formulas. Our tool is available at
+satisfiability of QBF. Our tool is available at
 \begin{center}
@@
 reduced from $x$ to $y$ acceptance marks. If the cell contains a sum of two
-numbers, the latter represents number of automata where the last call to QBF
-solver reached the time limit.}
+numbers, the latter represents number of automata where the attempt to reduce
+another mark has been unsuccessful due to a QBF solver timeout.}
 \label{tab:heatmap}

### 2023-05-30 paper/main.tex
modifications made together with Duri

@@
 \newcommand{\todo}[1]{\marginpar{\textcolor{red}{#1}}}
+\newcommand{\fixme}[1]{\textcolor{red}{#1}}
 %\newcommand{\todo}[1]{}
@@
 
-TELA have attracted a lot of attention during the last five
+TELA have attracted a lot of attention during the last %five
 years~\cite{baier.19.atva,major.19.atva,muller.17.gandalf,renkin.20.atva}.
@@
 automatically handle all automata with traditional acceptance
-conditions. TELA can be obtained for example by translating LTL
-formulas with tools \texttt{ltl2dela} (known as
-Delag)~\cite{muller.17.gandalf} or \texttt{ltl2dgra} (known as
-Rabinizer~4)~\cite{kretinsky.18.cav} of the Owl library,
-\texttt{ltl2tgba} of the Spot library~\cite{duret.16.atva2}, or
+conditions. TELA can be obtained for example by translating formulas
+of \emph{linear temporal logic (LTL)}~\cite{pnueli.77.focs} with tools
+\texttt{ltl2dela} (known as Delag)~\cite{muller.17.gandalf} or
+\texttt{ltl2dgra} (known as Rabinizer~4)~\cite{kretinsky.18.cav} of
+the Owl library, \texttt{ltl2tgba} of the Spot
+library~\cite{duret.16.atva2}, or
 \texttt{ltl3tela}~\cite{major.19.atva}. There are also algorithms
@@
 terms used in the paper. Section~\ref{sec:qbf} explains the
-construction of the three mentioned QBF. The reduction algorithm
-based on these formulas is presented in
+construction of the three mentioned quantified Boolean formulas. The
+reduction algorithm based on these formulas is presented in
 Section~\ref{sec:algo}. Section~\ref{sec:implementation} describes our
-tool \telatko implementing the reduction technique.
-Experimental results are shown in
-Section~\ref{sec:experiments}. Finally, Section~\ref{sec:conclusion}
-suggests other applications of our QBF-based reduction technique and
-closes the paper.
+tool \telatko implementing the reduction technique.  Experimental
+results are shown in Section~\ref{sec:experiments}. Finally,
+Section~\ref{sec:conclusion} suggests other applications of our
+QBF-based reduction technique and closes the paper.
 
@@
 
+By choosing an appropriate acceptance condition, one can easily
+represent many classical kinds of automata over infinite words. For
+example, a transition-based B\"uchi automaton is a TELA with
+$M = \{\tacc{1}{1}\}$, $\varphi = \Inf\tacc{1}{1}$, and the acceptance
+mark \tacc{1}{1} being placed on all accepting transitions.
+Similarly, a transition-based Rabin automaton with $k$ acceptance
+pairs has
+$M = \{\tacc{5}{1}, \tacc{3}{1'}, \ldots, \tacc{5}{k}, \tacc{3}{k'}\}$
+and
+$\varphi = (\Fin\tacc{5}{1} \land \Inf\tacc{3}{1'}) \lor \ldots \lor
+(\Fin\tacc{5}{k} \land \Inf\tacc{3}{k'})$.  \fixme{A state-based acceptance
+can be represented by TELA by placing the corresponding acceptance
+marks on all transitions leaving the accepting states.}
+
 \bigskip
@@
 
-\section{Construction of formulas}\label{sec:qbf}
+\section{Construction of quantified Boolean formulas}\label{sec:qbf}
 
@@
 \begin{itemize}
-\item a set $\oM$ of acceptance marks such that $|\oM|<|M|$,
+\item a set $\oM$ of acceptance marks satisfying $|\oM|<|M|$,
 \item an acceptance formula $\psi$ over $\oM$, and 
@@
   some $\oM'\subseteq\oM$)
-\item such that the automaton
-$\mB=(Q,\oM,\Sigma,\rem(\delta),q_I,\psi)$ is equivalent to $\mA$.
+  %
+  such that the automaton $\mB=(Q,\oM,\Sigma,\rem(\delta),q_I,\psi)$
+  is equivalent to $\mA$.
 \end{itemize}
@@
 attempt to reduce acceptance formulas. The presented method can be
-easily adapted to look for $\psi$ in CNF or in a different shape.
+easily adapted to look for $\psi$ in \emph{conjunctive normal form
+  (CNF)} or in a different shape.
 
@@
 $\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ are defined as
-follows. The subformula $\mcycle(\vec{e})$ will be defined as the last
-one and we present three versions of it.
+follows. \fixme{The subformula $\mcycle(\vec{e})$ will be defined as the last
+one and we present three versions of it.}
 
@@
 \begin{lemma}
-  A nonempty set $T\subseteq\delta$ is a cycle if and only if for each set of
-  states $S\subseteq Q$ it holds that
+  A nonempty set $T\subseteq\delta$ is a cycle if and only if for each
+  set of states $S\subseteq Q$ one of the following conditions holds.
   \begin{enumerate}[A.]
-  \item\label{en:1} all transitions in $T$ lead from a state in $S$ to a state in $S$ (i.e., $T\subseteq\delta_S$) or
-  \item\label{en:2} all transitions in $T$ lead from a state outside $S$ to a state outside $S$ (i.e., $T\subseteq\delta_{Q\smallsetminus S}$) or
+  \item\label{en:1} All transitions in $T$ lead from a state in $S$ to a state in $S$ (i.e., $T\subseteq\delta_S$).
+  \item\label{en:2} All transitions in $T$ lead from a state outside $S$ to a state outside $S$ (i.e., $T\subseteq\delta_{Q\smallsetminus S}$).
   \item\label{en:3} $T$ contains a transition leading from a state in $S$ to a state outside $S$ and a transition leading from a state outside
@@
 
-If some call of the function
-$\mathit{satisfiable}(\Phi_{j,C,K{-}1})$ in the first
-\textbf{while} loop of Algorithm~\ref{alg:single} does not return
-$\true$, then the name of the output automaton encodes the reason for
-it. In the case of a single level reduction, the name has the form
-\texttt{L$j$\_$k$\_$X$}, where $j$ is the considered level, $k=K-1$ is
-the number of acceptance marks considered by the formula, $X$ is either
-\texttt{U} if the formula is unsatisfiable or \texttt{T} if the solver
-did not decide within the time limit. If $X$ is \texttt{T}, a longer
-timeout may lead to further reductions. If the multi-level reduction
-is used, the automaton name contains the information from all
-levels. For example, the name `\texttt{L1\_5\_U L2\_3\_U L3\_1\_T}'
-means that level 1 reduced the number of marks to 6 (reduction to 5 is
-impossible on this level), level 2 reduced it to 4, and level 3 to 2
-as the QBF solver did not finish in the time limit when trying to
-reduce the number of marks to
+If some call of the function $\mathit{satisfiable}(\Phi_{j,C,K{-}1})$
+in the first \textbf{while} loop of Algorithm~\ref{alg:single} does
+not return $\true$, then the name of the output automaton (included in
+the generated HOA) encodes the reason for it. In the case of a single
+level reduction, the name has the form \texttt{L$j$\_$k$\_$X$}, where
+$j$ is the considered level, $k=K-1$ is the number of acceptance marks
+considered by the formula, $X$ is either \texttt{U} if the formula is
+unsatisfiable or \texttt{T} if the solver did not decide within the
+time limit. If $X$ is \texttt{T}, a longer timeout may lead to further
+reductions. If the multi-level reduction is used, the automaton name
+contains the information from all levels. For example, the name
+`\texttt{L1\_5\_U L2\_3\_U L3\_1\_T}' means that level 1 reduced the
+number of marks to 6 (reduction to 5 is impossible on this level),
+level 2 reduced it to 4, and level 3 to 2 as the QBF solver did not
+finish in the time limit when trying to reduce the number of marks to
 1.%when trying the automaton may be potentially reduced further.
@@
 %\todo{Upravit experimenty: pridat dalsi 2 sady automatu.}
-We have used the tool \texttt{genltl} of the Spot
+\fixme{We have used the tool \texttt{genltl} of the Spot
 library~\cite{duret.16.atva2} to generate LTL formulas from
-literature. The parametrized formulas were generated for all
+literature, using all available formula sets implemented in
+\texttt{genltl} 2.10.4.} The parametrized formulas were generated for all
 combinations of parameter values from 1 to 4. We used the tool
@@
 may be potentially further reduced by the reduction with a longer time
-limit.
+limit. \fixme{The reduction ratio drop in higher levels, e.g.~between level~1
+and level~2 for \texttt{ltl2dgra}, happens due to the QBF solver timeouts
+occuring earlier as the formula being solved in higher level is more
+complex. For \texttt{ltl2dela} and \texttt{ltl2dgra}, the automata sets
+do not contain any automaton where level~2 or level~3 achieves a better
+result than level~1. However, all levels contribute to the reductions in
+the multilevel setting.}
 

### 2023-05-30 paper/main.tex
progress in final version

@@
 represent many classical kinds of automata over infinite words. For
-example, a transition-based B\"uchi automaton is a TELA with
-$M = \{\tacc{1}{1}\}$, $\varphi = \Inf\tacc{1}{1}$, and the acceptance
-mark \tacc{1}{1} being placed on all accepting transitions.
-Similarly, a transition-based Rabin automaton with $k$ acceptance
-pairs has
-$M = \{\tacc{5}{1}, \tacc{3}{1'}, \ldots, \tacc{5}{k}, \tacc{3}{k'}\}$
-and
+example, a Büchi automaton can be represented as a TELA with the
+acceptance condition $\varphi = \Inf\tacc{1}{1}$ and the single mark
+$\tacc{1}{1}$ placed on all transitions leaving the accepting states
+of the Büchi automaton. Further, a Rabin automaton with $k$ acceptance
+pairs can be similarly represented as a TELA with acceptance condition
 $\varphi = (\Fin\tacc{5}{1} \land \Inf\tacc{3}{1'}) \lor \ldots \lor
-(\Fin\tacc{5}{k} \land \Inf\tacc{3}{k'})$.  \fixme{A state-based acceptance
-can be represented by TELA by placing the corresponding acceptance
-marks on all transitions leaving the accepting states.}
+(\Fin\tacc{5}{k} \land \Inf\tacc{3}{k'})$ and marks
+$M = \{\tacc{5}{1}, \tacc{3}{1'}, \ldots, \tacc{5}{k},
+\tacc{3}{k'}\}$.
 
@@
 where $\forall\vec{e}$ denotes the sequence composed of $\forall e_t$
-for all variables $e_t$ and the subformulas $\mcycle(\vec{e})$,
-$\msat_\varphi(\vec{e})$, and
-$\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$ are defined as
-follows. \fixme{The subformula $\mcycle(\vec{e})$ will be defined as the last
-one and we present three versions of it.}
+for all variables $e_t$. Now we define the subformulas
+$\msat_\varphi(\vec{e})$,
+$\msat_{C,K}(\vec{e},\vec{n},\vec{i},\vec{f})$, and three versions of
+$\mcycle(\vec{e})$.
 
@@
   \]  
-  Then there exists a function $\rem'$ that respects % the equivalence
-  $\sim_S$ (i.e.,
-  if $t_1\sim_S t_2$ then $\rem'(t_1)=\rem'(t_2)$) and
+  Then there exists a function $\rem'$ that respects the equivalence
+  $\sim_S$ (i.e., it assigns the same marks to equivalent transitions)
+  % if $t_1\sim_S t_2$ then $\mks(\rem'(t_1))=\mks(\rem'(t_2))$)
+  and
   \[
@@
   select one transition from each equivalence class of~$\sim_S$. For
-  every transition $t\in\delta_S$, by $\overline{t}$ we denote the
-  selected transition equivalent to $t$ and we define the function
-  $\rem'$ as $\rem'(t)=\rem(\overline{t})$. Note that we do not need
+  every transition $t=(p,a,M',q)\in\delta_S$, by $\overline{t}$ we
+  denote the selected transition equivalent to $t$ and we define the
+  function $\rem'$ such that
+  $\rem'(t)=(p,a,\mks(\rem(\overline{t})),q)$. Note that we do not need
   to discuss the value of $\rem'$ on transitions outside $\delta_S$ as
@@
   \begin{tikzpicture}[smallautomaton]
-    \begin{scope}[initial angle=90,xshift=-4cm]
+    \begin{scope}[initial angle=90,xshift=-4.5cm]
       \node[state,initial] (p) at (0,0) {};
@@
 
-    \begin{scope}[initial angle=90,densely dotted,black!35,xshift=4cm]
+    \begin{scope}[initial angle=90,densely dotted,black!35,xshift=4.5cm]
       \node[state,initial,solid,draw=black] (p) at (0,0) {};
@@
   \end{tikzpicture}
-  \caption{An automaton structure where the transition set forms a cycle
-    (left) and two sets $\Te$ that
-    are not cycles even if $\mcycle_2(\vec{e})$ holds (middle and
-    right). The transition labels and acceptance marks are not depicted.}
+  \caption{An automaton structure % where the transition set forms a cycle
+    (left) and two sets $\Te$ (middle and right) that are not
+    cycles even if $\mcycle_2(\vec{e})$ holds. The transition labels
+    and acceptance marks are not depicted.}
   \label{fig:inter}
@@
 \begin{lemma}
-  A nonempty set $T\subseteq\delta$ is a cycle if and only if for each
-  set of states $S\subseteq Q$ one of the following conditions holds.
+  A nonempty set $T\subseteq\delta$ is a cycle if and only if, for each
+  set of states $S\subseteq Q$, one of the following conditions holds.
   \begin{enumerate}[A.]
@@
 variables $s_q$ that are existentially quantified (when the formula is
-transformed to prenex normal form).
+transformed to prenex normal form) in the scope of universal
+quantification of variables $e_t$.
 
@@
 \subsection{Optimizations of formulas}
-
 Finally, we mention three simple optimizations of the formula
@@
 and the theorem prover Z3~\cite{demoura.08.tacas} to solve the
-satisfiability of QBF. Our tool is available at
+satisfiability of QBF transformed to prenex normal form (and not using
+CNF). Our tool is available at
 \begin{center}
@@
 
-%\todo{Upravit experimenty: pridat dalsi 2 sady automatu.}
-\fixme{We have used the tool \texttt{genltl} of the Spot
-library~\cite{duret.16.atva2} to generate LTL formulas from
-literature, using all available formula sets implemented in
-\texttt{genltl} 2.10.4.} The parametrized formulas were generated for all
-combinations of parameter values from 1 to 4. We used the tool
-\texttt{ltlfilt} of the Spot library to simplify the formulas and
-remove duplicates. Altogether, we have generated 348 unique LTL
+To evaluate our reduction technique, we considered all sets of LTL
+formulas from literature as they are provided by the tool
+\texttt{genltl} of the Spot library~\cite{duret.16.atva2}~2.10.4, where
+parametrized formulas were generated for all combinations of
+parameter values from 1 to 4. We used the tool \texttt{ltlfilt} of the
+Spot library to simplify the formulas and remove
+duplicates. Altogether, we have generated 348 unique LTL
 formulas. These have been translated to deterministic TELA by two
@@
 
-Table~\ref{tab:cumm-data} shows the cumulative numbers of marks in
-the input automata sets and after each reduction, together with the
+Table~\ref{tab:cumm-data} shows the cumulative numbers of marks in the
+input automata sets and after each reduction, together with the
 reduction ratio and total time spent by the considered reduction. The
@@
 the last query to QBF solver did not finish within the 30~seconds
-limit. The timeout of the last QBF query implies that the automaton
-may be potentially further reduced by the reduction with a longer time
-limit. \fixme{The reduction ratio drop in higher levels, e.g.~between level~1
-and level~2 for \texttt{ltl2dgra}, happens due to the QBF solver timeouts
-occuring earlier as the formula being solved in higher level is more
-complex. For \texttt{ltl2dela} and \texttt{ltl2dgra}, the automata sets
-do not contain any automaton where level~2 or level~3 achieves a better
-result than level~1. However, all levels contribute to the reductions in
-the multilevel setting.}
+limit. The timeout of the last QBF query means that the automaton may
+be potentially further reduced if a longer time limit is used. One can
+observe that a higher level sometimes achieves a smaller reduction
+than a lower level (e.g., compare level~1 and level~2 for
+\texttt{ltl2dgra}). This is caused by the QBF solver timeouts occuring
+earlier as formulas constructed by the higher level are more
+complex. For \texttt{ltl2dela} and \texttt{ltl2dgra}, the automata
+sets do not contain any automaton where level~2 or level~3 achieves a
+better result than level~1. However, all levels contribute to the
+reductions in the multi-level setting.
 
@@
 \begin{table}[b!]
-\caption{The effect of reduction.
+\caption{The effect of multi-level reduction.
 A cell on coordinates $(x,y)$ contains the number of automata that have been
 reduced from $x$ to $y$ acceptance marks. If the cell contains a sum of two
-numbers, the latter represents number of automata where the attempt to reduce
+numbers, the latter represents the number of automata where the attempt to reduce
 another mark has been unsuccessful due to a QBF solver timeout.}

### 2023-05-31 qbf/formula_preparation.py
reset

@@ -1,262 +1,0 @@
-from qbf.z3_formula import *
-from qbf.parser import clear_aut_edges
-import spot
-
-def update_run_info(K, res, tmp_mode):
-    if res == z3.unknown:
-        str_res = "T"
-    if res == unsat:
-        str_res = "U"
-    assert(str_res != "")
-    return "L{0}_{1}_{2} ".format(str(tmp_mode), str(K), str_res)
-
-
-def empty_aut(aut, original):
-    clear_aut_edges(aut)
-    aut.set_acceptance(0, spot.acc_code.t())
-    if not spot.are_equivalent(original, aut):
-        aut.set_acceptance(0, spot.acc_code.f())
-    aut.set_name(aut.get_name() + "F_0_S")
-    return aut
-
-def add_or_append(dict, key, value):
-    if key in dict:
-        dict[key].append(value)
-    else:
-        dict[key] = [value]
-
-
-def filter_out_representants( aut, representants_set, scc_edg, scc_state_info):
-    new_scc_edg = []
-    for i in range(len(scc_edg)):
-        new_scc_edg.append(list(filter(lambda e: aut.edge_number(e) in representants_set, scc_edg[i])))
-    
-    new_scc_state_info = []
-
-    for scc in scc_state_info:
-        for k,v in scc.items():
-            out_state = list(set(v[0]) & representants_set)
-            in_state = list(set(v[1]) & representants_set)
-            new_scc_state_info.append({k : [out_state, in_state]})
-    
-    return new_scc_edg, new_scc_state_info
-  
-
-def print_eq_edges(aut, eq_edges):
-    for d in eq_edges:
-        for key, val in d.items():
-            print(key, " : ", list(map(lambda e: aut.edge_number(e), val)))
-
-def dict_union(d1, d2):
-    """
-    Args:
-        d1 :: (acc. sets) : [edges]
-        d2 :: (acc. sets) : [edges]
-    
-    """
-    for key, val in d1.items():
-        if key in d2.keys():
-            d2[key] = d2.get(key, []) + d1.get(key, [])
-        else:
-            d2[key] = d1.get(key, [])
-    return d2
-
-def edge_dictionary(aut, mode, sccs_edges, scc_optimalization):
-    """
-
-    Args:
-        aut: spot::twa - automaton from input
-        sccs_edges [[edges of one SCC]]
-        scc_optimalization: Bool - Is the SCC optimalization turned on.
-
-    Returns: 
-        edge_dict: {acceptance set number : [numbers of edges in this set]}
[…]

### 2023-05-31 qbf/incremental.py

@@ -1,155 +1,0 @@
-from qbf.formula_preparation import edge_dictionary, update_run_info, try_evaluate0, get_formula, solver_init, filter_out_representants
-from qbf.parser import process_variables
-from telatko2.classes import *
-from z3 import *
-
-
-def restrict_f_vars(K, inner_edges_nums):
-    formula = []
-    for e in inner_edges_nums:
-        # these two are not equivalent?
-        # var1 = Bool("f_{0}_{1} ".format(str(e), str(K)))
-        var = Bool("f_" + str(e) + "_" + str(K))
-
-        formula.append(Not(var))
-    return And(formula)
-
-
-def restrict_pn_vars(C, K):
-    formula = []
-    for i in range(1, C + 1):
-
-        p = Bool("p_" + str(i) + "_" + str(K))
-        n = Bool("n_" + str(i) + "_" + str(K))
-        formula.append(Not(p))
-        formula.append(Not(n))
-    return And(formula)
-
-
-def restrict_C_vars(C, K):
-    formula = []
-    for i in range(1, K + 1):
-
-        p = Bool("p_" + str(C) + "_" + str(i))
-        n = Bool("n_" + str(C) + "_" + str(i))
-        formula.append(Not(p))
-        formula.append(Not(n))
-    return And(formula)
-
-
-def update_inc_solver(solver, currently_reduced, C, K, inner_edges_nums):
-    if currently_reduced == FormulaAtribute.K:
-        solver.push()
-        solver.add(restrict_f_vars(K, inner_edges_nums))
-        solver.add(restrict_pn_vars(C, K))
-
-        return solver
-    else:
-        solver.push()
-        solver.add(restrict_C_vars(C, K))
-        return solver
-    assert (False)
-
-
-def model_assert(model, K):
-    for t in model.decls():
-        if str(t)[0] == 'f' and str(t)[-1] == str(K):
-            assert (is_false(model[t]))
-
-
-def inc_loop(aut, tmp_mode,
-             scc_state_info, scc_edg, currently_reduced, original, formula_attr):
-
-    K = aut.get_acceptance().used_sets().count()
-    C = len(aut.get_acceptance().to_dnf().top_disjuncts())
-    orig_scc_edg = scc_edg
-    orig_scc_state_info = scc_state_info
-    if currently_reduced == FormulaAtribute.K:
-        K -= 1
-    else:
-        C -= 1
-
-    # edge_dict: {acc_set_num : [edge_nums]}
[…]

### 2023-05-31 qbf/parser.py

@@ -1,249 +1,0 @@
-from telatko2.classes import *
-from z3 import is_true
-from qbf.rejecting_edges import *
-
-
-def create_acc(variables):
-    """
-
-    Args:
-        variables [[{Variable}]]: list of lists of variables - each var in list has the same clause number.
-
-    Returns:
-        [[MarkType]] - list of list of MartType of the same structure as input
-
-    """
-    acc = []
-    for dis in variables:
-        disjunct = []
-        for con in dis:
-            if con.type == 'p':
-                disjunct.append(ACCMark(MarkType.Inf, con.set))
-            else:
-                disjunct.append(ACCMark(MarkType.Fin, con.set))
-
-        acc.append(disjunct)
-    return acc
-
-
-def string_formula(acc):
-    """
-    Returns string interpretation
-    Args:
-        acc [[MarkType]]:  acceptance condition
-
-    Returns: string - string interpretation of input acc
-
-    """
-    f = []
-    for dis in acc:
-        f.append('(')
-        for con in dis:
-            f.append(str(con))
-            if con is not dis[-1]:
-                f.append(" & ")
-        f.append(')')
-        if dis is not acc[-1]:
-            f.append(" | ")
-    return ''.join(f)
-
-
-def prepare_acc_vars(acc_vars):
-    """
-
-    Args:
-        acc_vars[string]: list of variables, that represent future acceptance condition
-        and were labeled as true by SAT solver
-
-    Returns:
-        [[Variable]] - each sublist contains {Variable}s with the same conjunct number
-
-    """
-    condition_vars_class = list(
-        map(lambda var: Variable(var.split('_')), acc_vars))
-    condition_vars_class.sort(key=lambda x: x.clause, reverse=False)
-    clause_num = condition_vars_class[0].clause
-    clause_max_num = condition_vars_class[-1].clause
-    var_list = []
-    while clause_num != (clause_max_num + 1):
-        var_list.append(
-            list(
-                filter(
-                    lambda x: x.clause == clause_num,
-                    condition_vars_class)))
-
-        clause_num += 1
-    return var_list
-
-
-
-def print_edges(aut):
-    for edge in aut.edges():
-        print("edges[{e}].src={src}, edges[{e}].dst={dst}, edges[{e}].acc={acc}".format(
[…]

### 2023-05-31 qbf/q_main.py

@@ -1,157 +1,0 @@
-import random
-from threading import current_thread
-from qbf.parser import *
-from qbf.incremental import inc_loop
-from qbf.not_incremental import not_incr_loop
-from qbf.formula_preparation import try_evaluate0
-
-
-def scc_info(aut):
-    """
-
-    Args:
-        aut:
-
-    Returns: scc_state_info[{state_number : [edges out of the state], [edges in the state]}]
-              scc_edg [[numbers of inner edges of one SCC]]
-
-    """
-    scc_state_inf = []
-
-    # [[edges in one scc]]
-    scc_edges = []
-
-    si = spot.scc_info(aut)
-
-    for i in range(si.scc_count()):
-        states = si.states_of(i)
-        if si.is_trivial(i):
-            continue
-
-        if si.check_scc_emptiness(i):
-            # if rejecting ignore edges (for now)
-            states = si.states_of(i)
-            continue
-
-        state_dict = {}
-        for s in states:
-            state_dict[s] = [[], []]
-        edges = si.inner_edges_of(i)
-
-        scc_edges.append(list(edges))
-
-        for e in edges:
-
-            state_dict[e.src][0].append(aut.edge_number(e))
-            state_dict[e.dst][1].append(aut.edge_number(e))
-
-        scc_state_inf.append(state_dict)
-
-    return scc_state_inf, scc_edges
-
-
-def count_sets(sets):
-    counter = 0
-    for s in sets:
-        counter += 1
-    return counter
-
-
-def resolve_formula_atributes(minimized_atribute, C, K):
-    if minimized_atribute == 'clauses':
-
-        return FormulaAtribute.C
-    else:
-
-        return FormulaAtribute.K
-
-
-def play(aut, formula_attr):
-    aut.set_name("")
-
-    spot.cleanup_acceptance_here(aut)
-
-    if aut.get_acceptance().used_sets().count(
-    ) < 1 or aut.prop_state_acc() == spot.trival.yes_value:
-        return aut
-
-    original = spot.automaton(aut.to_str())
-
-    # scc_edg [[nums of edges of one scc]]
-    # scc state info [{state num : [[num of edge of which is the state source
-    # of][num of edge of which is the state destination of]]}]
-    scc_state_info, scc_edg = scc_info(aut)
-
-    if formula_attr.incremental:
-        for i in range(formula_attr.tmp_level, formula_attr.level + 1):
-            if aut.get_acceptance().used_sets().count() <= 1:
[…]

### 2023-05-31 qbf/rejecting_edges.py

@@ -1,46 +1,0 @@
-import spot
-from telatko2.classes import MarkType
-
-
-def clear_aut_edges(aut):
-    for e in aut.edges():
-        e.acc = spot.mark_t()
-
-
-def list_of_rejecting_edges(aut):
-
-    rejecting_edges = []
-    si = spot.scc_info(aut)
-
-    for i in range(si.scc_count()):
-
-        if si.is_trivial(i):
-            continue
-
-        if si.check_scc_emptiness(i):
-            for e in si.inner_edges_of(i):
-                rejecting_edges.append(e)
-
-    return rejecting_edges
-
-
-def rejecting_sccs_edges(aut, acc):
-    """
-    Takes all edges from original automaton and makes them rejecting for the new acceptance condition.
-    (As side effect has to clean all edges of the automaton.)
-    """
-
-    rejecting_edges = list_of_rejecting_edges(aut)
-
-    # returns cpp pair <bool, spot.mark_t>, if bool is False, there is no rejecting combination
-    # there should be rejecting combination since we are dealing with
-    # rejecting SCCs
-    rejecting_marks = spot.acc_cond(spot.acc_code(acc)).unsat_mark()
-
-    # fails if there is no unsat combination of marks
-    assert (rejecting_marks[0])
-    # removes all acceptance marks from edges
-    clear_aut_edges(aut)
-
-    for e in rejecting_edges:
-        e.acc = rejecting_marks[1]

### 2023-05-31 sat23/CNF/experiments.py

@@ -1,20 +1,0 @@
-#!/usr/bin/env python3
-
-import spot
-import os
-import sys
-
-def main(argv):
-    telatko = "/home/xschwar3/telatko/telatko.py"
-    
-    os.environ['SPOT_TMPDIR'] = './autcross'
-    os.environ['SPOT_TMPKEEP'] = '1'
-    patterns_path = '/home/xschwar3/telatko/sat23/patterns' 
-    tools_order = ['ltl2tgba_det', 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det']
-    for tool in tools_order:
-        print(tool)
-        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
-
-
-if __name__ == "__main__":
-    main(sys.argv[1:])

### 2023-05-31 statistics/parse_statistics.py

@@ -1,111 +1,0 @@
-#!/usr/bin/env python3
-
-import argparse
-import sys
-import numpy as np
-import pandas as pd
-import csv
-
-def parse_csv(input_file):
-    csv_file = pd.read_csv(input_file)
-
-    # filter undesirable columns
-    """
-    filtered_columns_csv = csv_file.drop(columns=[ 'input.source','input.name', 'input.ap','input.edges', 'input.transitions', 'input.scc', 'input.nondetstates'
-                , 'input.nondeterministic', 'input.alternating', 'exit_status', 'exit_code', 'output.ap', 'output.edges','output.transitions',
-                'output.nondetstates', 'output.scc', 'output.nondeterministic', 'output.alternating'])
-    """
-    # sum filtered columns
-    sum_filtered_columns = csv_file.sum(axis=0)
-    df = pd.DataFrame(sum_filtered_columns)
-    return df
-    print(df.at['output.acc_sets', 0], np.round((df.at['time', 0]), 3))
-    return df.at['output.acc_sets', 0], np.round((df.at['time', 0]), 3)
-
-
-
-def add_input(data_frame, dict):
-    dict["Acceptance sets"].append(data_frame.at['input.acc_sets', 0])
-    dict["% reduction"].append(0)
-    dict["Time"].append(0)
-
-def add_output(data_frame, dict):
-    intput_sets = data_frame.at['input.acc_sets', 0]
-    output_sets = data_frame.at['output.acc_sets', 0]
-    dict["Acceptance sets"].append(output_sets)
-    dict["% reduction"].append(np.round(100 - (100*output_sets / intput_sets), 1) )
-    dict["Time"].append(np.round((data_frame.at['time', 0]), 3))
-
-def to_string(output, mrkdwn):
-    if output is not None:
-        f = open(output, 'a')
-        f.write(mrkdwn + "\n\n" )
-        f.close()
-    else:
-        print(mrkdwn)
-
-
-
-def main(argv):
-    parser = argparse.ArgumentParser()
-    parser.add_argument("-L1", "--level1", help="CSV file of statistics level1", default=None)
-    parser.add_argument("-L2", "--level2", help="CSV file of statistics level1", default=None)
-    parser.add_argument("-L3", "--level3", help="CSV file of statistics level1", default=None)
-    parser.add_argument("-G", "--gradual", help="CSV file of statistics G", default=None)
-    parser.add_argument("-O", "--outfile", help="File to print output to.", default=None)
-
-    args = parser.parse_args()
-
[…]

### 2023-05-31 stats_gen/scc.sh

@@ -1,37 +1,0 @@
-#!/bin/sh
-
-RAB=./../automata/rab4.hoa
-LTL3TELA=./../automata/ltl3tela.hoa
-LTL2TGBA=./../automata/ltl2tgba.hoa
-
-CSV_FILE=./../statistics/new/scc
-
-
-
-autcross --file=$RAB --tool='python3 ./../telatko.py -F%H -O%O -L 1 -S' --language-preserved --csv=$CSV_FILE'/rab4.l1.csv' 
-
-
-autcross --file=$RAB --tool='python3 ./../telatko.py -F%H -O%O -L 2 -S' --language-preserved --csv=$CSV_FILE'/rab4.l2.csv'
-
-autcross --file=$RAB --tool='python3 ./../telatko.py -F%H -O%O -L 3 -S' --language-preserved --csv=$CSV_FILE'/rab4.l3.csv'
-
-autcross --file=$RAB --tool='python3 ./../telatko.py -F%H -O%O -L 3 -G -S' --language-preserved --csv=$CSV_FILE'/rab4.g.csv'
-
-
-autcross --file=$LTL3TELA --tool='python3 ./../telatko.py -F%H -O%O -L 1 -S' --language-preserved --csv=$CSV_FILE'/ltl3tela.l1.csv'
-
-autcross --file=$LTL3TELA --tool='python3 ./../telatko.py -F%H -O%O -L 2 -S' --language-preserved --csv=$CSV_FILE'/ltl3tela.l2.csv'
-
-autcross --file=$LTL3TELA --tool='python3 ./../telatko.py -F%H -O%O -L 3 -S' --language-preserved --csv=$CSV_FILE'/ltl3tela.l3.csv'
-autcross --file=$LTL3TELA --tool='python3 ./../telatko.py -F%H -O%O -L 3 -G -S' --language-preserved --csv=$CSV_FILE'/ltl3tela.g.csv'
-
-
-
-autcross --file=$LTL2TGBA --tool='python3 ./../telatko.py -F%H -O%O -L 1 -S' --language-preserved --csv=$CSV_FILE'/ltl2tgba.l1.csv'
-
-autcross --file=$LTL2TGBA --tool='python3 ./../telatko.py -F%H -O%O -L 2 -S' --language-preserved --csv=$CSV_FILE'/ltl2tgba.l2.csv'
-
[…]

### 2023-05-31 telatko.py

@@ -1,104 +1,0 @@
-#!/usr/bin/env python3
-
-import spot
-import sys
-import argparse
-import os
-
-from telatko2.classes import Switches
-from qbf.q_main import play, print_aut
-from telatko2.playground import process_automaton
-
-
-def main(argv):
-    parser = argparse.ArgumentParser()
-    parser.add_argument(
-        "-F",
-        "--autfile",
-        help="File containing automata in HOA format.")
-    parser.add_argument(
-        "-O",
-        "--outfile",
-        help="File to print output to.",
-        default=None)
-    parser.add_argument(
-        "-L",
-        "--level",
-        help="Level of simplification.",
-        type=int,
-        default=0)
-    parser.add_argument(
-        "-T",
-        "--timeout",
-        help="Kill QBF solver after inserted number of seconds.", type=int, default=50)
-    parser.add_argument(
-        "-S",
-        "--scc",
-        help="Use SCC optimization.", action='store_true')
-    parser.add_argument(
-        "-C",
-        "--minimize_clauses",
-        help="Turn on reduction of clauses.", action='store_true')
-
-    parser.add_argument(
-        "-I",
-        "--incremental",
-        help="Use incremental solving.", action='store_true')
-    parser.add_argument(
-        "-G",
-        "--gradual",
-        help="Levels will gradually grow from L1 up to specified level.", action='store_true')
-    args = parser.parse_args()
-
-    if not args.autfile:
-        print("No automata to process.", file=sys.stderr)
-
-    # if no arguments given, default variant is gradually go up to level 3
-    if args.level == 0 and not args.gradual and not args.incremental and not args.scc:
-        args.gradual = True
-        args.level = 3
-        args.incremental = True
-        args.scc = True
-    elif args.level == 0:
-        args.gradual = True
-        args.level = 3
-
-    if args.level > 3:
-        args.level = 3
-
-    aut = spot.automata(args.autfile)
-    formula_attributes = Switches(args)
-    # formula_attributes.print_settings()
-
-    for a in aut:
-
-        origin = spot.automaton(a.to_str())
-        try:
-            spot.cleanup_acceptance_here(a)
-            formula_attributes.set_C(a)
-
-            if a.get_acceptance().used_sets().count() == 0:
-                auto = a
-            else:
-                auto = play(a, formula_attributes)
-
-            if args.outfile:
-                print_aut(auto, args.outfile, "a")
-            else:
-                print_aut(auto, None, " ")
[…]

### 2023-05-31 telatko2/classes.py

@@ -1,602 +1,0 @@
-#!/usr/bin/python3
-
-
-import spot
-import enum
-import sys
-import argparse
-from scipy.optimize import linear_sum_assignment
-import numpy as np
-import copy
-import os
-import subprocess
-import itertools
-from math import ceil, log2
-
-
-class Variable:
-    """
-    Class that contains information retrieved from SAT-result.
-     Variable in this class formulates shape of new acceptance condition.
-    """
-
-    def __init__(self, var):
-        self.type = var[0]
-        self.clause = int(var[1])
-        self.set = int(var[2].split('=')[0])
-
-
-class Switches:
-    """
-    Class containing overwiev of switches set by user.
-    """
-
-    def __init__(self, args):
-        self.level = args.level
-        self.tmp_level = 1 if args.gradual else args.level
-        self.min_clauses = args.minimize_clauses
-        self.timeout = args.timeout
-        self.scc_optimization = args.scc
-        self.incremental = args.incremental
-        self.gradual = args.gradual
-        self.C = None
-
-    def set_C(self, aut):
-        self.C = len(
-            aut.get_acceptance().to_dnf().top_disjuncts())
-
-    def print_settings(self):
-        print("Global Level ", self.level)
-        print("Current Level ", self.tmp_level)
-        print("Minimize clauses ", self.min_clauses)
-        print("Timeout ", self.timeout)
-        print("SCC optimization ", self.scc_optimization)
-        print("Incremental ", self.incremental)
-        print("Gradual ", self.gradual)
-
-
-class SATformula:
-    """
-    Represents boolean formula, retrieved from the input automata, that will be given to SAT-solver.
-
-    """
-
-    def __init__(self, bool):
-        self.formula = bool
-        self.subformula = []
-        self.negation = False
-        self.imperativ = False
-
-    def __len__(self):
-        if not self.subformula:
-            return 1
-        else:
-            return sum(map(len, self.subformula))
-
-    def __str__(self):
-
-        if not self.subformula:
-
-            return ''.join(" " + self.formula + " ")
-
-        elif len(self.subformula) == 1:
-            if self.negation and self.imperativ:
-
-                return ''.join('!(' + str(self.subformula[0]) + ')')
-            return ''.join('(' + str(self.subformula[0]) + ')')
-
-        else:
-            if self.negation and self.imperativ:
-                return '!(' + str(self.subformula[0]) + self.formula + ''.join(str(
[…]

### 2023-05-31 telatko2/playground.py

@@ -1,309 +1,0 @@
-
-import random
-
-
-from telatko2.merge import *
-from telatko2.restore_equivalence import *
-import copy
-from qbf.parser import clear_aut_edges
-
-
-def try_eval(aut, acc, scc):
-    """
-
-    """
-    m_all_edges = set(scc_everywhere(aut, scc))  # set of marks that include
-    # all transitions of SCC
-
-    eval_f = False  # proc
-    for dis in acc.formula:
-        eval_dis = True
-        for con in dis:
-            val = eval_set(aut, con, scc, m_all_edges)
-            if eval_dis is False or val is False:
-                eval_dis = False
-            elif (eval_dis is None or eval_dis is True) and val is None:
-                eval_dis = None
-        if eval_dis is True:
-            acc.sat = True
-            acc.formula = []
-            return
-        elif (eval_f is None or eval_f is False) and eval_dis is None:
-            eval_f = None
-    acc.set_sat(eval_f)
-    if acc.sat is not None:
-        acc.formula = []
-
-
-def weak_eval(aut, acc, scc, weak):
-    if weak:
-        if scc.is_rejecting():  # all cycles in scc are rejecting
-            acc.set_sat(False)
-        else:
-            acc.set_sat(True)
-        acc.formula = []
-        return True
-    return False
-
-
-def print_aut(aut, output, m):
-    if output is not None:
-        f = open(output, m)
-        f.write(aut.to_str() + '\n')
-        f.close()
-    else:
-        print(aut.to_str())
-
-
-def get_acc(aut):
-    """
-    Returns PACC (dnf) if dnf formula is shorter else PACC_CNF(cnf)
-    """
-    dnf_acc = ACC_DNF(aut.get_acceptance().to_dnf())
-    # return dnf_acc
-    cnf_acc = ACC_CNF(aut.get_acceptance().to_cnf())
-
-    #print((cnf_acc),"\n", (dnf_acc))
-
-    if (dnf_acc.acc_len() < cnf_acc.acc_len()) or (dnf_acc.acc_len()
-                                                   == cnf_acc.acc_len() and len(dnf_acc) <= len(cnf_acc)):
-        # print("DNF")
-        return dnf_acc
-    else:
-        #acc = PACC_CNF(cnf_acc)
-        # print("CNF")
-        return cnf_acc
-
-
-def get_accs(aut):
-    """
-    Simplifies acceptance condition in DNF for every SCC
-    Args:
-        aut:
-
[…]

### 2023-05-31 telatko2/restore_equivalence.py

@@ -1,246 +1,0 @@
-
-from telatko2.classes import *
-from telatko2.simplify import *
-from telatko2.merge import get_set_nums
-
-
-def make_true(aut, scc, merged_acc):
-    scc_clean_up_edges(aut, ACC_DNF(""), scc)
-
-    min_inf = len(merged_acc[0])
-    cheap_dis = merged_acc[0]
-    for dis in merged_acc.formula:
-        if all(con.type == MarkType.Fin for con in dis):
-            return
-        count = 0
-        for con in dis:
-            if con.type == MarkType.Inf:
-                count += 1
-        if count < min_inf:
-            min_inf = count
-            cheap_dis = dis
-
-    add_m = []
-    for con in cheap_dis:
-        if con.type == MarkType.Inf:
-            add_m.append(con.num)
-
-    if not add_m:
-        return
-    for s in scc.states():
-        for e in aut.out(s):
-            if e.dst in scc.states():
-                for m in add_m:
-                    e.acc.set(m)
-
-
-def make_false(aut, scc, merged_acc):
-    scc_clean_up_edges(aut, ACC_DNF(""), scc)
-    add_m = []
-    for dis in merged_acc.formula:
-        if all(con.type == MarkType.Fin for con in dis):
-            add_m.append(dis[0].num)
-
-    if not add_m:
-        return
-    for s in scc.states():
-        for e in aut.out(s):
-            if e.dst in scc.states():
-                for m in add_m:
-                    e.acc.set(m)
-
-
-def make_one_true(aut, scc, one):
-    """
-    Make ACCMark true in particular expression - Inf on every transition
-    :param aut:
-    :param scc:
-    :param one: ACCMark
-    :return:
-    """
-    if one.type == MarkType.Inf:
-        for s in scc.states():
-            for e in aut.out(s):
-                if e.dst in scc.states():
-                    e.acc.set(one.num)
-
-
-def make_one_false(aut, scc, one):
-    """
-    Make ACCMark true in particular expression - Fin on every transition
-    :param aut:
-    :param scc:
-    :param one: ACCMark
-    :return:
-    """
-    if one.type == MarkType.Fin:
-        for s in scc.states():
-            for e in aut.out(s):
-                if e.dst in scc.states():
-                    e.acc.set(one.num)
-
-
-def make_expr_true(aut, scc, expr):
-    """
-    Make con/dis true
[…]

### 2023-05-31 qbf/formula_preparation.py
update

@@ -1,0 +1,275 @@
+from qbf.z3_formula import *
+from qbf.parser import clear_aut_edges
+import spot
+from qbf.limboole_formula import *
+from telatko2.classes import Solver_result
+from qbf.solver_manipulation.solver_functions import z3_result, query
+
+
+
+
+def update_run_info(K, res, tmp_mode):
+    if res == Solver_result.timeout:
+        str_res = "T"
+    if res == Solver_result.unsat:
+        str_res = "U"
+    assert (str_res != "")
+    return "L{0}_{1}_{2} ".format(str(tmp_mode), str(K), str_res)
+
+
+def empty_aut(aut, original):
+    clear_aut_edges(aut)
+    aut.set_acceptance(0, spot.acc_code.t())
+    if not spot.are_equivalent(original, aut):
+        aut.set_acceptance(0, spot.acc_code.f())
+    aut.set_name(aut.get_name() + "F_0_S")
+    return aut
+
+
+def add_or_append(dict, key, value):
+    if key in dict:
+        dict[key].append(value)
+    else:
+        dict[key] = [value]
+
+
+def filter_out_representants(aut, representants_set, scc_edg, scc_state_info):
+    new_scc_edg = []
+    for i in range(len(scc_edg)):
+        new_scc_edg.append(
+            list(
+                filter(
+                    lambda e: aut.edge_number(e) in representants_set,
+                    scc_edg[i])))
+
+    new_scc_state_info = []
+
+    for scc in scc_state_info:
+        for k, v in scc.items():
+            out_state = list(set(v[0]) & representants_set)
+            in_state = list(set(v[1]) & representants_set)
+            new_scc_state_info.append({k: [out_state, in_state]})
+
+    return new_scc_edg, new_scc_state_info
+
+
+def print_eq_edges(aut, eq_edges):
+    for d in eq_edges:
+        for key, val in d.items():
+            print(key, " : ", list(map(lambda e: aut.edge_number(e), val)))
+
+
+def dict_union(d1, d2):
+    """
+    Args:
+        d1 :: (acc. sets) : [edges]
+        d2 :: (acc. sets) : [edges]
+
+    """
+    for key, val in d1.items():
+        if key in d2.keys():
+            d2[key] = d2.get(key, []) + d1.get(key, [])
+        else:
+            d2[key] = d1.get(key, [])
+    return d2
+
+
+def edge_dictionary(aut, mode, sccs_edges, scc_optimalization):
+    """
+
+    Args:
[…]

### 2023-05-31 qbf/incremental.py

@@ -1,0 +1,181 @@
+from qbf.formula_preparation import edge_dictionary, update_run_info, try_evaluate0, get_formula, solver_init, filter_out_representants, z3_result
+from qbf.parser import process_variables
+from telatko2.classes import *
+from z3 import *
+
+
+def restrict_f_vars(K, inner_edges_nums):
+    formula = []
+    for e in inner_edges_nums:
+        # these two are not equivalent?
+        # var1 = Bool("f_{0}_{1} ".format(str(e), str(K)))
+        var = Bool("f_" + str(e) + "_" + str(K))
+
+        formula.append(Not(var))
+    return And(formula)
+
+
+def restrict_pn_vars(C, K):
+    formula = []
+    for i in range(1, C + 1):
+
+        p = Bool("p_" + str(i) + "_" + str(K))
+        n = Bool("n_" + str(i) + "_" + str(K))
+        formula.append(Not(p))
+        formula.append(Not(n))
+    return And(formula)
+
+
+def restrict_C_vars(C, K):
+    formula = []
+    for i in range(1, K + 1):
+
+        p = Bool("p_" + str(C) + "_" + str(i))
+        n = Bool("n_" + str(C) + "_" + str(i))
+        formula.append(Not(p))
+        formula.append(Not(n))
+    return And(formula)
+
+
+def update_inc_solver(solver, currently_reduced, C, K, inner_edges_nums):
+    if currently_reduced == FormulaAtribute.K:
+        solver.push()
+        solver.add(restrict_f_vars(K, inner_edges_nums))
+        solver.add(restrict_pn_vars(C, K))
+
+        return solver
+    else:
+        solver.push()
+        solver.add(restrict_C_vars(C, K))
+        return solver
+    assert (False)
+
+
+def model_assert(model, K):
+    for t in model.decls():
+        if str(t)[0] == 'f' and str(t)[-1] == str(K):
+            assert (is_false(model[t]))
+
+
+def inc_loop(
+        aut,
+        tmp_mode,
+        scc_state_info,
+        scc_edg,
+        currently_reduced,
+        original,
+        formula_attr):
+
+    K = aut.get_acceptance().used_sets().count()
+    C = len(aut.get_acceptance().to_dnf().top_disjuncts())
+    orig_scc_edg = scc_edg
+    orig_scc_state_info = scc_state_info
+    if currently_reduced == FormulaAtribute.K:
+        K -= 1
+    else:
+        C -= 1
+
[…]

### 2023-05-31 qbf/limboole_formula.py

@@ -1,0 +1,334 @@
+from telatko2.classes import SATformula
+import copy
+
+
+class Boole_f_ctor:
+    def __init__(self, edge_dict, scc_edg, scc_edges_nums,
+                 scc_state_info, C, K, mode):
+        self.univ_vars = []
+        self.exist_vars = []
+        self.edge_dict = edge_dict
+        self.scc_edg = scc_edg
+        self.scc_edges_nums = scc_edges_nums
+
+        self.scc_state_info = scc_state_info
+        self.inner_edges_nums = [
+            num for one_scc_edges_nums in scc_edges_nums for num in one_scc_edges_nums]
+        self.C = C
+        self.K = K
+        self.mode = mode  # level of simplification
+
+    def add_inf_fin(self, impl):
+        inf_not_fin = self.inf_is_not_fin_clause()
+        inf_or_fin = self.inf_or_fin_f()
+        formula = SATformula("&", [impl, inf_not_fin, inf_or_fin])
+        return formula
+
+    def get_qbf_formula(self, aut):
+
+        # adds universally quantified variables
+        self.quant_all()
+
+        if self.mode == 3:
+
+            self.w_quant()
+
+            laso = self.laso_cycles(aut)
+            in_out = self.in_n_out()
+            one_scc = self.one_scc_f()
+            laso = SATformula("&", [laso, in_out, one_scc])
+
+        elif self.mode == 2:
+            laso = self.laso_f(aut)
+
+        else:
+            laso = self.one_scc_f()
+
+        # old = self.old_formula(ACC_DNF(aut.get_acceptance().to_dnf()))
+        old = self.rec_old_original_shape(aut.get_acceptance(), 'c', 0)
+        # old.sanity_checks()
+
+        new = self.new_formula()
+
+        if old is not None:
+            eq = self.equivalence_f(old, new)
+        else:
+            eq = new
+        impl = SATformula("->", [laso, eq])
+        formula = self.add_inf_fin(impl)
+
+        if self.exist_vars:
+            formula = SATformula(
+                "?",
+                subf_list=[formula],
+                quantifier=True,
+                quantified_vars=self.exist_vars)
+            #formula = Exists(self.exist_vars, formula)
+        formula = SATformula(
+            "#",
+            subf_list=[formula],
+            quantifier=True,
+            quantified_vars=self.univ_vars)
+        #formula = ForAll(self.univ_vars, formula)
+
+        return formula
[…]

### 2023-05-31 qbf/not_incremental.py

@@ -1,0 +1,88 @@
+from qbf.formula_preparation import edge_dictionary, update_run_info, try_evaluate0, empty_aut, get_formula, filter_out_representants, query
+from qbf.parser import process_variables
+from telatko2.classes import *
+from z3 import *
+
+
+def not_incr_loop(
+        aut,
+        tmp_mode,
+        scc_state_info,
+        scc_edg,
+        currently_reduced,
+        original,
+        formula_attr):
+
+    K = aut.get_acceptance().used_sets().count()
+    C = len(aut.get_acceptance().to_dnf().top_disjuncts())
+    orig_scc_edg = scc_edg
+    orig_scc_state_info = scc_state_info
+    if currently_reduced == FormulaAtribute.K:
+        K -= 1
+    else:
+        C -= 1
+
+    while C > 0 and K > 0:
+
+        if aut.get_acceptance().used_sets().count(
+        ) < 1 or aut.prop_state_acc() == spot.trival.yes_value:
+            return aut
+
+        edge_dict, scc_equiv_edges, representants = edge_dictionary(
+            aut, tmp_mode, scc_edg, formula_attr.scc_optimization)
+
+        if tmp_mode == 1:
+            scc_edg, scc_state_info = filter_out_representants(
+                aut, set(representants), orig_scc_edg, orig_scc_state_info)
+
+        formula = get_formula(
+            aut,
+            scc_state_info,
+            C,
+            K,
+            tmp_mode,
+            edge_dict,
+            scc_edg,
+            formula_attr)
+
+        scc_edg = orig_scc_edg
+        scc_state_info = orig_scc_state_info
+
+        result, model, var_mapping = query(formula, formula_attr)
+
+        if result == Solver_result.sat:
+
+            process_variables(
+                aut,
+                model,
+                scc_equiv_edges,
+                tmp_mode,
+                formula_attr.solver, var_mapping)
+
+            """
+            if not spot.are_equivalent(original, aut):
+                print("NOT EQUIVALENT! C=", C, " K=", K)
+                return aut
+            """
+
+        else:
+
+            if currently_reduced == FormulaAtribute.K:
+                aut.set_name(
+                    aut.get_name() +
+                    update_run_info(
+                        K,
+                        result,
+                        tmp_mode))
+            if aut.get_acceptance().used_sets().count() <= 1:
+                return try_evaluate0(aut, original)
+            return aut
+
[…]

### 2023-05-31 qbf/optimization.py

@@ -1,0 +1,371 @@
+
+from telatko2.classes import *
+from qbf.formula import *
+import copy
+
+
+def get_mark_edg(aut, edges, edges_translator):
+    """
+        Returns: dictionary {acc_mark : [real_edge_numbers]}
+    """
+    dict = {}
+    counter = 1
+    for e in edges:
+        edges_translator[aut.edge_number(e)] = counter
+        for m in e.acc.sets():
+            if m in dict:
+                dict[m].append(aut.edge_number(e))
+            else:
+                dict[m] = [aut.edge_number(e)]
+        counter += 1
+    return dict
+
+
+def one_fin2(clause, acc_set, edges_translator):
+    """
+        Create formula in shape of :
+        n_ck -> (&_1<=t<=T (!f_tk | !e_t)
+        Args:
+            clause: num of clause
+            acc_set: num of acc set
+
+
+        Returns:
+
+        """
+    c = str(clause)
+    k = str(acc_set)
+    impl = SATformula('->')
+    impl.add_subf(SATformula('n_' + c + '_' + k))
+
+    conjunction = SATformula('&')
+    for t in edges_translator:
+        disj = SATformula('|')
+        disj.add_subf(SATformula('!e_' + str(edges_translator[t])))
+        disj.add_subf(SATformula('!f_' + str(t) + '_' + k))
+
+        conjunction.add_subf(disj)
+    impl.add_subf(conjunction)
+    return impl
+
+
+def one_inf2(clause, acc_set, edges_translator):
+    """
+
+    :param clause: num of clause c
+    :param acc_set: num off acc set k
+    :param edges_translator: {num_of_edge_in_aut : num_of_edge_in_SATformula}
+
+    :return:
+    """
+
+    c = str(clause)
+    k = str(acc_set)
+    impl = SATformula('->')
+    impl.add_subf(SATformula('p_' + c + '_' + k))
+
+    disjunction = SATformula('|')
+    for t in edges_translator:
+        conj = SATformula('&')
+        conj.add_subf(SATformula('e_' + str(edges_translator[t])))
+        conj.add_subf(SATformula('f_' + str(t) + '_' + k))
+
+        disjunction.add_subf(conj)
+    impl.add_subf(disjunction)
+    return impl
+
+
+def new_formula2(edges_translator, C, K):
+    """
+    Creates requirements for variables for new, short acceptance formula of automaton
+    :param inner_edges: all inner edges of SCCs of automaton
+    :param C:
[…]

### 2023-05-31 qbf/parser.py

@@ -1,0 +1,288 @@
+from telatko2.classes import *
+from z3 import is_true
+from qbf.rejecting_edges import *
+import re
+
+
+def create_acc(variables):
+    """
+
+    Args:
+        variables [[{Variable}]]: list of lists of variables - each var in list has the same clause number.
+
+    Returns:
+        [[MarkType]] - list of list of MartType of the same structure as input
+
+    """
+    acc = []
+    for dis in variables:
+        disjunct = []
+        for con in dis:
+            if con.type == 'p':
+                disjunct.append(ACCMark(MarkType.Inf, con.set))
+            else:
+                disjunct.append(ACCMark(MarkType.Fin, con.set))
+
+        acc.append(disjunct)
+    return acc
+
+
+def string_formula(acc):
+    """
+    Returns string interpretation
+    Args:
+        acc [[MarkType]]:  acceptance condition
+
+    Returns: string - string interpretation of input acc
+
+    """
+    f = []
+    for dis in acc:
+        f.append('(')
+        for con in dis:
+            f.append(str(con))
+            if con is not dis[-1]:
+                f.append(" & ")
+        f.append(')')
+        if dis is not acc[-1]:
+            f.append(" | ")
+    return ''.join(f)
+
+
+def prepare_acc_vars(acc_vars):
+    """
+
+    Args:
+        acc_vars[string]: list of variables, that represent future acceptance condition
+        and were labeled as true by SAT solver
+
+    Returns:
+        [[Variable]] - each sublist contains {Variable}s with the same conjunct number
+
+    """
+    condition_vars_class = list(
+        map(lambda var: Variable(var.split('_')), acc_vars))
+    condition_vars_class.sort(key=lambda x: x.clause, reverse=False)
+    clause_num = condition_vars_class[0].clause
+    clause_max_num = condition_vars_class[-1].clause
+    var_list = []
+    while clause_num != (clause_max_num + 1):
+        var_list.append(
+            list(
+                filter(
+                    lambda x: x.clause == clause_num,
+                    condition_vars_class)))
+
+        clause_num += 1
+    return var_list
+
+
+def print_edges(aut):
+    for edge in aut.edges():
+        print("edges[{e}].src={src}, edges[{e}].dst={dst}, edges[{e}].acc={acc}".format(
[…]

### 2023-05-31 qbf/q_main.py

@@ -1,0 +1,159 @@
+import random
+from threading import current_thread
+from qbf.parser import *
+from qbf.incremental import inc_loop
+from qbf.not_incremental import not_incr_loop
+from qbf.formula_preparation import try_evaluate0
+
+
+def scc_info(aut):
+    """
+
+    Args:
+        aut:
+
+    Returns: scc_state_info[{state_number : [edges out of the state], [edges in the state]}]
+              scc_edg [[numbers of inner edges of one SCC]]
+
+    """
+    scc_state_inf = []
+
+    # [[edges in one scc]]
+    scc_edges = []
+
+    si = spot.scc_info(aut)
+
+    for i in range(si.scc_count()):
+        states = si.states_of(i)
+        if si.is_trivial(i):
+            continue
+
+        if si.check_scc_emptiness(i):
+            # if rejecting ignore edges (for now)
+            states = si.states_of(i)
+            continue
+
+        state_dict = {}
+        for s in states:
+            state_dict[s] = [[], []]
+        edges = si.inner_edges_of(i)
+
+        scc_edges.append(list(edges))
+
+        for e in edges:
+
+            state_dict[e.src][0].append(aut.edge_number(e))
+            state_dict[e.dst][1].append(aut.edge_number(e))
+
+        scc_state_inf.append(state_dict)
+
+    return scc_state_inf, scc_edges
+
+
+def count_sets(sets):
+    counter = 0
+    for s in sets:
+        counter += 1
+    return counter
+
+
+def resolve_formula_atributes(minimized_atribute, C, K):
+    if minimized_atribute == 'clauses':
+
+        return FormulaAtribute.C
+    else:
+
+        return FormulaAtribute.K
+
+
+def play(aut, formula_attr):
+    aut.set_name("")
+
+    spot.cleanup_acceptance_here(aut)
+
+    if aut.get_acceptance().used_sets().count() < 1:
+
+        return aut
+
+    original = spot.automaton(aut.to_str())
+
+    # scc_edg [[nums of edges of one scc]]
+    # scc state info [{state num : [[num of edge of which is the state source
+    # of][num of edge of which is the state destination of]]}]
+    scc_state_info, scc_edg = scc_info(aut)
+    tmp_l = formula_attr.tmp_level
+    if formula_attr.incremental:
+        for i in range(tmp_l, formula_attr.level + 1):
+            if aut.get_acceptance().used_sets().count() <= 1:
+                aut = try_evaluate0(aut, original)
[…]

### 2023-05-31 qbf/rejecting_edges.py

@@ -1,0 +1,46 @@
+import spot
+from telatko2.classes import MarkType
+
+
+def clear_aut_edges(aut):
+    for e in aut.edges():
+        e.acc = spot.mark_t()
+
+
+def list_of_rejecting_edges(aut):
+
+    rejecting_edges = []
+    si = spot.scc_info(aut)
+
+    for i in range(si.scc_count()):
+
+        if si.is_trivial(i):
+            continue
+
+        if si.check_scc_emptiness(i):
+            for e in si.inner_edges_of(i):
+                rejecting_edges.append(e)
+
+    return rejecting_edges
+
+
+def rejecting_sccs_edges(aut, acc):
+    """
+    Takes all edges from original automaton and makes them rejecting for the new acceptance condition.
+    (As side effect has to clean all edges of the automaton.)
+    """
+
+    rejecting_edges = list_of_rejecting_edges(aut)
+
+    # returns cpp pair <bool, spot.mark_t>, if bool is False, there is no rejecting combination
+    # there should be rejecting combination since we are dealing with
+    # rejecting SCCs
+    rejecting_marks = spot.acc_cond(spot.acc_code(acc)).unsat_mark()
+
+    # fails if there is no unsat combination of marks
+    assert (rejecting_marks[0])
+    # removes all acceptance marks from edges
+    clear_aut_edges(aut)
+
+    for e in rejecting_edges:
+        e.acc = rejecting_marks[1]

### 2023-05-31 qbf/solver_manipulation/encode.py

@@ -1,0 +1,73 @@
+
+
+import subprocess
+from telatko2.classes import SATformula, safe_file
+import re
+
+
+def encode_qdimacs(boole_formula_path: str) -> str:
+    """
+        return :str name of qdimacs file
+    """
+    print(f"booleguru formula path : {boole_formula_path} ")
+
+    try:
+
+        cp = subprocess.run(["./../solvers/booleguru/build/booleguru",
+                             boole_formula_path,
+                             "--qdimacs"],
+                            universal_newlines=True,
+                            stdout=subprocess.PIPE,
+                            stderr=subprocess.PIPE,
+                            timeout=20)
+
+    except subprocess.TimeoutExpired:
+        print("expired")
+        return
+    except BaseException:
+        print("Booleguru error")
+        return
+
+    qdimacs_file_path = "./qbf/solver_manipulation/formulas_files/qbf_formula.qdimacs"
+
+    safe_file(qdimacs_file_path, cp.stdout, "w")
+
+    return qdimacs_file_path
+
+
+def decode_qdimacs(qdimacs_path: str):
+    """
+        Booleguru encodes mapping of original variables to numbers
+        into the comments in the beginning of the file.
+    """
+    re_pattern = "c.*"
+    f = open(qdimacs_path, "r")
+    qdim_query = [line for line in f.readlines() if re.match(re_pattern, line)]
+    f.close()
+
+    return qdim_query
+
+
+def create_formula10():
+    """
+        ?t ?u #x #y (a & b & c & (d|e)) -> ( f & g & h)
+    """
+    left = SATformula("&", [SATformula("a"), SATformula("b"), SATformula(
+        "c"), SATformula("|", [SATformula("d"), SATformula("e")])])
+    left.negate()
+    right = SATformula(
+        "&", [SATformula("f"), SATformula("g"), SATformula("h")])
+    c = SATformula("->", [left, right])
+    prefix = SATformula("#", [c], False, True, ["x", "y"])
+    prefix2 = SATformula("?", [prefix], False, True, ["t", "u"])
+
+    return prefix2
+
+
+def test():
+
+    f = create_formula10()
+    encode_qdimacs(f)
+
+
+# test()

### 2023-05-31 qbf/solver_manipulation/solver_functions.py

@@ -1,0 +1,122 @@
+import subprocess
+from z3 import *
+import time
+from telatko2.classes import Solver_result, SATformula, safe_file
+from qbf.solver_manipulation.encode import encode_qdimacs, decode_qdimacs
+from transformations.qcir import *
+from typing import List
+
+
+COUNTER = 0
+
+
+def booleguru_translate(out_path):
+    global COUNTER
+
+    try:
+        cp = subprocess.run(["./../solvers/booleguru/build/booleguru",
+                             "./qbf_formula.boole",
+                             "--qdimacs"],
+                            universal_newlines=True,
+                            stdout=subprocess.PIPE,
+                            stderr=subprocess.PIPE,
+                            timeout=3)
+        cp2 = subprocess.run(
+            [
+                "./../solvers/booleguru/build/booleguru",
+                "./qbf_formula.boole",
+                "--qcir"],
+            universal_newlines=True,
+            stdout=subprocess.PIPE,
+            stderr=subprocess.PIPE,
+            timeout=3)
+    except subprocess.TimeoutExpired:
+        print("expired")
+        return
+
+    safe_file(f"{out_path}/qdimacs/telatko{COUNTER}.qdimacs", cp.stdout, "w")
+    safe_file(f"{out_path}/qcir/telatko{COUNTER}.qcir", cp2.stdout, "w")
+
+    COUNTER += 1
+
+
+def z3_result(result, model=None):
+
+    if result == sat:
+        return Solver_result.sat, model, None
+    elif result == z3.unknown:
+        return Solver_result.timeout, None, None
+    elif result == unsat:
+        return Solver_result.unsat, None, None
+    assert(False)
+
+
+def query(formula, formula_attr):
+    formula_file_path = "./qbf/solver_manipulation/formulas_files/qbf_formula.boole"
+
+    if formula_attr.solver == "z3":
+        solver = Solver()
+        solver.add(formula)
+        solver.set("timeout", 1000 * formula_attr.timeout)
+        solver.push()
+        result = solver.check()
+        model = None
+        if result == sat:
+            model = solver.model()
+        return z3_result(result, model)
+    elif formula_attr.solver == "quabs":
+
+        qcir_c = Qcir_ctor(formula)
+        qcir_c.to_qcir("./qbf/solver_manipulation/formulas_files/formula.qcir")
+        return Solver_result.unsat, None
+    elif formula_attr.solver == "caqe":
+        safe_file(formula_file_path, str(formula), "w")
[…]

### 2023-05-31 qbf/z3_formula.py

@@ -1,0 +1,352 @@
+from z3 import *
+from telatko2.classes import ACC, ACC_DNF, MarkType
+
+
+class Z3_f_ctor:
+    def __init__(self, edge_dict, scc_edg, scc_edges_nums,
+                 scc_state_info, C, K, mode):
+        self.univ_vars = []
+        self.exist_vars = []
+        self.edge_dict = edge_dict
+        self.scc_edg = scc_edg
+        self.scc_edges_nums = scc_edges_nums
+
+        self.scc_state_info = scc_state_info
+        self.inner_edges_nums = [
+            num for one_scc_edges_nums in scc_edges_nums for num in one_scc_edges_nums]
+        self.C = C
+        self.K = K
+        self.mode = mode  # level of simplification
+
+    def add_inf_fin(self, impl):
+        inf_not_fin = self.inf_is_not_fin_clause()
+        inf_or_fin = self.inf_or_fin_f()
+        inf_not_fin = simplify(inf_not_fin)
+        inf_or_fin = simplify(inf_or_fin)
+        formula = And(impl, inf_not_fin, inf_or_fin)
+        return formula
+
+    def get_qbf_formula(self, aut):
+
+        # adds universally quantified variables
+        self.quant_all()
+
+        if self.mode == 3:
+
+            self.w_quant()
+
+            laso = self.laso_cycles(aut)
+            laso = simplify(laso)
+            in_out = self.in_n_out()
+            one_scc = self.one_scc_f()
+            laso = And(laso, in_out, one_scc)
+            # least_one_edge = self.disjunct_formula( self.inner_edges_nums)
+            # laso = And(laso, least_one_edge)
+
+        elif self.mode == 2:
+            laso = self.laso_f(aut)
+
+        else:
+            one_scc = self.one_scc_f()
+            laso = one_scc
+
+        # old = self.old_formula(ACC_DNF(aut.get_acceptance().to_dnf()))
+        old = self.rec_old_original_shape(aut.get_acceptance(), 'c', 0)
+
+        new = self.new_formula()
+
+        new = simplify(new)
+
+        laso = simplify(laso)
+
+        if old is not None:
+            old = simplify(old)
+            eq = self.equivalence_f(old, new)
+        else:
+            eq = new
+        impl = Implies(laso, eq)
+        formula = self.add_inf_fin(impl)
+        if self.exist_vars:
+            formula = Exists(self.exist_vars, formula)
+
[…]

### 2023-05-31 statistics/sat23-camera-ready/current-z3-version/patterns/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - patterns")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = '/home/xschwar3/telatko/statistics/sat23-camera-ready/patterns'
+    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-05-31 statistics/sat23-camera-ready/current-z3-version/random/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - random")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = '/home/xschwar3/telatko/statistics/sat23-camera-ready/random'
+    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/random-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-05-31 statistics/sat23-camera-ready/patterns/translate.py

@@ -1,0 +1,38 @@
+import spot
+import os
+import re
+from ltlcross_runner import LtlcrossRunner, ResultsAnalyzer
+
+
+def main(argv):
+    tools = {
+    'ltl2tgba_det': 'ltl2tgba -D -G %f > %O ',
+    'ltl3tela_det': '/home/xschwar3/ltl3tela-master/ltl3tela  -D1 -f %f > %O',
+    'ltl2tgba': 'ltl2tgba %f > %O ',
+    'ltl3tela': '/home/xschwar3/ltl3tela-master/ltl3tela -f %f > %O',
+    'delag': '/home/xschwar3/owl/bin/owl ltl2dela -f %f > %O',
+    'rabinizer4': '/home/xschwar3/owl/bin/owl ltl2dgra -f %f > %O',
+    }
+
+    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+
+    pattern_ltl = 'pattern.ltl'
+    pattern_aut = 'pattern.csv'
+    lcr = LtlcrossRunner(tools, formulas = [pattern_ltl], res_filename = pattern_aut)
+    lcr.run_ltlcross(automata = True, timeout = '60')
+    ra = ResultsAnalyzer(rand_aut)
+    ra.parse_results()
+    ltlf_translated = {}
+
+    for tool in tools_order:
+        fs = ra.values.filter(items = [('acc', tool)]).dropna()
+
+        ltlf_translated[tool] = len(fs[fs[('acc', tool)] > 1])
+
+        fd = open('pattern-' + tool + '.hoa', 'w')
+        for (ix, _), _ in fs.iterrows():
+            fd.write(ra.aut_for_id(ix, tool).to_str() + '\n')
+        fd.close()
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-05-31 statistics/sat23-camera-ready/random/translate.py

@@ -1,0 +1,38 @@
+import spot
+import os
+import re
+from ltlcross_runner import LtlcrossRunner, ResultsAnalyzer
+
+
+def main(argv):
+    tools = {
+    'ltl2tgba_det': 'ltl2tgba -D -G %f > %O ',
+    'ltl3tela_det': '/home/xschwar3/ltl3tela-master/ltl3tela  -D1 -f %f > %O',
+    'ltl2tgba': 'ltl2tgba %f > %O ',
+    'ltl3tela': '/home/xschwar3/ltl3tela-master/ltl3tela -f %f > %O',
+    'delag': '/home/xschwar3/owl/bin/owl ltl2dela -f %f > %O',
+    'rabinizer4': '/home/xschwar3/owl/bin/owl ltl2dgra -f %f > %O',
+    }
+
+    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+
+    rand_ltl = 'random.ltl'
+    rand_aut = 'random.csv'
+    lcr = LtlcrossRunner(tools, formulas = [rand_ltl], res_filename = rand_aut)
+    lcr.run_ltlcross(automata = True, timeout = '60')
+    ra = ResultsAnalyzer(rand_aut)
+    ra.parse_results()
+    ltlf_translated = {}
+
+    for tool in tools_order:
+        fs = ra.values.filter(items = [('acc', tool)]).dropna()
+
+        ltlf_translated[tool] = len(fs[fs[('acc', tool)] > 1])
+
+        fd = open('random-' + tool + '.hoa', 'w')
+        for (ix, _), _ in fs.iterrows():
+            fd.write(ra.aut_for_id(ix, tool).to_str() + '\n')
+        fd.close()
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-05-31 statistics/sat23-camera-ready/sat22-z3-version/patterns/experiments.py

@@ -1,0 +1,20 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    telatko = "/home/xschwar3/Desktop/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = '/home/xschwar3/telatko/statistics/sat23-camera-ready/patterns'
+    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-05-31 statistics/sat23-camera-ready/sat22-z3-version/random/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("sat22 telatko z3 version - random")
+    telatko = "/home/xschwar3/Desktop/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = '/home/xschwar3/telatko/statistics/sat23-camera-ready/random'
+    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/random-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-05-31 statistics/sat23/SAT22_version/experiments.py

@@ -1,0 +1,20 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    telatko = "/home/xschwar3/Desktop/telatko/telatko.py"
+    
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = '/home/xschwar3/telatko/sat23/patterns' 
+    tools_order = ['ltl2tgba_det', 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-05-31 telatko.py

@@ -1,0 +1,126 @@
+#!/usr/bin/env python3
+
+import spot
+import sys
+import argparse
+import os
+
+from telatko2.classes import Switches
+from qbf.q_main import play, print_aut
+from telatko2.playground import process_automaton
+
+
+def parser(argv):
+    parser = argparse.ArgumentParser()
+    parser.add_argument(
+        "-F",
+        "--autfile",
+        help="File containing automata in HOA format.")
+    parser.add_argument(
+        "-O",
+        "--outfile",
+        help="File to print output to.",
+        default=None)
+    parser.add_argument(
+        "-L",
+        "--level",
+        help="Level of simplification.",
+        type=int,
+        default=0)
+    parser.add_argument(
+        "-T",
+        "--timeout",
+        help="Kill QBF solver after inserted number of seconds.",
+        type=int,
+        default=50)
+    parser.add_argument(
+        "-S",
+        "--scc",
+        help="Use SCC optimization.", action='store_true')
+    parser.add_argument(
+        "-C",
+        "--minimize_clauses",
+        help="Turn on reduction of clauses.", action='store_true')
+
+    parser.add_argument(
+        "-I",
+        "--incremental",
+        help="Use incremental solving.", action='store_true')
+    parser.add_argument(
+        "-G",
+        "--gradual",
+        help="Levels will gradually grow from L1 up to specified level.",
+        action='store_true')
+
+    parser.add_argument(
+        "-M",
+        "--solver",
+        help="Choose solver",
+        choices=['limboole', 'z3', 'quabs', 'caqe'],
+        default='z3')
+    args = parser.parse_args()
+    return args
+
+
+def main(argv):
+    args = parser(argv)
+
+    if not args.autfile:
+        print("No automata to process.", file=sys.stderr)
+
+    # if no arguments given, default variant is gradually go up to level 3
+    if args.level == 0 and not args.gradual and not args.incremental and not args.scc and args.solver == "z3":
+        args.gradual = True
+        args.level = 3
+        args.incremental = True
+        args.scc = True
+
+    elif args.incremental and args.solver == "limboole":
+        print("Incremental solving is possible only with Z3 solver.")
+        return
+
+    elif args.level == 0:
+        args.gradual = True
+        args.level = 3
+
+    if args.level > 3:
+        args.level = 3
+
+    aut = spot.automata(args.autfile)
+    formula_attributes = Switches(args)
+    # formula_attributes.print_settings()
+
[…]

### 2023-05-31 telatko2/classes.py

@@ -1,0 +1,755 @@
+#!/usr/bin/python3
+
+
+import spot
+import enum
+import sys
+import argparse
+from scipy.optimize import linear_sum_assignment
+import numpy as np
+import copy
+import os
+import subprocess
+import itertools
+from math import ceil, log2
+
+
+class Solver_result(enum.Enum):
+    """
+        Result of QBF query.
+    """
+    sat = 1
+    unsat = 2
+    timeout = 3
+
+
+class Variable:
+    """
+    Class that contains information retrieved from SAT-result.
+     Variable in this class formulates shape of new acceptance condition.
+    """
+
+    def __init__(self, var):
+        self.type = var[0]
+        self.clause = int(var[1])
+        self.set = int(var[2].split('=')[0])
+
+
+class Switches:
+    """
+    Class containing overwiev of switches set by user.
+    """
+
+    def __init__(self, args):
+        self.level = args.level
+        self.tmp_level = 1 if args.gradual else args.level
+        self.min_clauses = args.minimize_clauses
+        self.timeout = args.timeout
+        self.scc_optimization = args.scc
+        self.incremental = args.incremental
+        self.gradual = args.gradual
+        self.C = None
+        self.solver = args.solver  # z3 or limboole
+
+    def set_C(self, aut):
+        self.C = len(
+            aut.get_acceptance().to_dnf().top_disjuncts())
+
+    def print_settings(self):
+        print("Global Level ", self.level)
+        print("Current Level ", self.tmp_level)
+        print("Minimize clauses ", self.min_clauses)
+        print("Timeout ", self.timeout)
+        print("SCC optimization ", self.scc_optimization)
+        print("Incremental ", self.incremental)
+        print("Gradual ", self.gradual)
+        print("Solver ", self.solver)
+
+
+class SATformula:
+    """
+    Represents boolean formula, retrieved from the input automata, that will be given to SAT-solver.
+
+    """
+
+    def __init__(
+            self,
+            op,
+            subf_list: list = [],
+            negate: bool = False,
+            quantifier=False,
+            quantified_vars=[]):
+        self.operator = op
+        self.subformula = subf_list
+        self.negation = negate
+        self.quantifier = quantifier
+        self.quantified_vars = quantified_vars
+
+    def __len__(self):
+        if not self.subformula:
+            return 1
+        else:
+            return sum(map(len, self.subformula))
+
+    def __str__(self):
+
+        if self.quantifier:
+            if self.subformula:
+                return self.operator + \
[…]

### 2023-05-31 telatko2/merge.py

@@ -1,0 +1,586 @@
+
+from telatko2.classes import *
+from telatko2.simplify import *
+import copy
+import warnings
+warnings.filterwarnings("ignore", category=np.VisibleDeprecationWarning)
+
+
+def shift_fst_acc(aut, acc, scc, m):
+    next_m = m + 1
+    log = {}
+    #
+
+    for dis in acc.formula:
+        for con in dis:
+            if con.num in log:
+                con.num = log[con.num]
+            else:
+                add_dupl_marks(aut, scc, con.num, next_m)
+                log[con.num] = next_m
+                con.num = next_m
+                next_m += 1
+    scc_clean_up_edges(aut, acc, scc)
+
+
+def count_cost(dis1, dis2):
+    inf, fin = 0, 0
+    for con in dis1:
+        if con.type == MarkType.Inf:
+            inf -= 1
+        else:
+            fin -= 1
+    for con in dis2:
+        if con.type == MarkType.Inf:
+            inf += 1
+        else:
+            fin += 1
+    if inf < 0:
+        inf = 0
+    if fin < 0:
+        fin = 0
+    return inf + fin
+
+
+def make_matrix(acc1, acc2):
+    m = []
+    for i in range(len(acc1)):
+        row = []
+        for j in range(len(acc1)):
+            if j >= len(acc2):
+                row.append(0)
+            else:
+                row.append(count_cost(acc1[i], acc2[j]))
+        m.append(row)
+    return m
+
+
+def replace_marks(aut, scc, origin_m, new_m):
+    for s in scc.states():
+        for e in aut.out(s):
+            if e.dst in scc.states() and origin_m in e.acc.sets():
+                e.acc.clear(origin_m)
+                if new_m not in e.acc.sets():
+                    e.acc.set(new_m)
+
+
+def longest_expr_in_acc(acc):
+    max = 0
+    for expr in acc.formula:
+        if len(expr) > max:
+            max = len(expr)
+    return max
+
+
+def partition(accs, sccs, low, high):
+    """
+    Formula is in dnf, len == number of disjuncts
+    Args:
+        accs: list of acceptance formulas
+        sccs: list of sccs corressponding to accs
+        low: int
+        high: int
+
+    Returns:
+
+    """
+    i = (low - 1)
+    pivot = len(accs[high])
+
+    for j in range(low, high):
+        if len(accs[j]) > pivot:
+            i += 1
[…]

### 2023-05-31 telatko2/playground.py

@@ -1,0 +1,309 @@
+
+import random
+
+
+from telatko2.merge import *
+from telatko2.restore_equivalence import *
+import copy
+from qbf.parser import clear_aut_edges
+
+
+def try_eval(aut, acc, scc):
+    """
+
+    """
+    m_all_edges = set(scc_everywhere(aut, scc))  # set of marks that include
+    # all transitions of SCC
+
+    eval_f = False  # proc
+    for dis in acc.formula:
+        eval_dis = True
+        for con in dis:
+            val = eval_set(aut, con, scc, m_all_edges)
+            if eval_dis is False or val is False:
+                eval_dis = False
+            elif (eval_dis is None or eval_dis is True) and val is None:
+                eval_dis = None
+        if eval_dis is True:
+            acc.sat = True
+            acc.formula = []
+            return
+        elif (eval_f is None or eval_f is False) and eval_dis is None:
+            eval_f = None
+    acc.set_sat(eval_f)
+    if acc.sat is not None:
+        acc.formula = []
+
+
+def weak_eval(aut, acc, scc, weak):
+    if weak:
+        if scc.is_rejecting():  # all cycles in scc are rejecting
+            acc.set_sat(False)
+        else:
+            acc.set_sat(True)
+        acc.formula = []
+        return True
+    return False
+
+
+def print_aut(aut, output, m):
+    if output is not None:
+        f = open(output, m)
+        f.write(aut.to_str() + '\n')
+        f.close()
+    else:
+        print(aut.to_str())
+
+
+def get_acc(aut):
+    """
+    Returns PACC (dnf) if dnf formula is shorter else PACC_CNF(cnf)
+    """
+    dnf_acc = ACC_DNF(aut.get_acceptance().to_dnf())
+    # return dnf_acc
+    cnf_acc = ACC_CNF(aut.get_acceptance().to_cnf())
+
+    # print((cnf_acc),"\n", (dnf_acc))
+
+    if (dnf_acc.acc_len() < cnf_acc.acc_len()) or (dnf_acc.acc_len()
+                                                   == cnf_acc.acc_len() and len(dnf_acc) <= len(cnf_acc)):
+        # print("DNF")
+        return dnf_acc
+    else:
+        # acc = PACC_CNF(cnf_acc)
+        # print("CNF")
+        return cnf_acc
+
+
+def get_accs(aut):
+    """
+    Simplifies acceptance condition in DNF for every SCC
+    Args:
+        aut:
+
[…]

### 2023-05-31 telatko2/restore_equivalence.py

@@ -1,0 +1,246 @@
+
+from telatko2.classes import *
+from telatko2.simplify import *
+from telatko2.merge import get_set_nums
+
+
+def make_true(aut, scc, merged_acc):
+    scc_clean_up_edges(aut, ACC_DNF(""), scc)
+
+    min_inf = len(merged_acc[0])
+    cheap_dis = merged_acc[0]
+    for dis in merged_acc.formula:
+        if all(con.type == MarkType.Fin for con in dis):
+            return
+        count = 0
+        for con in dis:
+            if con.type == MarkType.Inf:
+                count += 1
+        if count < min_inf:
+            min_inf = count
+            cheap_dis = dis
+
+    add_m = []
+    for con in cheap_dis:
+        if con.type == MarkType.Inf:
+            add_m.append(con.num)
+
+    if not add_m:
+        return
+    for s in scc.states():
+        for e in aut.out(s):
+            if e.dst in scc.states():
+                for m in add_m:
+                    e.acc.set(m)
+
+
+def make_false(aut, scc, merged_acc):
+    scc_clean_up_edges(aut, ACC_DNF(""), scc)
+    add_m = []
+    for dis in merged_acc.formula:
+        if all(con.type == MarkType.Fin for con in dis):
+            add_m.append(dis[0].num)
+
+    if not add_m:
+        return
+    for s in scc.states():
+        for e in aut.out(s):
+            if e.dst in scc.states():
+                for m in add_m:
+                    e.acc.set(m)
+
+
+def make_one_true(aut, scc, one):
+    """
+    Make ACCMark true in particular expression - Inf on every transition
+    :param aut:
+    :param scc:
+    :param one: ACCMark
+    :return:
+    """
+    if one.type == MarkType.Inf:
+        for s in scc.states():
+            for e in aut.out(s):
+                if e.dst in scc.states():
+                    e.acc.set(one.num)
+
+
+def make_one_false(aut, scc, one):
+    """
+    Make ACCMark true in particular expression - Fin on every transition
+    :param aut:
+    :param scc:
+    :param one: ACCMark
+    :return:
+    """
+    if one.type == MarkType.Fin:
+        for s in scc.states():
+            for e in aut.out(s):
+                if e.dst in scc.states():
+                    e.acc.set(one.num)
+
+
+def make_expr_true(aut, scc, expr):
+    """
+    Make con/dis true
[…]

### 2023-05-31 transformations/cnf_transform.py

@@ -1,0 +1,354 @@
+from telatko2.classes import SATformula
+from copy import deepcopy
+from transformations.tseitin import *
+from typing import Callable
+import subprocess
+import sys
+sys.path.append('..')
+
+
+class Transform:
+    def __init__(self, formula):
+        self.formula = formula
+
+    def __str__(self):
+        return str(self.formula)
+
+    def eliminate_implication_init(self):
+        self.eliminate_implication_rec(self.formula)
+
+    def eliminate_implication(self):
+        self.eliminate_implication_init()
+
+    def eliminate_implication_rec(self, ast):
+        #print(f"ast before {ast}")
+        if ast.operator == "->":
+            assert(len(ast.subformula) == 2)
+            ast.set_op("|")
+            ast.subformula[0].negate()
+        #print(f"ast after {ast}")
+        for subf in ast.subformula_list():
+            self.eliminate_implication_rec(subf)
+
+    def eliminate_equivalence(self):
+        self.eliminate_equivalence_init()
+
+    def eliminate_equivalence_init(self):
+        self.eliminate_equivalence_rec(self.formula)
+
+    def eliminate_equivalence_rec(self, ast):
+        #print(f"ast before {ast}")
+        if ast.operator == "<->":
+            assert(len(ast.subformula) == 2)
+            ast.set_op("&")
+
+            left = deepcopy(ast.subformula[0])
+            right = deepcopy(ast.subformula[1])
+            right.negate()
+            l_notr = SATformula("|", [deepcopy(left), deepcopy(right)])
+            right.negate()
+            left.negate()
+            notl_r = SATformula("|", [deepcopy(left), deepcopy(right)])
+            ast.subformula = [l_notr, notl_r]
+
+        #print(f"ast after {ast}")
+        for subf in ast.subformula_list():
+            self.eliminate_equivalence_rec(subf)
+
+    def distribute_not(self):
+        self.distribute_not_init()
+
+    def distribute_not_init(self):
+        self.distribute_not_rec(self.formula)
+        # self.formula.print_info()
+
+    def distribute_not_rec(self, ast):
+
+        #assert(ast.get_op() == ("&" or "|") )
+        # print(ast)
+
+        if ast.is_negated() and not ast.is_empty():
+            if ast.get_op() == "&":
+                ast.set_op("|")
+            else:
+                ast.set_op("&")
+            ast.negate()
+            for subf in ast.subformula_list():
[…]

### 2023-05-31 transformations/qcir.py

@@ -1,0 +1,115 @@
+from telatko2.classes import SATformula, safe_file
+from transformations.cnf_transform import *
+import sys
+sys.path.append('..')
+
+
+class Qcir_ctor:
+    def __init__(self, formula):
+        self.formula = formula
+        self.new_var_num = 1
+        self.free_vars = set()
+        self.body = []  # will contain qcir lines representing body of a formula
+        self.var_dict = {}
+
+    def to_nnf(self):
+        t = Transform(self.formula)
+        t.negation_nf(True)
+
+    def get_fresh_var(self) -> str:
+        """
+            Return str fresh variable t_n
+            and increment self.new_var - smallest unused var num
+        """
+        self.new_var_num += 1
+        return self.new_var_num - 1
+
+    def write_headline(self, file_handle):
+        file_handle.write("#QCIR-G14\n")
+
+    def add_var_to_dict(self, var: str, negated: bool = False) -> str:
+        if var in self.var_dict:
+            return str(-1 * self.var_dict[var]
+                       if negated else self.var_dict[var])
+        self.var_dict[var] = self.get_fresh_var()
+
+        return str(-1 * self.var_dict[var] if negated else self.var_dict[var])
+
+    def encode_prefix(self, prefix_ast: SATformula, file_handle):
+        if prefix_ast.operator == "#":
+            file_handle.write(
+                f"forall({','.join(map(self.add_var_to_dict, prefix_ast.quantified_vars))})\n")
+        else:
+            file_handle.write(
+                f"exists({','.join(map(self.add_var_to_dict, prefix_ast.quantified_vars))})\n")
+        for child in prefix_ast.subformula:
+            self.encode_prefix(child, file_handle)
+
+    def encode_output(self, file_handle, var: str):
+        """
+            Returns main variable
+        """
+
+        file_handle.write(f"output({var})\n")
+
+    def encode_body_rec(self, ast: SATformula) -> str:
+        if ast.is_leaf():
+            # we need to keep check of free variables, every var beginning on
+            # other letter is quantified
+            if ast.operator[0] in ['f', 'p', 'n']:
+                #print(f"adding free var {ast} first letter {ast.operator[0]} ")
+                self.free_vars.add(self.add_var_to_dict(ast.operator))
+            return self.add_var_to_dict(ast.operator, ast.negation)
+
+        children_vars = []
+
[…]

### 2023-05-31 transformations/qdimacs.py

@@ -1,0 +1,160 @@
+from transformations.cnf_transform import *
+from telatko2.classes import SATformula
+import sys
+sys.path.append('..')
+
+
+class Qdimacs_transform:
+    def __init__(self, formula):
+        self.boole_formula = formula
+        self.mapping = {}
+        self.qdim_body: [str] = []
+        self.qdim_prefix = []
+        self.new_var_num = 1
+        self.preamble = None  # (number of atoms, number of clauses)
+
+    def check_cnf(self) -> bool:
+        if not self.boole_formula.check_cnf():
+            print("Formula is not in CNF")
+            return False
+        return True
+
+    # create qdimacs formula
+
+    def create_atom_var(self, var: str, negated: bool = False) -> int:
+        """
+            var : variable in boole format
+            Get variable from map (or create mapping it).
+        """
+
+        n = 0
+
+        if var in self.mapping:
+            n = self.mapping[var]
+        else:
+            self.mapping[var] = self.new_var_num
+            n = self.new_var_num
+            self.new_var_num += 1
+        return n * -1 if negated else n
+
+    def create_qdim_disjunct(self, ast: SATformula):
+
+        assert(ast.get_op() == "|")
+        disj_list = []
+
+        for child in ast.subformula_list():
+            var_num = self.create_atom_var(child.get_op(), child.is_negated())
+            disj_list.append(var_num)
+        self.qdim_body.append(disj_list)
+
+    def create_qdim_body(self, ast: SATformula):
+
+        if not self.check_cnf():
+            return
+
+        if ast.is_leaf():
+            # creates qdim variable and returns it
+            var_num = self.create_atom_var(
+                var=ast.get_op(), negated=ast.is_negated())
+            self.qdim_body.append([var_num])
+
+        elif ast.get_op() == "|":
+            # formula of the form : (a | b | c)
+            self.create_qdim_disjunct(ast)
+        else:
+            # formula of the fotm (a | b) & (c | d) & e
+            for child in ast.subformula_list():
+                self.create_qdim_body(child)
+
+    def create_qdim_quant_part(self, quant_ast: SATformula):
+        if not quant_ast:
+            # no quantified prefix (quant_ast == None)
+            return
+        assert(quant_ast.is_quantified())
[…]

### 2023-05-31 transformations/tseitin.py

@@ -1,0 +1,187 @@
+from copy import deepcopy
+from telatko2.classes import SATformula
+import sys
+sys.path.append('..')
+
+
+class Tseitin_transform:
+    """Transforms SATformula into Tseitin's encoding."""
+
+    def __init__(self, sat_formula):
+        self.old_formula = sat_formula
+        self.tseitin_formula = SATformula("&", [])
+        self.prefix = None
+        self.new_var_index = 0
+
+    def new_variable(self):
+
+        var = SATformula("h_" + str(self.new_var_index), [])
+        self.new_var_index += 1
+        return var
+
+    def skip_prefix(self, formula):
+
+        while formula.is_quantified():
+            # each quantifier has only one successor
+
+            assert(len(formula.subformula_list()) == 1)
+            prefix = deepcopy(formula)
+            prefix.delete_subformulas()
+            self.add_prefix(prefix)
+            formula = formula.pop_subformula()
+
+        return formula
+
+    def quantify_tseitin_vars(self):
+
+        if self.prefix:
+
+            self.prefix.last_node().add_subf(
+                SATformula(
+                    "?",
+                    subf_list=[],
+                    negate=False,
+                    quantifier=True,
+                    quantified_vars=[
+                        f"h_{i}" for i in range(
+                            self.new_var_index)]))
+        # if there is no prefix, all variables are outermost exist quantified
+        # variables
+
+    def transform_init(self):
+
+        self.prefix, prefix_free_formula = self.old_formula.get_prefix_and_nonprefix_part()
+        if prefix_free_formula.is_leaf():
+            self.tseitin_formula = prefix_free_formula
+        else:
+            help_root = self.new_variable()
+            self.tseitin_formula.add_subf(help_root)
+            self.transform_rec(prefix_free_formula, SATformula(
+                "<->", [deepcopy(help_root)]))
+
+    def transform(self):
+        self.transform_init()
+
+        self.quantify_tseitin_vars()
+
+    def transform_rec(self, old_root, new_tseitin_ekv):
+        """
+        old_root - root of original subformula
+        new_tseitin_ekv - SATformula("<->", [h_i])
+        """
+
+        assert(not old_root.quantifier)
+
+        new_root = SATformula(old_root.get_op(), [])
+        if old_root.is_negated():
[…]

### 2023-06-01 paper/main.tex
progress mainly in Section 3 of main.tex

@@
 \keywords{Emerson-Lei automata, TELA, automata reduction, QBF, telatko}
-%\funding{This work has been supported by the Czech Science Foundation grant GA19-24397S.}
+\funding{T.~Schwarzová received funding from
+the European Union’s Horizon Europe program under the Grant Agreement
+No.\ 101087529.}
 
@@
 
-TELA have attracted a lot of attention during the last %five
+TELA have attracted a lot of attention during the last few
 years~\cite{baier.19.atva,major.19.atva,muller.17.gandalf,renkin.20.atva}.
@@
 \end{itemize}
-To ensure that the modified automaton is equivalent to the original one,
-we actually look for $\psi$ and $\rem$ such that each run
+% To simplify the task, 
+% To ensure that the modified automaton is equivalent to the original one,
+We will actually look for $\psi$ and $\rem$ such that each run
 $\pi=t_0t_1t_2\ldots$ of $\mA$ is accepting if and only if the run
-$\rem(t_0)\rem(t_1)\rem(t_2)\ldots$ of $\mB$ is accepting.
-
+$\rem(t_0)\rem(t_1)\rem(t_2)\ldots$ of $\mB$ is accepting. This
+requirement clearly guarantees the equiavalence of $\mA$ and $\mB$. We
+note that preserving the acceptance of individual runs is not a
+necessary condition for language equivalence of $\mA$ and $\mB$ and
+there exist automata where relaxing this requirement can lead to
+a bigger reduction of acceptance marks (see
+Figure~\ref{fig:counter}). However, looking for $\psi$ and $\rem$
+preserving the acceptance of individual runs makes the problem easier
+as we can, for example, ignore the labelling of transitions by the
+elements of $\Sigma$.
+\begin{figure}[t]
+  \centering
+  \begin{tikzpicture}[smallautomaton]
+    \begin{scope}[initial angle=180,xshift=-3.5cm]
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (2,0) {};
+      \path[->,auto]
+      (p) edge[out=120,in=60,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
+      (p) edge[out=-60,in=-120,looseness=10] pic {acc={1}{1}} pic {l=$b$} (p)
+      (p) edge[out=30,in=150] pic {l=$a$} (q)
+      (q) edge[out=-150,in=-30] pic {l=$b$} (p);
+
+      \node[acclabel] at (1,-1.7)
+      {$\Inf\tacc{0}{0}\wedge\Inf\tacc{1}{1}$};
+    \end{scope}
+    \begin{scope}[initial angle=180,xshift=3.5cm]
+      \node[state,initial] (p) at (0,0) {};
+      \node[state] (q) at (2,0) {};
+      \path[->,auto]
+      (p) edge[out=120,in=60,looseness=10] pic {l=$a$} (p)
+      (p) edge[out=-60,in=-120,looseness=10] pic {l=$b$} (p)
+      (p) edge[out=30,in=150] pic {acc={0}{0}} pic {l=$a$} (q)
+      (q) edge[out=-150,in=-30] pic {l=$b$} (p);
+
+      \node[acclabel] at (1,-1.7)
+      {$\Inf\tacc{0}{0}$};
+    \end{scope}
+  \end{tikzpicture}
+  \smallskip
+
+  \caption{Both automata accept the same language of words
+    $w\in\{a,b\}^\omega$ that contain infinitely many occurences of
+    both $a$ and $b$. Each accepting run of the left automaton has to
+    contain infinitely many occurrences of both transitions looping on
+    the initial state. Hence, there does not exist any automaton with
+    the same accepting runs as the left automaton and less than two
+    accetpance marks.}
+  \label{fig:counter}
+\end{figure}
+
+\fixme{TADY TO HONZA JESTE UPRAVUJE}
 We start with several simple observations. Acceptance of a run $\pi$
@@
 % $\pi$.
+\begin{lemma}
+  A set $T\subseteq\delta$ is a cycle if and only if
+  there is a run $\pi$ such that $T=\minf(\pi)$.
+\end{lemma}
+
+As $\minf(\pi)$ is the set of transitions that appear infinitely many
+times in $\pi$, there has to be a suffix $\pi'$ of $\pi$ containing
+only transitions of $\minf(\pi)$. Let $p$ be the first state of
+$\pi'$. As each transition of $\pi'$ appears infinitely many times in
+$\pi$ and thus also in $\pi'$, there has to be a finite prefix $\rho$
+of $\pi'$ such that $\rho$ is a path from $p$ to $p$ that contains all
+transitions of $\minf(\pi)$. In other words, the set $\minf(\pi)$ of a
+run $\pi$ is always a cycle. Moreover, as we assume that 
+%\footnote{We can also prove that for each cycle $T$ there exists a run $\pi$ such that $T=\minf(\pi)$.}
+
+\fixme{ODTUD DAL JE TO PUVODNI, ale taky se to trochu bude menit!}
+
 Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and vice

### 2023-06-01 statistics/sat23-camera-ready/patterns/translate.py
Experiments ben

@@ -2,7 +2,7 @@ import spot
 import os
 import re
 from ltlcross_runner import LtlcrossRunner, ResultsAnalyzer
-
+import sys
 
 def main(argv):
     tools = {
@@ -14,11 +14,11 @@ def main(argv):
     'rabinizer4': '/home/xschwar3/owl/bin/owl ltl2dgra -f %f > %O',
     }
 
-    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+    tools_order = [ 'ltl2tgba_det']
 
-    pattern_ltl = 'pattern.ltl'
-    pattern_aut = 'pattern.csv'
-    lcr = LtlcrossRunner(tools, formulas = [pattern_ltl], res_filename = pattern_aut)
+    rand_ltl = 'patterns.ltl'
+    rand_aut = 'patterns.csv'
+    lcr = LtlcrossRunner(tools, formulas = [rand_ltl], res_filename = rand_aut)
     lcr.run_ltlcross(automata = True, timeout = '60')
     ra = ResultsAnalyzer(rand_aut)
     ra.parse_results()
@@ -29,7 +29,7 @@ def main(argv):
 
         ltlf_translated[tool] = len(fs[fs[('acc', tool)] > 1])
 
-        fd = open('pattern-' + tool + '.hoa', 'w')
+        fd = open('random-' + tool + '.hoa', 'w')
         for (ix, _), _ in fs.iterrows():
             fd.write(ra.aut_for_id(ix, tool).to_str() + '\n')
         fd.close()

### 2023-06-01 telatko.py

@@ -109,9 +109,9 @@ def main(argv):
             else:
                 print_aut(auto, None, " ")
 
-            if not spot.are_equivalent(origin, auto):
-                print("NOT EQUIVALENT!")
-                assert (False)
+           # if not spot.are_equivalent(origin, auto):
+           #    print("NOT EQUIVALENT!")
+           #     assert (False)
 
         # except BaseException as err:
             # print(f"Unexpected {err=}, {type(err)=}")

### 2023-06-01 qbf/solver_manipulation/solver_functions.py
else branch in old rec formula

@@ -69,7 +69,8 @@ def query(formula, formula_attr):
         qcir_c = Qcir_ctor(formula)
         qcir_c.to_qcir("./qbf/solver_manipulation/formulas_files/formula.qcir")
         return Solver_result.unsat, None
-    elif formula_attr.solver == "caqe":
+        
+    elif formula_attr.solver == "booleguru + caqe":
         safe_file(formula_file_path, str(formula), "w")
         formula_file_path = encode_qdimacs(formula_file_path)
         tool_path = "./../solvers/caqe/target/release/caqe"

### 2023-06-01 qbf/z3_formula.py

@@ -60,8 +60,10 @@ class Z3_f_ctor:
         laso = simplify(laso)
 
         if old is not None:
-            old = simplify(old)
+
+
             eq = self.equivalence_f(old, new)
+            simplify(eq)
         else:
             eq = new
         impl = Implies(laso, eq)
@@ -158,6 +160,8 @@ class Z3_f_ctor:
             Returns encoded literal.
         """
 
+
+
         inf, fin = acc.used_inf_fin_sets()
         used_set = acc.used_sets()
 
@@ -167,6 +171,7 @@ class Z3_f_ctor:
 
         edges_vars = []
 
+
         # spot anomaly, cant extract number of mark the standard way
         # mark_num = -1
         for m in used_set.sets():
@@ -175,8 +180,11 @@ class Z3_f_ctor:
 
         if mark_num in self.edge_dict:
 
+
             for e in list(set(self.edge_dict[mark_num]) & set(
                     self.inner_edges_nums)):
+
+
                 var = Bool("e_" + str(e))
                 edges_vars.append(var)
 
@@ -192,9 +200,19 @@ class Z3_f_ctor:
             if fin:
                 literal_formula = Not(literal_formula)
 
-        return literal_formula
+
+            return literal_formula
+        if inf:
+            #  Inf(∅) = False
+            return False
+        else:
+            #  Fin(∅) = True
+            return True
+
 
     def rec_old_original_shape(self, acc, prev_type, level):
+        
+
 
         if acc.top_disjuncts() == acc.top_conjuncts():
             return self.encode_acc_mark(acc)

### 2023-06-01 telatko.py

@@ -30,9 +30,7 @@ def parser(argv):
     parser.add_argument(
         "-T",
         "--timeout",
-        help="Kill QBF solver after inserted number of seconds.",
-        type=int,
-        default=50)
+        help="Kill QBF solver after inserted number of seconds.", type=int, default=50)
     parser.add_argument(
         "-S",
         "--scc",
@@ -49,15 +47,18 @@ def parser(argv):
     parser.add_argument(
         "-G",
         "--gradual",
-        help="Levels will gradually grow from L1 up to specified level.",
-        action='store_true')
+
+        help="Levels will gradually grow from L1 up to specified level.", action='store_true')
 
     parser.add_argument(
         "-M",
         "--solver",
         help="Choose solver",
-        choices=['limboole', 'z3', 'quabs', 'caqe'],
+        choices=['limboole', 'z3', 'quabs', 'caqe', 'booleguru + caqe'],
         default='z3')
+
+
+
     args = parser.parse_args()
     return args
 
@@ -69,16 +70,13 @@ def main(argv):
         print("No automata to process.", file=sys.stderr)
 
     # if no arguments given, default variant is gradually go up to level 3
+
     if args.level == 0 and not args.gradual and not args.incremental and not args.scc and args.solver == "z3":
+
         args.gradual = True
         args.level = 3
         args.incremental = True
         args.scc = True
-
-    elif args.incremental and args.solver == "limboole":
-        print("Incremental solving is possible only with Z3 solver.")
-        return
-
     elif args.level == 0:
         args.gradual = True
         args.level = 3
@@ -98,9 +96,7 @@ def main(argv):
             formula_attributes.set_C(a)
 
             if a.get_acceptance().used_sets().count() == 0:
-
                 auto = a
-                auto.set_name("F_0_S")
             else:
                 auto = play(a, formula_attributes)
 
@@ -109,9 +105,9 @@ def main(argv):
             else:
                 print_aut(auto, None, " ")
 
-            if not spot.are_equivalent(origin, auto):
-                print("NOT EQUIVALENT!")
-                assert (False)
+            #if not spot.are_equivalent(origin, auto):
+            #    print("NOT EQUIVALENT!")
+            #    assert (False)
 
         # except BaseException as err:
             # print(f"Unexpected {err=}, {type(err)=}")

### 2023-06-01 paper/main.tex

@@
 $\rem(t_0)\rem(t_1)\rem(t_2)\ldots$ of $\mB$ is accepting. This
-requirement clearly guarantees the equiavalence of $\mA$ and $\mB$. We
-note that preserving the acceptance of individual runs is not a
-necessary condition for language equivalence of $\mA$ and $\mB$ and
-there exist automata where relaxing this requirement can lead to
-a bigger reduction of acceptance marks (see
-Figure~\ref{fig:counter}). However, looking for $\psi$ and $\rem$
-preserving the acceptance of individual runs makes the problem easier
-as we can, for example, ignore the labelling of transitions by the
-elements of $\Sigma$.
+requirement clearly guarantees the equiavalence of $\mA$ and $\mB$,
+but it is not a necessary condition for the equivalence. Indeed, there
+exist automata where relaxing this requirement can lead to a bigger
+reduction of acceptance marks (see Figure~\ref{fig:counter}). However,
+looking for $\psi$ and $\rem$ that preserve the acceptance of individual
+runs makes the problem easier as we can, for example, ignore the
+labelling of transitions by the elements of $\Sigma$.
 \begin{figure}[t]
@@
 
-  \caption{Both automata accept the same language of words
-    $w\in\{a,b\}^\omega$ that contain infinitely many occurences of
-    both $a$ and $b$. Each accepting run of the left automaton has to
-    contain infinitely many occurrences of both transitions looping on
-    the initial state. Hence, there does not exist any automaton with
-    the same accepting runs as the left automaton and less than two
-    accetpance marks.}
+  \caption{The left automaton accepts the words % $w\in\{a,b\}^\omega$
+    that contain infinitely many occurences of both $a$ and $b$. Each
+    accepting run of the left automaton has to contain infinitely many
+    occurrences of both transitions looping on the initial
+    state. Hence, there does not exist any automaton with the same
+    accepting runs as the left automaton and less than two accetpance
+    marks. The right automaton accepts the same language using one
+    acceptance mark and a different set of accepting runs.}
   \label{fig:counter}
@@
 
-\fixme{TADY TO HONZA JESTE UPRAVUJE}
-We start with several simple observations. Acceptance of a run $\pi$
-is fully determined by the set $\minf(\pi)$.
-% of acceptance marks that
-% appear infinitely often on $\pi$, which is in turn fully determined by
-% the set $\minf(\pi)$ of transitions appearing infinitely often in
-% $\pi$.
+Our reduction method is based on two facts. First, the acceptance of a
+run $\pi$ is fully determined by $\minf(\pi)$. Second, each set
+$\minf(\pi)$ is a cycle and vice versa.
+
+% Before we We start with a reformulation of our requirement on $\psi$
+% and $\rem$.  Recall that a run $\pi$ of $\mA$ is accepting if and
+% only iff $\minf(\pi)$ satisfies $\varphi$.  Further, the following
+% lemma says that each set $\minf(\pi)$ is a cycle and vice versa.
+
 \begin{lemma}
   A set $T\subseteq\delta$ is a cycle if and only if
-  there is a run $\pi$ such that $T=\minf(\pi)$.
+  there is a run $\pi$ such that $\minf(\pi)=T$.
 \end{lemma}
-
-As $\minf(\pi)$ is the set of transitions that appear infinitely many
-times in $\pi$, there has to be a suffix $\pi'$ of $\pi$ containing
-only transitions of $\minf(\pi)$. Let $p$ be the first state of
-$\pi'$. As each transition of $\pi'$ appears infinitely many times in
-$\pi$ and thus also in $\pi'$, there has to be a finite prefix $\rho$
-of $\pi'$ such that $\rho$ is a path from $p$ to $p$ that contains all
-transitions of $\minf(\pi)$. In other words, the set $\minf(\pi)$ of a
-run $\pi$ is always a cycle. Moreover, as we assume that 
-%\footnote{We can also prove that for each cycle $T$ there exists a run $\pi$ such that $T=\minf(\pi)$.}
-
-\fixme{ODTUD DAL JE TO PUVODNI, ale taky se to trochu bude menit!}
-
-Further, each set $\minf(\pi)$ for a run $\pi$ is a cycle and vice
-versa. Hence, our reduction method is looking for a new acceptance
-formula $\psi$ and function $\rem$ such that for each cycle
-$T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and only
-if $\rem(T)$ satisfies $\psi$. This can be roughly denoted by the
+\begin{proof}
+  To prove the direction `$\Longrightarrow$', we assume that $T$ is a
+  cycle. The definition says that there exists a path $\tau$ from a
+  state $p$ to the same state containing each transition of $T$ at
+  least once and no transition outside $T$. As our automata contain
+  only reachable states, there exists a path $\rho$ from the initial
+  state $q_I$ to $p$ or $p=q_I$ and we set $\rho=\varepsilon$. The
+  infinite sequence $\pi=\rho.\tau^\omega$ is a run satisfying
+  $\minf(\pi)=T$.
+
+  To prove the direction `$\Longleftarrow$', we consider a run
+  $\pi$.  As $\minf(\pi)$ is the set of transitions that appear
+  infinitely many times in $\pi$, there has to be a suffix $\pi'$ of
+  $\pi$ containing only transitions of $\minf(\pi)$. Let $p$ be the
+  first state of $\pi'$. As each transition of $\pi'$ appears
+  infinitely many times in $\pi$ and thus also in $\pi'$, there has to
+  be a finite prefix $\rho$ of $\pi'$ such that $\rho$ is a path from
+  $p$ to $p$ that contains all transitions of $\minf(\pi)$. In other
+  words, the set $\minf(\pi)$ is a cycle.
+\end{proof}
+
+Hence, our goal can be reformulated as follows. We look for a new
+acceptance formula $\psi$ and a function $\rem$ such that for each
+cycle $T\subseteq\delta$, it holds that $T$ satisfies $\varphi$ if and
+only if $\rem(T)$ satisfies $\psi$. This can be roughly denoted by the
 formula

### 2023-06-01 statistics/sat23-camera-ready/current-z3-version/patterns/delag/experiments.py
Prepared statistics generation.

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - patterns")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = './'
+    tools_order = [ 'delag']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-06-01 statistics/sat23-camera-ready/current-z3-version/patterns/ltl2tgba/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - patterns")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = './'
+    tools_order = [ 'ltl2tgba']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-06-01 statistics/sat23-camera-ready/current-z3-version/patterns/ltl2tgba_det/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - patterns ltl2tgba_det")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = './'
+    tools_order = [ 'ltl2tgba_det']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-06-01 statistics/sat23-camera-ready/current-z3-version/patterns/ltl3tela/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - patterns ltl3tela")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = './'
+    tools_order = [  'ltl3tela']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-06-01 statistics/sat23-camera-ready/current-z3-version/patterns/ltl3tela_det/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - patterns ltl3tela_det")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = './'
+    tools_order = [ 'ltl3tela_det']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-06-01 statistics/sat23-camera-ready/current-z3-version/patterns/rabinizer4/experiments.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python3
+
+import spot
+import os
+import sys
+
+def main(argv):
+    print("Current telatko z3 version - patterns rabinizer4")
+    telatko = "/home/xschwar3/telatko/telatko.py"
+
+    os.environ['SPOT_TMPDIR'] = './autcross'
+    os.environ['SPOT_TMPKEEP'] = '1'
+    patterns_path = './'
+    tools_order = [  'rabinizer4']
+    for tool in tools_order:
+        print(tool)
+        os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/patterns-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")
+
+
+if __name__ == "__main__":
+    main(sys.argv[1:])

### 2023-06-01 telatko.py

@@ -105,10 +105,11 @@ def main(argv):
             else:
                 print_aut(auto, None, " ")
 
-
-            #if not spot.are_equivalent(origin, auto):
-            #    print("NOT EQUIVALENT!")
-            #    assert (False)
+            """
+            if not spot.are_equivalent(origin, auto):
+                print("NOT EQUIVALENT!")
+                assert (False)
+            """
 
 
         # except BaseException as err:

### 2023-06-02 paper/main.tex
first part of section 6 updated

@@
 \author{Tereza Schwarzov\'{a}}{%Faculty of Informatics,
-  Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{}{}
+  Masaryk University, Brno, Czech Republic}{xschwar3@mail.muni.cz}{https://orcid.org/0009-0001-9016-5399}{}
 \author{Jan Strej\v{c}ek}{%Faculty of Informatics,
@@
   technique reduced some acceptance marks in automata produced by all
-  the tools. On automata with more than one acceptance mark produced
+  the tools. \fixme{On automata with more than one acceptance mark produced
   by tools Delag and Rabinizer~4 of the Owl library, our technique reduced
-  29.1\% and 38.5\% of acceptance marks, respectively.
+  29.1\% and 38.5\% of acceptance marks, respectively.}
   % The achieved
@@
 Rabinizer~4~\cite{kretinsky.18.cav} (both included in the Owl
-library), Spot~\cite{duret.16.atva2} and
+library), Spot~\cite{duret.16.atva2}, and
 \texttt{ltl3tela}~\cite{major.19.atva}.  While the reduction is
@@
 
-To evaluate our reduction technique, we considered all sets of LTL
-formulas from literature as they are provided by the tool
-\texttt{genltl} of the Spot library~\cite{duret.16.atva2}~2.10.4, where
-parametrized formulas were generated for all combinations of
-parameter values from 1 to 4. We used the tool \texttt{ltlfilt} of the
-Spot library to simplify the formulas and remove
-duplicates. Altogether, we have generated 348 unique LTL
-formulas. These have been translated to deterministic TELA by two
+To evaluate our reduction technique, we applied \telatko to automata
+produced by the following process. We started with two sets of LTL formulas.
+\begin{itemize}
+\item One set contains all LTL formulas from \emph{literature} that
+  are provided by the tool \texttt{genltl} of the Spot
+  library~\cite{duret.16.atva2}~2.10.4. For parametrized formula
+  patterns, we consider instances for all combinations of parameter
+  values from 1 to 4.
+\item The second set consists of 400 random LTL formulas with 4 atomic
+  propositions. These formulas were generated by the tool
+  \texttt{randltl} of the Spot library.
+\end{itemize}
+On both these sets, we applied the tool \texttt{ltlfilt} of the Spot
+library to simplify the formulas and remove duplicates and formulas
+equivalent to $\true$ and $\false$. After these steps, we had 348
+LTL formulas from literature and 366 random formulas.
+%
+Formulas from both sets have been translated to nondeterministic TELA
+by two state-of-the-art translators, namely \texttt{ltl2tgba} (used
+with option \texttt{-G}) from the Spot library~\cite{duret.16.atva2}
+and \texttt{ltl3tela}~\cite{major.19.atva}, and to deterministic TELA
+by \texttt{ltl3tela} with option \texttt{-D1} and by another two
 state-of-the-art translators from the Owl
@@
 Delag)~\cite{muller.17.gandalf} and \texttt{ltl2dgra} (known as
-Rabinizer~4)~\cite{kretinsky.18.cav}, and to nondeterministic TELA by
-another two state-of-the-art translators, namely \texttt{ltl2tgba}
-from the Spot library~\cite{duret.16.atva2} and
-\texttt{ltl3tela}~\cite{major.19.atva}. Three of the considered
-translators failed on some formulas: it usually reached a timeout of
-60 seconds or it produced an automaton that cannot be parsed by the
-Spot library.
+Rabinizer~4)~\cite{kretinsky.18.cav}. Some translators failed on some
+formulas: it usually reached a timeout of 60 seconds or it produced an
+automaton that cannot be parsed by the Spot library.
 % produced by the individual tools contains less than 348 automata as not
@@
 % seconds.
-Further, we have removed automata with 0 or 1 acceptance marks as
+Further, we have removed automata with 0 or 1 acceptance mark as
 there is a little point in reducing these. Table
-\ref{tab:ltltranslators} shows the exact versions and homepages of
-used translators. For each translator, it also provides the number of
-fails, the number of produced automata with less than two marks, and
-the number of automata with at least two marks. The last number is
-typeset in bold as these automata are actually used for the
-experimental evaluation of our reduction technique.
+\ref{tab:ltltranslators} shows the exact versions of the
+translators. For each set of formulas and for each translator, it also
+provides the number of fails, the number of produced automata with
+less than two marks, and the number of automata with at least two
+marks. The numbers of automata with at least two marks are typeset in
+bold as these automata are actually used for the experimental
+evaluation of our reduction technique.
 
 \begin{table}[t]
-  \caption{LTL to automata translators used for the experimental
-    evaluation and the numbers of fails and successfully constructed
-    automata with less than two marks and with at least two marks.}
+  \caption{Considered LTL to automata translators and the numbers of
+    fails and successfully constructed automata with at most one mark and with at least two marks for each translator and set of
+    formulas.}
 \label{tab:ltltranslators}
@@
 %
-\begin{tabular}{lrrrr}
+% \begin{tabular}{lrrrr}
+%   \toprule
+%   &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
+%   tool &\!\!\!version/homepage & ~~~~fails & less than 2 marks & \textbf{at least 2 marks}\\
+%   \midrule
+%   \texttt{ltl2dela} & 21.0\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
+%        & 5 & 214 & \textbf{129} \\
+%   \texttt{ltl2dgra} & 21.0\footnotemark[1]
+%        & 12 & 102 & \textbf{234}\\
+%   \texttt{ltl2tgba} & 2.10.4\tablefootnote{\url{https://spot.lrde.epita.fr}}
+%        & 0 & 278 & \textbf{70}\\
+%   \texttt{ltl3tela} & 2.2.0\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
+%        & 18 & 239 & \textbf{91}\\
+%   \bottomrule
+% \end{tabular}
+%
+
+\setlength{\tabcolsep}{5.7pt}
+\begin{tabular}{lrrrrrrr}
   \toprule
-  &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
-  tool &\!\!\!version/homepage & ~~~~fails & less than 2 marks & \textbf{at least 2 marks}\\
+  && \multicolumn{3}{c}{\!348 formulas from literature\!} & \multicolumn{3}{c}{366 random formulas}\\
+  \cmidrule(lr){3-5} \cmidrule(lr){6-8}
+  &&& \multicolumn{2}{c}{automata with} && \multicolumn{2}{c}{automata with}\\
+  \cmidrule(lr){4-5} \cmidrule(lr){7-8}
+  &&& at most & \textbf{at least} && at most & \textbf{at least}\\[-.8ex]
+  \multicolumn{2}{l}{translator ~~~ (version)$^{\text{web}}$} & ~~~~fails & 1 mark & \!\!\!\!\textbf{2 marks} & ~~~~fails & 1 mark & \!\!\!\!\textbf{2 marks}\\
   \midrule 
-  \texttt{ltl2dela} & 21.0\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
-       & 5 & 214 & \textbf{129} \\
-  \texttt{ltl2dgra} & 21.0\footnotemark[1]
-       & 12 & 102 & \textbf{234}\\
-  \texttt{ltl2tgba} & 2.10.4\tablefootnote{\url{https://spot.lrde.epita.fr}}
-       & 0 & 278 & \textbf{70}\\
-  \texttt{ltl3tela} & 2.2.0\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
-       & 18 & 239 & \textbf{91}\\
+  \texttt{ltl2tgba -G} & (2.10.4)\tablefootnote{\url{https://spot.lrde.epita.fr}}
+  & 0 & 278 & \textbf{70} & ? & ? & \textbf{?}\\
+  \texttt{ltl3tela} & (2.2.0)\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
+  & 18 & 239 & \textbf{91} & ? & ? & \textbf{?}\\
+  \texttt{ltl3tela -D1} & (2.2.0)\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
+  & ? & ? & \textbf{?} & ? & ? & \textbf{?}\\ 
+  \texttt{ltl2dela} & (21.0)\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
+       & 5 & 214 & \textbf{129} & ? & ? & \textbf{?} \\
+  \texttt{ltl2dgra} & (21.0)\footnotemark[4]
+       & 12 & 102 & \textbf{234} & ? & ? & \textbf{?}\\
   \bottomrule
@@
 
+\fixme{Honza je tu.}
 To each automata set (that is, the automata with at least two marks

### 2023-06-02 paper/main.tex
polishing in Table 1

@@
 \begin{table}[t]
-  \caption{Considered LTL to automata translators and the numbers of
-    fails and successfully constructed automata with at most one mark and with at least two marks for each translator and set of
+  \caption{Considered translators and the numbers of \emph{fails} and
+    successfully constructed automata with \emph{at most 1 mark} and
+    with \emph{at least 2 marks} for each translator and set of
     formulas.}
@@
   & 18 & 239 & \textbf{91} & ? & ? & \textbf{?}\\
-  \texttt{ltl3tela -D1} & (2.2.0)\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
+  \texttt{ltl3tela -D1} & (2.2.0)\footnotemark[2]
   & ? & ? & \textbf{?} & ? & ? & \textbf{?}\\ 
@@
        & 5 & 214 & \textbf{129} & ? & ? & \textbf{?} \\
-  \texttt{ltl2dgra} & (21.0)\footnotemark[4]
+  \texttt{ltl2dgra} & (21.0)\footnotemark[3]
        & 12 & 102 & \textbf{234} & ? & ? & \textbf{?}\\

### 2023-06-02 qbf/formula_preparation.py
Adjusted L2

@@ -162,6 +162,7 @@ def scc_unoptimized_formula(
             list(map(lambda e: aut.edge_number(e), scc_edg_list)))
     f_creator = None
     if solver == "z3":
+        print("z3 formula")
         f_creator = Z3_f_ctor(
             edge_dict,
             scc_edg,
@@ -230,6 +231,7 @@ def scc_optimized_formula(aut, scc_state_info, C,
 
 
 def run_with_limited_time(aut1, aut2, time):
+    
     safe_file("a1", aut1.to_str(), "w")
     safe_file("a2", aut2.to_str(), "w")
 

### 2023-06-02 qbf/incremental.py

@@ -114,14 +114,14 @@ def inc_loop(
         C += 1
 
     while C > 0 and K > 0:
-        
 
         if aut.get_acceptance().used_sets().count(
         ) < 1 or aut.prop_state_acc() == spot.trival.yes_value:
             return aut
-
+        
         solver.check()
         res = solver.check()
+
 
         result, _, _ = z3_result(res)
 

### 2023-06-02 qbf/limboole_formula.py

@@ -244,6 +244,7 @@ class Boole_f_ctor:
         return SATformula("&", eq_list)
 
     def disjunct_formula_list(self, scc_edges_nums):
+        ptint(f" scc edges nums {scc_edges_nums}")
         alist = []
         for i in scc_edges_nums:
             dis = self.disjunct_formula(i)
@@ -271,10 +272,10 @@ class Boole_f_ctor:
 
         # edges that are true are in one SCC and none of edges from other SCCs is
         # true
-        # one_scc = self.one_scc_f()
-        least_one_edge = self.disjunct_formula(self.inner_edges_nums)
+        one_scc = self.one_scc_f()
+        #least_one_edge = self.disjunct_formula(self.inner_edges_nums)
 
-        laso = SATformula("&", [in_out, least_one_edge])
+        laso = SATformula("&", [in_out, one_scc])
         return laso
 
     def laso_cycles(self, aut):

### 2023-06-02 qbf/z3_formula.py

@@ -211,7 +211,7 @@ class Z3_f_ctor:
 
 
     def rec_old_original_shape(self, acc, prev_type, level):
-        
+
 
 
         if acc.top_disjuncts() == acc.top_conjuncts():
@@ -309,14 +309,16 @@ class Z3_f_ctor:
 
     def laso_f(self, aut):
 
+
         in_out = self.in_n_out()
 
         # edges that are true are in one SCC and none of edges from other SCCs is
         # true
-        # one_scc = self.one_scc_f()
-        least_one_edge = self.disjunct_formula(self.inner_edges_nums)
+        one_scc = self.one_scc_f()
+        #least_one_edge = self.disjunct_formula(self.inner_edges_nums)
+        
 
-        laso = And(in_out, least_one_edge)
+        laso = And(in_out, one_scc)
         return laso
 
     def laso_cycles(self, aut):

### 2023-06-02 statistics/sat23-camera-ready/sat22-z3-version/random/experiments.py

@@ -11,7 +11,7 @@ def main(argv):
     os.environ['SPOT_TMPDIR'] = './autcross'
     os.environ['SPOT_TMPKEEP'] = '1'
     patterns_path = '/home/xschwar3/telatko/statistics/sat23-camera-ready/random'
-    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det', 'ltl2tgba_det']
+    tools_order = [ 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl3tela_det']
     for tool in tools_order:
         print(tool)
         os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/random-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")

### 2023-06-02 telatko.py

@@ -114,6 +114,7 @@ def main(argv):
 
 
 
+
         # except BaseException as err:
             # print(f"Unexpected {err=}, {type(err)=}")
 

### 2023-06-02 paper/main.tex
suggestion of a new layout of Table 2

@@
 %
-
 \setlength{\tabcolsep}{5.7pt}
@@
 
-\fixme{Honza je tu.}
 To each automata set (that is, the automata with at least two marks
-produced by an individual LTL translator), we have applied all
-single-level reductions and the multi-level reduction, always with
-incremental approach and without reducing the number of cubes. We
-do not reduce the number of cubes as our primary aim is to reduce the
-number of acceptance marks.
+produced by an individual LTL translator for one of the two formula
+sets), we have applied all single-level reductions and the multi-level
+reduction, always with incremental approach and without reducing the
+number of cubes. We do not reduce the number of cubes as our primary
+aim is to reduce the number of acceptance marks.
 %\todo{Vysvetlit, proc neredukujeme pocet cubes?}
@@
 
-Table~\ref{tab:cumm-data} shows the cumulative numbers of marks in the
-input automata sets and after each reduction, together with the
-reduction ratio and total time spent by the considered reduction. The
-column \textit{solver timeout} shows the number of automata for which
-the last query to QBF solver did not finish within the 30~seconds
-limit. The timeout of the last QBF query means that the automaton may
-be potentially further reduced if a longer time limit is used. One can
-observe that a higher level sometimes achieves a smaller reduction
-than a lower level (e.g., compare level~1 and level~2 for
-\texttt{ltl2dgra}). This is caused by the QBF solver timeouts occuring
-earlier as formulas constructed by the higher level are more
-complex. For \texttt{ltl2dela} and \texttt{ltl2dgra}, the automata
-sets do not contain any automaton where level~2 or level~3 achieves a
-better result than level~1. However, all levels contribute to the
-reductions in the multi-level setting.
+For each automata set identified by the translator and the set of
+formulas, Table~\ref{tab:cumm-data} shows the cumulative numbers of
+marks in the input automata sets and after each reduction, together
+with the reduction ratio and total time spent by the considered
+reduction. The column \textit{solver timeout} shows the number of
+automata for which the last query to QBF solver did not finish within
+the 30~seconds limit. The timeout of the last QBF query means that the
+automaton may be potentially further reduced if a longer time limit is
+used. \fixme{Zkontrolovat zbytek odstavce podle aktualnich vysledku,
+  specifikovat formula set.} One can observe that a higher level
+sometimes achieves a smaller reduction than a lower level (e.g.,
+compare level~1 and level~2 for \texttt{ltl2dgra}). This is caused by
+the QBF solver timeouts occuring earlier as formulas constructed by
+the higher level are more complex. For \texttt{ltl2dela} and
+\texttt{ltl2dgra}, the automata sets do not contain any automaton
+where level~2 or level~3 achieves a better result than
+level~1. However, all levels contribute to the reductions in the
+multi-level setting.
 
@@
   \caption{The cumulative numbers of acceptance marks before and after
-    reduction for various reductions and automata sets. The column
-    \emph{reduction} shows the cumulative percentage of saved
-    acceptance marks and \emph{time} reports the cumulative reduction
-    time. The column \textit{solver timeout} indicates the number of
-    instances where the last call to the QBF solver timed out.  }
+    reduction for various reduction levels and automata sets given by
+    the translator and formula set. The column \emph{reduction} shows
+    the cumulative percentage of saved acceptance marks and
+    \emph{time} reports the cumulative reduction time. The column
+    \textit{solver timeout} indicates the number of instances where
+    the last call to the QBF solver timed out.  }
 \label{tab:cumm-data}
 \centering
-\setlength{\tabcolsep}{10pt}
-\input{table-data.tex}
+\setlength{\tabcolsep}{3.9pt}
+\newcommand{\mybox}[1]{\fbox{\begin{minipage}{4.7cm}\centering#1\end{minipage}}}
+\begin{tabular}{lcrrrrrrrr}
+  \toprule
+  && \multicolumn{4}{c}{\makecell{reduction of automata from\\[-.8ex] formulas from literature}}
+  & \multicolumn{4}{c}{\makecell{reduction of automata from\\[-.8ex] random formulas}}
+  \\ \cmidrule(lr){3-6} \cmidrule(lr){7-10}
+  translator & \makecell{reduction\\[-1ex] level}
+  & marks & \makecell{reduction\\[-1ex] [\%]} & \makecell{time\\[-1ex] [s]} & \makecell{solver\\[-1ex] timeout} 
+  & marks & \makecell{reduction\\[-1ex] [\%]} & \makecell{time\\[-1ex] [s]} & \makecell{solver\\[-1ex] timeout} \\
+  \midrule
+%  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
+  \multicolumn{2}{l}{\texttt{ltl2tgba -G}} 
+  & \multicolumn{4}{c}{\mybox{198 marks in 70 automata}}
+  & \multicolumn{4}{c}{\mybox{XYZ marks in XYZ automata}}\\
+  & 1
+  & 198 & 0.0~~~ & 49.3 & 0~~~
+  &   ? &   ?~~~ &    ? & ?~~~ \\                     
+  & 2
+  & 198 & 0.0~~~ & 78.9 & 0~~~
+  &   ? &   ?~~~ &    ? & ?~~~ \\                     
+  & 3
+  & 189 & 4.5~~~ & 1222.8 & 7~~~
+  &   ? &   ?~~~ &    ? & ?~~~ \\                     
+  & multi
+  & 189 & 4.5~~~ & 1259.1 & 6~~~
+  &   ? &   ?~~~ &    ? & ?~~~ \\                    
+  \midrule
+%  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
+  \multicolumn{2}{l}{\texttt{ltl3tela}} 
+  & \multicolumn{4}{c}{\mybox{348 marks in 91 automata}}
+  & \multicolumn{4}{c}{\mybox{XYZ marks in XYZ automata}}\\
+  & 1
+  & 332 & 4.6~~~ & 465.9 & 12~~~
+  &   ? &   ?~~~ &     ? &  ?~~~ \\                     
+  & 2
+  & 334 & 4.0~~~ & 553.9 & 15~~~
+  &   ? &   ?~~~ &     ? &  ?~~~ \\                     
+  & 3
+  & 326 & 6.3~~~ & 691.8 & 18~~~
+  &   ? &   ?~~~ &     ? &  ?~~~ \\                     
+  & multi
+  & 323 & 7.2~~~ & 1545.9 & 18~~~
+  &   ? &   ?~~~ &    ? & ?~~~ \\                    
+  \midrule
+  &&&& \vdots \\
+% \\
+% level 1 &
+% 387~ & 26.0~~~~~~ & 755.4~ &  18~~~
+% \\
+% &
+% level 2 &
+% 390~ & 25.4~~~~~~ & 1024.5~ &  20~~~
+% \\
+% &
+% level 3 &
+% 393~ & 24.9~~~~~~ & 7273.3~ &  25~~~
+% \\
+% &
+% multi &
+% 371~ & 29.1~~~~~~ & 9111.4~ &  21~~~
+% \\
+% \midrule
+% \multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\234 automata\\882 acc.~marks}}
+% &
+% level 1 &
+% 552~ & 37.4~~~~~~ & 738.7~ & 14~~~
+% \\
+% &
+% level 2 &
+% 569~ & 35.5~~~~~~ & 907.3~ &  18~~~
+% \\
+% &
+% level 3 &
+% 561~ & 36.4~~~~~~ & 1163.4~ &  21~~~
+% \\
+% &
+% multi &
+% 542~ & 38.5~~~~~~ & 2320.7~ &  21~~~
+% \\
+% \midrule
+% \multirow{4}{*}{\makecell{\texttt{ltl3tela}\\91 automata\\348 acc.~marks}}
+% &
+% level 1 &
+% 332~ & 4.6~~~~~~ & 465.9~ &  12~~~
+% \\
+% &
+% level 2 &
+% 334~ & 4.0~~~~~~ & 553.9~ &  15~~~
+% \\
+% &
+% level 3 &
+% 326~ & 6.3~~~~~~ & 691.8~ &  18~~~
+% \\
+% &
+% multi &
+% 323~ & 7.2~~~~~~ & 1545.9~ & 18~~~
+% \\
+\bottomrule                                                                     
+\end{tabular}
 \end{table}
 
+\fixme{Upravit zbytek textu.}
 Table~\ref{tab:heatmap} shows the number of automata from all

### 2023-06-03 qbf/formula_preparation.py
Experiments ben with fixed L2

@@ -162,7 +162,7 @@ def scc_unoptimized_formula(
             list(map(lambda e: aut.edge_number(e), scc_edg_list)))
     f_creator = None
     if solver == "z3":
-        print("z3 formula")
+        
         f_creator = Z3_f_ctor(
             edge_dict,
             scc_edg,

### 2023-06-03 qbf/limboole_formula.py

@@ -244,7 +244,7 @@ class Boole_f_ctor:
         return SATformula("&", eq_list)
 
     def disjunct_formula_list(self, scc_edges_nums):
-        ptint(f" scc edges nums {scc_edges_nums}")
+        
         alist = []
         for i in scc_edges_nums:
             dis = self.disjunct_formula(i)

### 2023-06-03 statistics/sat23-camera-ready/current-z3-version/patterns/ltl2tgba_det/experiments.py

@@ -5,7 +5,7 @@ import os
 import sys
 
 def main(argv):
-    print("Current telatko z3 version - patterns ltl2tgba_det)
+    print("Current telatko z3 version - patterns ltl2tgba_det")
     telatko = "/home/xschwar3/telatko/telatko.py"
 
     os.environ['SPOT_TMPDIR'] = './autcross'

### 2023-06-03 statistics/sat23-camera-ready/current-z3-version/random/experiments.py

@@ -11,7 +11,7 @@ def main(argv):
     os.environ['SPOT_TMPDIR'] = './autcross'
     os.environ['SPOT_TMPKEEP'] = '1'
     patterns_path = '/home/xschwar3/telatko/statistics/sat23-camera-ready/random'
-    tools_order = ['ltl3tela_det', 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela', 'ltl2tgba_det']
+    tools_order = ['ltl3tela_det', 'delag', 'rabinizer4', 'ltl2tgba', 'ltl3tela']
     for tool in tools_order:
         print(tool)
         os.system(f"autcross --no-checks --csv=patterns-telatko-{tool}.csv -F {patterns_path}/random-{tool}.hoa 'autfilt %H > %O' '{telatko} -F %H -T 30 -L 1 -I -O %O' '{telatko} -F %H -T 30 -L 2 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -O %O' '{telatko} -F %H -T 30 -L 3 -I -G -O %O' > autcross/{tool}.log 2>&1")

### 2023-06-03 paper/main.tex

@@
 \author{Juraj Major}{%Faculty of Informatics,
-  Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{}{}
+  Masaryk University, Brno, Czech Republic}{major@fi.muni.cz}{https://orcid.org/0009-0007-1871-9047}{}
 \authorrunning{T. Schwarzov\'{a}, J. Strej\v{c}ek, and J. Major}
@@
 $\rem(t_0)\rem(t_1)\rem(t_2)\ldots$ of $\mB$ is accepting. This
-requirement clearly guarantees the equiavalence of $\mA$ and $\mB$,
+requirement clearly guarantees the equivalence of $\mA$ and $\mB$,
 but it is not a necessary condition for the equivalence. Indeed, there
@@
     state. Hence, there does not exist any automaton with the same
-    accepting runs as the left automaton and less than two accetpance
+    accepting runs as the left automaton and less than two acceptance
     marks. The right automaton accepts the same language using one

### 2023-06-03 paper/main.tex
polishing Tables 1 and 2 with Terka

@@
   \cmidrule(lr){4-5} \cmidrule(lr){7-8}
-  &&& at most & \textbf{at least} && at most & \textbf{at least}\\[-.8ex]
-  \multicolumn{2}{l}{translator ~~~ (version)$^{\text{web}}$} & ~~~~fails & 1 mark & \!\!\!\!\textbf{2 marks} & ~~~~fails & 1 mark & \!\!\!\!\textbf{2 marks}\\
+  \multicolumn{2}{l}{translator ~~~ (version)$^{\text{web}}$}
+  & ~~~~fails & \makecell[r]{at most\\[-1ex] 1 mark} & \makecell[r]{\textbf{at least}\\[-1ex] \textbf{2 marks}}
+  & ~~~~fails & \makecell[r]{at most\\[-1ex] 1 mark} & \makecell[r]{\textbf{at least}\\[-1ex] \textbf{2 marks}}\\
+  % &&& at most & \textbf{at least} && at most & \textbf{at least}\\[-.8ex]
+  % \multicolumn{2}{l}{translator ~~~ (version)$^{\text{web}}$} & ~~~~fails & 1 mark & \!\!\!\!\textbf{2 marks} & ~~~~fails & 1 mark & \!\!\!\!\textbf{2 marks}\\
   \midrule 
@@
   \texttt{ltl2dela} & (21.0)\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
-       & 5 & 214 & \textbf{129} & ? & ? & \textbf{?} \\
+  & 5 & 214 & \textbf{129} & ? & ? & \textbf{?}\\
   \texttt{ltl2dgra} & (21.0)\footnotemark[3]
@@
   \toprule
-  && \multicolumn{4}{c}{\makecell{reduction of automata from\\[-.8ex] formulas from literature}}
-  & \multicolumn{4}{c}{\makecell{reduction of automata from\\[-.8ex] random formulas}}
+  && \multicolumn{4}{c}{\makecell{reduction of marks in automata\\[-.8ex] from formulas from literature}}
+  & \multicolumn{4}{c}{\makecell{reduction of marks in automata\\[-.8ex] from random formulas}}
   \\ \cmidrule(lr){3-6} \cmidrule(lr){7-10}

### 2023-06-03 qbf/formula_preparation.py
Current experiments - except delag

@@ -237,7 +237,7 @@ def run_with_limited_time(aut1, aut2, time):
 
     try:
         cp = subprocess.run(["python3",
-                             "./are_eq.py",
+                             "/home/xschwar3/telatko/are_eq.py",
                              "-F1", "./a1", "-F2", "a2"],
                             universal_newlines=True,
                             stdout=subprocess.PIPE,
@@ -247,10 +247,10 @@ def run_with_limited_time(aut1, aut2, time):
         print("expired")
         return False
 
-    if cp.stdout == "NOT-EQUIVALENT\n":
-        return False
-    else:
+    if cp.stdout == "EQUIVALENT\n":
         return True
+    else:
+        return False
 
 
 

### 2023-06-03 qbf/incremental.py

@@ -118,10 +118,9 @@ def inc_loop(
         if aut.get_acceptance().used_sets().count(
         ) < 1 or aut.prop_state_acc() == spot.trival.yes_value:
             return aut
-        
-        solver.check()
-        res = solver.check()
+        #print(f"L = {tmp_mode} C= {C} K = {K} call solver")
 
+        res = solver.check()
 
         result, _, _ = z3_result(res)
 

### 2023-06-03 statistics/sat23-camera-ready/random/translate.py

@@ -20,13 +20,14 @@ def main(argv):
     rand_aut = 'random.csv'
     lcr = LtlcrossRunner(tools, formulas = [rand_ltl], res_filename = rand_aut)
     lcr.run_ltlcross(automata = True, timeout = '60')
+    
     ra = ResultsAnalyzer(rand_aut)
     ra.parse_results()
     ltlf_translated = {}
 
     for tool in tools_order:
         fs = ra.values.filter(items = [('acc', tool)]).dropna()
-
+        
         ltlf_translated[tool] = len(fs[fs[('acc', tool)] > 1])
 
         fd = open('random-' + tool + '.hoa', 'w')

### 2023-06-03 telatko.py

@@ -105,7 +105,7 @@ def main(argv):
             else:
                 print_aut(auto, None, " ")
 
-            """
+            """ 
             if not spot.are_equivalent(origin, auto):
                 print("NOT EQUIVALENT!")
                 assert (False)

### 2023-06-03 paper/main.tex
minor changes

@@
   \toprule
-  && \multicolumn{3}{c}{\!348 formulas from literature\!} & \multicolumn{3}{c}{366 random formulas}\\
+  && \multicolumn{3}{c}{\!348 formulas from literature\!} & \multicolumn{3}{c}{335 random formulas}\\
   \cmidrule(lr){3-5} \cmidrule(lr){6-8}
@@
   \texttt{ltl2tgba -G} & (2.10.4)\tablefootnote{\url{https://spot.lrde.epita.fr}}
-  & 0 & 278 & \textbf{70} & ? & ? & \textbf{?}\\
+  & 0 & 278 & \textbf{70} & 0 & 320 & \textbf{15}\\
   \texttt{ltl3tela} & (2.2.0)\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
-  & 18 & 239 & \textbf{91} & ? & ? & \textbf{?}\\
+  & 18 & 239 & \textbf{91} & 0 & 286 & \textbf{49}\\
   \texttt{ltl3tela -D1} & (2.2.0)\footnotemark[2]
-  & ? & ? & \textbf{?} & ? & ? & \textbf{?}\\ 
+  & 20 & 247 & \textbf{81} & 0 & 291 & \textbf{44}\\ 
   \texttt{ltl2dela} & (21.0)\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
-  & 5 & 214 & \textbf{129} & ? & ? & \textbf{?}\\
+  & 5 & 214 & \textbf{129} & 0 & 246 & \textbf{89}\\
   \texttt{ltl2dgra} & (21.0)\footnotemark[3]
-  & 12 & 102 & \textbf{234} & ? & ? & \textbf{?}\\
+  & 12 & 102 & \textbf{234} & 0 & 130 & \textbf{205}\\
   \bottomrule

### 2023-06-03 qbf/formula_preparation.py

@@ -162,7 +162,7 @@ def scc_unoptimized_formula(
             list(map(lambda e: aut.edge_number(e), scc_edg_list)))
     f_creator = None
     if solver == "z3":
-        
+
         f_creator = Z3_f_ctor(
             edge_dict,
             scc_edg,
@@ -231,7 +231,7 @@ def scc_optimized_formula(aut, scc_state_info, C,
 
 
 def run_with_limited_time(aut1, aut2, time):
-    
+
     safe_file("a1", aut1.to_str(), "w")
     safe_file("a2", aut2.to_str(), "w")
 
@@ -246,11 +246,14 @@ def run_with_limited_time(aut1, aut2, time):
     except subprocess.TimeoutExpired:
         print("expired")
         return False
+    print(f"Response: {cp.stdout}")
 
-    if cp.stdout == "NOT-EQUIVALENT\n":
-        return False
-    else:
+    if cp.stdout == "EQUIVALENT\n":
+        print("here")
+
         return True
+    else:
+        return False
 
 
 

### 2023-06-03 qbf/incremental.py

@@ -118,9 +118,12 @@ def inc_loop(
         if aut.get_acceptance().used_sets().count(
         ) < 1 or aut.prop_state_acc() == spot.trival.yes_value:
             return aut
-        
+        print(f"C = {C}, K = {K} solver call")
         solver.check()
+        print("___")
         res = solver.check()
+        print("solver call done")
+
 
 
         result, _, _ = z3_result(res)

### 2023-06-03 telatko.py

@@ -105,11 +105,12 @@ def main(argv):
             else:
                 print_aut(auto, None, " ")
 
-            """
-            if not spot.are_equivalent(origin, auto):
-                print("NOT EQUIVALENT!")
-                assert (False)
-            """
+	   
+            #if not spot.are_equivalent(origin, auto):
+            #   print("NOT EQUIVALENT!")
+            #   assert (False)
+           
+
 
 
 

### 2023-06-03 paper/main.tex
new numbers (exept for patterns-delag)

@@
 equivalent to $\true$ and $\false$. After these steps, we had 348
-LTL formulas from literature and 366 random formulas.
+LTL formulas from literature and 335 random formulas.
 %
@@
   & \multicolumn{4}{c}{\mybox{198 marks in 70 automata}}
-  & \multicolumn{4}{c}{\mybox{XYZ marks in XYZ automata}}\\
+  & \multicolumn{4}{c}{\mybox{32 marks in 15 automata}}\\
   & 1
-  & 198 & 0.0~~~ & 49.3 & 0~~~
-  &   ? &   ?~~~ &    ? & ?~~~ \\                     
+  & 198 & 0.0~~~ & 48.5 & 0~~~
+  &   32 &   0.0~~~ &    8.4 & 0~~~ \\                     
   & 2
-  & 198 & 0.0~~~ & 78.9 & 0~~~
-  &   ? &   ?~~~ &    ? & ?~~~ \\                     
+  & 198 & 0.0~~~ & 65.2 & 0~~~
+  &   31 &   3.1~~~ &    9.4 & 0~~~ \\                     
   & 3
-  & 189 & 4.5~~~ & 1222.8 & 7~~~
-  &   ? &   ?~~~ &    ? & ?~~~ \\                     
+  & 189 & 4.5~~~ & 409.8 & 7~~~
+  &   26 &   18.8~~~ &    43.9 & 1~~~ \\                     
   & multi
-  & 189 & 4.5~~~ & 1259.1 & 6~~~
-  &   ? &   ?~~~ &    ? & ?~~~ \\                    
+  & 189 & 4.5~~~ & 427.6 & 7~~~
+  &   26 &   18.8~~~ &    44.9 & 1~~~ \\                    
   \midrule
@@
   & \multicolumn{4}{c}{\mybox{348 marks in 91 automata}}
-  & \multicolumn{4}{c}{\mybox{XYZ marks in XYZ automata}}\\
+  & \multicolumn{4}{c}{\mybox{120 marks in 49 automata}}\\
+  & 1
+  & 332 & 4.6~~~ & 530.3 & 13~~~
+  &   101 &   15.8~~~ &     32.4 &  0~~~ \\                     
+  & 2
+  & 334 & 4.0~~~ & 551.0 & 14~~~
+  &   100 &   16.7~~~ &     32.3 &  0~~~ \\                     
+  & 3
+  & 326 & 6.3~~~ & 698.5 & 18~~~
+  &   95 &   20.8~~~ &     66.9 &  1~~~ \\                     
+  & multi
+  & 319 & 8.3~~~ & 1619.2 & 18~~~
+  &   95 &   20.8~~~ &    73.4 & 1~~~ \\                    
+  \midrule
+  %  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
+  \multicolumn{2}{l}{\texttt{ltl3tela -D1}} 
+  & \multicolumn{4}{c}{\mybox{272 marks in 81 automata}}
+  & \multicolumn{4}{c}{\mybox{97 marks in 44 automata}}\\
+  & 1
+  & 272 & 0.0~~~ & 383.1 & 9~~~
+  &   95 &   2.1~~~ &     23.8 &  0~~~ \\                     
+  & 2
+  & 272 & 0.0~~~ & 386.6 & 10~~~
+  &   95 &   2.1~~~ &     24.6 &  0~~~ \\                     
+  & 3
+  & 272 & 0.0~~~ & 950.2 & 14~~~
+  &   92 &   5.2~~~ &     54.1 &  0~~~ \\                     
+  & multi
+  & 272 & 0.0~~~ & 1659.6 & 16~~~
+  &   92 &   5.2~~~ &    67.2 & 1~~~ \\                    
+  \midrule
+  %  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
+  \multicolumn{2}{l}{\texttt{ltl2dgra}} 
+  & \multicolumn{4}{c}{\mybox{882 marks in 234 automata}}
+  & \multicolumn{4}{c}{\mybox{491 marks in 205 automata}}\\
+  & 1
+  & 544 & 38.3~~~ & 859.1 & 14~~~
+  &   293 &   40.3~~~ &     275.6 &  0~~~ \\                     
+  & 2
+  & 554 & 37.2~~~ & 1073.5 & 17~~~
+  &   280 &   43.0~~~ &     283.9 &  0~~~ \\                     
+  & 3
+  & 553 & 37.3~~~ & 1349.7 & 22~~~
+  &   267 &   45.6~~~ &     433.6 &  3~~~ \\                     
+  & multi
+  & 535 & 39.3~~~ & 2434.0 & 23~~~
+  &   264 &   46.2~~~ &    411.7 & 2~~~ \\                    
+  \midrule
+  %  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
+  \multicolumn{2}{l}{\texttt{ltl2dela}} 
+  & \multicolumn{4}{c}{\mybox{348 marks in 91 automata}}
+  & \multicolumn{4}{c}{\mybox{234 marks in 89 automata}}\\
   & 1
   & 332 & 4.6~~~ & 465.9 & 12~~~
-  &   ? &   ?~~~ &     ? &  ?~~~ \\                     
+  &   154 &   34.2~~~ &     89.0 &  0~~~ \\                     
   & 2
   & 334 & 4.0~~~ & 553.9 & 15~~~
-  &   ? &   ?~~~ &     ? &  ?~~~ \\                     
+  &   153 &   34.6~~~ &     123.6 &  0~~~ \\                     
   & 3
   & 326 & 6.3~~~ & 691.8 & 18~~~
-  &   ? &   ?~~~ &     ? &  ?~~~ \\                     
+  &   149 &   36.3~~~ &     172.7 &  2~~~ \\                     
   & multi
   & 323 & 7.2~~~ & 1545.9 & 18~~~
-  &   ? &   ?~~~ &    ? & ?~~~ \\                    
-  \midrule
-  &&&& \vdots \\
+  &   148 &   36.8~~~ &    219.8 & 2~~~ \\
+
 % \\

### 2023-06-04 paper/main.tex
Updated delag data in paper.
moving tables and figures in Section 6
Added heatmap2
polishing mainly in Section 6

@@
   % also decreasing speed of their satisfiability solving.
-  We evaluated
-  our reduction technique on TELA produced by state-of-the-art tools
-  of the libraries Owl and Spot and by the tool \texttt{ltl3tela}. The
-  technique reduced some acceptance marks in automata produced by all
-  the tools. \fixme{On automata with more than one acceptance mark produced
-  by tools Delag and Rabinizer~4 of the Owl library, our technique reduced
-  29.1\% and 38.5\% of acceptance marks, respectively.}
+  We evaluated our reduction technique on TELA produced by
+  state-of-the-art tools of the libraries Owl and Spot and by the tool
+  \texttt{ltl3tela}. The technique reduced some acceptance marks in
+  automata produced by all the tools. On automata with more than one
+  acceptance mark obtained by translation of LTL formulas from
+  literature with tools Delag and Rabinizer~4, our
+  technique reduced 27.7\% and 39.3\% of acceptance marks,
+  respectively. The reduction was even higher on automata from random
+  formulas.
   % The achieved
@@
 and the theorem prover Z3~\cite{demoura.08.tacas} to solve the
-satisfiability of QBF transformed to prenex normal form (and not using
-CNF). Our tool is available at
+satisfiability of QBF transformed to prenex (non-CNF) normal form. Our
+tool is available at
 \begin{center}
@@
 and \texttt{ltl3tela}~\cite{major.19.atva}, and to deterministic TELA
-by \texttt{ltl3tela} with option \texttt{-D1} and by another two
+by \texttt{ltl3tela} with option \texttt{-D1} and by two
 state-of-the-art translators from the Owl
@@
 % seconds.
-Further, we have removed automata with 0 or 1 acceptance mark as
-there is a little point in reducing these. Table
-\ref{tab:ltltranslators} shows the exact versions of the
-translators. For each set of formulas and for each translator, it also
-provides the number of fails, the number of produced automata with
-less than two marks, and the number of automata with at least two
-marks. The numbers of automata with at least two marks are typeset in
-bold as these automata are actually used for the experimental
-evaluation of our reduction technique.
+Further, we have removed automata with 0 or 1 acceptance mark as there
+is a little point in reducing these. Table \ref{tab:ltltranslators}
+shows the exact versions of the translators. For each translator and
+each set of formulas, the table also provides the number of fails, the
+number of produced automata with less than two marks, and the number
+of automata with at least two marks. The numbers of automata with at
+least two marks are typeset in bold as these automata are actually
+used for the experimental evaluation of our reduction technique.
 
@@
 
-For each automata set identified by the translator and the set of
-formulas, Table~\ref{tab:cumm-data} shows the cumulative numbers of
-marks in the input automata sets and after each reduction, together
-with the reduction ratio and total time spent by the considered
-reduction. The column \textit{solver timeout} shows the number of
-automata for which the last query to QBF solver did not finish within
-the 30~seconds limit. The timeout of the last QBF query means that the
-automaton may be potentially further reduced if a longer time limit is
-used. \fixme{Zkontrolovat zbytek odstavce podle aktualnich vysledku,
-  specifikovat formula set.} One can observe that a higher level
-sometimes achieves a smaller reduction than a lower level (e.g.,
-compare level~1 and level~2 for \texttt{ltl2dgra}). This is caused by
-the QBF solver timeouts occuring earlier as formulas constructed by
-the higher level are more complex. For \texttt{ltl2dela} and
-\texttt{ltl2dgra}, the automata sets do not contain any automaton
-where level~2 or level~3 achieves a better result than
-level~1. However, all levels contribute to the reductions in the
-multi-level setting.
-
-\begin{table}[t]
+\begin{table}[tp]
   \caption{The cumulative numbers of acceptance marks before and after
@@
   %  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
+  \multicolumn{2}{l}{\texttt{ltl2dela}} 
+  & \multicolumn{4}{c}{\mybox{523 marks in 129 automata}}
+  & \multicolumn{4}{c}{\mybox{234 marks in 89 automata}}\\
+  & 1
+  & 386 & 26.2~~~ & 811.4 & 18~~~
+  &   154 &   34.2~~~ &     89.0 &  0~~~ \\                     
+  & 2
+  & 391 & 25.2~~~ & 1071.6 & 19~~~
+  &   153 &   34.6~~~ &     123.6 &  0~~~ \\                     
+  & 3
+  & 397 & 24.1~~~ & 7326.8 & 26~~~
+  &   149 &   36.3~~~ &     172.7 &  2~~~ \\                     
+  & multi
+  & 378 & 27.7~~~ & 9186.0 & 24~~~
+  &   148 &   36.8~~~ &    219.8 & 2~~~ \\
+  \midrule
+  %  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
   \multicolumn{2}{l}{\texttt{ltl2dgra}} 
@@
   &   264 &   46.2~~~ &    411.7 & 2~~~ \\                             
-  \midrule
-  %  \multirow{5}{*}{\makecell{\texttt{ltl2tgba -G}}} &
-  \multicolumn{2}{l}{\texttt{ltl2dela}} 
-  & \multicolumn{4}{c}{\mybox{348 marks in 91 automata}}
-  & \multicolumn{4}{c}{\mybox{234 marks in 89 automata}}\\
-  & 1
-  & 332 & 4.6~~~ & 465.9 & 12~~~
-  &   154 &   34.2~~~ &     89.0 &  0~~~ \\                     
-  & 2
-  & 334 & 4.0~~~ & 553.9 & 15~~~
-  &   153 &   34.6~~~ &     123.6 &  0~~~ \\                     
-  & 3
-  & 326 & 6.3~~~ & 691.8 & 18~~~
-  &   149 &   36.3~~~ &     172.7 &  2~~~ \\                     
-  & multi
-  & 323 & 7.2~~~ & 1545.9 & 18~~~
-  &   148 &   36.8~~~ &    219.8 & 2~~~ \\
-
-% \\
-% level 1 &
-% 387~ & 26.0~~~~~~ & 755.4~ &  18~~~
-% \\
-% &
-% level 2 &
-% 390~ & 25.4~~~~~~ & 1024.5~ &  20~~~
-% \\
-% &
-% level 3 &
-% 393~ & 24.9~~~~~~ & 7273.3~ &  25~~~
-% \\
-% &
-% multi &
-% 371~ & 29.1~~~~~~ & 9111.4~ &  21~~~
-% \\
-% \midrule
-% \multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\234 automata\\882 acc.~marks}}
-% &
-% level 1 &
-% 552~ & 37.4~~~~~~ & 738.7~ & 14~~~
-% \\
-% &
-% level 2 &
-% 569~ & 35.5~~~~~~ & 907.3~ &  18~~~
-% \\
-% &
-% level 3 &
-% 561~ & 36.4~~~~~~ & 1163.4~ &  21~~~
-% \\
-% &
-% multi &
-% 542~ & 38.5~~~~~~ & 2320.7~ &  21~~~
-% \\
-% \midrule
-% \multirow{4}{*}{\makecell{\texttt{ltl3tela}\\91 automata\\348 acc.~marks}}
-% &
-% level 1 &
-% 332~ & 4.6~~~~~~ & 465.9~ &  12~~~
-% \\
-% &
-% level 2 &
-% 334~ & 4.0~~~~~~ & 553.9~ &  15~~~
-% \\
-% &
-% level 3 &
-% 326~ & 6.3~~~~~~ & 691.8~ &  18~~~
-% \\
-% &
-% multi &
-% 323~ & 7.2~~~~~~ & 1545.9~ & 18~~~
-% \\
   \bottomrule                                                                     
@@
 
-\fixme{Upravit zbytek textu.}
+For each automata set identified by the translator and the set of
+formulas, Table~\ref{tab:cumm-data} shows the cumulative numbers of
+marks in the input automata set and after each reduction, together
+with the reduction ratio and total time spent by the considered
+reduction. The column \textit{solver timeout} shows the number of
+automata for which the last query to QBF solver did not finish within
+the 30~seconds limit. The timeout of the last QBF query means that the
+automaton may be potentially further reduced if a longer time limit is
+used. One can observe that a higher level sometimes achieves a smaller
+reduction than a lower level (e.g., compare level~1 and level~2 for
+\texttt{ltl3tela} on automata coming from formulas from
+literature). This is caused by the QBF solver timeouts occuring
+earlier as formulas constructed by the higher level are more
+complex. The automata sets produced by \texttt{ltl2dela} and
+\texttt{ltl2dgra}, on formulas from literature do not contain any
+automaton where level~2 or level~3 achieves a better result than
+level~1. However, all levels contribute to the reductions in the
+multi-level setting.
+
+\fixme{Upravit zbytek textu a popis Table~3.}
 Table~\ref{tab:heatmap} shows the number of automata from all
@@
 
-\begin{table}[b!]
+\begin{table}[t]
 \caption{The effect of multi-level reduction.
@@
 \centering
-\input{table-heatmap.tex}
+\input{table-heatmap2.tex}
 \end{table}

### 2023-06-04 paper/main.tex
Plots
heatmap tables polished

@@
 by two state-of-the-art translators, namely \texttt{ltl2tgba} (used
-with option \texttt{-G}) from the Spot library~\cite{duret.16.atva2}
-and \texttt{ltl3tela}~\cite{major.19.atva}, and to deterministic TELA
-by \texttt{ltl3tela} with option \texttt{-D1} and by two
-state-of-the-art translators from the Owl
-library~\cite{kretinsky.18.atva}, namely \texttt{ltl2dela} (known as
-Delag)~\cite{muller.17.gandalf} and \texttt{ltl2dgra} (known as
-Rabinizer~4)~\cite{kretinsky.18.cav}. Some translators failed on some
-formulas: it usually reached a timeout of 60 seconds or it produced an
-automaton that cannot be parsed by the Spot library.
+with option \texttt{-G} to get generic TELA) from the Spot
+library~\cite{duret.16.atva2} and
+\texttt{ltl3tela}~\cite{major.19.atva}, and to deterministic TELA by
+\texttt{ltl3tela} with option \texttt{-D1} and by two state-of-the-art
+translators from the Owl library~\cite{kretinsky.18.atva}, namely
+\texttt{ltl2dela} (known as Delag)~\cite{muller.17.gandalf} and
+\texttt{ltl2dgra} (known as Rabinizer~4)~\cite{kretinsky.18.cav}. Some
+translators failed on some formulas: it usually reached a timeout of
+60 seconds or it produced an automaton that cannot be parsed by the
+Spot library.
 % produced by the individual tools contains less than 348 automata as not
@@
 \begin{table}[tp]
-  \caption{The cumulative numbers of acceptance marks before and after
-    reduction for various reduction levels and automata sets given by
-    the translator and formula set. The column \emph{reduction} shows
-    the cumulative percentage of saved acceptance marks and
-    \emph{time} reports the cumulative reduction time. The column
-    \textit{solver timeout} indicates the number of instances where
-    the last call to the QBF solver timed out.  }
+  \caption{For each automata set identified by the translator and the
+    set of formulas, the table provides the cumulative number of
+    acceptance marks before any reduction (in the box), after
+    reduction of individual levels and after multi-level reduction
+    (column \emph{marks}). The column \emph{reduction} shows the
+    percentage of saved acceptance marks and \emph{time} reports the
+    cumulative reduction time in seconds. The column \textit{solver
+      timeout} indicates the number of instances where the last call
+    to the QBF solver timed out.  }
 \label{tab:cumm-data}
@@
 
-\fixme{Upravit zbytek textu a popis Table~3.}
-Table~\ref{tab:heatmap} shows the number of automata from all
-four sets that have a given number of original acceptance marks and a
-given number of acceptance marks after multi-level reduction. The
-table indicates that in many cases only 1 or 2~marks can be
-saved. However, the achieved reduction is substantial for some
-automata with a higher number of original acceptance marks. For
-example, in 24 cases, we have reduced 7 or more
-acceptance marks to only 4.
+Table~\ref{tab:heatmap} shows the number of automata from all five
+sets constructed from formulas from literature that have a given
+number of original acceptance marks and a given number of acceptance
+marks after multi-level reduction. The table indicates that in many
+cases only 1 or 2~marks can be saved. However, the achieved reduction
+is substantial for some automata with a higher number of original
+acceptance marks. For example, in 26 cases, we have reduced 7 or more
+acceptance marks to only 4 or less.
 % Moreover, there were many timeouts of the last QBF query for manof these automata.
@@
 % reduced 8 or more acceptance marks to only 4.
+Table~\ref{tab:heatmap-random} shows the same information for all
+considered automata constructed from random formulas.
 
 \begin{table}[t]
-\caption{The effect of multi-level reduction.
-A cell on coordinates $(x,y)$ contains the number of automata that have been
-reduced from $x$ to $y$ acceptance marks. If the cell contains a sum of two
-numbers, the latter represents the number of automata where the attempt to reduce
-another mark has been unsuccessful due to a QBF solver timeout.}
+  \caption{The effect of multi-level reduction on automata constructed
+    from formulas from literature.  A cell on coordinates $(x,y)$
+    contains the number of automata that have been reduced from $x$ to
+    $y$ acceptance marks. If the cell contains a sum of two numbers,
+    the latter represents the number of automata where the attempt to
+    reduce another mark has been unsuccessful due to a QBF solver
+    timeout.}
 \label{tab:heatmap}
@@
 \centering
-\input{table-heatmap2.tex}
+\input{table-heatmap.tex}
+\end{table}
+
+\begin{table}[t]
+\caption{The effect of multi-level reduction on automata constructed
+    from random formulas. The meaning of each cell is the same as in Table~\ref{tab:heatmap}.}
+\label{tab:heatmap-random}
+\setlength{\tabcolsep}{2.5pt}
+\centering
+\input{table-heatmap-random.tex}
 \end{table}
@@
 a relatively small number of complicated automata.
+
+\ref{fig:quantile-random}
 
@@
 \input{quantile.tex}
+\end{figure}
+
+\begin{figure}[t]
+  \caption{Quantile random.}
+\label{fig:quantile-random}
+\bigskip
+\centering
+\input{quantile-random.tex}
 \end{figure}

### 2023-06-05 paper/main.tex
final version finished?

@@
 
-Figure~\ref{fig:quantile} presents the time spent by multi-level
-reduction on individual automata of each automata set. The chart
-shows a pleasing finding that for every set, most automata are reduced
-in under a second and the high cumulative running times are caused by 
-a relatively small number of complicated automata.
-
-\ref{fig:quantile-random}
-
-\begin{figure}[t]
-  \caption{Performance of \telatko on individual automata of each automata sets.
-    The graph shows the time ($y$ axis) needed by \telatko
-    to process the $x$-th automaton of the set, where automata of each set are ordered by their processing time.
+\begin{figure}[p]
+  \centering
+  \input{quantile.tex}
+  \bigskip\bigskip
+  \input{quantile-random.tex}
+  \bigskip
+  \caption{Performance of \telatko on individual automata of each automata set.
+    Automata sets constructed from formulas from literature are in the upper graph,
+    automata sets constructed from random formulas are in the lower graph.
+    Each line shows the time ($y$ axis) needed by \telatko to process the
+    $x$-th automaton of the set, where automata in the set are ordered by
+    their processing time.
     % The $x$ axis represents the number of automata from each automata
@@
   }
-\label{fig:quantile}
-\bigskip
-\centering
-\input{quantile.tex}
-\end{figure}
-
-\begin{figure}[t]
-  \caption{Quantile random.}
-\label{fig:quantile-random}
-\bigskip
-\centering
-\input{quantile-random.tex}
-\end{figure}
+\label{fig:quantile}\end{figure}
+
+Figure~\ref{fig:quantile} presents the time spent by multi-level
+reduction on individual automata of each automata set. The charts
+show a pleasing finding that for every set, most automata are reduced
+in under 5 seconds and the high cumulative running times are caused by 
+a relatively small number of complicated automata.
 
@@
 automata produced by all considered state-of-the-art LTL to automata
-translators on formulas from literature. Further, it clearly shows
-that the translators of the Owl library are significantly less
-economical with acceptance marks than the other two translators.
+translators. Further, it clearly shows that the translators of the Owl
+library are significantly less economical with acceptance marks than
+the other two translators.
 
@@
 
-In the future, we plan to improve the reduction process (e.g., use the
-bisection method to reduce the number of acceptance marks more
-efficiently) and study whether the reduction of acceptance condition
-can enable further reductions or simplifications of automata
-structures.
+% In the future, we plan to improve the reduction process (e.g., use the
+% bisection method to reduce the number of acceptance marks more
+% efficiently) and study whether the reduction of acceptance condition
+% can enable further reductions or simplifications of automata
+% structures.
 

### 2023-06-05 paper/main.tex
more polishing towards expected format

@@
 \bibliographystyle{plainurl}% the mandatory bibstyle
+\nolinenumbers
 
@@
 \usepackage{amsmath,amssymb}
-\usepackage{enumerate}
+%\usepackage{enumerate}
 \usepackage{tablefootnote}
@@
 %}}}
+
+\EventEditors{Meena Mahajan and Friedrich Slivovsky}
+\EventNoEds{2}
+\EventLongTitle{26th International Conference on Theory and Applications of Satisfiability Testing (SAT 2023)}
+\EventShortTitle{SAT 2023}
+\EventAcronym{SAT}
+\EventYear{2023}
+\EventDate{July 4--8, 2023}
+\EventLocation{Alghero, Italy}
+\EventLogo{}
+\SeriesVolume{271}
+\ArticleNo{22}
 
@@
 \section{Introduction}
-
-% Automata over infinite words (also known as $\omega$-automata) are now
-% used in specification, verification, analysis, monitoring, and
-% synthesis of various systems with infinite behaviour. The first and
-% the most common kind of $\omega$-automata is Büchi
-% automaton~\cite{buchi.60.clmps} introduced in 1960. As deterministic
-% Büchi automata are less expressive than nondeterministic ones,
-% deterministic $\omega$-automata are typically considered with other
-% acceptance condition like Rabin, Streett, Muller, or parity. In 1987,
-% Emerson and Lei~\cite{emerson.87.scp} introduced $\omega$-automata
-% where the acceptance condition can be an arbitrary combination of
-% acceptance primitives saying that a certain set of states should be
-% visited finitely often or infinitely often. In 2015, the same kind of
-% acceptance condition was implemented in the \emph{Hanoi omega-automata
-%   format (HOAF)}~\cite{babiak.15.cav}, with the difference that finiteness
 
@@
 
-% \begin{algorithm}[t]
-%   \DontPrintSemicolon
-%   \SetKwFunction{FextendBackward}{extendBackward}
-%   \SetKwProg{PI}{Procedure}{}{}
-%   \PI{$\mathit{SingleLevelReduction}(\mA,j,\mathit{reduceC})$}{
-%     \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, level $j\in\{1,2,3\}$, $\mathit{reduceC}\in\{\true,\false\}$}
-%     \KwOut{an equivalent TELA with the same structure as $\mA$ and with at most as many acceptance marks as in $\mA$} %with the same or lower number of acceptance~marks}
-%     \;
-%   %   \If{$L(\mA)=\emptyset$}{
-%   % %      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'=\{(p,a,\emptyset,q)\mid(p,a,M',q)\in\delta\}$
-%   %       \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all acceptance marks removed
-%   %     }
-%     $C_\mA\gets$ the number of cubes in the acceptance formula of $\mA$ transformed to DNF\;
-%     $K_\mA\gets$ the number of acceptance marks in $\mA$\;
-%     $C\gets C_\mA$\;
-%     $K\gets K_\mA$\;
-%     $\mathit{stop}\gets\false$\;
-%     \While(\tcp*[f]{reduction of the number of marks}){$\neg\mathit{stop}~\wedge~K>0$}{
-%       $\Phi\gets\Phi^\mathit{opt}_{j,C,K{-}1}$ for $\mA$\;
-%       \leIf{$\mathit{satisfiable}(\Phi)$}{$K\gets K{-}1$}{$\mathit{stop}\gets\true$}
-%     }
-%     \If(\tcp*[f]{check whether the condition $\false$ fits}){$K=1~\wedge~\mathit{stop}$}{
-%       $\Phi\gets\Phi^\mathit{opt}_{j,0,0}$ for $\mA$\; 
-%       \If{$\mathit{satisfiable}(\Phi)$}{
-%         \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all marks removed
-%       }
-%     }  
-%     \If(\tcp*[f]{reduction of the number of cubes}){$\mathit{reduceC}$}{
-%       $\mathit{stop}\gets\false$\;
-%       \While{$\neg\mathit{stop}~\wedge~C>1$}{
-%         $\Phi\gets\Phi_{j,C-1,K}$ for $\mA$\;
-%         \leIf{$\mathit{satisfiable}(\Phi)$}{$C\gets C{-}1$}{$\mathit{stop}\gets\true$}
-%       }
-%     }
-%     \eIf{$K=K_\mA~\wedge~C=C_\mA$}{
-%       \KwRet $\mA$
-%     }{
-%       compute $\rem$ and $\psi$ from a model of $\Phi_{j,C,K}$\;
-%       \KwRet $(Q,N_K,\Sigma,q_I,\rem(\delta),\psi)$
-%     }
-%   }\;
-%   \caption{The single-level reduction procedure}
-%   \label{alg:single}
-% \end{algorithm}
-
 \begin{algorithm}[t]
@@
 finish in the time limit when trying to reduce the number of marks to
-1.%when trying the automaton may be potentially reduced further.
-
-% that does not . 
-% The parameters of the last query of the QBF solver are encoded in the name of the automaton.
-
-% The name of the automaton is in the following format:
-% \begin{center}
-%   \texttt{level\_marks\_solver}
-% \end{center}
-% where 
-% \begin{itemize}
-%     \item \texttt{level} $\in \{L1, L2, L3, T, F\}$ and it gives us the information about which level of simplification was applied (or in case of $T$ and $F$ if the acceptance formula can be $\true$ or $\false$).
-%     \item \texttt{marks} $\in \mathbf{N}$ representing the number of marks we built the formula for.
-%     \item \texttt{solver} $\in \{U, T, S\}$ where $U$ means that the query was unsatisfiable, $T$ means the solver timeouted and $S$ means that the simplification is satisfiable.
-% \end{itemize}
-
-% Note that we are only interested in the result of the last query, therefore we get the information that the simplification was satisfiable only if the result automaton has zero acceptance marks.
-Meaning that the acceptance condition is either $\true$ or $\false$.
+1.
 
@@
 \centering
-% \begin{tabular}{lr@{~~~~~~}lr}
-% \toprule
-% tool        & version&homepage & ~~~automata \\
-% \midrule
-% \texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
-% \texttt{ltl2dgra} & 21.0 & & 234\\
-% \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
-% \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
-% \bottomrule
-% \end{tabular}
-%
-% \begin{tabular}{lrrrr}
-%   \toprule
-%   &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
-%   tool &\!\!\!version/homepage & ~~~~fails & less than 2 marks & \textbf{at least 2 marks}\\
-%   \midrule
-%   \texttt{ltl2dela} & 21.0\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
-%        & 5 & 214 & \textbf{129} \\
-%   \texttt{ltl2dgra} & 21.0\footnotemark[1]
-%        & 12 & 102 & \textbf{234}\\
-%   \texttt{ltl2tgba} & 2.10.4\tablefootnote{\url{https://spot.lrde.epita.fr}}
-%        & 0 & 278 & \textbf{70}\\
-%   \texttt{ltl3tela} & 2.2.0\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
-%        & 18 & 239 & \textbf{91}\\
-%   \bottomrule
-% \end{tabular}
-%
 \setlength{\tabcolsep}{5.7pt}
@@
 \centering
-\setlength{\tabcolsep}{3.9pt}
+\setlength{\tabcolsep}{3.7pt}
 \newcommand{\mybox}[1]{\fbox{\begin{minipage}{4.7cm}\centering#1\end{minipage}}}
@@
 
-% Pridat do experimentu nejaky graf, ktery ukaze, jak rychle se nektere
-% automaty redukovaly na minimum? Nebo spis ukazat vliv timeoutu na
-% redukce?  (zvysovat timeout od 1s do 50s a vzdy ukazat miru
-% redukce). Ty redukce ukazovat nejak jinak nez jen celkovym cislem
-% (boxplot? Scatterplot?).
-
-% Zkusit upravit redukci tak, aby vysledkem bylo TGBA a zkusit dale
-% redukovat akceptacni znacky TGBA automatu redukovanych Spotem?
-
-% Chceme ukazovat, ze redukce znacek muze nekdy umoznit dalsi redukci stavu?
-
 %}}}
@@
 
-%{{{ standard reductions
-
-\section{Reduction based on acceptance marks relations}\label{sec:standard}
-
-This section presents a reduction technique based on relations between
-acceptance sets restricted to individual SCCs. It employs the fact
-that acceptance of a run always depends only on a single SCC, namely
-the SCC that contains the whole run except some finite prefix.
-
-Let $\mA$ be a TELA with the acceptance condition $\varphi$. We assume
-that $\varphi$ is in DNF. This technique has two steps.
-\begin{enumerate}
-\item For each SCC $S$, we make a copy $\varphi_S$ of $\varphi$ and
-  simplify it according to relations between acceptance sets
-  restricted to $\delta_S$. 
-\item We collect the simplified acceptance conditions
-  $\varphi_{S_1},\ldots,\varphi_{S_k}$ of all SCCs and merge them into
-  one acceptance condition $\varphi'$. Additionally, we sometimes add
-  acceptance marks to all transitions of some SCCs to make the
-  resulting automaton with acceptance formula $\varphi'$ equivalent to
-  the original automaton.
-\end{enumerate}
-
-% Let $\mathcal{A}$ be an original automaton and let $\psi$ be an
-% acceptance condition of $\mathcal{A}$. Every run of $\mathcal{A}$ has
-% an infinite suffix that takes place within one SCC $S_{i}$. Thus the
-% evaluation of $\psi$ depends purely on the SCC $S_{i}$ and we can
-% optimize the acceptance condition for each SCC separately. This
-% optimization consists of removing redundant terms from acceptance
-% condition and relabeling of acceptance marks on the transitions. The
-% state and transition structure of the SCC does not change. This way we
-% obtain set of simplified acceptance conditions that we merge into new
-% acceptance formula $\psi'$. The automaton with relabeled transitons
-% and new acceptance condition $\psi'$ we denote $\mathcal{A}'$.
-% Finally, we ensure that automaton $\mathcal{A'}$ is equivalent to the
-% original automaton $\mathcal{A}$.
-
-\subsection{Simplification of $\varphi_S$ for an SCC $S$}\label{ssec:step1}
-Recall that $\varphi_S$ is in DNF. Hence, we can represent it as a set
-$\varphi_S=\{D_1,\ldots,D_k\}$ of disjunctive clauses.
-
-Since formula $\psi_{i}$ is in DNF, we can represent it as a set of disjuncts $\overline{\psi_{i}} = \{D_1, D_2, \ldots D_k\}$.
-
-Let $D_{k} \in \overline{\psi_{i}}$ be a disjunct of formula $\overline{\psi_{i}}$ and $C_{j} \in D_{k}$ a conjunct of disjunct $D_{k}$.
-Furthermore let  $\tacc{0}{i}, \tacc{4}{j}, \tacc{8}{k}, \tacc{2}{l} \in M$ be distinct acceptance marks that occur in $\mathcal{A}$.
-
-In the next section, some properties of an automaton allow us to substitute a Boolean value $\true$ or $\false$ for a particular term of the formula $\psi_i$.
-The consequences of this substitution are divided into a number of cases. We represent it on the set-format of formula we just defined.
-
-\begin{itemize}
-    \item If $C_{j}$ is substituted by $\true$, the conjunct is omitted from $D_{k}$. Thus $D_{k} = D_{k} \setminus \{C_{j}\}$.
-    \item If $C_{j}$ is substituted by $\false$, the conjunct causes that the whole $D_{k}$ evaluates to $\false$. Thus $\overline{\psi_{i}} = \overline{\psi_{i}} \setminus \{D_{k}\}$.
-
-\end{itemize}
-
-That being said, we can define the reduction techniques.
-
-\subsubsection{Acceptance formula modifications\\} 
-\label{subsection:acc_modif}
-Since we optimize for each SCC separately, the acceptance formula corresponding to SCC $S_{i}$ can contain terms with acceptance marks that are not present on the transitions of $S_{i}$.
-Let $\tacc{8}{k} \in M$ be an acceptance mark that is not present on any edge of $S_{i}$.
-Then any term that contains this acceptance mark in $\overline{\psi_{i}}$ can be immediately substituted with a boolean value and thus removed from $\overline{\psi_{i}}$.
-\medskip
-
-Therefore acceptance mark $\tacc{8}{k}$ is not visited in any run of $S_{i}$ and thus every term in form $\Fin\tacc{8}{k}$ is substituted with $\true$ and every term $\Inf\tacc{8}{k}$ is substituted with $\false$.
-
-On the contrary, let $\tacc{4}{j}$ be an acceptance mark that is present on every transition of $S_{i}$.
-That means that at least one transition with $\tacc{4}{j}$ is visited infinitely often by a run of $\mathcal{S_{i}}$.
-Thus every term $\Fin\tacc{4}{j}$ is substituted by $\false$ and every term in form $\Inf\tacc{4}{j}$ is substituted by $\true$.
-
-\subsubsection{$\Inf$ conjuncts\\}
-\label{subsection:simpl_inf}
-Reduction of a conjunct of $\Inf\tacc{8}{k}$ form  is based on the inclusion of the sets of edges labeled with $\tacc{8}{k}$ and set of edges labeled with $\tacc{4}{j}$.
-If all transitions that contain an acceptance mark $\tacc{8}{k}$ also contain $\tacc{4}{j}$, we can remove $\Inf\tacc{4}{j}$ from all disjuncts, where it occurs together with $\Inf\tacc{8}{k}$.
-More formally, if the following conditions hold:
-\begin{itemize}
-    \item $Z(\tacc{8}{k}) \subseteq Z(\tacc{4}{j})$
-    \item $\Inf\tacc{8}{k}, \Inf\tacc{4}{j} \in D_{k}$
-\end{itemize}
-then we can we can substitute $\Inf\tacc{4}{j}$ with Boolean value $\true$. 
-This modification does not change the language because it does not affect accepting runs of $S_{i}$.
-The transitions can create an accepting run that has the potential to satisfy $D_{k}$ only if they are labeled with both  $\tacc{8}{k}$  and $\tacc{4}{j}$.
-Since $Z(\tacc{8}{k}) \subseteq Z(\tacc{4}{j})$, one can notice that that if a run satisfies $\Inf\tacc{8}{k}$ then  $\Inf\tacc{4}{j}$ is satisfied as well.
-Therefore if we remove $\Inf \tacc{4}{j}$ from $\overline{\psi_{i}}$, this modification does not affect the language.
-
-\subsubsection{$\Fin$ conjuncts\\}
-Similarly as in previous section, this reduction is based on inclusion.
-If following conditions hold:
-\begin{itemize}
-    \item $Z(\tacc{8}{k}) \subseteq Z(\tacc{4}{j})$
-    \item $\Fin\tacc{8}{k}, \Fin\tacc{4}{j} \in D_{k}$
-\end{itemize}
-then we can we can substitute $\Fin\tacc{8}{k}$ with boolean value $\true$.
-This modification does not change the language because from if $\tacc{4}{j}$ is visited finitely often, then also $\tacc{8}{k}$ is visited finitely often.
-\medskip
-
-We can merge two conjuncts into one if the conjuncts always occur in the same disjuncts in $\Fin$ form.
-We can do so regardless of the position of the acceptance marks that are contained in the conjuncts.
-If the following condition is met:
-\begin{itemize}
-    \item $\forall D_{k}: \Fin\tacc{8}{k} \in D_{k} \Longleftrightarrow \Fin\tacc{4}{j} \in D_{k}$
-\end{itemize}
-We introduce a fresh acceptance mark $\tacc{2}{l}$, such that $Z(\tacc{2}{l}) = Z(\tacc{8}{k}) \cup Z(\tacc{4}{j})$.
- Finally, we can update $D_{k}$:
-\begin{align*}
-    D_{k} = (D_{k} \setminus \{ \Fin\tacc{8}{k}, \Fin\tacc{4}{j} \}) \cup \{\Fin\tacc{2}{l}\}
-\end{align*}
-
-This update basically means that $\Fin\tacc{8}{k}, \Fin\tacc{4}{j}$ are substituted by $\true$ and new conjunct $\Fin\tacc{2}{l}$ is added to $D_{k}$.
-The acceptance mark $\tacc{2}{j}$ is placed on every edge of the SCC $S_{i}$ that is labeled with $\tacc{8}{k}$ or $\tacc{4}{j}$.
-This modification does not change the language because it does not change the acceptance runs of $S_{i}$. It only relabels the acceptance marks on edges that can be seen finitely often.
-
-\subsubsection{$\Fin \land \Inf$ conjuncts\\}
-There are two cases when we can reduce one of the pair of $\Inf\tacc{0}{i}$ and $\Fin\tacc{2}{l}$ conjuncts.
-
-At first, if every transition labeled with  $\tacc{0}{i}$ is also labeled with $\tacc{2}{l}$, the conjunction of $\Inf\tacc{0}{i} \land \Fin\tacc{2}{l}$ is never $\true$ because every run that visits $\tacc{2}{l}$ also visits $\tacc{0}{i}$.
-Therefore we can reduce the whole disjunct $D_{k}$.
-If the following conditions are met:
-\begin{itemize}
-    \item $Z(\tacc{0}{i}) \subseteq Z(\tacc{2}{l})$
-    \item $\Inf\tacc{0}{i}, \Fin\tacc{2}{l} \in D_{k}$
-\end{itemize}
-We substitute the whole disjunct $D_{k}$ for $\false$. In other words $\overline{\psi_{i}} = \overline{\psi_{i}} \setminus \{D_{K}\}$.
-
-Finally we can simplify $\overline{\psi_{i}}$, if the edges labeled with $\tacc{0}{i}$ and edges labeled with $\tacc{2}{l}$ are complementary.
-More formally if:
-\begin{itemize}
-    \item  $\Inf\tacc{0}{i}$ and $\Fin\tacc{2}{l} \in D_{k}$
-    \item $Z(\tacc{0}{i}) \cap Z(\tacc{2}{l}) = \emptyset \land Z(\tacc{0}{i}) \cup Z(\tacc{2}{l}) = \delta_{S_{i}} $
-\end{itemize}
-Then we can substitute $\true$ for $\Inf\tacc{0}{i}$.
-This reduction does not change the language because if during a run of $S_{i}$ all transitions labeled with $\tacc{2}{l}$ are visited only finitely often, then necessarily at least one transition labeled with $\tacc{0}{i}$ is visited infinitely often.
-Therefore removing $\Inf\tacc{0}{i}$ from $\overline{\psi_{i}}$ does not affect the language, since the validity of $\Fin\tacc{2}{l}$ implies the validity of $\Inf\tacc{0}{i}$.
-
-\subsection{Merge of $\psi_{i}$}
-After applying reduction techniques from the previous section, each $S_{i}$ has its own acceptance condition $\psi_{i}$.
-The goal of this section is to merge these acceptance conditions into one formula $\psi'$ with emphasis on obtaining a formula with the minimal number of acceptance sets.
-
-At first, the algorithm finds $\psi_{i}$ with the greatest number of disjuncts and states it as a base of the new acceptance condition $\psi'$.
-Now the algorithm continues by the successive merging of $\psi'$ with the unmerged formula $\psi_{i}$ that has the greatest number of disjuncts.
-This repeats until all formulae $\psi_{i}$ are merged.
-This process updates the form of $\psi'$ until it reaches its final form.
-The process of merging two formulae consists of two phases.
-In the first phase a suitable pairing of disjuncts is found and in the second phase, these disjuncts are merged.
-
-\begin{enumerate}
-	\item Let $\psi'$ denote the future acceptance condition of $\mathcal{A}'$ and $\psi_{i}$ is the acceptance condition of $S_{i}$ chosen to be merged.
-Then the algorithm uses \emph{linear sum assignment} to determine which disjunct of $\psi_{i}$ is merged with which disjunct of $\psi'$.
-The process of using \emph{linear sum assignment} and the pairing procedure is described in detail in work by T.
-Šťastná \cite{stastna.19.bc}.
-    \item Let $D_{K} \in \overline{\psi_{i}}$ be a disjunct paired with $D_{L} \in \overline{\psi'}$.
-All conjuncts of $D_{L}$ are initially labeled as unused.
-Now the algorithm maps every conjunct $C_{k} \in D_{K}$ to a suitable conjunct $C_{l} \in D_{L}$ and labels $C_{l}$ as used.
-\footnote{The indexes differ to emphasize the fact that they are not equal.} A conjunct $C_{l}$ is suitable for $C_{k}$ is it is on the same type ($\Fin$ or $\Inf$), it is labeled as unused.
-If no suitable $C_{l}$ is found, the algorithm adds $C_{k}$ to disjunct $D_{L}$ and marks it as used.
-\end{enumerate}
-
-In an acceptance mark $m_{l} \in M$ occurs in $\psi'$ in more than one conjunct, we need to check whether all of these conjuncts are mapped to conjuncts form $\psi_{i}$ with the same acceptance mark $m_{k} \in M$.
-If not, we resolve this conflict by replacing all additional occurrences of $m_{l}$ with fresh acceptance mark $m_{n}$ and we place $m_{n}$ on the exact transition of $\mathcal{A}'$ where $m_{l}$ is placed.
-Finally, in the end, the algorithm removes every acceptance mark that is not present in $\psi'$ from the edge of $\mathcal{A}$.
-
-\medskip
-
-\begin{example}
-Consider $\overline{\psi'} =  \{ \{ \Fin\tacc{0}{0}, \Inf\tacc{1}{1} \} , \{\Fin\tacc{0}{0} \} \}$ \\ and $\overline{\psi_{i}} = \{  \{\Fin\tacc{4}{4} \}, \{ \Fin\tacc{2}{2}, \Inf\tacc{3}{3} \} \} $.
-We obtain the pairing $(1,2), (2,1)$, meaning that the first disjunct of $\overline{\psi_{i}}$ is merged wit the second disjunct of $\overline{\psi'}$ and second disjunct of $\overline{\psi_{i}}$ is merged with the first disjunct of $\overline{\psi'}$.
-Then $\Fin\tacc{0}{0}$ is mapped to $\Fin\tacc{2}{2}$, $\Inf\tacc{1}{1} $ is mapped to $\Inf\tacc{3}{3}$ and $\Fin\tacc{0}{0}$ is again mapped to $\Fin\tacc{4}{4}$ which causes a conflict.
-Therefore we need to replace the second occurrence of $\tacc{0}{0}$ with fresh mark $\tacc{5}{5}$.
-The set-representation of formula $\psi'$ after merging is $\psi' =  \{ \{ \Fin\tacc{0}{0}, \Inf\tacc{1}{1} \} , \{\Fin\tacc{5}{5} \} \}$.
-
-\end{example}
-
-Unfortunately, sometimes when the optimal mapping was not found, the procedure produces an automaton with a more complex acceptance condition than the acceptance condition of the input automaton.
-Therefore we propose and implement an optimization, which prevents such a behavior and in general, produces smaller acceptance condition than the original one.
-
-\begin{itemize}
-    \item The selection of base of $\psi'$ is based on two keys.
-The primary key is cardinality (number of clauses) of $\overline{\psi_{i}}$, the secondary key is cardinality of clauses (number of terms in clause).
-    \item We order the disjunct in every $\overline{\psi_{i}}$ in descending order, where the key is the cardinality of a particular disjunct.
-    \item The DNF conversion of the input acceptance condition can lead to an exponential blowup of the formula, where some acceptance marks occur more than once.
-Since the reduction of acceptance sets is basically denoting the redundant ones as $\true$ or $\false$, we can as well easily obtain simplified acceptance formula in CNF (we simply denote the same acceptance sets as $\true$ or $\false$ as we did in DNF).
-According to the shape of the input acceptance condition, we choose the suitable (shorter) normal form and perform the merging algorithm on it.
-\end{itemize}
-By choosing the CNF over DNF in the cases where CNF is more natural (shorter), we prevent the formula to contain some acceptance marks more than once (at least, we do not create them by conversion of $\psi$ to DNF).
-This way, we prevent adding new acceptance marks to the formula when resolving conflicts of acceptance marks which leads to better results.
-If the original formula contains an acceptance mark that is present in more than one disjunct, then by ordering the disjuncts in every $\overline{\psi_{i}}$, the \emph{linear sum assignment} returns more convenient pairing.
-Meaning that if \emph{linear sum assignment} finds two equal assignments, it respects the order and the disjuncts with the  acceptance mark that occurs more than once are paired with disjuncts that also contain acceptance mark that occurs more than once.
-
-\begin{example}
-	Consider an automaton $\mathcal{A}$ in Figure \ref{fig:L1_before}.
-In this example, we demonstrate the simplification procedure described in Section \ref{section:simplification} and obtain simplified acceptance conditions $\psi_{i}$ for each SCC $S_{i}$.
-The simplified acceptance conditions $\overline{\psi_{i}}$ are displayed in the set-format we introduced earlier because we make use of it afterward during the merge.
-Then we merge these formulae into the acceptance condition of the simplified automaton $\mathcal{A}'$.
-We enumerate the SCCs $S_{1}, S_{2}, S_{3}$.
-(The indices correspond with the numbers inside the states in Figure \ref{fig:L1_before}.) And we assign an accepting condition $\psi_{1}$ to $S_{1}$, $\psi_{2}$ to $S_{2}$ and $\psi_{3}$ to $S_{3}$.
-    \begin{enumerate}
-	    \item We simplify $\psi_{1}$ according to the placement of acceptance marks on the edges in $S_{1}$.
-At first we notice, that the formula $\overline{\psi_{1}}$ contains acceptance marks that are not present on the edges of $S_{1}$.
-Therefore modify $\overline{\psi_{1}}$ as described in Subsection \ref{subsection:acc_modif}.
-We substitute $\true$ for $\Fin\tacc{3}{3}$ and $\false$ for $\Inf\tacc{1}{1}$ and $\Inf\tacc{0}{0}$.
-(This substitution deletes the whole disjunct $D_{1}$.) Then we notice that the acceptance marks are in a position which allows us to perform the simplification described in first part of subsection 3.1.4.
-We substitute $\true $ for $\Inf\tacc{4}{4}$ and obtain the final form of $\overline{\psi_{1}} = \{\{ \Fin\tacc{2}{2}\} \}$.
-	    \item Similarly as in previous case, we notice that acceptance marks $\tacc{2}{2}$, $\tacc{3}{3}$ and $\tacc{4}{4}$ are not present on the edges of $S_{2}$.
-Therefore we substitute $\true$ for both $\Fin\tacc{2}{2} \text{ and } \Fin\tacc{3}{3}$ and $\false$ for $\Inf\tacc{4}{4}$ (and thus remove the whole disjunct $D_{2}$).
-Then we use the procedure described in Subsection \ref{subsection:simpl_inf} and remove $\Inf\tacc{1}{1}$ from the disjunct $D_{1}$ and obtain $\overline{\psi_{2}} = \{ \{\Inf\tacc{0}{0}\} \}$.
-        \item Finally, we yet again remove marks that are not present in $S_{3}$ and perform the simplification described in subsection 3.1.3.
-We substitute $\true $ for $\Fin\tacc{2}{2} \text{ and } \Fin\tacc{3}{3}$ and add new term $\Fin\tacc{5}{5}$ into $D_{2}$ and place $\tacc{5}{5}$ on the edges that are labeled with $\tacc{2}{2}$ or $\tacc{3}{3}$.
-The simplified acceptance condition is $\overline{\psi_{3}} = \{\{ \Fin\tacc{5}{5}, \Inf\tacc{4}{4}\}\}$.
-    \end{enumerate}
-
-	Now we proceed to the merge of the formulae of the SCCs $\psi_{i}$.
-As a base of the new acceptance condition $\psi'$ we choose the formula $\psi_{3}$ because it has the highest number of disjuncts compared to the other formulae $\psi_i$, so $\overline{\psi'}=\{ \{ \Fin\tacc{5}{5}, \Inf\tacc{4}{4} \}\}$.
-Now we merge  $\psi_{1}$ and $\psi_{2}$ with $\psi'$.
-The order is in this particular case irrelevant because $\overline{\psi_{1}}$ has the same number of disjuncts as $\overline{\psi_{2}}$ (and both disjuncts have the same number of conjuncts).
-So we merge $\overline{\psi_{1}}$ at first because of its lower index.
-Since both $\overline{\psi'}$ and $\overline{\psi_{1}}$ have only one disjunct, the only possible pairing that \emph{linear sum assignment
-    } can give us is $(0,0)$, meaning that $\{ \Fin\tacc{2}{2}\}$ is merged with $\{ \Fin\tacc{5}{5}, \Inf\tacc{4}{4}\}$.
-Then we need to map an unused conjunct  $\Fin\tacc{5}{5}$ to a conjunct of the same type which is $\Fin\tacc{2}{2}$.
-Similarly in case of $\overline{\psi_{2}}$ we map $\Inf\tacc{4}{4}$ on $\Inf\tacc{0}{0}$.
-This way we produce $\mathcal{A}'$ displayed in Figure  \ref{fig:L1_middle} with the acceptance condition $ \Fin\tacc{5}{5} \land \Inf\tacc{4}{4}$.
-Observe that the automaton $\mathcal{A}'$ is not equivalent to the original $\mathcal{A}$.
-For example in Figure \ref{fig:L1_middle} is the highlighted loop not accepting but in the original automaton in Figure \ref{fig:L1_before} the loop accepting is.
-To reach the equivalence between the two automata, we need to \emph{restore equivalence} which is described in the next section.
-\end{example}
-
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-			\node[flabel] at (2,5) {$( \Inf\tacc{0}{0} \land \Inf\tacc{1}{1}) \lor (\Fin\tacc{2}{2} \land \Fin\tacc{3}{3} \land \Inf\tacc{4}{4})$};
-
-			\node[state, initial] (x) at (0,0) {0};
-			\node[state] (y) at (2,2) {1};
-			\node[state] (z) at (2,-1.5) {2};
-			\node[state] (w) at (5,-1.5) {3};
-
-			\path[->]
-				(x) edge[]  (y)
-				(x) edge[]  (z)
-				(z) edge[] (w)
-
-            	(z) edge[loop, in = 75, out = 105, looseness=6] (z)
-            	(z) edge[loop, in = 72, out = 108, looseness=9] (z)
-            	(z) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc1] {1}(z)
-            	(z) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc0, pos=0.6] {0} node[accset,collacc1, pos=0.4] {1}(z)
-
-            	(w) edge[loop, in = 75, out = 105, looseness=6] (w)
-            	(w) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc3] {3}(w)
-            	(w) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc2] {2}(w)
-            	(w) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4}(w)
-
-            	(y) edge[loop, in = 75, out = 105, looseness=6] node[accset,collacc4] {4}(y)
-            	(y) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc4] {4}(y)
-            	(y) edge[loop, in = 70, out = 110, looseness=12] node[accset,collacc2] {2} (y)
-            	(y) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc2] {2}(y)
-
-			;
-		\end{tikzpicture}
-    \caption{TELA before simplification Level 1}
-    \label{fig:L1_before}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-			\node[flabel] at (2,5) {$ \Fin\tacc{5}{5} \land \Inf\tacc{4}{4}$};
-
-			\node[state, initial] (x) at (0,0) {0};
-			\node[state] (y) at (2,2) {1};
-			\node[state] (z) at (2,-1.5) {2};
-			\node[state] (w) at (5,-1.5) {3};
-
-			\path[->]
-				(x) edge[]  (y)
-				(x) edge[]  (z)
-				(z) edge[] (w)
-
-            	(z) edge[loop, in = 75, out = 105, looseness=6] (z)
-            	(z) edge[loop, in = 72, out = 108, looseness=9] (z)
-            	(z) edge[loop, in = 70, out = 110, looseness=12]  (z)
-            	(z) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4} (z)
-
-            	(w) edge[loop, in = 75, out = 105, looseness=6] (w)
-            	(w) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc5] {5}(w)
-            	(w) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc5] {5}(w)
-            	(w) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4}(w)
-
-            	(y) edge[loop, in = 75, out = 105, looseness=6, color=red] (y)
-            	(y) edge[loop, in = 72, out = 108, looseness=9] (y)
-            	(y) edge[loop, in = 70, out = 110, looseness=12] node[accset,collacc5] {5} (y)
-            	(y) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc5] {5}(y)
-
-			;
-		\end{tikzpicture}
-    \caption{TELA after simplification and merging in Level 1}
-    \label{fig:L1_middle}
-\end{figure}
-
-\subsection{Restoring equivalence}
-Replacing the input acceptance condition with $\psi'$ in $\mathcal{A}$ might change the language, so it is no longer equivalent to the language recognized by the original automaton $\mathcal{A}$. 
- By now, each SCC $S_{i}$ is adjusted to the acceptance condition $\psi_{i}$ which might be different than $\psi'$. 
- Therefore, for each SCC we need to restore equivalence with the new acceptance condition $\psi'$.
- The idea is that the only terms of $\psi'$ that can affect accepting runs on SCC $S_{i}$ are the ones that have been mapped on some terms in $\psi_{i}$.
-From now on, we refer to them as \emph{used terms}.
-We need to make sure that the other terms (the ones that have not been mapped on any term of $\psi_{i}$) can't affect any accepting run of $\mathcal{A}'$.
-From now on, we call these terms \emph{unused terms}.
-We influence the evaluation of the  \emph{unused terms} in the following way:
-\begin{itemize}
-     \item If we want to make term in the $\Fin$ form always $\false$, we place it's acceptance mark on every edge of the SCC.
-     \item If we want to make term in the $\Inf$ form always $\true$, we place it's acceptance mark on every edge of the SCC.
-\end{itemize}
- Keep in mind that by now, any acceptance mark that is not in $\psi'$ is not present on any edge of $\mathcal{A}'$.
-Therefore any $\Fin$ term is already always $\true$ if its acceptance mark is not present on any edge of the SCC.
-Dually, $\Inf$ term is already always $\false$ if it is not present on any edge of the SCC.
-\medskip
- Now we distinguish the cases when $\psi'$ is in DNF or $\psi'$ is in CNF. 
-\begin{itemize}
-     \item If $\psi'$ is in DNF, we need to make $\false $ every disjunct that does not contain any \emph{used term}.
-Further, every \emph{unused term} that occurs in the same disjunct as any \emph{used term} needs to be made $\true$.
-     \item Dually, if $\psi'$ is in CNF, every conjunct that does not contain any \emph{used term} needs to be made $\true$ and every \emph{unused term} that occurs in a conjunct with any \emph{used term} needs to be made $\false$.
-
-\end{itemize}
-
-\begin{example}
-	Consider the example from the Figure \ref{fig:L1_middle}, where the DNF acceptance condition $\psi' = (\Fin\tacc{5}{5} \lor \Inf\tacc{4}{4})$ and the acceptance condition of SCC $S_{1}$ $\psi_{1} = (\Fin\tacc{2}{2})$.
-The term $\Fin\tacc{5}{5}$ is mapped to $\Fin\tacc{2}{2}$.
-Then we need to enforce that the \emph{unused term} $\Inf\tacc{4}{4}$ has no effect on the evaluation of $\psi'$ by the run with the infinite suffix contained within the SCC $S_{1}$.
-Otherwise the result will not be equivalent (see the highlited loop in Figure \ref{fig:L1_middle}).
-Therefore we make term $\Inf\tacc{4}{4}$ $\true$ by adding $\tacc{4}{4}$ on every edge of the SCC $S_{i}$.
-The result can be seen in Figure \ref{fig:L1_done}.
-\end{example}
-
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[automaton]
-			\node[flabel] at (2,5) {$ \Fin\tacc{5}{5} \land \Inf\tacc{4}{4}$};
-
-			\node[state, initial] (x) at (0,0) {0};
-			\node[state] (y) at (2,2) {1};
-			\node[state] (z) at (2,-1.5) {2};
-			\node[state] (w) at (5,-1.5) {3};
-
-			\path[->]
-				(x) edge[]  (y)
-				(x) edge[]  (z)
-				(z) edge[] (w)
-
-            	(z) edge[loop, in = 75, out = 105, looseness=6] (z)
-            	(z) edge[loop, in = 72, out = 108, looseness=9] (z)
-            	(z) edge[loop, in = 70, out = 110, looseness=12]  (z)
-            	(z) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4} (z)
-
-            	(w) edge[loop, in = 75, out = 105, looseness=6] (w)
-            	(w) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc5] {5}(w)
-            	(w) edge[loop, in = 70, out = 110, looseness=12]  node[accset,collacc5] {5}(w)
-            	(w) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4] {4}(w)
-
-            	(y) edge[loop, in = 75, out = 105, looseness=6] node[accset,collacc4] {4} (y)
-            	(y) edge[loop, in = 72, out = 108, looseness=9] node[accset,collacc4] {4}(y)
-            	(y) edge[loop, in = 70, out = 110, looseness=12] node[accset,collacc4, pos=0.4] {4} node[accset,collacc5, pos=0.6] {5}(y)
-            	(y) edge[loop, in = 65, out = 115, looseness=14] node[accset,collacc4, pos=0.4] {4} node[accset,collacc5, pos=0.6] {5}(y)
-
-			;
-		\end{tikzpicture}
-    \caption{Equivalent TELA after simplification Level 1}
-    \label{fig:L1_done}
-\end{figure}
-
-%}}}
-
 %%% Local Variables:

### 2023-06-05 paper/main.tex
spell check

@@
   \caption{The left automaton accepts the words % $w\in\{a,b\}^\omega$
-    that contain infinitely many occurences of both $a$ and $b$. Each
+    that contain infinitely many occurrences of both $a$ and $b$. Each
     accepting run of the left automaton has to contain infinitely many
@@
 Now we describe the construction of the QBF in detail. The
-construction is parametrized by two integers $C,K\ge 0$, where $K$ is
+construction is parameterized by two integers $C,K\ge 0$, where $K$ is
 the desired number of acceptance marks and $C$ is the number of cubes
@@
   are provided by the tool \texttt{genltl} of the Spot
-  library~\cite{duret.16.atva2}~2.10.4. For parametrized formula
+  library~\cite{duret.16.atva2}~2.10.4. For parameterized formula
   patterns, we consider instances for all combinations of parameter
@@
 \texttt{ltl3tela} on automata coming from formulas from
-literature). This is caused by the QBF solver timeouts occuring
+literature). This is caused by the QBF solver timeouts occurring
 earlier as formulas constructed by the higher level are more
@@
     Each line shows the time ($y$ axis) needed by \telatko to process the
-    $x$-th automaton of the set, where automata in the set are ordered by
+    $x$-Th automaton of the set, where automata in the set are ordered by
     their processing time.

### 2023-06-05 paper/main.tex
fix of the spell check

@@
     Each line shows the time ($y$ axis) needed by \telatko to process the
-    $x$-Th automaton of the set, where automata in the set are ordered by
+    $x^\textrm{th}$ automaton of the set, where automata in the set are ordered by
     their processing time.

### 2023-06-05 paper/main.tex
Martin's suggestins reflected

@@
 library to simplify the formulas and remove duplicates and formulas
-equivalent to $\true$ and $\false$. After these steps, we had 348
-LTL formulas from literature and 335 random formulas.
+equivalent to $\true$ and $\false$. After these steps, we had 348~LTL
+formulas from literature and 335 random formulas.
 %
@@
 \texttt{ltl2dgra} (known as Rabinizer~4)~\cite{kretinsky.18.cav}. Some
-translators failed on some formulas: it usually reached a timeout of
-60 seconds or it produced an automaton that cannot be parsed by the
+translators failed on some formulas: they usually reached a timeout of
+60 seconds or produced an automaton that cannot be parsed by the
 Spot library.
@@
 
-To each automata set (that is, the automata with at least two marks
-produced by an individual LTL translator for one of the two formula
-sets), we have applied all single-level reductions and the multi-level
+To all automata,
+% To each automata set (that is, the automata with at least two marks
+% produced by an individual LTL translator for one of the two formula
+% sets),
+we have applied all single-level reductions and the multi-level
 reduction, always with incremental approach and without reducing the
@@
 experiments have been run on a computer with Intel$^\text{®}$ Core™ i7-8700
-processor and 32 GB of memory running Ubuntu 20.04.4. We used the tool
+processor and 32~GB of memory running Ubuntu 20.04.4. We used the tool
 \texttt{autcross} of the Spot library to get the statistics of the
@@
 complex. The automata sets produced by \texttt{ltl2dela} and
-\texttt{ltl2dgra}, on formulas from literature do not contain any
+\texttt{ltl2dgra} on formulas from literature do not contain any
 automaton where level~2 or level~3 achieves a better result than
@@
 
-Table~\ref{tab:heatmap} shows the number of automata from all five
-sets constructed from formulas from literature that have a given
-number of original acceptance marks and a given number of acceptance
-marks after multi-level reduction. The table indicates that in many
-cases only 1 or 2~marks can be saved. However, the achieved reduction
-is substantial for some automata with a higher number of original
+Table~\ref{tab:heatmap} shows the number of all automata constructed
+from formulas from literature that have a given number of original
+acceptance marks and a given number of acceptance marks after
+multi-level reduction. The table indicates that in many cases only 1
+or 2~marks can be saved. However, the achieved reduction is
+substantial for some automata with a higher number of original
 acceptance marks. For example, in 26 cases, we have reduced 7 or more
@@
 Table~\ref{tab:heatmap-random} shows the same information for all
-considered automata constructed from random formulas.
+automata constructed from random formulas.
 
@@
   \bigskip
-  \caption{Performance of \telatko on individual automata of each automata set.
+  \caption{Running times of \telatko on individual automata of each automata set.
     Automata sets constructed from formulas from literature are in the upper graph,

### 2023-06-05 paper/main.tex
last touch :-)

@@
 
-Table~\ref{tab:heatmap} shows the number of all automata constructed
-from formulas from literature that have a given number of original
-acceptance marks and a given number of acceptance marks after
-multi-level reduction. The table indicates that in many cases only 1
-or 2~marks can be saved. However, the achieved reduction is
-substantial for some automata with a higher number of original
-acceptance marks. For example, in 26 cases, we have reduced 7 or more
-acceptance marks to only 4 or less.
+Table~\ref{tab:heatmap} shows the effect of multi-level reduction to
+the number of acceptance marks in individual automata constructed from
+formulas from literature.
+% that have a given number of original
+% acceptance marks and a given number of acceptance marks after
+% multi-level reduction.
+The table indicates that in many cases only 1 or 2~marks can be
+saved. However, the achieved reduction is substantial for some
+automata with a higher number of original acceptance marks. For
+example, in 26 cases, we have reduced 7 or more acceptance marks to
+only 4 or less.
 % Moreover, there were many timeouts of the last QBF query for manof these automata.
@@
 % reduced 8 or more acceptance marks to only 4.
-Table~\ref{tab:heatmap-random} shows the same information for all
+Table~\ref{tab:heatmap-random} shows the same information for
 automata constructed from random formulas.
@@
 \begin{table}[t]
-  \caption{The effect of multi-level reduction on automata constructed
-    from formulas from literature.  A cell on coordinates $(x,y)$
-    contains the number of automata that have been reduced from $x$ to
-    $y$ acceptance marks. If the cell contains a sum of two numbers,
-    the latter represents the number of automata where the attempt to
-    reduce another mark has been unsuccessful due to a QBF solver
-    timeout.}
+  \caption{The effect of multi-level reduction on all considered
+    automata constructed from formulas from literature.  A cell on
+    coordinates $(x,y)$ contains the number of automata that have been
+    reduced from $x$ to $y$ acceptance marks. If the cell contains a
+    sum of two numbers, the latter represents the number of automata
+    where the attempt to reduce another mark has been unsuccessful due
+    to a QBF solver timeout.}
 \label{tab:heatmap}
@@
 \begin{table}[t]
-\caption{The effect of multi-level reduction on automata constructed
-    from random formulas. The meaning of each cell is the same as in Table~\ref{tab:heatmap}.}
+  \caption{The effect of multi-level reduction on all considered
+    automata constructed from random formulas. The meaning of each
+    cell is the same as in Table~\ref{tab:heatmap}.}
 \label{tab:heatmap-random}
